---
title: "Oklahoma Induced Seismicity: Causal ETAS Analysis"
subtitle: "Bivariate ETAS with Partition Sensitivity"
author: "PPDisentangle"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 7
    fig-dpi: 150
    self-contained: true
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    fig-width: 7
    fig-height: 5
    fig-dpi: 300
    keep-tex: true
    documentclass: article
    geometry:
      - margin=1in
    echo: false
    include-in-header:
      text: |
        \usepackage{longtable}[=v4.13]
execute:
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: false
library(PPDisentangle)
library(spatstat)
library(sf)
library(tigris)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(fig.align = "center", out.width = "100%")
`%||%` <- function(a, b) if (!is.null(a)) a else b

SCRIPT_DIR <- getwd()
DATA_DIR <- file.path(SCRIPT_DIR,
                       "oklahoma_induced_seismicity_data_regional20150318")
candidate_out_dirs <- unique(c(
  file.path(dirname(dirname(SCRIPT_DIR)), "output", "oklahoma"),
  file.path(dirname(dirname(SCRIPT_DIR)), "output"),
  file.path(SCRIPT_DIR, "output")
))
result_patterns <- c("^oklahoma_results(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_job([0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_sensitivity(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_sensitivity_job([0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_bootstrap(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_bootstrap_job([0-9]+)?\\.rds$")
candidate_result_files <- unlist(lapply(candidate_out_dirs, function(d) {
  if (!dir.exists(d)) return(character(0))
  unique(unlist(lapply(result_patterns, function(pat) {
    list.files(d, pattern = pat, full.names = TRUE)
  })))
}), use.names = FALSE)
existing_result_files <- candidate_result_files[file.exists(candidate_result_files)]
if (length(existing_result_files) == 0) {
  stop("No Oklahoma results RDS file found in expected output directories.")
}
result_meta <- data.frame(
  file = existing_result_files,
  base = basename(existing_result_files),
  mtime = as.POSIXct(file.info(existing_result_files)$mtime, origin = "1970-01-01"),
  stringsAsFactors = FALSE
)
result_meta$stage <- ifelse(
  grepl("pre_sensitivity", result_meta$base, fixed = TRUE), "pre_sensitivity",
  ifelse(grepl("pre_bootstrap", result_meta$base, fixed = TRUE), "pre_bootstrap", "full")
)
result_meta$stage_rank <- match(result_meta$stage, c("pre_sensitivity", "pre_bootstrap", "full"))

# Infer run id from suffixes so we can stay within the newest run:
# - preferred: _job<id>
# - fallback: trailing _<digits> before .rds
# - else: untagged
job_id <- sub(".*_job([0-9]+).*", "\\1", result_meta$base)
job_id[!grepl("_job[0-9]+", result_meta$base)] <- NA_character_
num_id <- sub(".*_([0-9]+)\\.rds$", "\\1", result_meta$base)
num_id[!grepl("_[0-9]+\\.rds$", result_meta$base)] <- NA_character_
result_meta$run_key <- ifelse(
  !is.na(job_id),
  paste0("job:", job_id),
  ifelse(!is.na(num_id), paste0("num:", num_id), "untagged")
)

run_latest_idx <- tapply(seq_len(nrow(result_meta)), result_meta$run_key, function(idx) {
  idx[which.max(result_meta$mtime[idx])]
})
run_latest_rows <- result_meta[as.integer(run_latest_idx), , drop = FALSE]
newest_run_key <- run_latest_rows$run_key[which.max(run_latest_rows$mtime)]

# Select highest available stage from the newest run (full > pre_bootstrap > pre_sensitivity).
run_pool <- result_meta[result_meta$run_key == newest_run_key, , drop = FALSE]
best_stage <- max(run_pool$stage_rank, na.rm = TRUE)
stage_pool <- run_pool[run_pool$stage_rank == best_stage, , drop = FALSE]
result_file <- stage_pool$file[which.max(stage_pool$mtime)]
OUT_DIR <- dirname(result_file)
candidate_plot_dirs <- unique(file.path(candidate_out_dirs, "plots"))

results <- readRDS(result_file)
meta_for_cfg <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
cfg <- if (!is.null(results$config)) results$config else list()
if (is.null(cfg$ETAS_M0) || length(cfg$ETAS_M0) == 0) {
  cfg$ETAS_M0 <- meta_for_cfg$design$min_mag
}
if (is.null(cfg$FIXED_STRUCTURAL)) cfg$FIXED_STRUCTURAL <- list()

# Older pre-bootstrap RDS files omitted pp_data/counties/full config; rebuild from CSV.
report_rebuilt <- NULL
if (is.null(results$pp_data) || is.null(results$counties)) {
  support_path <- file.path(SCRIPT_DIR, "oklahoma_report_support.R")
  if (!file.exists(support_path)) {
    stop("Missing ", support_path, " — cannot rebuild pp_data for this checkpoint.")
  }
  source(support_path, local = TRUE)
  report_rebuilt <- oklahoma_report_rebuild_pp_and_counties(DATA_DIR, etas_m0 = cfg$ETAS_M0)
  results$pp_data <- report_rebuilt$pp_data
  results$counties <- report_rebuilt$counties
}
cty <- results$counties
# Back-fill config keys expected by the narrative / tables (narrow legacy checkpoints).
if (is.null(cfg$n_counties) && !is.null(cty$n_counties)) cfg$n_counties <- cty$n_counties
if (is.null(cfg$n_treated) && !is.null(cty$n_treated)) cfg$n_treated <- cty$n_treated
pd0 <- results$pp_data
if (!is.null(pd0)) {
  if (is.null(cfg$n_pre) || (length(cfg$n_pre) == 1L && is.na(cfg$n_pre))) {
    cfg$n_pre <- nrow(as.data.frame(pd0$pp_pre))
  }
  if (is.null(cfg$n_post) || (length(cfg$n_post) == 1L && is.na(cfg$n_post))) {
    cfg$n_post <- nrow(as.data.frame(pd0$pp_post))
  }
  if ((is.null(cfg$n_pre_holdout) || (length(cfg$n_pre_holdout) == 1L && is.na(cfg$n_pre_holdout))) &&
      !is.null(pd0$pp_pre_holdout)) {
    cfg$n_pre_holdout <- nrow(as.data.frame(pd0$pp_pre_holdout))
  }
  if (is.null(cfg$n_pre_total) || (length(cfg$n_pre_total) == 1L && is.na(cfg$n_pre_total))) {
    cfg$n_pre_total <- (cfg$n_pre_holdout %||% 0L) + (cfg$n_pre %||% 0L)
  }
}
# Default ATE horizon if missing (days post-treatment in design)
if (is.null(cfg$ATE_WINDOW_DAYS) || (length(cfg$ATE_WINDOW_DAYS) == 1L && is.na(cfg$ATE_WINDOW_DAYS))) {
  t_star <- as.POSIXct(meta_for_cfg$design$t_star_utc, tz = "UTC")
  post_end <- as.POSIXct(meta_for_cfg$design$post_end_utc, tz = "UTC")
  cfg$ATE_WINDOW_DAYS <- as.integer(ceiling(as.numeric(difftime(post_end, t_star, units = "days"))))
}

using_pre_bootstrap <- isTRUE(identical(results$metadata$stage, "pre_bootstrap")) ||
  isTRUE(identical(results$metadata$stage, "pre_sensitivity")) ||
  (!is.null(results$checkpoint) && isTRUE(identical(results$checkpoint$stage, "pre_bootstrap"))) ||
  (!is.null(results$checkpoint) && isTRUE(identical(results$checkpoint$stage, "pre_sensitivity"))) ||
  grepl("pre_sensitivity", basename(result_file), fixed = TRUE) ||
  grepl("pre_bootstrap", basename(result_file), fixed = TRUE)

resolve_plot_file <- function(filename) {
  candidates <- file.path(candidate_plot_dirs, filename)
  existing <- candidates[file.exists(candidates)]
  if (length(existing) > 0) return(existing[which.max(file.info(existing)$mtime)])
  stem <- sub("\\.[^.]+$", "", filename)
  ext <- sub("^.*(\\.[^.]+)$", "\\1", filename)
  ext <- if (identical(ext, filename)) "" else ext
  tagged_pat <- paste0("^", gsub("([.|()\\^{}+$*?\\[\\]\\\\])", "\\\\\\1", stem), "_job[0-9]+", gsub("([.|()\\^{}+$*?\\[\\]\\\\])", "\\\\\\1", ext), "$")
  tagged <- unlist(lapply(candidate_plot_dirs, function(d) {
    if (!dir.exists(d)) return(character(0))
    list.files(d, pattern = tagged_pat, full.names = TRUE)
  }), use.names = FALSE)
  tagged <- tagged[file.exists(tagged)]
  if (length(tagged) == 0) return(NA_character_)
  tagged[which.max(file.info(tagged)$mtime)]
}

render_timestamp <- format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")

# KDE/config scalars may be NULL, list-wrapped, or non-numeric after saveRDS / merges.
fmt_scaled_num <- function(x, digits = 2) {
  if (is.null(x)) return("N/A")
  xv <- suppressWarnings(as.numeric(unlist(x, use.names = FALSE)))
  xv <- xv[is.finite(xv)]
  if (length(xv) < 1L) return("N/A")
  as.character(round(xv[[1]], digits))
}

has_fit <- function(fit_name) {
  !is.null(results[[fit_name]])
}
decode_enabled <- isTRUE(cfg$RUN_DECODE) || has_fit("fitG") || has_fit("fitH")

coerce_pp_df <- function(obj) {
  if (is.null(obj)) return(data.frame())

  # Handle spatstat point-pattern objects.
  if (inherits(obj, "ppp")) {
    out <- data.frame(x = obj$x, y = obj$y)
    if (!is.null(obj$marks)) {
      marks_df <- as.data.frame(obj$marks)
      out <- cbind(out, marks_df, stringsAsFactors = FALSE)
    }
    return(out)
  }

  # Standard data-frame-like objects.
  if (is.data.frame(obj)) return(as.data.frame(obj))
  if (is.matrix(obj)) return(as.data.frame(obj))

  # Common list structure with coordinate/time vectors.
  if (is.list(obj) && all(c("x", "y") %in% names(obj))) {
    n <- min(length(obj$x), length(obj$y))
    out <- data.frame(
      x = as.numeric(obj$x)[seq_len(n)],
      y = as.numeric(obj$y)[seq_len(n)]
    )
    if ("t" %in% names(obj)) out$t <- as.numeric(obj$t)[seq_len(min(n, length(obj$t)))]
    for (nm in setdiff(names(obj), c("x", "y", "t"))) {
      v <- obj[[nm]]
      if (length(v) >= n && is.atomic(v)) out[[nm]] <- v[seq_len(n)]
    }
    return(out)
  }

  # Last fallback.
  out <- tryCatch(as.data.frame(obj), error = function(e) NULL)
  if (is.null(out)) {
    stop("Could not coerce point-pattern object to data frame for plotting.")
  }
  out
}
```

# Introduction

**Report rendered at:** `r render_timestamp`

**Results source:** `r basename(result_file)` `r if (using_pre_bootstrap) "(pre-bootstrap checkpoint)" else "(full results)"`

On 18 March 2015 the Oklahoma Corporation Commission (OCC) issued
directive **AOI_20150318**, requiring operators within a defined Area of
Interest (AOI) in central Oklahoma to reduce wastewater disposal volumes.
This report estimates the **all-or-nothing causal effect** of the
directive on earthquake rates using the **PPDisentangle** framework.

## Estimand

The target quantity is the **all-or-nothing total treatment effect**:
the expected difference in total earthquake counts between
a world in which *every* county receives the directive
($\mathbf{Z}=(1,\dots,1)$) and a world in which *no* county receives it
($\mathbf{Z}=(0,\dots,0)$):

$$
\Delta_{\text{AoN}}
= \bigl\{
   \mathbb{E}\!\bigl[N(\mathbf{1})\bigr]
  -\mathbb{E}\!\bigl[N(\mathbf{0})\bigr]
\bigr\},
$$

where $N(\mathbf{z})$ denotes the total event count under treatment
assignment $\mathbf{z}$.

## Models

The analysis fits four bivariate ETAS models on the **county partition**:

| Label | Model | Background | Labels |
|:---:|:---|:---|:---|
| **B** | Naive bivariate ETAS | Homogeneous | Location-based |
| **D** | SEM bivariate ETAS | Homogeneous | SEM-corrected |
| **E** | Naive bivariate ETAS | KDE (non-parametric) | Location-based |
| **F** | SEM bivariate ETAS | KDE (non-parametric) | SEM-corrected |

Additionally, fits B and D are repeated under alternative **partitioning
schemes** to assess sensitivity to spatial resolution.

# Study Design

```{r design}
#| label: design
meta <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
```

**Treatment.** OCC directive AOI_20150318: mandatory reduction in
wastewater disposal volumes within the AOI.

**Tessellation.** The `r cty$n_counties` Oklahoma counties (US Census
Bureau via `tigris`). Counties whose centroid falls inside the AOI are
treated.

**Treated counties** (`r cty$n_treated`):
`r paste(cty$treated_names, collapse = ", ")`.

**Temporal windows.**

- Pre-treatment: `r meta$design$pre_start_utc` to `r meta$design$t_star_utc`
- Post-treatment: `r meta$design$t_star_utc` to `r meta$design$post_end_utc`

**Magnitude threshold.** $m_0 =$ `r cfg$ETAS_M0`.

**Event counts.**
- Pre-treatment total: `r if (!is.null(cfg$n_pre_total)) cfg$n_pre_total else cfg$n_pre`
- Pre-treatment holdout (first 50%): `r if (!is.null(cfg$n_pre_holdout)) cfg$n_pre_holdout else NA`
- Pre-treatment used for estimation/carryover (second 50%): `r cfg$n_pre`
- Post-treatment: `r cfg$n_post`
- Total used in SEM estimation (used pre + post): `r cfg$n_pre + cfg$n_post`

# Data

```{r load-data}
#| label: load-data
ev_all <- fread(file.path(DATA_DIR, "events_all.csv"))
t_star_utc <- as.POSIXct(meta$design$t_star_utc, tz = "UTC")
ev_all[, time_utc := as.POSIXct(time_utc, tz = "UTC")]
ev_all[, t_days := as.numeric(difftime(time_utc, t_star_utc, units = "days"))]
ev_all <- ev_all[mag >= cfg$ETAS_M0]
pp_pre  <- results$pp_data$pp_pre
pp_post <- results$pp_data$pp_post
pp_pre <- coerce_pp_df(pp_pre)
pp_post <- coerce_pp_df(pp_post)
if (!("t" %in% names(pp_pre))) pp_pre$t <- numeric(nrow(pp_pre))
if (!("t" %in% names(pp_post))) pp_post$t <- numeric(nrow(pp_post))
```

## Magnitude Distribution

```{r fig-mag}
#| label: fig-mag
#| fig-cap: "Magnitude distribution of events above the completeness threshold."
ggplot(ev_all, aes(x = mag)) +
  geom_histogram(binwidth = 0.1, fill = "#4292c6", color = "white",
                 boundary = cfg$ETAS_M0) +
  geom_vline(xintercept = cfg$ETAS_M0, linetype = "dashed", color = "red") +
  annotate("text", x = cfg$ETAS_M0 + 0.15, y = Inf, vjust = 2,
           label = paste0("m0 = ", cfg$ETAS_M0), color = "red", size = 4) +
  labs(x = "Magnitude", y = "Count") +
  theme_minimal(base_size = 13)
```

## Temporal Profile

```{r fig-temporal}
#| label: fig-temporal
#| fig-cap: "Cumulative event count; dashed line marks the OCC directive."
ev_all[, cum_n := seq_len(.N)]
ggplot(ev_all, aes(x = t_days, y = cum_n)) +
  geom_step(color = "#2171b5", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("text", x = 5, y = max(ev_all$cum_n) * 0.1,
           label = "OCC directive\n(18 Mar 2015)", color = "red", hjust = 0, size = 3.5) +
  labs(x = "Days since treatment", y = "Cumulative events") +
  theme_minimal(base_size = 13)
```

## County Partition

```{r fig-partition}
#| label: fig-partition
#| fig-cap: "Oklahoma county-level partition. Red counties are treated (centroid inside AOI)."
options(tigris_use_cache = TRUE)
counties_sf <- counties(state = "OK", cb = TRUE, year = 2022)
counties_sf <- st_transform(counties_sf, 5070)
counties_sf <- st_make_valid(counties_sf)

aoi_sf <- st_read(file.path(DATA_DIR, "occ_aoi_layer_2.geojson"), quiet = TRUE)
aoi_sf <- st_transform(aoi_sf, 5070) |> st_make_valid()
aoi_union <- st_union(aoi_sf)

county_centroids <- st_centroid(counties_sf)
inside_aoi <- lengths(st_within(county_centroids, aoi_union)) > 0
counties_sf$treatment <- ifelse(inside_aoi, "treated", "control")

ggplot() +
  geom_sf(data = counties_sf, aes(fill = treatment),
          color = "grey40", linewidth = 0.4) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    name = "Assignment") +
  geom_sf(data = aoi_sf, fill = NA, color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Oklahoma County Partition",
       subtitle = paste0("Treated: ", paste(cty$treated_names, collapse = ", "))) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 9))

counties_km <- counties_sf
st_geometry(counties_km) <- st_geometry(counties_sf) / 1000
st_crs(counties_km) <- NA
```

## Point Patterns

```{r fig-pp}
#| label: fig-pp
#| fig-cap: "Pre-treatment (left) and post-treatment (right) events overlaid on the county partition."
#| fig-height: 5
xs <- c(pp_pre$x, pp_post$x)
ys <- c(pp_pre$y, pp_post$y)
xs <- xs[is.finite(xs)]
ys <- ys[is.finite(ys)]
if (length(xs) == 0L || length(ys) == 0L) {
  bb_map <- sf::st_bbox(counties_km)
  xlims <- as.numeric(bb_map[c("xmin", "xmax")])
  ylims <- as.numeric(bb_map[c("ymin", "ymax")])
} else {
  xlims <- range(xs, na.rm = TRUE)
  ylims <- range(ys, na.rm = TRUE)
}

base_map <- ggplot() +
  geom_sf(data = counties_km, aes(fill = treatment),
          colour = "grey60", linewidth = 0.25) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    guide = "none") +
  coord_sf(xlim = xlims, ylim = ylims, datum = NA, expand = FALSE) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = "none")

p1 <- base_map +
  geom_point(data = pp_pre, aes(x = x, y = y, alpha = t),
             colour = "#2166ac", size = 0.8) +
  scale_alpha_continuous(range = c(0.15, 0.8), guide = "none") +
  labs(title = "Pre-treatment", x = "X (km)", y = "Y (km)")

pp_post_plot <- pp_post
process_col <- dplyr::case_when(
  "location_process" %in% names(pp_post_plot) ~ "location_process",
  "inferred_process" %in% names(pp_post_plot) ~ "inferred_process",
  "process" %in% names(pp_post_plot) ~ "process",
  TRUE ~ NA_character_
)
nr_post <- nrow(pp_post_plot)
if (nr_post == 0L) {
  pp_post_plot$Process <- factor(character(0), levels = c("control", "treated"))
} else if (is.na(process_col)) {
  pp_post_plot$Process <- factor(rep("control", nr_post), levels = c("control", "treated"))
} else {
  pp_post_plot$Process <- factor(as.character(pp_post_plot[[process_col]]),
                                 levels = c("control", "treated"))
}
p2 <- base_map +
  geom_point(data = pp_post_plot,
             aes(x = x, y = y, colour = Process, alpha = t),
             size = 0.8) +
  scale_colour_manual(values = c(control = "#2166ac", treated = "#b2182b"),
                      guide = "none") +
  scale_alpha_continuous(range = c(0.3, 1), guide = "none") +
  labs(title = "Post-treatment", x = "X (km)", y = "Y (km)")

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

# Model

## ETAS Conditional Intensity

$$
\lambda(t,x,y) = \frac{\mu}{|S|} + \sum_{j:\,t_j<t} A\,e^{\alpha_m(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

with Omori--Utsu temporal kernel $g$ and power-law spatial kernel $f$.

The kernels used in the fits are:

$$
g(\Delta t) = \frac{p-1}{c}\left(1+\frac{\Delta t}{c}\right)^{-p}, \qquad \Delta t > 0
$$

and

$$
f(\Delta x,\Delta y \mid m_j) =
\frac{q-1}{\pi\,d(m_j)}\,
\left(1+\frac{r^2}{d(m_j)}\right)^{-q},
\quad r^2 = \Delta x^2 + \Delta y^2,
\quad d(m_j)=D\exp\{\gamma(m_j-m_0)\}.
$$

## Bivariate Extension

$$
\lambda_k(t,x,y) = \frac{\mu_k}{|S_k|}\,W_k(x,y) + \sum_{l\in\{0,1\}}\sum_{\substack{j:\,t_j<t\\Z_j=l}} A_{kl}\,e^{\alpha_{m,kl}(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

Off-diagonal terms $A_{01}$ (treated$\to$control) and $A_{10}$
(control$\to$treated) capture cross-excitation. For the KDE background
models (E, F), $W_k(x,y)$ is the normalized kernel density estimate from
control pre-treatment data; for the homogeneous models (B, D),
$W_k \equiv 1$.

## Non-Parametric Background Rate

```{r tbl-kde}
#| label: tbl-kde
kde <- results$kde_info
if (is.null(kde)) {
  kde_df <- data.frame(
    Setting = "KDE summary",
    Value = paste0("Not stored in this results file. ",
                   if (using_pre_bootstrap) {
                     "Re-run oklahoma_analysis.R through the pre-bootstrap checkpoint (current script saves kde_info), or use full oklahoma_results.rds."
                   } else {
                     "Use a full results RDS that includes kde_info."
                   }),
    stringsAsFactors = FALSE
  )
} else {
  kde_holdout <- if (!is.null(kde$n_pre_holdout)) kde$n_pre_holdout else NA
  kde_used <- if (!is.null(kde$n_pre_used)) kde$n_pre_used else NA
  kde_df <- data.frame(
    Setting = c("KDE training events (held-out control pre)",
                "Held-out pre-treatment events (all processes)",
                "Pre-treatment events used for estimation",
                "Bandwidth (km)", "Estimated triggering range (km)", "Method"),
    Value = c(kde$n_training, kde_holdout, kde_used,
              fmt_scaled_num(kde$bw_sigma), fmt_scaled_num(kde$trigger_range_km),
              "Diggle cross-validation"))
}
kable(kde_df, align = "lc",
      caption = "KDE background rate estimation details.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

Only the **second 50%** of pre-treatment events are used for
estimation/carryover in both homogeneous and inhomogeneous SEM fits.
The **first 50%** is held out for fairness; in the inhomogeneous case,
its control events are used to estimate the KDE background via
`bw.diggle`.

```{r fig-kde-bg}
#| label: fig-kde-bg
#| fig-cap: "Estimated inhomogeneous background rate from held-out control pre-treatment events."
bg_plot_file <- resolve_plot_file("background_rate_kde.png")
if (file.exists(bg_plot_file)) {
  knitr::include_graphics(bg_plot_file)
}
```

## Fixed Parameters

```{r tbl-fixed}
#| label: tbl-fixed
fixed_cfg <- cfg$FIXED_STRUCTURAL
if (is.null(fixed_cfg) || length(fixed_cfg) == 0) {
  fixed_df <- data.frame(
    Parameter = "None",
    Value = "Estimated (not fixed)",
    Role = "All ETAS parameters are estimated from data",
    stringsAsFactors = FALSE
  )
  cap <- "Fixed structural parameters: none for this run."
} else {
  role_map <- c(
    c = "Omori onset delay",
    p = "Omori decay exponent",
    D = "Spatial scale",
    gamma = "Mag-spatial scaling",
    q = "Spatial decay exponent"
  )
  fixed_df <- data.frame(
    Parameter = names(fixed_cfg),
    Value = as.numeric(unlist(fixed_cfg)),
    Role = unname(role_map[names(fixed_cfg)]),
    stringsAsFactors = FALSE
  )
  cap <- "Fixed structural parameters."
}
kable(fixed_df, align = "lcl", caption = cap) |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## SEM Configuration

- **Outer iterations:** `r cfg$SEM_N_ITER`
- **Inner iterations:** `r cfg$SEM_INNER_ITER`
- **SEM inner proposals (`SEM_INNER_PROPS`):** `r if (!is.null(cfg$SEM_INNER_PROPS)) cfg$SEM_INNER_PROPS else cfg$SEM_N_LABELLINGS`
- **SEM labellings (`SEM_N_LABELLINGS`):** `r cfg$SEM_N_LABELLINGS`
- **Change factor:** `r cfg$SEM_CHANGE_FACTOR`

# Results — County Partition

```{r extract}
#| label: extract
get_fit_by_letter <- function(letter, legacy_key = NULL) {
  if (!is.null(results$fits_named) && !is.null(results$fits_named[[letter]])) {
    return(results$fits_named[[letter]])
  }
  if (!is.null(legacy_key) && !is.null(results[[legacy_key]])) return(results[[legacy_key]])
  NULL
}

fit_A_obj <- get_fit_by_letter("A", "fitB")
fit_B_obj <- get_fit_by_letter("B", "fitD")
fit_C_obj <- get_fit_by_letter("C", "fitE")
fit_D_obj <- get_fit_by_letter("D", "fitF")
fit_E_obj <- get_fit_by_letter("E")
fit_F_obj <- get_fit_by_letter("F")
fit_G_obj <- get_fit_by_letter("G")
fit_H_obj <- get_fit_by_letter("H")
fit_I_obj <- get_fit_by_letter("I", "fitG")
fit_J_obj <- get_fit_by_letter("J", "fitH")

A_par <- if (!is.null(fit_A_obj)) fit_A_obj$params else NULL
B_par <- if (!is.null(fit_B_obj)) fit_B_obj$params else NULL
C_par <- if (!is.null(fit_C_obj)) fit_C_obj$params else NULL
D_par <- if (!is.null(fit_D_obj)) fit_D_obj$params else NULL
E_par <- if (!is.null(fit_E_obj)) fit_E_obj$params else NULL
F_par <- if (!is.null(fit_F_obj)) fit_F_obj$params else NULL
G_par <- if (!is.null(fit_G_obj)) fit_G_obj$params else NULL
H_par <- if (!is.null(fit_H_obj)) fit_H_obj$params else NULL
I_par <- if (!is.null(fit_I_obj)) fit_I_obj$params else NULL
J_par <- if (!is.null(fit_J_obj)) fit_J_obj$params else NULL

bc <- function(par, p) if (p=="mu") par[["mu_0"]] else par[[paste0(p,"_00")]]
bt <- function(par, p) if (p=="mu") par[["mu_1"]] else par[[paste0(p,"_11")]]
safe_bc <- function(par, p) if (is.null(par)) NA_real_ else bc(par, p)
safe_bt <- function(par, p) if (is.null(par)) NA_real_ else bt(par, p)

model_defs <- list(
  list(id = "A", fit = "fits_named$A", label = "A: Naive", long = "A: Naive Biv", par = A_par),
  list(id = "B", fit = "fits_named$B", label = "B: SEM", long = "B: SEM Biv", par = B_par),
  list(id = "C", fit = "fits_named$C", label = "C: Naive+KDE (ctrl-fixed)", long = "C: Naive+KDE (ctrl-fixed)", par = C_par),
  list(id = "D", fit = "fits_named$D", label = "D: SEM+KDE (ctrl-fixed)", long = "D: SEM+KDE (ctrl-fixed)", par = D_par),
  list(id = "E", fit = "fits_named$E", label = "E: Naive+KDE (all-free)", long = "E: Naive+KDE (all-free)", par = E_par),
  list(id = "F", fit = "fits_named$F", label = "F: SEM+KDE (all-free)", long = "F: SEM+KDE (all-free)", par = F_par),
  list(id = "G", fit = "fits_named$G", label = "G: Naive+KDE (struct-fixed)", long = "G: Naive+KDE (struct-fixed)", par = G_par),
  list(id = "H", fit = "fits_named$H", label = "H: SEM+KDE (struct-fixed)", long = "H: SEM+KDE (struct-fixed)", par = H_par),
  list(id = "I", fit = "fits_named$I", label = "I: Decode", long = "I: Decode Biv", par = I_par),
  list(id = "J", fit = "fits_named$J", label = "J: Decode+KDE", long = "J: Decode+KDE", par = J_par)
)
param_models <- Filter(function(m) !is.null(m$par), model_defs)

fit_fill_colors <- c(
  "A: Naive Bivariate" = "#fc8d62",
  "B: SEM Bivariate" = "#8da0cb",
  "C: Naive Biv+KDE (ctrl-fixed)" = "#e78ac3",
  "D: SEM Biv+KDE (ctrl-fixed)" = "#a6d854",
  "E: Naive Biv+KDE (all-free)" = "#fdb462",
  "F: SEM Biv+KDE (all-free)" = "#80b1d3",
  "G: Naive Biv+KDE (struct-fixed)" = "#fb9a99",
  "H: SEM Biv+KDE (struct-fixed)" = "#b3de69",
  "I: Decode Bivariate" = "#66c2a5",
  "J: Decode Biv+KDE" = "#ffd92f"
)
fit_line_colors <- c(
  "A: Naive Bivariate" = "#d95f02",
  "B: SEM Bivariate" = "#7570b3",
  "C: Naive Biv+KDE (ctrl-fixed)" = "#c51b7d",
  "D: SEM Biv+KDE (ctrl-fixed)" = "#4d9221",
  "E: Naive Biv+KDE (all-free)" = "#e6550d",
  "F: SEM Biv+KDE (all-free)" = "#3182bd",
  "G: Naive Biv+KDE (struct-fixed)" = "#e31a1c",
  "H: SEM Biv+KDE (struct-fixed)" = "#33a02c",
  "I: Decode Bivariate" = "#1b9e77",
  "J: Decode Biv+KDE" = "#b8860b"
)

get_sem_metric_trace <- function(fit_obj) {
  if (is.null(fit_obj)) return(NULL)
  sem_obj <- if (!is.null(fit_obj$fit) && !is.null(fit_obj$fit$adaptive)) {
    fit_obj$fit
  } else if (!is.null(fit_obj$sem) && !is.null(fit_obj$sem$adaptive)) {
    fit_obj$sem
  } else {
    NULL
  }
  if (is.null(sem_obj) || is.null(sem_obj$adaptive)) return(NULL)
  ad <- sem_obj$adaptive
  if (!is.null(ad$metrics) && length(ad$metrics) > 0) {
    return(as.numeric(ad$metrics))
  }
  # Backward compatibility: reconstruct selected metric per iter as max over proposal metrics.
  if (!is.null(ad$all_metrics) && length(ad$all_metrics) > 0) {
    vals <- vapply(ad$all_metrics, function(v) {
      if (is.null(v) || length(v) == 0) return(NA_real_)
      max(as.numeric(v), na.rm = TRUE)
    }, numeric(1))
    vals[is.infinite(vals)] <- NA_real_
    return(vals)
  }
  NULL
}

# Lightweight wiring audit to ensure report labels map to the intended objects.
build_fit_wiring_audit <- function(results_obj) {
  defs <- data.frame(
    Label = c("A", "B", "C", "D", "E", "F", "G", "H"),
    `Result key` = c("fits_named$A", "fits_named$B", "fits_named$C", "fits_named$D",
                     "fits_named$E", "fits_named$F", "fits_named$G", "fits_named$H"),
    `Background` = c("Homogeneous", "Homogeneous", "KDE", "KDE", "KDE", "KDE", "KDE", "KDE"),
    `Labelling source` = c("Location", "SEM", "Location", "SEM", "Location", "SEM", "Location", "SEM"),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  defs$`Present in RDS` <- vapply(defs$Label, function(lbl) {
    !is.null(results_obj$fits_named) && !is.null(results_obj$fits_named[[lbl]])
  }, logical(1))
  defs
}
```

```{r tbl-fit-wiring-audit}
#| label: tbl-fit-wiring-audit
wiring_df <- build_fit_wiring_audit(results)
kable(wiring_df, align = "lcccc",
      caption = "Model wiring audit for the county-partition report tables/figures.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Marginal Parameters

```{r tbl-params}
#| label: tbl-params
beta_proxy <- suppressWarnings(as.numeric(unlist(cfg$BETA_GR, use.names = FALSE)))
beta_proxy <- beta_proxy[is.finite(beta_proxy)]
beta_proxy <- if (length(beta_proxy) > 0) beta_proxy[[1]] else 2.3

eta_proxy <- function(mu, A, alpha_m, beta_gr = beta_proxy) {
  if (!all(is.finite(c(mu, A, alpha_m, beta_gr))) || beta_gr <= 0) return(NA_real_)
  A * exp(alpha_m / beta_gr)
}

param_rows <- do.call(rbind, lapply(param_models, function(m) {
  mu_c <- bc(m$par, "mu"); A_c <- bc(m$par, "A"); a_c <- bc(m$par, "alpha_m")
  mu_t <- bt(m$par, "mu"); A_t <- bt(m$par, "A"); a_t <- bt(m$par, "alpha_m")
  eta_c <- eta_proxy(mu_c, A_c, a_c)
  eta_t <- eta_proxy(mu_t, A_t, a_t)
  rbind(
    data.frame(
      Fit = m$long, Process = "Control",
      mu = mu_c, A = A_c, alpha_m = a_c,
      branch_proxy = eta_c,
      mu_branch_proxy = if (is.finite(mu_c) && is.finite(eta_c)) mu_c * eta_c else NA_real_,
      stringsAsFactors = FALSE
    ),
    data.frame(
      Fit = m$long, Process = "Treated",
      mu = mu_t, A = A_t, alpha_m = a_t,
      branch_proxy = eta_t,
      mu_branch_proxy = if (is.finite(mu_t) && is.finite(eta_t)) mu_t * eta_t else NA_real_,
      stringsAsFactors = FALSE
    )
  )
}))

param_rows$Fit <- factor(param_rows$Fit, levels = vapply(param_models, function(m) m$long, character(1)))
param_rows <- param_rows[order(param_rows$Fit, param_rows$Process), , drop = FALSE]
rownames(param_rows) <- NULL
fit_end_rows <- which(param_rows$Process == "Treated")

tbl_k <- kable(
  param_rows,
  digits = 4,
  align = "llrrrrr",
  col.names = c("Fit", "Process", "mu", "A", "alpha_m", "Branch proxy (A*exp(alpha/beta))", "mu * branch proxy"),
  caption = paste0(
    "Marginal ETAS parameters and branching proxies by fit. ",
    "Double separators mark fit boundaries. ",
    "Branch proxy uses beta=", signif(beta_proxy, 4), "."
  ),
  booktabs = TRUE,
  linesep = "",
  hline_after = fit_end_rows
)

tbl_k |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE,
                font_size = 8, latex_options = c("scale_down", "hold_position")) |>
  row_spec(fit_end_rows, extra_css = "border-bottom: 3px double #444;")
```

## Cross-excitation

```{r tbl-cross}
#| label: tbl-cross
cn <- c("A_01","alpha_m_01","A_10","alpha_m_10")
cl <- c("$A_{01}$ (treat$\\to$ctrl)","$\\alpha_{m,01}$",
        "$A_{10}$ (ctrl$\\to$treat)","$\\alpha_{m,10}$")
cross_tbl <- data.frame(Parameter = cl)
for (m in param_models) {
  cross_tbl[[m$id]] <- sapply(cn, function(nm) m$par[[nm]])
}
rownames(cross_tbl) <- NULL
kable(cross_tbl, digits = 4, escape = FALSE,
      align = paste0("l", paste(rep("c", length(param_models)), collapse = "")),
      col.names = c("", vapply(param_models, function(m) m$label, character(1))),
      caption = "Cross-excitation parameters for available models.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Parameter Comparison

```{r fig-params}
#| label: fig-params
#| fig-cap: "Marginal parameter estimates across bivariate fits."
#| fig-height: 5
param_rows <- lapply(param_models, function(m) {
  rbind(
    data.frame(Fit = m$long, Process = "Control",
               mu = bc(m$par, "mu"), A = bc(m$par, "A"), alpha_m = bc(m$par, "alpha_m")),
    data.frame(Fit = m$long, Process = "Treated",
               mu = bt(m$par, "mu"), A = bt(m$par, "A"), alpha_m = bt(m$par, "alpha_m"))
  )
})
param_long <- do.call(rbind, param_rows)
pm <- pivot_longer(param_long, cols=c(mu,A,alpha_m), names_to="Parameter", values_to="Value")
pm <- pm[is.finite(pm$Value), , drop = FALSE]
pm$Parameter <- factor(pm$Parameter, levels=c("mu","A","alpha_m"), labels=c("mu","A","alpha[m]"))
pm$Fit <- factor(pm$Fit, levels = vapply(param_models, function(m) m$long, character(1)))
ggplot(pm, aes(x=Fit, y=Value, colour=Process)) +
  geom_point(size=3, position=position_dodge(0.4)) +
  facet_wrap(~Parameter, scales="free_y", labeller=label_parsed) +
  scale_colour_manual(values=c(Control="#377eb8",Treated="#e41a1c")) +
  labs(y="Estimate", x=NULL) +
  theme_minimal(base_size=13) +
  theme(axis.text.x=element_text(angle=30,hjust=1), strip.text=element_text(face="bold"), legend.position="top")
```

## SEM Convergence

```{r fig-flips}
#| label: fig-flips
#| fig-cap: "SEM label flips per relabelling iteration for all county SEM fits (raw and smoothed traces)."
#| fig-height: 4
build_flip_df <- function(avg_flips, max_flips) {
  if (is.null(avg_flips) || is.null(max_flips)) return(NULL)
  n <- min(length(avg_flips), length(max_flips))
  if (n == 0) return(NULL)
  df <- data.frame(
    iteration = seq_len(n),
    average_flips = as.numeric(avg_flips[seq_len(n)]),
    max_metric_flips = as.numeric(max_flips[seq_len(n)])
  )
  # Add smooth trend overlays so directional convergence is easier to inspect.
  if (n >= 5) {
    avg_low <- stats::lowess(df$iteration, df$average_flips, f = 0.15, iter = 1)
    max_low <- stats::lowess(df$iteration, df$max_metric_flips, f = 0.15, iter = 1)
    df$average_flips_smooth <- stats::approx(avg_low$x, avg_low$y, xout = df$iteration, rule = 2)$y
    df$max_metric_flips_smooth <- stats::approx(max_low$x, max_low$y, xout = df$iteration, rule = 2)$y
  } else {
    df$average_flips_smooth <- df$average_flips
    df$max_metric_flips_smooth <- df$max_metric_flips
  }
  df
}

plot_flip_df <- function(df, ttl) {
  if (is.null(df) || nrow(df) == 0) return(NULL)
  ggplot(df, aes(x = iteration)) +
    geom_line(aes(y = average_flips, color = "Average flips"), alpha = 0.45) +
    geom_line(aes(y = max_metric_flips, color = "Max metric flips"), alpha = 0.45) +
    geom_line(aes(y = average_flips_smooth, color = "Average flips (smoothed)"), linewidth = 1.05) +
    geom_line(aes(y = max_metric_flips_smooth, color = "Max metric flips (smoothed)"), linewidth = 1.05) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    labs(title = ttl, x = "Iteration", y = "Number of flips") +
    theme_minimal() +
    scale_color_manual(name = "Metric", values = c(
      "Average flips" = "#6baed6",
      "Max metric flips" = "#fb6a4a",
      "Average flips (smoothed)" = "#08519c",
      "Max metric flips (smoothed)" = "#a50f15"
    ))
}

plots <- list()
sem_fit_convergence <- list(
  list(obj = if (!is.null(fit_B_obj)) fit_B_obj$fit else NULL, lab = "B: SEM Biv"),
  list(obj = if (!is.null(fit_D_obj)) fit_D_obj$fit else NULL, lab = "D: SEM Biv+KDE (ctrl-fixed)"),
  list(obj = if (!is.null(fit_F_obj)) fit_F_obj$fit else NULL, lab = "F: SEM Biv+KDE (all-free)"),
  list(obj = if (!is.null(fit_H_obj)) fit_H_obj$fit else NULL, lab = "H: SEM Biv+KDE (struct-fixed)")
)
for (nm in sem_fit_convergence) {
  if (!is.null(nm$obj) && !is.null(nm$obj$adaptive)) {
    df <- build_flip_df(nm$obj$adaptive$average_flips, nm$obj$adaptive$max_metric_flips)
    p <- plot_flip_df(df, nm$lab)
    if (!is.null(p)) plots[[length(plots)+1]] <- p
  }
}
if (length(plots) >= 2) {
  ncol_use <- if (length(plots) >= 4) 2 else length(plots)
  do.call(gridExtra::grid.arrange, c(plots, ncol = ncol_use))
} else if (length(plots) == 1) {
  print(plots[[1]])
} else {
  plot.new()
  text(0.5, 0.5, "No SEM flip traces found for county SEM fits (B/D/F/H).")
}
```

```{r fig-sem-adaptive-loglik}
#| label: fig-sem-adaptive-loglik
#| fig-cap: "Adaptive SEM selected-proposal score progression (county SEM fits). This trace is stochastic and is not expected to be monotone."
#| fig-height: 4
sem_metric_rows <- list()
for (info in list(
  list(obj = fit_B_obj, lab = "B: SEM Biv"),
  list(obj = fit_D_obj, lab = "D: SEM Biv+KDE (ctrl-fixed)"),
  list(obj = fit_F_obj, lab = "F: SEM Biv+KDE (all-free)"),
  list(obj = fit_H_obj, lab = "H: SEM Biv+KDE (struct-fixed)")
)) {
  tr <- get_sem_metric_trace(info$obj)
  if (!is.null(tr) && length(tr) > 0) {
    sem_metric_rows[[length(sem_metric_rows) + 1]] <- data.frame(
      Iteration = seq_along(tr),
      LogLik = as.numeric(tr),
      Model = info$lab
    )
  }
}
if (length(sem_metric_rows) > 0) {
  sem_metric_df <- do.call(rbind, sem_metric_rows)
  sem_metric_df <- sem_metric_df[is.finite(sem_metric_df$LogLik), , drop = FALSE]
  sem_metric_df <- sem_metric_df |>
    dplyr::group_by(.data$Model) |>
    dplyr::arrange(.data$Iteration, .by_group = TRUE) |>
    dplyr::mutate(RunningBest = cummax(.data$LogLik)) |>
    dplyr::ungroup()
  ggplot(sem_metric_df, aes(x = Iteration, y = LogLik, colour = Model)) +
    geom_line(linewidth = 0.9) +
    geom_line(aes(y = RunningBest), linewidth = 0.7, linetype = "dashed", alpha = 0.7) +
    scale_colour_manual(values = c(
      "B: SEM Biv" = "#6a3d9a",
      "D: SEM Biv+KDE (ctrl-fixed)" = "#e31a1c",
      "F: SEM Biv+KDE (all-free)" = "#1f78b4",
      "H: SEM Biv+KDE (struct-fixed)" = "#ff7f00"
    )) +
    labs(x = "Adaptive SEM iteration", y = "Selected SEM score (solid) / running best (dashed)") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top")
} else {
  plot.new()
  text(0.5, 0.5, "No adaptive SEM log-likelihood trace found in this results file.")
}
```

```{r fig-selected-loglik}
#| label: fig-selected-loglik
#| fig-cap: "Decode selected log-likelihood progression across relabelling iterations (when Decode models are present)."
#| fig-height: 5
metric_rows <- list()
if (!is.null(fit_I_obj) && !is.null(fit_I_obj$decode) && !is.null(fit_I_obj$decode$trace$objective)) {
  metric_rows[[length(metric_rows) + 1]] <- data.frame(
    Iteration = seq_along(fit_I_obj$decode$trace$objective),
    Objective = as.numeric(fit_I_obj$decode$trace$objective),
    Model = "I: Decode Biv"
  )
}
if (!is.null(fit_J_obj) && !is.null(fit_J_obj$decode) && !is.null(fit_J_obj$decode$trace$objective)) {
  metric_rows[[length(metric_rows) + 1]] <- data.frame(
    Iteration = seq_along(fit_J_obj$decode$trace$objective),
    Objective = as.numeric(fit_J_obj$decode$trace$objective),
    Model = "J: Decode Biv+KDE"
  )
}
if (length(metric_rows) > 0) {
  metric_df <- do.call(rbind, metric_rows)
  metric_df <- metric_df[is.finite(metric_df$Objective), , drop = FALSE]
  ggplot(metric_df, aes(x = Iteration, y = Objective, colour = Model)) +
    geom_line(linewidth = 0.8, alpha = 0.9) +
    scale_colour_manual(values = c(
      "I: Decode Biv" = "#1b9e77",
      "J: Decode Biv+KDE" = "#b8860b"
    )) +
    labs(x = "Iteration", y = "Selected objective (log-likelihood scale)") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top")
}
```

```{r fig-decode-flips}
#| label: fig-decode-flips
#| fig-cap: "Decode relabelling totals (county models)."
decode_counts <- data.frame(
  Model = c("I: Decode Bivariate", "J: Decode Biv+KDE"),
  Flips = c(
    if (!is.null(fit_I_obj) && !is.null(fit_I_obj$decode$n_flips)) fit_I_obj$decode$n_flips else NA_real_,
    if (!is.null(fit_J_obj) && !is.null(fit_J_obj$decode$n_flips)) fit_J_obj$decode$n_flips else NA_real_
  )
)
decode_counts <- decode_counts[is.finite(decode_counts$Flips), , drop = FALSE]
if (nrow(decode_counts) > 0) {
  ggplot(decode_counts, aes(x = Model, y = Flips, fill = Model)) +
    geom_col(width = 0.6, alpha = 0.9) +
    scale_fill_manual(values = c("I: Decode Bivariate" = "#66c2a5",
                                 "J: Decode Biv+KDE" = "#ffd92f")) +
    labs(x = NULL, y = "Number of relabelled points") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 20, hjust = 1))
}
```

## SEM Relabelling: Confusion Matrices

```{r tbl-confusion}
#| label: tbl-confusion
build_cm <- function(sem_obj, label) {
  if (is.null(sem_obj)) return(NULL)
  lab <- sem_obj$adaptive$adaptive_labelling
  post <- lab[lab$t >= 0, ]
  if (is.null(post$location_process) || is.null(post$inferred_process)) return(NULL)
  tab <- table(Location = post$location_process, Inferred = post$inferred_process)
  as.data.frame.matrix(tab)
}

build_cm_from_labelling <- function(lab) {
  if (is.null(lab)) return(NULL)
  post <- lab[lab$t >= 0, ]
  if (is.null(post$location_process) || is.null(post$inferred_process)) return(NULL)
  tab <- table(Location = post$location_process, Inferred = post$inferred_process)
  as.data.frame.matrix(tab)
}

cm_D <- build_cm(if (!is.null(fit_B_obj)) fit_B_obj$fit else NULL, "B")
cm_F <- build_cm(if (!is.null(fit_D_obj)) fit_D_obj$fit else NULL, "D")
cm_H <- build_cm(if (!is.null(fit_F_obj)) fit_F_obj$fit else NULL, "F")
cm_K <- build_cm(if (!is.null(fit_H_obj)) fit_H_obj$fit else NULL, "H")
cm_G <- build_cm_from_labelling(if (!is.null(fit_I_obj) && !is.null(fit_I_obj$decode)) fit_I_obj$decode$labelling else NULL)
cm_J <- build_cm_from_labelling(if (!is.null(fit_J_obj) && !is.null(fit_J_obj$decode)) fit_J_obj$decode$labelling else NULL)
```

### Fit B — SEM Bivariate ETAS

```{r tbl-cm-D}
#| label: tbl-cm-D
if (!is.null(cm_D)) {
  n_post_D <- sum(cm_D)
  cm_D_pct <- round(cm_D / n_post_D * 100, 1)
  cm_display_D <- cm_D
  for (rn in rownames(cm_D)) for (cn in colnames(cm_D)) {
    cm_display_D[rn, cn] <- paste0(cm_D[rn, cn], "  (", cm_D_pct[rn, cn], "%)")
  }
  kable(cm_display_D,
        caption = paste0("Fit B: Location vs SEM-inferred label (n = ", n_post_D,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_D)))
}
```

### Fit D — SEM Bivariate + KDE Background

```{r tbl-cm-F}
#| label: tbl-cm-F
if (!is.null(cm_F)) {
  n_post_F <- sum(cm_F)
  cm_F_pct <- round(cm_F / n_post_F * 100, 1)
  cm_display_F <- cm_F
  for (rn in rownames(cm_F)) for (cn in colnames(cm_F)) {
    cm_display_F[rn, cn] <- paste0(cm_F[rn, cn], "  (", cm_F_pct[rn, cn], "%)")
  }
  kable(cm_display_F,
        caption = paste0("Fit D: Location vs SEM-inferred label (n = ", n_post_F,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_F)))
}
```

### Fit F — SEM Bivariate + KDE (All-Free)

```{r tbl-cm-F-allfree}
#| label: tbl-cm-F-allfree
if (!is.null(cm_H)) {
  n_post_H <- sum(cm_H)
  cm_H_pct <- round(cm_H / n_post_H * 100, 1)
  cm_display_H <- cm_H
  for (rn in rownames(cm_H)) for (cn in colnames(cm_H)) {
    cm_display_H[rn, cn] <- paste0(cm_H[rn, cn], "  (", cm_H_pct[rn, cn], "%)")
  }
  kable(cm_display_H,
        caption = paste0("Fit F: Location vs SEM-inferred label (n = ", n_post_H,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_H)))
}
```

### Fit H — SEM Bivariate + KDE (Structural Fixed)

```{r tbl-cm-H-struct}
#| label: tbl-cm-H-struct
if (!is.null(cm_K)) {
  n_post_K <- sum(cm_K)
  cm_K_pct <- round(cm_K / n_post_K * 100, 1)
  cm_display_K <- cm_K
  for (rn in rownames(cm_K)) for (cn in colnames(cm_K)) {
    cm_display_K[rn, cn] <- paste0(cm_K[rn, cn], "  (", cm_K_pct[rn, cn], "%)")
  }
  kable(cm_display_K,
        caption = paste0("Fit H: Location vs SEM-inferred label (n = ", n_post_K,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_K)))
}
```

### Fit I — Decode Bivariate ETAS

```{r tbl-cm-G}
#| label: tbl-cm-G
if (!is.null(cm_G)) {
  n_post_G <- sum(cm_G)
  cm_G_pct <- round(cm_G / n_post_G * 100, 1)
  cm_display_G <- cm_G
  for (rn in rownames(cm_G)) for (cn in colnames(cm_G)) {
    cm_display_G[rn, cn] <- paste0(cm_G[rn, cn], "  (", cm_G_pct[rn, cn], "%)")
  }
  kable(cm_display_G,
        caption = paste0("Fit I: Location vs Decode-inferred label (n = ", n_post_G,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_G)))
}
```

### Fit J — Decode Bivariate + KDE Background

```{r tbl-cm-H}
#| label: tbl-cm-H
if (!is.null(cm_J)) {
  n_post_H <- sum(cm_J)
  cm_H_pct <- round(cm_J / n_post_H * 100, 1)
  cm_display_H <- cm_J
  for (rn in rownames(cm_J)) for (cn in colnames(cm_J)) {
    cm_display_H[rn, cn] <- paste0(cm_J[rn, cn], "  (", cm_H_pct[rn, cn], "%)")
  }
  kable(cm_display_H,
        caption = paste0("Fit J: Location vs Decode-inferred label (n = ", n_post_H,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_J)))
}
```

### Relabelling Summary

```{r tbl-cm-summary}
#| label: tbl-cm-summary
sum_rows <- list()
for (info in list(list(cm = cm_D, lab = "B: SEM Bivariate"),
                  list(cm = cm_F, lab = "D: SEM Bivariate+KDE (ctrl-fixed)"),
                  list(cm = cm_H, lab = "F: SEM Bivariate+KDE (all-free)"),
                  list(cm = cm_K, lab = "H: SEM Bivariate+KDE (struct-fixed)"),
                  list(cm = cm_G, lab = "I: Decode Bivariate"),
                  list(cm = cm_J, lab = "J: Decode Biv+KDE"))) {
  cm <- info$cm
  if (is.null(cm)) next
  n_total <- sum(cm)
  ctrl_to_treat <- if ("treated" %in% colnames(cm) && "control" %in% rownames(cm)) cm["control","treated"] else 0
  treat_to_ctrl <- if ("control" %in% colnames(cm) && "treated" %in% rownames(cm)) cm["treated","control"] else 0
  n_flipped <- ctrl_to_treat + treat_to_ctrl
  sum_rows[[length(sum_rows)+1]] <- data.frame(
    Model = info$lab,
    `Post events` = n_total,
    `Ctrl to Treat` = ctrl_to_treat,
    `Treat to Ctrl` = treat_to_ctrl,
    `Total relabelled` = n_flipped,
    `Pct relabelled` = paste0(round(n_flipped / n_total * 100, 1), "%"),
    check.names = FALSE)
}
if (length(sum_rows) > 0) {
  kable(do.call(rbind, sum_rows),
        caption = "Summary of SEM relabelling across models.",
        align = "lccccc") |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
}
```

# All-or-Nothing Treatment Effect

## County Partition

```{r tbl-ate-biv}
#| label: tbl-ate-biv
n_cty <- cfg$n_counties
ate_days <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 365

biv_rows <- list(
  list(lab = "A: Naive Bivariate", ate = if (!is.null(fit_A_obj)) fit_A_obj$ate else NULL, louis = NULL),
  list(lab = "B: SEM Bivariate",   ate = if (!is.null(fit_B_obj)) fit_B_obj$ate else NULL, louis = if (!is.null(fit_B_obj)) fit_B_obj$louis else NULL),
  list(lab = "C: Naive Biv+KDE (ctrl-fixed)",   ate = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL, louis = NULL),
  list(lab = "D: SEM Biv+KDE (ctrl-fixed)",     ate = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL, louis = if (!is.null(fit_D_obj)) fit_D_obj$louis else NULL),
  list(lab = "E: Naive Biv+KDE (all-free)",     ate = if (!is.null(fit_E_obj)) fit_E_obj$ate else NULL, louis = NULL),
  list(lab = "F: SEM Biv+KDE (all-free)",       ate = if (!is.null(fit_F_obj)) fit_F_obj$ate else NULL, louis = if (!is.null(fit_F_obj)) fit_F_obj$louis else NULL),
  list(lab = "G: Naive Biv+KDE (struct-fixed)", ate = if (!is.null(fit_G_obj)) fit_G_obj$ate else NULL, louis = NULL),
  list(lab = "H: SEM Biv+KDE (struct-fixed)",   ate = if (!is.null(fit_H_obj)) fit_H_obj$ate else NULL, louis = if (!is.null(fit_H_obj)) fit_H_obj$louis else NULL))
if (!is.null(fit_I_obj) && !is.null(fit_I_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "I: Decode Bivariate", ate = fit_I_obj$ate, louis = NULL)
}
if (!is.null(fit_J_obj) && !is.null(fit_J_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "J: Decode Biv+KDE", ate = fit_J_obj$ate, louis = NULL)
}

boot_obj <- results$bootstrap_ate
boot_E_df <- NULL
boot_F_df <- NULL
if (!is.null(boot_obj) &&
    !is.null(boot_obj$fit_E) &&
    !is.null(boot_obj$fit_E$replicate_summary) &&
    nrow(boot_obj$fit_E$replicate_summary) > 0) {
  boot_E_df <- boot_obj$fit_E$replicate_summary
}
if (!is.null(boot_obj) &&
    !is.null(boot_obj$fit_F) &&
    !is.null(boot_obj$fit_F$replicate_summary) &&
    nrow(boot_obj$fit_F$replicate_summary) > 0) {
  boot_F_df <- boot_obj$fit_F$replicate_summary
}

get_boot_summary <- function(model_label) {
  vals <- NULL
  if (model_label == "C: Naive Biv+KDE (ctrl-fixed)" && !is.null(boot_E_df)) vals <- boot_E_df$ate_total_mean
  if (model_label == "D: SEM Biv+KDE (ctrl-fixed)" && !is.null(boot_F_df)) vals <- boot_F_df$ate_total_mean
  if (is.null(vals)) return(NULL)
  vals <- vals[is.finite(vals)]
  if (length(vals) < 2) return(NULL)
  list(
    sd = stats::sd(vals, na.rm = TRUE),
    lo = as.numeric(stats::quantile(vals, probs = 0.025, na.rm = TRUE)),
    hi = as.numeric(stats::quantile(vals, probs = 0.975, na.rm = TRUE)),
    n = length(vals)
  )
}

get_total_effect_sim <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  sim_df <- ate_obj$all_nothing_sim
  if ("total_saved" %in% names(sim_df)) return(sim_df$total_saved)
  if ("total_effect" %in% names(sim_df)) return(sim_df$total_effect)
  # Backward compatibility with old result files.
  sim_df$ATE * n_tiles
}

get_expected_totals <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  sim_df <- ate_obj$all_nothing_sim
  if (all(c("c_total", "t_total") %in% names(sim_df))) {
    return(list(c_total = sim_df$c_total, t_total = sim_df$t_total))
  }
  # Backward compatibility with old result files.
  list(c_total = sim_df$c_mean * n_tiles, t_total = sim_df$t_mean * n_tiles)
}

format_partition_label <- function(part_id) {
  if (is.na(part_id) || is.null(part_id)) return(part_id)
  if (part_id == "county") return("County")
  if (part_id == "aoi_region") return("AOI region")
  m <- regexec("^grid_([0-9.]+)R$", part_id)
  g <- regmatches(part_id, m)[[1]]
  if (length(g) == 2) {
    mult <- sub("\\.0$", "", g[2])
    return(paste0("Grid (", mult, "x triggering range)"))
  }
  part_id
}

ate_biv <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  expected <- get_expected_totals(r$ate, n_tiles = n_cty)
  sim_totals <- get_total_effect_sim(r$ate, n_tiles = n_cty)
  c_mean <- mean(expected$c_total, na.rm = TRUE)
  t_mean <- mean(expected$t_total, na.rm = TRUE)
  saved_diff_mean <- c_mean - t_mean
  sim_sd_total  <- sd(sim_totals, na.rm = TRUE)
  base_row <- data.frame(
    Model = r$lab,
    `E[N(control everywhere)]` = round(c_mean, 0),
    `E[N(actual treatment regime)]` = round(t_mean, 0),
    `E[N(treated everywhere)]` = if (!is.null(r$ate$analytic) && !is.null(r$ate$analytic$eta_treat)) {
      round(as.numeric(r$ate$analytic$eta_treat) * n_cty, 0)
    } else { NA_real_ },
    `Saved = control - actual` = round(saved_diff_mean, 0),
    `SD(saved)` = round(sim_sd_total, 0),
    `Spillover` = round(ifelse(!is.null(r$ate$total_spillover), r$ate$total_spillover, r$ate$ATE_spillover * n_cty), 0),
    check.names = FALSE
  )
  if (using_pre_bootstrap) return(base_row)
  boot_sum <- get_boot_summary(r$lab)
  boot_sd <- if (!is.null(boot_sum)) round(boot_sum$sd, 0) else NA
  boot_ci <- if (!is.null(boot_sum)) {
    sprintf("[%d, %d]", round(boot_sum$lo, 0), round(boot_sum$hi, 0))
  } else {
    "NA"
  }
  base_row$`Bootstrap SD` <- boot_sd
  base_row$`Bootstrap 95% CI` <- boot_ci
  base_row
}))
kable(ate_biv,
      col.names = if (using_pre_bootstrap) {
        c("Model", "E[N(control everywhere)]", "E[N(actual treatment regime)]", "E[N(treated everywhere)]",
          paste0("Saved = control - actual / ", n_cty, " counties / ", ate_days, "d"),
          "SD(saved)", paste0("Spillover / ", ate_days, "d"))
      } else {
        c("Model", "E[N(control everywhere)]", "E[N(actual treatment regime)]", "E[N(treated everywhere)]",
          paste0("Saved = control - actual / ", n_cty, " counties / ", ate_days, "d"),
          "SD(saved)", "Bootstrap SD", "Bootstrap 95% CI",
          paste0("Spillover / ", ate_days, "d"))
      },
      caption = if (using_pre_bootstrap) {
        paste0("All-or-nothing expected counts over ", ate_days, " days with saved shown as E[N(control everywhere)] - E[N(actual treatment regime)] (pre-bootstrap checkpoint).")
      } else {
        paste0("All-or-nothing expected counts over ", ate_days, " days with saved shown as E[N(control everywhere)] - E[N(actual treatment regime)].")
      },
      align = if (using_pre_bootstrap) "lcccccc" else "lcccccccc") |>
  kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE)
if (using_pre_bootstrap) {
  cat("\nBootstrap results are not shown because the newest available source is a pre-bootstrap checkpoint.\n")
} else if (!is.null(boot_E_df) || !is.null(boot_F_df)) {
  labs <- c()
  if (!is.null(boot_E_df)) labs <- c(labs, "C (Naive Biv+KDE ctrl-fixed)")
  if (!is.null(boot_F_df)) labs <- c(labs, "D (SEM Biv+KDE ctrl-fixed)")
  cat(sprintf("\nBootstrap uncertainty is available for: %s.\n", paste(labs, collapse = ", ")))
}
```

```{r fig-ate-biv}
#| label: fig-ate-biv
#| fig-cap: "Simulation distribution of the all-or-nothing total saved over the post-treatment horizon."
#| fig-height: 4
biv_sim <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  data.frame(Model = r$lab, TotalATE = get_total_effect_sim(r$ate, n_tiles = n_cty))
}))
if (!is.null(biv_sim) && nrow(biv_sim) > 0) {
  biv_means <- aggregate(TotalATE ~ Model, biv_sim, mean)
  model_levels <- unique(as.character(biv_sim$Model))
  fill_vals <- fit_fill_colors[names(fit_fill_colors) %in% model_levels]
  line_vals <- fit_line_colors[names(fit_line_colors) %in% model_levels]
  ggplot(biv_sim, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_vline(data = biv_means, aes(xintercept = TotalATE, colour = Model),
               linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
    scale_fill_manual(values = fill_vals) +
    scale_colour_manual(values = line_vals) +
    labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes saved /" ~ .(ate_days) ~ "days)"),
         y = "Simulation count",
         title = "All-or-Nothing Total Saved: Bivariate Models") +
    theme_minimal(base_size = 13) +
    theme(legend.position = "top",
          plot.title = element_text(hjust = 0.5, face = "bold"))
}
```

# Partition Sensitivity

The analysis is repeated under alternative spatial partitions to assess
sensitivity to spatial resolution:

- **Grid partitions** with cell diameters equal to 1, 2, and 5
  times the estimated triggering range
  (`r fmt_scaled_num(if (!is.null(results$kde_info)) results$kde_info$trigger_range_km else NULL)` km).
- **AOI region** partition: the AOI polygon defines one treated region,
  everything else is control (2 tiles total).

For each partition, both the naive bivariate + KDE (C-type) and
SEM bivariate + KDE (D-type) models are fitted.

```{r fig-partition-all}
#| label: fig-partition-all
#| fig-cap: "Partition maps for all partition schemes evaluated in the run."
partition_ids <- names(results$partition_results)
partition_plot_files <- character(0)
partition_plot_labels <- character(0)
for (pid in partition_ids) {
  pfile <- resolve_plot_file(paste0("partition_map_", pid, ".png"))
  if (file.exists(pfile)) {
    partition_plot_files <- c(partition_plot_files, pfile)
    partition_plot_labels <- c(partition_plot_labels, format_partition_label(pid))
  }
}
if (length(partition_plot_files) > 0) {
  knitr::kable(
    data.frame(`Panels (top to bottom)` = partition_plot_labels),
    align = "l"
  )
  knitr::include_graphics(partition_plot_files)
}
```

```{r tbl-partition-sensitivity}
#| label: tbl-partition-sensitivity
pr <- results$partition_results
ap <- results$ate_partitions

part_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (info$label == "county") {
    list(ate_E = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL,
         ate_F = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL)
  } else {
    ap[[pname]]
  }

  for (fit_type in list(
    list(lab = "Naive Biv+KDE", par = info$E_params, ate = ate_info$ate_E, sem_obj = NULL),
    list(lab = "SEM Biv+KDE",   par = info$F_params, ate = ate_info$ate_F, sem_obj = info$semF)
  )) {
    expected <- if (!is.null(fit_type$ate)) {
      get_expected_totals(fit_type$ate, n_tiles = info$n_tiles)
    } else { NULL }
    sim_total <- if (!is.null(fit_type$ate)) {
      round(mean(get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles), na.rm = TRUE), 0)
    } else { NA }
    n_relabel <- if (!is.null(fit_type$sem_obj) &&
                     !is.null(fit_type$sem_obj$adaptive$adaptive_labelling)) {
      lp <- fit_type$sem_obj$adaptive$adaptive_labelling
      lp <- lp[lp$t >= 0, , drop = FALSE]
      sum(lp$location_process != lp$inferred_process, na.rm = TRUE)
    } else { 0 }
    c_mean <- if (!is.null(expected)) round(mean(expected$c_total, na.rm = TRUE), 0) else NA
    t_mean <- if (!is.null(expected)) round(mean(expected$t_total, na.rm = TRUE), 0) else NA

    part_rows[[length(part_rows) + 1]] <- data.frame(
      PartitionID = info$label,
      Partition = format_partition_label(info$label),
      Model = fit_type$lab,
      Tiles = info$n_tiles,
      Treated = info$n_treated,
      `Relabelled points` = n_relabel,
      `E[N(0)]` = c_mean,
      `E[N(1)]` = t_mean,
      mu_0 = round(fit_type$par[["mu_0"]], 4),
      mu_1 = round(fit_type$par[["mu_1"]], 4),
      A_0 = round(fit_type$par[["A_00"]], 4),
      A_1 = round(fit_type$par[["A_11"]], 4),
      alpha_m_0 = round(fit_type$par[["alpha_m_00"]], 4),
      alpha_m_1 = round(fit_type$par[["alpha_m_11"]], 4),
      `Total Saved` = sim_total,
      check.names = FALSE)
  }
}

if (length(part_rows) > 0) {
  part_df <- do.call(rbind, part_rows)
  # Keep county first, then all remaining partitions alphabetically by ID.
  part_df$PartitionID <- factor(part_df$PartitionID,
                                levels = c("county",
                                           sort(setdiff(unique(as.character(part_df$PartitionID)),
                                                        "county"))))
  part_df <- part_df[order(part_df$PartitionID, part_df$Model), ]
  part_df$Partition <- factor(as.character(part_df$Partition),
                              levels = unique(as.character(part_df$Partition)))
  part_df_print <- subset(part_df, select = -PartitionID)
  kable(part_df_print, align = "llccccccccccccc",
        caption = paste0("Partition comparison over ", ate_days, " days: E[N(0)], E[N(1)], marginal ETAS parameters, and total saved.")) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                  font_size = 9, latex_options = c("scale_down", "hold_position"))
}
```

```{r fig-partition-params}
#| label: fig-partition-params
#| fig-cap: "Marginal ETAS parameters (mu, A, alpha_m) across partition schemes. County is highlighted to compare against grid-based partitions."
#| fig-height: 6
if (exists("part_df") && nrow(part_df) > 0) {
  plot_df <- part_df
  plot_df$is_county <- as.character(plot_df$Partition) == "County"

  param_long <- rbind(
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "control", mu = plot_df$mu_0, A = plot_df$A_0, alpha_m = plot_df$alpha_m_0),
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "treated", mu = plot_df$mu_1, A = plot_df$A_1, alpha_m = plot_df$alpha_m_1)
  )
  param_long <- pivot_longer(param_long, cols = c(mu, A, alpha_m),
                             names_to = "Parameter", values_to = "Value")
  param_long$Process <- factor(param_long$Process, levels = c("control", "treated"))

  param_long$Parameter <- factor(
    param_long$Parameter,
    levels = c("mu", "A", "alpha_m"),
    labels = c(expression(mu), expression(A), expression(alpha[m]))
  )

  ggplot(param_long, aes(x = Partition, y = Value, colour = Model, group = Model)) +
    geom_line(alpha = 0.45) +
    geom_point(aes(size = is_county), alpha = 0.9) +
    scale_size_manual(values = c(`FALSE` = 2.1, `TRUE` = 3.6), guide = "none") +
    scale_colour_manual(values = c("Naive Biv+KDE" = "#e78ac3", "SEM Biv+KDE" = "#a6d854")) +
    facet_grid(Parameter ~ Process, scales = "free_y", labeller = label_parsed) +
    labs(x = "Partition scheme", y = "Estimate", colour = "Model") +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      axis.text.x = element_text(angle = 30, hjust = 1),
      strip.text = element_text(face = "bold")
    )
}
```

```{r fig-partition-ate}
#| label: fig-partition-ate
#| fig-cap: "Partition-sensitivity total-saved distributions (same units and x-axis scale as Section 6.1)."
if (exists("biv_sim") && !is.null(biv_sim) && nrow(biv_sim) > 0) {
  biv_xlim <- range(biv_sim$TotalATE, na.rm = TRUE)
} else {
  biv_xlim <- NULL
}

part_sim_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (info$label == "county") {
    list(ate_E = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL,
         ate_F = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL)
  } else {
    ap[[pname]]
  }
  for (fit_type in list(list(lab = "Naive Biv+KDE", ate = ate_info$ate_E),
                        list(lab = "SEM Biv+KDE",   ate = ate_info$ate_F))) {
    if (!is.null(fit_type$ate) && !is.null(fit_type$ate$all_nothing_sim)) {
      part_sim_rows[[length(part_sim_rows) + 1]] <- data.frame(
        Partition = info$label,
        Model = fit_type$lab,
        TotalATE = get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles)
      )
    }
  }
}

if (length(part_sim_rows) > 0) {
  part_sim <- do.call(rbind, part_sim_rows)
  part_sim$PartitionID <- part_sim$Partition
  part_sim$Partition <- vapply(part_sim$Partition, format_partition_label, character(1))
  part_sim$PartitionID <- factor(part_sim$PartitionID,
                                 levels = c("county",
                                            sort(setdiff(unique(as.character(part_sim$PartitionID)),
                                                         "county"))))
  part_order <- unique(as.character(part_sim[order(part_sim$PartitionID), "Partition"]))
  part_sim$Partition <- factor(part_sim$Partition, levels = part_order)
  p_part <- ggplot(part_sim, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 25, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    facet_wrap(~Partition, scales = "fixed") +
    scale_fill_manual(values = c("Naive Biv+KDE" = "#e78ac3", "SEM Biv+KDE" = "#a6d854")) +
    labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes saved /" ~ .(ate_days) ~ "days)"),
         y = "Simulation count") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          strip.text = element_text(face = "bold"))
  if (!is.null(biv_xlim) && all(is.finite(biv_xlim))) {
    p_part <- p_part + coord_cartesian(xlim = biv_xlim)
  }
  print(p_part)
}
```

## KDE Bandwidth Sensitivity (County Inhomogeneous E/F)

County-level inhomogeneous fits are re-run across KDE bandwidths:
Diggle, Diggle$\\times$2, Diggle$\\times$5, and Diggle$\\times$10.

```{r tbl-kde-bandwidth-sensitivity}
#| label: tbl-kde-bandwidth-sensitivity
bws <- results$kde_bandwidth_sensitivity
if (!is.null(bws) && length(bws) > 0) {
  bw_rows <- list()
  bw_order <- c("diggle", "digglex2", "digglex5", "digglex10")
  bw_label <- function(x) {
    switch(as.character(x),
           diggle = "Diggle",
           digglex2 = "Diggle x2",
           digglex5 = "Diggle x5",
           digglex10 = "Diggle x10",
           x)
  }
  for (nm in names(bws)) {
    info <- bws[[nm]]
    if (is.null(info)) next
    for (fit_type in list(
      list(model = "Naive Biv+KDE", par = info$E_params, ate = info$ate_E, relab = 0),
      list(model = "SEM Biv+KDE",   par = info$F_params, ate = info$ate_F,
           relab = if (!is.null(info$n_relabel)) info$n_relabel else NA)
    )) {
      if (is.null(fit_type$par)) next
      expected <- if (!is.null(fit_type$ate)) get_expected_totals(fit_type$ate, n_tiles = n_cty) else NULL
      sim_total <- if (!is.null(fit_type$ate)) {
        round(mean(get_total_effect_sim(fit_type$ate, n_tiles = n_cty), na.rm = TRUE), 0)
      } else { NA }
      c_mean <- if (!is.null(expected)) round(mean(expected$c_total, na.rm = TRUE), 0) else NA
      t_mean <- if (!is.null(expected)) round(mean(expected$t_total, na.rm = TRUE), 0) else NA
      bw_rows[[length(bw_rows) + 1]] <- data.frame(
        BandwidthID = nm,
        Bandwidth = bw_label(nm),
        Model = fit_type$model,
        Tiles = n_cty,
        Treated = cfg$n_treated,
        `Relabelled points` = fit_type$relab,
        `E[N(0)]` = c_mean,
        `E[N(1)]` = t_mean,
        mu_0 = round(fit_type$par[["mu_0"]], 4),
        mu_1 = round(fit_type$par[["mu_1"]], 4),
        A_0 = round(fit_type$par[["A_00"]], 4),
        A_1 = round(fit_type$par[["A_11"]], 4),
        alpha_m_0 = round(fit_type$par[["alpha_m_00"]], 4),
        alpha_m_1 = round(fit_type$par[["alpha_m_11"]], 4),
        `Total Saved` = sim_total,
        check.names = FALSE
      )
    }
  }
  if (length(bw_rows) > 0) {
    bw_df <- do.call(rbind, bw_rows)
    bw_df$BandwidthID <- factor(bw_df$BandwidthID, levels = bw_order)
    bw_df <- bw_df[order(bw_df$BandwidthID, bw_df$Model), ]
    bw_df$Bandwidth <- factor(as.character(bw_df$Bandwidth),
                              levels = unique(as.character(bw_df$Bandwidth)))
    bw_print <- subset(bw_df, select = -BandwidthID)
    kable(bw_print, align = "llccccccccccccc",
          caption = paste0("County inhomogeneous KDE bandwidth sensitivity over ", ate_days,
                           " days: E[N(0)], E[N(1)], marginal ETAS parameters, and total saved.")) |>
      kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                    font_size = 9, latex_options = c("scale_down", "hold_position"))
  }
}
```

# Interpretation

```{r summaries, include=FALSE}
n_cty_interp <- cfg$n_counties
ate_days_interp <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 100

safe_mu_ratio <- function(par) {
  if (is.null(par)) return(NA_real_)
  denom <- safe_bc(par, "mu")
  numer <- safe_bt(par, "mu")
  if (!is.finite(denom) || is.na(denom) || abs(denom) < .Machine$double.eps) return(NA_real_)
  numer / denom
}
mu_ratio_C <- safe_mu_ratio(C_par)
mu_ratio_D <- safe_mu_ratio(D_par)
mu_ratio_E <- safe_mu_ratio(E_par)
mu_ratio_F <- safe_mu_ratio(F_par)
mu_ratio_G <- safe_mu_ratio(G_par)
mu_ratio_H <- safe_mu_ratio(H_par)

total_C_1yr <- if (!is.null(fit_C_obj) && !is.null(fit_C_obj$ate)) round(mean(get_total_effect_sim(fit_C_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_D_1yr <- if (!is.null(fit_D_obj) && !is.null(fit_D_obj$ate)) round(mean(get_total_effect_sim(fit_D_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_E_1yr <- if (!is.null(fit_E_obj) && !is.null(fit_E_obj$ate)) round(mean(get_total_effect_sim(fit_E_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_F_1yr <- if (!is.null(fit_F_obj) && !is.null(fit_F_obj$ate)) round(mean(get_total_effect_sim(fit_F_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_G_1yr <- if (!is.null(fit_G_obj) && !is.null(fit_G_obj$ate)) round(mean(get_total_effect_sim(fit_G_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_H_1yr <- if (!is.null(fit_H_obj) && !is.null(fit_H_obj$ate)) round(mean(get_total_effect_sim(fit_H_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_J_1yr <- if (!is.null(fit_J_obj) && !is.null(fit_J_obj$ate)) round(mean(get_total_effect_sim(fit_J_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA

effect_direction <- function(x, tol = 10) {
  if (is.na(x)) return("unavailable")
  if (x > tol) return("worse under treatment")
  if (x < -tol) return("beneficial under treatment")
  "close to no detectable difference"
}
```

**Inhomogeneous background rate focus.** Fits E and F use a
spatially inhomogeneous background estimated from the first half of
control pre-treatment data. This provides an independent estimate of
the baseline seismicity pattern without using post-treatment data.

**Background contrast (inhomogeneous variants).** The treated-to-control
background ratio $\mu_1/\mu_0$ is:
`r paste(sprintf("C=%.1fx, D=%.1fx, E=%.1fx, F=%.1fx, G=%.1fx, H=%.1fx",
                 mu_ratio_C, mu_ratio_D, mu_ratio_E, mu_ratio_F, mu_ratio_G, mu_ratio_H), collapse = "")`.

**All-or-nothing total saved (`r ate_days_interp`-day horizon).**

```{r interp-effects, results='asis'}
effect_lines <- c(
  paste0("- **Fit C (naive+KDE, ctrl-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_C_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit D (SEM+KDE, ctrl-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_D_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit E (naive+KDE, all-free):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_E_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit F (SEM+KDE, all-free):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_F_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit G (naive+KDE, struct-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_G_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit H (SEM+KDE, struct-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_H_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  {
    sem_totals <- c(total_D_1yr, total_F_1yr, total_H_1yr)
    sem_totals <- sem_totals[is.finite(sem_totals)]
    if (length(sem_totals) > 0L) {
      paste0("- **Interpretation:** Across KDE variants, SEM fits range from **",
             effect_direction(min(sem_totals)),
             "** to **",
             effect_direction(max(sem_totals)),
             "**.")
    } else {
      "- **Interpretation:** KDE SEM variant effects are unavailable in this results file."
    }
  }
)
if (!is.na(total_J_1yr)) {
  effect_lines <- c(effect_lines, paste0("- **Fit J (Decode+KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_J_1yr, "** total earthquakes saved / ", ate_days_interp, " days"))
}
cat(paste(effect_lines, collapse = "\n"), "\n")
```

**Partition sensitivity.** The alternative partition schemes (grids at
different scales and the AOI-as-single-region partition) show how the
treatment effect estimate varies with spatial resolution. Coarser
partitions aggregate more spatial heterogeneity, while finer grids
may better capture localized effects but have fewer events per cell.

# Configuration

```{r tbl-config}
#| label: tbl-config
scalar_chr <- function(x, default = "N/A") {
  if (is.null(x)) return(default)
  xv <- unlist(x, use.names = FALSE)
  if (length(xv) < 1L) return(default)
  x1 <- xv[[1]]
  if (is.na(x1)) return(default)
  as.character(x1)
}
ate_win <- scalar_chr(cfg$ATE_WINDOW_DAYS)
sem_inner_props <- scalar_chr(cfg$SEM_INNER_PROPS %||% cfg$SEM_N_LABELLINGS)
grid_vals <- suppressWarnings(as.numeric(unlist(cfg$grid_diameters, use.names = FALSE)))
grid_vals <- grid_vals[is.finite(grid_vals)]
grid_str <- if (length(grid_vals) > 0L) {
  paste(round(grid_vals, 2), "km", collapse = ", ")
} else { "N/A" }
cdf <- data.frame(
  Setting = c("Magnitude threshold","GR \u03b2","Counties (total / treated)",
              "SEM outer iters","SEM inner iters","SEM inner proposals","SEM labellings",
              "SEM change factor","Decode enabled","Decode iterations","Fitting window (days)",
              "ATE simulation window (days)","Simulation replicates",
              "Spatial scale D (km)", "Estimated triggering range (km)", "Grid diameters",
              "Test mode"),
  Value = c(scalar_chr(cfg$ETAS_M0), scalar_chr(cfg$BETA_GR),
            paste0(scalar_chr(cfg$n_counties), " / ", scalar_chr(cfg$n_treated)),
            scalar_chr(cfg$SEM_N_ITER), scalar_chr(cfg$SEM_INNER_ITER), sem_inner_props, scalar_chr(cfg$SEM_N_LABELLINGS),
            scalar_chr(cfg$SEM_CHANGE_FACTOR),
            scalar_chr(cfg$RUN_DECODE %||% decode_enabled),
            scalar_chr(cfg$DECODE_ITER),
            if (!is.null(cfg$windowT_post)) paste(unlist(cfg$windowT_post, use.names = FALSE), collapse=" \u2013 ") else "N/A",
            ate_win, scalar_chr(cfg$ATE_N_SIMS),
            fmt_scaled_num(cfg$spatial_scale_D),
            fmt_scaled_num(if (!is.null(results$kde_info)) results$kde_info$trigger_range_km else NULL),
            grid_str,
            scalar_chr(cfg$TEST_MODE)),
  stringsAsFactors = FALSE
)
kable(cdf, align="lc", caption="Analysis configuration.") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)
```

# References

- Ogata, Y. (1988). Statistical models for earthquake occurrences.
  *JASA*, 83(401), 9--27.
- Zhuang, J., Ogata, Y. & Vere-Jones, D. (2002). Stochastic
  declustering of space-time earthquake occurrences. *JASA*, 97(458),
  369--380.
- Helmstetter, A. & Sornette, D. (2003). Importance of direct and
  indirect triggered seismicity in the ETAS model of seismicity.
  *Geophysical Research Letters*, 30(11), 1576.
- Langenbruch, C. & Zoback, M. D. (2016). How will induced seismicity
  in Oklahoma respond to decreased saltwater injection rates? *Science
  Advances*, 2(11), e1601542.

---
*Report generated by PPDisentangle.*
