---
title: "Oklahoma Induced Seismicity: Causal ETAS Analysis"
subtitle: "Point estimates (pre-bootstrap) · univariate G–J counts + bivariate C/D primary"
author: "PPDisentangle"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 7
    fig-dpi: 150
    self-contained: true
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    fig-width: 7
    fig-height: 5
    fig-dpi: 300
    keep-tex: true
    documentclass: article
    geometry:
      - margin=1in
    echo: false
    include-in-header:
      text: |
        \usepackage{longtable}[=v4.13]
execute:
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: false
SCRIPT_DIR <- getwd()
REPO_ROOT <- if (basename(SCRIPT_DIR) == "oklahoma" &&
                  basename(dirname(SCRIPT_DIR)) == "inst") {
  normalizePath(file.path(SCRIPT_DIR, "..", ".."), winslash = "/", mustWork = FALSE)
} else {
  normalizePath(SCRIPT_DIR, winslash = "/", mustWork = FALSE)
}
# Prefer the local package source so report rendering matches the repo
# (installed PPDisentangle can lag behind and miss helpers like pp_output_path).
if (file.exists(file.path(REPO_ROOT, "DESCRIPTION")) &&
    requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(REPO_ROOT, quiet = TRUE, export_all = FALSE, helpers = FALSE, attach_testthat = FALSE)
} else if (requireNamespace("PPDisentangle", quietly = TRUE)) {
  library(PPDisentangle)
} else {
  stop("PPDisentangle is not available via local source or installed package.")
}
library(spatstat)
library(sf)
library(tigris)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(fig.align = "center", out.width = "100%")
`%||%` <- function(a, b) if (!is.null(a)) a else b

DATA_DIR <- file.path(SCRIPT_DIR,
                       "oklahoma_induced_seismicity_data_regional20150318")
candidate_out_dirs <- unique(c(
  PPDisentangle::pp_output_path("oklahoma", repo_root = REPO_ROOT)
))
explicit_result_file <- trimws(Sys.getenv("OK_RESULTS_FILE", ""))
if (nzchar(explicit_result_file) && !file.exists(explicit_result_file)) {
  stop("OK_RESULTS_FILE was provided but does not exist: ", explicit_result_file)
}
# Prefer the canonical publication snapshot when present. Scenario job RDS
# files (univ/obs/biv) can be newer but are not the default main-report source.
canonical_for_paper <- unlist(lapply(candidate_out_dirs, function(d) {
  fp <- file.path(d, "for_paper.rds")
  if (file.exists(fp)) fp else character(0)
}), use.names = FALSE)
result_patterns <- c("^oklahoma_results(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_job([0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_bootstrap(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_bootstrap_job([0-9]+)?\\.rds$",
                     "^for_paper\\.rds$")
candidate_result_files <- unlist(lapply(candidate_out_dirs, function(d) {
  if (!dir.exists(d)) return(character(0))
  unique(unlist(lapply(result_patterns, function(pat) {
    list.files(d, pattern = pat, full.names = TRUE)
  })))
}), use.names = FALSE)
existing_result_files <- candidate_result_files[file.exists(candidate_result_files)]
if (length(existing_result_files) == 0 && length(canonical_for_paper) == 0) {
  stop("No Oklahoma results RDS file found in expected output directories.")
}
result_file <- if (nzchar(explicit_result_file)) {
  normalizePath(explicit_result_file, winslash = "/", mustWork = TRUE)
} else if (length(canonical_for_paper) > 0) {
  normalizePath(canonical_for_paper[[1]], winslash = "/", mustWork = TRUE)
} else {
  result_meta <- data.frame(
    file = existing_result_files,
    base = basename(existing_result_files),
    mtime = as.POSIXct(file.info(existing_result_files)$mtime, origin = "1970-01-01"),
    stringsAsFactors = FALSE
  )
  result_meta$stage <- ifelse(
    grepl("pre_bootstrap", result_meta$base, fixed = TRUE), "pre_bootstrap", "full"
  )
  result_meta$stage_rank <- match(result_meta$stage, c("pre_bootstrap", "full"))
  job_id <- sub(".*_job([0-9]+).*", "\\1", result_meta$base)
  job_id[!grepl("_job[0-9]+", result_meta$base)] <- NA_character_
  result_meta$run_key <- ifelse(!is.na(job_id), paste0("job:", job_id), "untagged")
  run_latest_idx <- tapply(seq_len(nrow(result_meta)), result_meta$run_key, function(idx) {
    idx[which.max(result_meta$mtime[idx])]
  })
  run_latest_rows <- result_meta[as.integer(run_latest_idx), , drop = FALSE]
  newest_run_key <- run_latest_rows$run_key[which.max(run_latest_rows$mtime)]
  run_pool <- result_meta[result_meta$run_key == newest_run_key, , drop = FALSE]
  best_stage <- max(run_pool$stage_rank, na.rm = TRUE)
  stage_pool <- run_pool[run_pool$stage_rank == best_stage, , drop = FALSE]
  stage_pool$file[which.max(stage_pool$mtime)]
}
OUT_DIR <- dirname(result_file)
candidate_plot_dirs <- unique(file.path(candidate_out_dirs, "plots"))

results <- readRDS(result_file)

# Remap legacy A–L lettering (E/F = all_free primary, duplicate C/D slot, K/L univ homog)
# onto the current A–J scheme (C/D = all_free primary; no K/L).
normalize_oklahoma_fit_letters <- function(res) {
  fn <- res$fits_named
  if (is.null(fn) || !is.list(fn)) return(res)
  e_txt <- if (!is.null(fn$E)) {
    paste(c(fn$E$method, fn$E$label, fn$E$id), collapse = " ")
  } else {
    ""
  }
  e_is_all_free <- grepl("all[_ -]?free", e_txt, ignore.case = TRUE)
  old_scheme <- !is.null(fn$L) || (isTRUE(e_is_all_free) && !is.null(fn$C) && !is.null(fn$E))
  if (!isTRUE(old_scheme)) return(res)
  old <- fn
  remap_one <- function(obj, letter) {
    if (is.null(obj)) return(NULL)
    obj$letter <- letter
    obj
  }
  res$fits_named <- list(
    A = remap_one(old$A, "A"),
    B = remap_one(old$B, "B"),
    C = remap_one(old$E, "C"),
    D = remap_one(old$F, "D"),
    E = remap_one(old$G, "E"),
    F = remap_one(old$H, "F"),
    G = remap_one(old$I, "G"),
    H = remap_one(old$J, "H"),
    I = remap_one(old$K, "I"),
    J = remap_one(old$L, "J")
  )
  boot <- res$bootstrap_ate
  if (is.list(boot)) {
    if (!is.null(boot$fit_E) && is.null(boot$fit_C)) boot$fit_C <- boot$fit_E
    if (!is.null(boot$fit_F) && is.null(boot$fit_D)) boot$fit_D <- boot$fit_F
    res$bootstrap_ate <- boot
  }
  if (!is.null(res$fit_name_map)) {
    res$fit_name_map <- list(
      A = "fits_named$A", B = "fits_named$B",
      C = "fits_named$C", D = "fits_named$D",
      E = "fits_named$E", F = "fits_named$F",
      G = "fits_named$G", H = "fits_named$H",
      I = "fits_named$I", J = "fits_named$J"
    )
  }
  res
}
results <- normalize_oklahoma_fit_letters(results)

meta_for_cfg <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
cfg <- if (!is.null(results$config)) results$config else list()
if (is.null(cfg$ETAS_M0) || length(cfg$ETAS_M0) == 0) {
  cfg$ETAS_M0 <- meta_for_cfg$design$min_mag
}
if (is.null(cfg$FIXED_STRUCTURAL)) cfg$FIXED_STRUCTURAL <- list()

expected_main_inner <- suppressWarnings(as.integer(trimws(Sys.getenv("OK_SEM_INNER_ITER", ""))))
expected_sens_inner <- suppressWarnings(as.integer(trimws(Sys.getenv("OK_SENS_SEM_INNER_ITER", ""))))
expected_boot_inner <- suppressWarnings(as.integer(trimws(Sys.getenv("OK_BOOT_SEM_INNER_ITER", ""))))
if (is.finite(expected_main_inner) && !is.na(expected_main_inner) &&
    !identical(as.integer(cfg$SEM_INNER_ITER), as.integer(expected_main_inner))) {
  stop(sprintf(
    "Report/results mismatch: cfg$SEM_INNER_ITER=%s but OK_SEM_INNER_ITER=%s (file: %s)",
    as.character(cfg$SEM_INNER_ITER), as.character(expected_main_inner), result_file
  ))
}
if (is.finite(expected_sens_inner) && !is.na(expected_sens_inner) &&
    !identical(as.integer(cfg$SENS_SEM_INNER_ITER), as.integer(expected_sens_inner))) {
  stop(sprintf(
    "Report/results mismatch: cfg$SENS_SEM_INNER_ITER=%s but OK_SENS_SEM_INNER_ITER=%s (file: %s)",
    as.character(cfg$SENS_SEM_INNER_ITER), as.character(expected_sens_inner), result_file
  ))
}
if (is.finite(expected_boot_inner) && !is.na(expected_boot_inner) &&
    !identical(as.integer(cfg$BOOT_SEM_INNER_ITER), as.integer(expected_boot_inner))) {
  stop(sprintf(
    "Report/results mismatch: cfg$BOOT_SEM_INNER_ITER=%s but OK_BOOT_SEM_INNER_ITER=%s (file: %s)",
    as.character(cfg$BOOT_SEM_INNER_ITER), as.character(expected_boot_inner), result_file
  ))
}

# Older pre-bootstrap RDS files omitted pp_data/counties/full config; rebuild from CSV.
report_rebuilt <- NULL
if (is.null(results$pp_data) || is.null(results$counties)) {
  support_path <- file.path(SCRIPT_DIR, "oklahoma_report_support.R")
  if (!file.exists(support_path)) {
    stop("Missing ", support_path, " — cannot rebuild pp_data for this checkpoint.")
  }
  source(support_path, local = TRUE)
  report_rebuilt <- oklahoma_report_rebuild_pp_and_counties(DATA_DIR, etas_m0 = cfg$ETAS_M0)
  results$pp_data <- report_rebuilt$pp_data
  results$counties <- report_rebuilt$counties
}
cty <- results$counties
# Back-fill config keys expected by the narrative / tables (narrow legacy checkpoints).
if (is.null(cfg$n_counties) && !is.null(cty$n_counties)) cfg$n_counties <- cty$n_counties
if (is.null(cfg$n_treated_counties) && !is.null(cty$n_treated)) cfg$n_treated_counties <- cty$n_treated
if (is.null(cfg$PRIMARY_PARTITION) || !nzchar(as.character(cfg$PRIMARY_PARTITION[[1]]))) {
  cfg$PRIMARY_PARTITION <- "county"
}
if (is.null(cfg$n_tiles) || !is.finite(suppressWarnings(as.numeric(cfg$n_tiles[[1]])))) {
  cfg$n_tiles <- cfg$n_counties
}
if (is.null(cfg$n_treated) && !is.null(cty$n_treated)) cfg$n_treated <- cty$n_treated
primary_partition_id <- as.character(cfg$PRIMARY_PARTITION[[1]])
pd0 <- results$pp_data
if (!is.null(pd0)) {
  if (is.null(cfg$n_pre) || (length(cfg$n_pre) == 1L && is.na(cfg$n_pre))) {
    cfg$n_pre <- nrow(as.data.frame(pd0$pp_pre))
  }
  if (is.null(cfg$n_post) || (length(cfg$n_post) == 1L && is.na(cfg$n_post))) {
    cfg$n_post <- nrow(as.data.frame(pd0$pp_post))
  }
  if ((is.null(cfg$n_pre_holdout) || (length(cfg$n_pre_holdout) == 1L && is.na(cfg$n_pre_holdout))) &&
      !is.null(pd0$pp_pre_holdout)) {
    cfg$n_pre_holdout <- nrow(as.data.frame(pd0$pp_pre_holdout))
  }
  if (is.null(cfg$n_pre_total) || (length(cfg$n_pre_total) == 1L && is.na(cfg$n_pre_total))) {
    cfg$n_pre_total <- (cfg$n_pre_holdout %||% 0L) + (cfg$n_pre %||% 0L)
  }
}
# Default ATE horizon if missing (days post-treatment in design)
if (is.null(cfg$ATE_WINDOW_DAYS) || (length(cfg$ATE_WINDOW_DAYS) == 1L && is.na(cfg$ATE_WINDOW_DAYS))) {
  t_star <- as.POSIXct(meta_for_cfg$design$t_star_utc, tz = "UTC")
  post_end <- as.POSIXct(meta_for_cfg$design$post_end_utc, tz = "UTC")
  cfg$ATE_WINDOW_DAYS <- as.integer(ceiling(as.numeric(difftime(post_end, t_star, units = "days"))))
}

bootstrap_absent <- is.null(results$bootstrap_ate) ||
  (is.list(results$bootstrap_ate) &&
     is.null(results$bootstrap_ate$fit_C) &&
     is.null(results$bootstrap_ate$fit_D))
using_pre_bootstrap <- isTRUE(identical(results$metadata$stage, "pre_bootstrap")) ||
  isTRUE(identical(results$metadata$stage, "pre_sensitivity")) ||
  (!is.null(results$checkpoint) && isTRUE(identical(results$checkpoint$stage, "pre_bootstrap"))) ||
  (!is.null(results$checkpoint) && isTRUE(identical(results$checkpoint$stage, "pre_sensitivity"))) ||
  grepl("pre_sensitivity", basename(result_file), fixed = TRUE) ||
  grepl("pre_bootstrap", basename(result_file), fixed = TRUE) ||
  isTRUE(bootstrap_absent)

path_rel_to <- function(path, start) {
  path <- normalizePath(path, winslash = "/", mustWork = FALSE)
  start <- normalizePath(start, winslash = "/", mustWork = FALSE)
  path_seg <- strsplit(path, "/", fixed = TRUE)[[1L]]
  start_seg <- strsplit(start, "/", fixed = TRUE)[[1L]]
  path_seg <- path_seg[nzchar(path_seg)]
  start_seg <- start_seg[nzchar(start_seg)]
  n <- min(length(path_seg), length(start_seg))
  i <- 1L
  while (i <= n && identical(path_seg[[i]], start_seg[[i]])) i <- i + 1L
  rel_parts <- c(rep("..", length(start_seg) - i + 1L), path_seg[i:length(path_seg)])
  if (length(rel_parts) < 1L) "." else paste(rel_parts, collapse = "/")
}

resolve_plot_file <- function(filename) {
  plot_dirs <- unique(c(file.path(OUT_DIR, "plots"), candidate_plot_dirs))
  candidates <- file.path(plot_dirs, filename)
  existing <- candidates[file.exists(candidates)]
  if (length(existing) > 0) {
    abs_path <- existing[which.max(file.info(existing)$mtime)]
    return(path_rel_to(abs_path, SCRIPT_DIR))
  }
  stem <- sub("\\.[^.]+$", "", filename)
  ext <- sub("^.*(\\.[^.]+)$", "\\1", filename)
  ext <- if (identical(ext, filename)) "" else ext
  tagged_pat <- paste0("^", gsub("([.|()\\^{}+$*?\\[\\]\\\\])", "\\\\\\1", stem), "_job[0-9]+", gsub("([.|()\\^{}+$*?\\[\\]\\\\])", "\\\\\\1", ext), "$")
  tagged <- unlist(lapply(plot_dirs, function(d) {
    if (!dir.exists(d)) return(character(0))
    list.files(d, pattern = tagged_pat, full.names = TRUE)
  }), use.names = FALSE)
  tagged <- tagged[file.exists(tagged)]
  if (length(tagged) == 0) return(NA_character_)
  path_rel_to(tagged[which.max(file.info(tagged)$mtime)], SCRIPT_DIR)
}

render_timestamp <- format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")

# KDE/config scalars may be NULL, list-wrapped, or non-numeric after saveRDS / merges.
fmt_scaled_num <- function(x, digits = 2) {
  if (is.null(x)) return("N/A")
  xv <- suppressWarnings(as.numeric(unlist(x, use.names = FALSE)))
  xv <- xv[is.finite(xv)]
  if (length(xv) < 1L) return("N/A")
  as.character(round(xv[[1]], digits))
}

# NULL / none / non-positive SEM_T_TRUNC_DAYS means untruncated Omori (p > 1).
fmt_t_trunc_cfg <- function(x, digits = 4) {
  tryCatch(
    PPDisentangle:::.etas_format_t_trunc_days(x, digits = digits),
    error = function(e) "none"
  )
}

has_fit <- function(fit_name) {
  !is.null(results[[fit_name]])
}

coerce_pp_df <- function(obj) {
  if (is.null(obj)) return(data.frame())

  # Handle spatstat point-pattern objects.
  if (inherits(obj, "ppp")) {
    out <- data.frame(x = obj$x, y = obj$y)
    if (!is.null(obj$marks)) {
      marks_df <- as.data.frame(obj$marks)
      out <- cbind(out, marks_df, stringsAsFactors = FALSE)
    }
    return(out)
  }

  # Standard data-frame-like objects.
  if (is.data.frame(obj)) return(as.data.frame(obj))
  if (is.matrix(obj)) return(as.data.frame(obj))

  # Common list structure with coordinate/time vectors.
  if (is.list(obj) && all(c("x", "y") %in% names(obj))) {
    n <- min(length(obj$x), length(obj$y))
    out <- data.frame(
      x = as.numeric(obj$x)[seq_len(n)],
      y = as.numeric(obj$y)[seq_len(n)]
    )
    if ("t" %in% names(obj)) out$t <- as.numeric(obj$t)[seq_len(min(n, length(obj$t)))]
    for (nm in setdiff(names(obj), c("x", "y", "t"))) {
      v <- obj[[nm]]
      if (length(v) >= n && is.atomic(v)) out[[nm]] <- v[seq_len(n)]
    }
    return(out)
  }

  # Last fallback.
  out <- tryCatch(as.data.frame(obj), error = function(e) NULL)
  if (is.null(out)) {
    stop("Could not coerce point-pattern object to data frame for plotting.")
  }
  out
}
```

# Introduction

**Report rendered at:** `r render_timestamp`

**Results source:** `r basename(result_file)` `r if (using_pre_bootstrap) "(pre-bootstrap checkpoint)" else "(full results)"`

On 18 March 2015 the Oklahoma Corporation Commission (OCC) issued
directive **AOI_20150318**, requiring operators within a defined Area of
Interest (AOI) in central Oklahoma to reduce wastewater disposal volumes.
This report estimates the **all-or-nothing causal effect** of the
directive on earthquake rates using the **PPDisentangle** framework.
Counterfactual evaluations use the **full bivariate** ETAS law
(including cross-excitation).

## Estimand

The target quantity is the **all-or-nothing total treatment effect**
under the fitted bivariate law: the expected difference in total
earthquake counts between a world in which *every tile of the primary
tessellation* receives the directive ($\mathbf{Z}=(1,\dots,1)$) and a
world in which *no tile* receives it ($\mathbf{Z}=(0,\dots,0)$):

$$
\Delta_{\text{AoN}}
= \bigl\{
   \mathbb{E}\!\bigl[N(\mathbf{1})\bigr]
  -\mathbb{E}\!\bigl[N(\mathbf{0})\bigr]
\bigr\},
$$

where $N(\mathbf{z})$ denotes the total event count under treatment
assignment $\mathbf{z}$.

## Models

Primary tessellation (`r primary_partition_id`) publication pair:

| Label | Model | Background | Labels |
|:---:|:---|:---|:---|
| **C** | Naive bivariate ETAS | KDE (non-parametric) | Location-based |
| **D** | SEM bivariate ETAS | KDE (non-parametric) | SEM-corrected |

All fits fix $\gamma\equiv 0$ (magnitude-independent spatial scale $d(m)=D$).
```{r cd-only-scope}
#| echo: false
#| results: asis
if (isTRUE(cfg$CD_ONLY)) {
  cat(
    "This artifact is a **C/D-only** run: only the primary naive (C) and SEM (D) ",
    "KDE bivariate fits were estimated. Homogeneous A/B, structural E/F, ",
    "univariate G–J, bootstrap, and tessellation/bandwidth/t-trunc sensitivities ",
    "were not run. Sections below that depend on those outputs will say so.\n",
    sep = ""
  )
} else {
  cat(
    "**A/B** homogeneous bivariate (all-free except $\\gamma$); ",
    "**C/D** KDE all-free (primary); ",
    "**E/F** KDE with structural $(c,p,D,q)$ fixed from first-half control pre-treatment; ",
    "**G/H** univariate+KDE; **I/J** univariate homogeneous (no KDE). ",
    "Partition-sensitivity refits use the same C/D (all-free) specifications on ",
    "alternative tessellations.\n",
    sep = ""
  )
}
```

# Study Design

```{r design}
#| label: design
meta <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
```

**Treatment.** OCC directive AOI_20150318: mandatory reduction in
wastewater disposal volumes within the AOI.

**Primary tessellation.** `r primary_partition_id`
(`r cfg$n_tiles` tiles, `r cfg$n_treated` treated).
Default is a 1× triggering-range grid whose treated tiles are those whose
centroid falls inside the OCC AOI.

**County tessellation (sensitivity).** The `r cty$n_counties` Oklahoma
counties (US Census Bureau via `tigris`). Counties whose centroid falls
inside the AOI are treated.

**Treated counties** (`r cty$n_treated`):
`r paste(cty$treated_names, collapse = ", ")`.

**Temporal windows.**

- Pre-treatment: `r meta$design$pre_start_utc` to `r meta$design$t_star_utc`
- Post-treatment: `r meta$design$t_star_utc` to `r meta$design$post_end_utc`

**Magnitude threshold.** $m_0 =$ `r cfg$ETAS_M0`.

**Event counts.**
- Pre-treatment total: `r if (!is.null(cfg$n_pre_total)) cfg$n_pre_total else cfg$n_pre`
- Pre-treatment holdout (first 50%): `r if (!is.null(cfg$n_pre_holdout)) cfg$n_pre_holdout else NA`
- Pre-treatment used for estimation/carryover (second 50%): `r cfg$n_pre`
- Post-treatment: `r cfg$n_post`
- Total used in SEM estimation (used pre + post): `r cfg$n_pre + cfg$n_post`

# Data

```{r load-data}
#| label: load-data
ev_all <- fread(file.path(DATA_DIR, "events_all.csv"))
t_star_utc <- as.POSIXct(meta$design$t_star_utc, tz = "UTC")
ev_all[, time_utc := as.POSIXct(time_utc, tz = "UTC")]
ev_all[, t_days := as.numeric(difftime(time_utc, t_star_utc, units = "days"))]
ev_all <- ev_all[mag >= cfg$ETAS_M0]
pp_pre  <- results$pp_data$pp_pre
pp_post <- results$pp_data$pp_post
pp_pre <- coerce_pp_df(pp_pre)
pp_post <- coerce_pp_df(pp_post)
if (!("t" %in% names(pp_pre))) pp_pre$t <- numeric(nrow(pp_pre))
if (!("t" %in% names(pp_post))) pp_post$t <- numeric(nrow(pp_post))
```

## Magnitude Distribution

```{r fig-mag}
#| label: fig-mag
#| fig-cap: "Magnitude distribution of events above the completeness threshold."
ggplot(ev_all, aes(x = mag)) +
  geom_histogram(binwidth = 0.1, fill = "#4292c6", color = "white",
                 boundary = cfg$ETAS_M0) +
  geom_vline(xintercept = cfg$ETAS_M0, linetype = "dashed", color = "red") +
  annotate("text", x = cfg$ETAS_M0 + 0.15, y = Inf, vjust = 2,
           label = paste0("m0 = ", cfg$ETAS_M0), color = "red", size = 4) +
  labs(x = "Magnitude", y = "Count") +
  theme_minimal(base_size = 13)
```

## Temporal Profile

```{r fig-temporal}
#| label: fig-temporal
#| fig-cap: "Cumulative event count; dashed line marks the OCC directive."
ev_all[, cum_n := seq_len(.N)]
ggplot(ev_all, aes(x = t_days, y = cum_n)) +
  geom_step(color = "#2171b5", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("text", x = 5, y = max(ev_all$cum_n) * 0.1,
           label = "OCC directive\n(18 Mar 2015)", color = "red", hjust = 0, size = 3.5) +
  labs(x = "Days since treatment", y = "Cumulative events") +
  theme_minimal(base_size = 13)
```

## Tessellation Partition

```{r fig-partition}
#| label: fig-partition
#| fig-cap: "Primary tessellation used for C/D. Red tiles are treated (centroid inside the AOI). The dashed outline is the OCC AOI."
options(tigris_use_cache = TRUE)
counties_sf <- counties(state = "OK", cb = TRUE, year = 2022)
counties_sf <- st_transform(counties_sf, 5070)
counties_sf <- st_make_valid(counties_sf)

aoi_sf <- st_read(file.path(DATA_DIR, "occ_aoi_layer_2.geojson"), quiet = TRUE)
aoi_sf <- st_transform(aoi_sf, 5070) |> st_make_valid()
aoi_union <- st_union(aoi_sf)

county_centroids <- st_centroid(counties_sf)
inside_aoi <- lengths(st_within(county_centroids, aoi_union)) > 0
counties_sf$treatment <- ifelse(inside_aoi, "treated", "control")

counties_km <- counties_sf
st_geometry(counties_km) <- st_geometry(counties_sf) / 1000
st_crs(counties_km) <- NA
aoi_km <- aoi_sf
st_geometry(aoi_km) <- st_geometry(aoi_sf) / 1000
st_crs(aoi_km) <- NA

geom_path <- file.path(SCRIPT_DIR, "oklahoma_geometry.R")
if (file.exists(geom_path) && !exists("oklahoma_sf_to_owin_km", mode = "function")) {
  source(geom_path, local = FALSE)
}
support_path <- file.path(SCRIPT_DIR, "oklahoma_report_support.R")
if (file.exists(support_path) && !exists("oklahoma_report_grid_partition", mode = "function")) {
  source(support_path, local = FALSE)
}

ok_boundary <- st_union(counties_sf)
ok_boundary <- st_make_valid(ok_boundary)
win_km_part <- if (exists("oklahoma_sf_to_owin_km", mode = "function")) {
  oklahoma_sf_to_owin_km(ok_boundary)
} else {
  NULL
}
aoi_owin_part <- if (exists("oklahoma_sf_to_owin_km", mode = "function")) {
  tryCatch(oklahoma_sf_to_owin_km(aoi_union), error = function(e) NULL)
} else {
  NULL
}
trigger_km <- suppressWarnings(as.numeric(results$kde_info$trigger_range_km))
if (!is.finite(trigger_km) || trigger_km <= 0) {
  trigger_km <- suppressWarnings(as.numeric(cfg$spatial_scale_D))
}
grid_max_primary <- if (isTRUE(cfg$TEST_MODE)) 2000L else 5000L
primary_tess_kind <- "sf"
primary_tess_tiles <- counties_km
primary_tess_label <- if (exists("format_partition_label", mode = "function")) {
  format_partition_label(primary_partition_id)
} else if (identical(primary_partition_id, "grid_1.0R")) {
  "Grid (1x triggering range)"
} else {
  primary_partition_id
}
if (!identical(primary_partition_id, "county") &&
    exists("oklahoma_report_grid_partition", mode = "function") &&
    !is.null(win_km_part) && !is.null(aoi_owin_part)) {
  grid_diam <- if (identical(primary_partition_id, "grid_coarse")) {
    NA_real_
  } else {
    mult <- suppressWarnings(as.numeric(sub("^grid_([0-9.]+)R$", "\\1", primary_partition_id)))
    if (is.finite(mult) && is.finite(trigger_km)) mult * trigger_km else trigger_km
  }
  gp <- oklahoma_report_grid_partition(
    diameter = grid_diam,
    win = win_km_part,
    aoi_owin = aoi_owin_part,
    label = primary_partition_id,
    max_tiles = grid_max_primary
  )
  primary_tess_kind <- "rect"
  primary_tess_tiles <- oklahoma_report_tess_rects(gp$partition, gp$processes)
}

n_tess <- if (identical(primary_tess_kind, "rect")) {
  nrow(primary_tess_tiles)
} else {
  nrow(primary_tess_tiles)
}
n_tess_treated <- if (identical(primary_tess_kind, "rect")) {
  sum(primary_tess_tiles$treatment == "treated", na.rm = TRUE)
} else {
  sum(as.character(primary_tess_tiles$treatment) == "treated", na.rm = TRUE)
}
tess_subtitle <- sprintf(
  "%s: %d tiles (%d treated). Dashed line = AOI.",
  primary_tess_label, n_tess, n_tess_treated
)

if (identical(primary_tess_kind, "rect")) {
  ggplot() +
    geom_rect(
      data = primary_tess_tiles,
      aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = treatment),
      colour = "grey70", linewidth = 0.05
    ) +
    geom_sf(data = aoi_km, fill = NA, color = "red", linewidth = 0.7, linetype = "dashed") +
    scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                      name = "Assignment") +
    coord_sf(datum = NA, expand = FALSE) +
    labs(title = "Oklahoma Primary Tessellation",
         subtitle = tess_subtitle, x = "X (km)", y = "Y (km)") +
    theme_minimal(base_size = 13) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, size = 9))
} else {
  ggplot() +
    geom_sf(data = primary_tess_tiles, aes(fill = treatment),
            color = "grey40", linewidth = 0.4) +
    scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                      name = "Assignment") +
    geom_sf(data = aoi_km, fill = NA, color = "red", linewidth = 0.7, linetype = "dashed") +
    labs(title = "Oklahoma Primary Tessellation",
         subtitle = tess_subtitle, x = "X (km)", y = "Y (km)") +
    theme_minimal(base_size = 13) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, size = 9))
}
```

## Point Patterns

```{r fig-pp}
#| label: fig-pp
#| fig-cap: "Pre-treatment (left) and post-treatment (right) events overlaid on the primary tessellation."
#| fig-height: 5
xs <- c(pp_pre$x, pp_post$x)
ys <- c(pp_pre$y, pp_post$y)
xs <- xs[is.finite(xs)]
ys <- ys[is.finite(ys)]
if (length(xs) == 0L || length(ys) == 0L) {
  bb_map <- sf::st_bbox(counties_km)
  xlims <- as.numeric(bb_map[c("xmin", "xmax")])
  ylims <- as.numeric(bb_map[c("ymin", "ymax")])
} else {
  xlims <- range(xs, na.rm = TRUE)
  ylims <- range(ys, na.rm = TRUE)
}

if (exists("primary_tess_kind") && identical(primary_tess_kind, "rect") &&
    exists("primary_tess_tiles")) {
  base_map <- ggplot() +
    geom_rect(
      data = primary_tess_tiles,
      aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = treatment),
      colour = "grey70", linewidth = 0.05
    ) +
    scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                      guide = "none") +
    coord_sf(xlim = xlims, ylim = ylims, datum = NA, expand = FALSE) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          legend.position = "none")
} else {
  base_map <- ggplot() +
    geom_sf(data = counties_km, aes(fill = treatment),
            colour = "grey60", linewidth = 0.25) +
    scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                      guide = "none") +
    coord_sf(xlim = xlims, ylim = ylims, datum = NA, expand = FALSE) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          legend.position = "none")
}

p1 <- base_map +
  geom_point(data = pp_pre, aes(x = x, y = y, alpha = t),
             colour = "#2166ac", size = 0.8) +
  scale_alpha_continuous(range = c(0.15, 0.8), guide = "none") +
  labs(title = "Pre-treatment", x = "X (km)", y = "Y (km)")

pp_post_plot <- pp_post
process_col <- dplyr::case_when(
  "location_process" %in% names(pp_post_plot) ~ "location_process",
  "inferred_process" %in% names(pp_post_plot) ~ "inferred_process",
  "process" %in% names(pp_post_plot) ~ "process",
  TRUE ~ NA_character_
)
nr_post <- nrow(pp_post_plot)
if (nr_post == 0L) {
  pp_post_plot$Process <- factor(character(0), levels = c("control", "treated"))
} else if (is.na(process_col)) {
  pp_post_plot$Process <- factor(rep("control", nr_post), levels = c("control", "treated"))
} else {
  pp_post_plot$Process <- factor(as.character(pp_post_plot[[process_col]]),
                                 levels = c("control", "treated"))
}
p2 <- base_map +
  geom_point(data = pp_post_plot,
             aes(x = x, y = y, colour = Process, alpha = t),
             size = 0.8) +
  scale_colour_manual(values = c(control = "#2166ac", treated = "#b2182b"),
                      guide = "none") +
  scale_alpha_continuous(range = c(0.3, 1), guide = "none") +
  labs(title = "Post-treatment", x = "X (km)", y = "Y (km)")

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

# Model

## ETAS Conditional Intensity

$$
\lambda(t,x,y) = \frac{\mu}{|S|} + \sum_{j:\,t_j<t} A\,e^{\alpha_m(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

with Omori--Utsu temporal kernel $g$ and power-law spatial kernel $f$.

The kernels used in the fits are:

$$
g(\Delta t) = \frac{p-1}{c}\left(1+\frac{\Delta t}{c}\right)^{-p}, \qquad \Delta t > 0
$$

and

$$
f(\Delta x,\Delta y \mid m_j) =
\frac{q-1}{\pi\,D}\,
\left(1+\frac{r^2}{D}\right)^{-q},
\quad r^2 = \Delta x^2 + \Delta y^2.
$$

The spatial scale is **magnitude-independent**: we fix $\gamma\equiv 0$ in
every fit (equivalently $d(m)=D$), so larger events are more productive only
through $\alpha_m>0$, not through a wider spatial kernel.

## Bivariate Extension

$$
\lambda_k(t,x,y) = \frac{\mu_k}{|S_k|}\,W_k(x,y) + \sum_{l\in\{0,1\}}\sum_{\substack{j:\,t_j<t\\Z_j=l}} A_{kl}\,e^{\alpha_{m,kl}(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

Off-diagonal terms $A_{01}$ (treated$\to$control) and $A_{10}$
(control$\to$treated) capture cross-excitation. For the KDE background
models (C, D, G, H), $W_k(x,y)$ is the normalized kernel density estimate from
control pre-treatment data; for the homogeneous models (A, B, I, J),
$W_k \equiv 1$.

## Branching Matrix, Spectral Radius, and Explosiveness

The bivariate ETAS law admits a Poisson-cluster (branching) representation.
Averaging productivity over magnitudes under a Gutenberg–Richter law with
slope $\beta$ (here $\beta=`r signif(as.numeric(cfg$BETA_GR %||% 2.3), 3)`$)
gives the expected number of type-$k$ offspring of a type-$\ell$ parent,

$$
\eta_{k\ell}
=
A_{k\ell}\,\frac{\beta}{\beta-\alpha_{m,k\ell}},
\qquad
\alpha_{m,k\ell}<\beta
$$

(and $\eta_{k\ell}=\infty$ if $\alpha_{m,k\ell}\ge\beta$). These four
channel means form the **reproduction matrix**

$$
M
=
\begin{pmatrix}
\eta_{00} & \eta_{01} \\
\eta_{10} & \eta_{11}
\end{pmatrix}.
$$

Its **spectral radius** $\rho(M)=\max\{|\lambda|:\lambda\text{ eigenvalue of }M\}$
is the long-run growth factor of a typical cluster:

- $\rho(M)<1$ (**subcritical**): clusters die out almost surely; the expected
  total progeny of one immigrant is finite, equal to
  $\mathbf{1}^\top(I-M)^{-1}\mathbf{e}_\ell$ from a type-$\ell$ immigrant.
  Mean event counts over a finite horizon are well-defined, so all-or-nothing
  DAITE contrasts computed by forward simulation are meaningful.
- $\rho(M)=1$ (**critical**): expected cluster size diverges; finite-horizon
  counts remain random but have heavy tails.
- $\rho(M)>1$ (**explosive / supercritical**): clusters grow exponentially in
  expectation. Simulated catalogues routinely explode, and plug-in causal
  contrasts become numerically unstable and scientifically uninterpretable
  as finite expected savings.

In the independent (univariate) case there is no cross-excitation, so
$M=\mathrm{diag}(\eta_0,\eta_1)$ and $\rho(M)=\max(\eta_0,\eta_1)$: each
margin is checked separately. With cross-excitation, even two subcritical
self-excitation ratios can combine with large off-diagonal $\eta_{01},\eta_{10}$
to push $\rho(M)$ above one.

Fits in this analysis enforce a hard subcritical constraint
$\rho(M)<`r signif(as.numeric(cfg$ETAS_BRANCHING_MAX %||% 0.98), 3)`$
(and $\alpha_m>0$), projecting productivity when needed. Parameter tables
below report $\rho$ and an explosive flag for each fitted law.

## Likelihood Truncation in This Run

For ETAS likelihood evaluation in this project, temporal triggering is
truncated at a finite horizon `t_trunc` (days) when enabled. In code,
parent-child contributions with $\Delta t > t_{\text{trunc}}$ are dropped
from the triggering sum. The Omori part in the **log-intensity** is scaled by

$$
\text{temporal\_norm}
= 1 - \left(1 + \frac{t_{\text{trunc}}}{c}\right)^{-(p-1)},
$$

so $p$ directly controls how much triggering mass remains inside the truncation
window. Larger $p$ concentrates mass near zero lag (less tail), smaller $p$
leaves heavier temporal tails. With a finite horizon any $p>0$ is a valid
density. When truncation is off (`SEM_T_TRUNC_DAYS = none`), the untruncated
Omori law is used and the fit enforces $p>1$. Do not compare log-likelihoods
across different truncation settings.

The spatial exponent $q$ does **not** interact with temporal truncation directly.
It continues to govern spatial tail decay in
$f(r\mid m)\propto (1+r^2/d(m))^{-q}$, while truncation only limits temporal
history depth.

```{r tbl-truncation-params}
#| label: tbl-truncation-params
safe_num <- function(x) {
  v <- suppressWarnings(as.numeric(x))
  if (length(v) < 1L) return(NA_real_)
  v <- v[[1]]
  if (is.finite(v)) v else NA_real_
}
fs <- cfg$FIXED_STRUCTURAL
c_val <- if (!is.null(fs)) safe_num(fs[["c"]]) else NA_real_
p_val <- if (!is.null(fs)) safe_num(fs[["p"]]) else NA_real_
q_val <- if (!is.null(fs)) safe_num(fs[["q"]]) else NA_real_
ttr <- safe_num(cfg$SEM_T_TRUNC_DAYS)
rel_target <- safe_num(cfg$SEM_T_TRUNC_REL)
t_source <- if (!is.null(cfg$SEM_T_TRUNC_SOURCE)) as.character(cfg$SEM_T_TRUNC_SOURCE[[1]]) else "N/A"

temporal_mass_retained <- if (is.finite(ttr) && is.finite(c_val) && is.finite(p_val) && c_val > 0 && p_val > 1) {
  1 - (1 + ttr / c_val)^(-(p_val - 1))
} else NA_real_
tail_mass_dropped <- if (is.finite(temporal_mass_retained)) 1 - temporal_mass_retained else NA_real_
kernel_rel_at_ttr <- if (is.finite(ttr) && is.finite(c_val) && is.finite(p_val) && c_val > 0) {
  (1 + ttr / c_val)^(-p_val)
} else NA_real_

p_bound <- safe_num(cfg$ETAS_P_LOWER_BOUND)
trunc_df <- data.frame(
  Setting = c(
    "SEM t_trunc (days)",
    "SEM t_trunc source",
    "Untruncated Omori (p > 1)",
    "ETAS p lower bound",
    "Legacy auto-trunc rel level (unused)",
    "Structural c (Omori offset)",
    "Structural p (temporal exponent)",
    "Structural q (spatial exponent)",
    "Temporal mass retained up to t_trunc",
    "Temporal tail mass dropped beyond t_trunc",
    "Relative Omori kernel at t_trunc: ((t_trunc+c)/c)^(-p)"
  ),
  Value = c(
    fmt_t_trunc_cfg(cfg$SEM_T_TRUNC_DAYS),
    t_source,
    if (isTRUE(cfg$SEM_NO_TRUNCATION)) "yes" else if (is.finite(ttr)) "no" else "N/A",
    if (is.finite(p_bound)) sprintf("%.4f", p_bound) else "N/A",
    if (is.finite(rel_target)) sprintf("%.4f", rel_target) else "N/A",
    if (is.finite(c_val)) sprintf("%.4f", c_val) else "N/A",
    if (is.finite(p_val)) sprintf("%.4f", p_val) else "N/A",
    if (is.finite(q_val)) sprintf("%.4f", q_val) else "N/A",
    if (is.finite(temporal_mass_retained)) sprintf("%.4f", temporal_mass_retained) else "N/A",
    if (is.finite(tail_mass_dropped)) sprintf("%.4f", tail_mass_dropped) else "N/A",
    if (is.finite(kernel_rel_at_ttr)) sprintf("%.6f", kernel_rel_at_ttr) else "N/A"
  ),
  stringsAsFactors = FALSE
)
kable(trunc_df, align = "lc",
      caption = "Temporal truncation settings and implied Omori-tail diagnostics used in likelihood/simulation.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

The truncation-horizon grid below refits **Fit C** (naive bivariate + KDE) and
**Fit D** (SEM bivariate + KDE) at each `t_trunc`, then recomputes the
bivariate all-or-nothing contrast. Fit C is a single MLE so it is cheap
relative to SEM. Older results files that only stored Fit D still render
the D columns.

```{r prep-t-trunc-sensitivity}
#| label: prep-t-trunc-sensitivity
#| include: false
one_num <- function(x) {
  v <- suppressWarnings(as.numeric(unlist(x, use.names = FALSE)))
  if (length(v) < 1L) NA_real_ else v[[1]]
}
one_int <- function(x) {
  v <- suppressWarnings(as.integer(unlist(x, use.names = FALSE)))
  if (length(v) < 1L) NA_integer_ else v[[1]]
}
one_chr <- function(x) {
  if (is.null(x) || length(unlist(x, use.names = FALSE)) < 1L) NA_character_ else as.character(unlist(x, use.names = FALSE)[[1]])
}
trunc_sens <- results$t_trunc_sensitivity
trunc_sens_df <- NULL
trunc_sens_plot_df <- NULL
trunc_sens_has_c <- FALSE
trunc_sens_main_ttr <- one_num(cfg$SEM_T_TRUNC_DAYS)
if (!is.null(trunc_sens) && length(trunc_sens) >= 1L) {
  rows <- lapply(trunc_sens, function(x) {
    if (is.null(x)) return(NULL)
    data.frame(
      t_trunc_days = one_num(x$t_trunc_days),
      C_status = one_chr(x$C_status),
      D_status = one_chr(x$D_status %||% x$status),
      C_mean_saved = one_num(x$C_ate_mean_saved),
      C_sd_saved = one_num(x$C_ate_sd_saved),
      C_n_sims = one_int(x$C_ate_n_sims),
      D_mean_saved = one_num(x$ate_mean_saved),
      D_sd_saved = one_num(x$ate_sd_saved),
      D_n_sims = one_int(x$ate_n_sims),
      n_relabel = one_int(x$n_relabel),
      elapsed_min = one_num(x$elapsed_sec) / 60,
      stringsAsFactors = FALSE
    )
  })
  trunc_sens_df <- do.call(rbind, rows[!vapply(rows, is.null, logical(1))])
  if (!is.null(trunc_sens_df) && nrow(trunc_sens_df) >= 1L) {
    trunc_ord <- trunc_sens_df$t_trunc_days
    trunc_ord[!is.finite(trunc_ord)] <- Inf
    trunc_sens_df <- trunc_sens_df[order(trunc_ord), , drop = FALSE]
    trunc_sens_has_c <- any(is.finite(trunc_sens_df$C_mean_saved) |
                              (!is.na(trunc_sens_df$C_status) & nzchar(trunc_sens_df$C_status)))
    plot_rows <- list(
      data.frame(
        t_trunc_days = trunc_sens_df$t_trunc_days,
        Fit = "D: SEM biv+KDE",
        mean_saved = trunc_sens_df$D_mean_saved,
        sd_saved = trunc_sens_df$D_sd_saved,
        stringsAsFactors = FALSE
      )
    )
    if (isTRUE(trunc_sens_has_c)) {
      plot_rows <- c(
        list(data.frame(
          t_trunc_days = trunc_sens_df$t_trunc_days,
          Fit = "C: Naive biv+KDE",
          mean_saved = trunc_sens_df$C_mean_saved,
          sd_saved = trunc_sens_df$C_sd_saved,
          stringsAsFactors = FALSE
        )),
        plot_rows
      )
    }
    trunc_sens_plot_df <- do.call(rbind, plot_rows)
  } else {
    trunc_sens_df <- NULL
  }
}
```

```{r tbl-t-trunc-sensitivity}
#| label: tbl-t-trunc-sensitivity
if (is.null(trunc_sens_df)) {
  cat("No Fit C/D t_trunc sensitivity grid was saved in this results file.\n")
} else {
  cap <- if (isTRUE(trunc_sens_has_c)) {
    "Fit C (naive biv+KDE) and Fit D (SEM biv+KDE) all-or-nothing ATE vs temporal truncation horizon."
  } else {
    "Fit D (SEM biv+KDE) all-or-nothing ATE vs temporal truncation horizon."
  }
  cap <- paste0(cap, sprintf(" Main-run t_trunc=%s.", fmt_t_trunc_cfg(cfg$SEM_T_TRUNC_DAYS)))
  tbl <- if (isTRUE(trunc_sens_has_c)) {
    trunc_sens_df |>
      transmute(
        `t_trunc (days)` = ifelse(is.finite(t_trunc_days), sprintf("%.3g", t_trunc_days), "none"),
        `C status` = ifelse(is.na(C_status) | !nzchar(C_status), "NA", C_status),
        `C mean saved` = ifelse(is.finite(C_mean_saved), sprintf("%.1f", C_mean_saved), "NA"),
        `C SD` = ifelse(is.finite(C_sd_saved), sprintf("%.1f", C_sd_saved), "NA"),
        `D status` = ifelse(is.na(D_status) | !nzchar(D_status), "NA", D_status),
        `D mean saved` = ifelse(is.finite(D_mean_saved), sprintf("%.1f", D_mean_saved), "NA"),
        `D SD` = ifelse(is.finite(D_sd_saved), sprintf("%.1f", D_sd_saved), "NA"),
        Relabels = ifelse(is.finite(n_relabel), as.character(n_relabel), "NA"),
        `Elapsed (min)` = ifelse(is.finite(elapsed_min), sprintf("%.1f", elapsed_min), "NA")
      )
  } else {
    trunc_sens_df |>
      transmute(
        `t_trunc (days)` = ifelse(is.finite(t_trunc_days), sprintf("%.3g", t_trunc_days), "none"),
        Status = D_status,
        `Mean saved` = ifelse(is.finite(D_mean_saved), sprintf("%.1f", D_mean_saved), "NA"),
        `SD saved` = ifelse(is.finite(D_sd_saved), sprintf("%.1f", D_sd_saved), "NA"),
        `ATE sims` = ifelse(is.finite(D_n_sims), as.character(D_n_sims), "NA"),
        Relabels = ifelse(is.finite(n_relabel), as.character(n_relabel), "NA"),
        `Elapsed (min)` = ifelse(is.finite(elapsed_min), sprintf("%.1f", elapsed_min), "NA")
      )
  }
  knitr::kable(tbl, align = "lcccccccc", caption = cap, row.names = FALSE) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

```{r fig-t-trunc-sensitivity}
#| label: fig-t-trunc-sensitivity
#| fig-cap: "Fit C/D all-or-nothing ATE versus temporal truncation horizon."
if (!is.null(trunc_sens_plot_df) &&
    requireNamespace("ggplot2", quietly = TRUE) &&
    sum(is.finite(trunc_sens_plot_df$mean_saved)) >= 2L) {
  p_trunc <- ggplot2::ggplot(
    trunc_sens_plot_df[is.finite(trunc_sens_plot_df$mean_saved), , drop = FALSE],
    ggplot2::aes(x = t_trunc_days, y = mean_saved, colour = Fit)
  ) +
    ggplot2::geom_line(linewidth = 0.8) +
    ggplot2::geom_point(size = 2.4) +
    ggplot2::geom_errorbar(
      ggplot2::aes(
        ymin = mean_saved - sd_saved,
        ymax = mean_saved + sd_saved
      ),
      width = 0.4, alpha = 0.7,
      na.rm = TRUE
    ) +
    ggplot2::scale_colour_manual(
      values = c("C: Naive biv+KDE" = "#fdb462", "D: SEM biv+KDE" = "#80b1d3")
    ) +
    ggplot2::labs(
      title = "Fit C/D ATE vs temporal truncation",
      x = expression(t[trunc]~"(days)"),
      y = "Mean saved events (AoN)",
      colour = NULL
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(legend.position = "top")
  if (is.finite(trunc_sens_main_ttr)) {
    p_trunc <- p_trunc +
      ggplot2::geom_vline(
        xintercept = trunc_sens_main_ttr, linetype = "dashed", colour = "grey40"
      )
  }
  print(p_trunc)
}
```

## Non-Parametric Background Rate

```{r tbl-kde}
#| label: tbl-kde
kde <- results$kde_info
if (is.null(kde)) {
  kde_df <- data.frame(
    Setting = "KDE summary",
    Value = paste0("Not stored in this results file. ",
                   if (using_pre_bootstrap) {
                     "Re-run oklahoma_analysis.R through the pre-bootstrap checkpoint (current script saves kde_info), or use full oklahoma_results.rds."
                   } else {
                     "Use a full results RDS that includes kde_info."
                   }),
    stringsAsFactors = FALSE
  )
} else {
  kde_holdout <- if (!is.null(kde$n_pre_holdout)) kde$n_pre_holdout else NA
  kde_used <- if (!is.null(kde$n_pre_used)) kde$n_pre_used else NA
  kde_df <- data.frame(
    Setting = c("KDE training events (held-out control pre)",
                "Held-out pre-treatment events (all processes)",
                "Pre-treatment events used for estimation",
                "Bandwidth (km)", "Estimated triggering range (km)", "Method"),
    Value = c(kde$n_training, kde_holdout, kde_used,
              fmt_scaled_num(kde$bw_sigma), fmt_scaled_num(kde$trigger_range_km),
              if (!is.null(kde$bw_label)) kde$bw_label else if (!is.null(kde$bw_method)) kde$bw_method else "2 x Diggle CV"))
}
kable(kde_df, align = "lc",
      caption = "KDE background rate estimation details.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

Only the **second 50%** of pre-treatment events are used for
estimation/carryover in both homogeneous and inhomogeneous SEM fits.
The **first 50%** is held out for fairness; in the inhomogeneous case,
its control events are used to estimate the KDE background
(`r if (!is.null(kde) && !is.null(kde$bw_label)) kde$bw_label else if (!is.null(kde) && !is.null(kde$bw_method)) kde$bw_method else "configured bandwidth method"`).

```{r fig-kde-bg}
#| label: fig-kde-bg
#| fig-cap: "Estimated inhomogeneous background rate (KDE) from held-out control pre-treatment events."
#| fig-height: 6
bg_plot <- NULL
if (!is.null(results$plots) && !is.null(results$plots$background_rate_kde)) {
  bg_plot <- results$plots$background_rate_kde
}
if (inherits(bg_plot, "gg") || inherits(bg_plot, "ggplot")) {
  # Analysis stores this ggplot on the results RDS; PNG export is optional.
  if (exists("counties_km") && inherits(counties_km, "sf")) {
    county_xy <- tryCatch({
      as.data.frame(sf::st_coordinates(sf::st_cast(counties_km, "MULTILINESTRING")))
    }, error = function(e) NULL)
    if (!is.null(county_xy) && all(c("X", "Y", "L1") %in% names(county_xy))) {
      bg_plot <- bg_plot +
        geom_path(
          data = county_xy,
          aes(x = X, y = Y, group = L1),
          inherit.aes = FALSE,
          colour = "grey25",
          linewidth = 0.2,
          alpha = 0.7
        )
    }
  }
  print(bg_plot)
} else {
  bg_plot_file <- resolve_plot_file("background_rate_kde.png")
  if (is.character(bg_plot_file) && length(bg_plot_file) == 1L &&
      nzchar(bg_plot_file) && file.exists(bg_plot_file)) {
    knitr::include_graphics(bg_plot_file)
  } else {
    plot.new()
    text(0.5, 0.5, "KDE background-rate plot unavailable in this results file.")
  }
}
```

## KDE bandwidth sensitivity

Fit C (naive bivariate + KDE) and Fit D (SEM bivariate + KDE) are refit on an
isotropic Gaussian bandwidth grid at the same temporal truncation as the main
run. Saved events use the bivariate all-or-nothing contrast.

```{r prep-kde-bw-sensitivity}
#| label: prep-kde-bw-sensitivity
#| include: false
bw_sens <- results$kde_bandwidth_sensitivity
bw_sens_df <- NULL
bw_sens_plot_df <- NULL
bw_sens_has_c <- FALSE
ate_saved_from_obj <- function(ate_obj) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim$total_saved)) {
    return(list(mean = NA_real_, sd = NA_real_, n = NA_integer_))
  }
  saved <- suppressWarnings(as.numeric(ate_obj$all_nothing_sim$total_saved))
  saved <- saved[is.finite(saved)]
  list(
    mean = if (length(saved) > 0L) mean(saved) else NA_real_,
    sd = if (length(saved) > 1L) stats::sd(saved) else NA_real_,
    n = as.integer(length(saved))
  )
}
if (!is.null(bw_sens) && length(bw_sens) >= 1L) {
  rows <- lapply(bw_sens, function(x) {
    if (is.null(x)) return(NULL)
    d_mean <- one_num(x$ate_mean_saved)
    d_sd <- one_num(x$ate_sd_saved)
    d_n <- one_int(x$ate_n_sims)
    if (!is.finite(d_mean)) {
      d_stats <- ate_saved_from_obj(x$ate_F)
      d_mean <- d_stats$mean
      d_sd <- d_stats$sd
      d_n <- d_stats$n
    }
    c_mean <- one_num(x$C_ate_mean_saved)
    c_sd <- one_num(x$C_ate_sd_saved)
    c_n <- one_int(x$C_ate_n_sims)
    if (!is.finite(c_mean)) {
      c_stats <- ate_saved_from_obj(x$ate_E)
      c_mean <- c_stats$mean
      c_sd <- c_stats$sd
      c_n <- c_stats$n
    }
    data.frame(
      label = one_chr(x$label),
      sigma_km = one_num(x$sigma),
      C_mean_saved = c_mean,
      C_sd_saved = c_sd,
      C_n_sims = c_n,
      D_mean_saved = d_mean,
      D_sd_saved = d_sd,
      D_n_sims = d_n,
      n_relabel = one_int(x$n_relabel),
      stringsAsFactors = FALSE
    )
  })
  bw_sens_df <- do.call(rbind, rows[!vapply(rows, is.null, logical(1))])
  if (!is.null(bw_sens_df) && nrow(bw_sens_df) >= 1L) {
    bw_sens_df <- bw_sens_df[order(bw_sens_df$sigma_km, bw_sens_df$label), , drop = FALSE]
    bw_sens_has_c <- any(is.finite(bw_sens_df$C_mean_saved))
    plot_rows <- list(
      data.frame(
        sigma_km = bw_sens_df$sigma_km,
        Fit = "D: SEM biv+KDE",
        mean_saved = bw_sens_df$D_mean_saved,
        sd_saved = bw_sens_df$D_sd_saved,
        stringsAsFactors = FALSE
      )
    )
    if (isTRUE(bw_sens_has_c)) {
      plot_rows <- c(
        list(data.frame(
          sigma_km = bw_sens_df$sigma_km,
          Fit = "C: Naive biv+KDE",
          mean_saved = bw_sens_df$C_mean_saved,
          sd_saved = bw_sens_df$C_sd_saved,
          stringsAsFactors = FALSE
        )),
        plot_rows
      )
    }
    bw_sens_plot_df <- do.call(rbind, plot_rows)
  } else {
    bw_sens_df <- NULL
  }
}
```

```{r tbl-kde-bw-sensitivity}
#| label: tbl-kde-bw-sensitivity
if (is.null(bw_sens_df)) {
  cat("No Fit C/D KDE bandwidth sensitivity grid was saved in this results file.\n")
} else {
  cap <- if (isTRUE(bw_sens_has_c)) {
    "Fit C (naive biv+KDE) and Fit D (SEM biv+KDE) all-or-nothing ATE vs isotropic KDE bandwidth."
  } else {
    "Fit D (SEM biv+KDE) all-or-nothing ATE vs isotropic KDE bandwidth."
  }
  cap <- paste0(cap, sprintf(" Main-run t_trunc=%s.", fmt_t_trunc_cfg(cfg$SEM_T_TRUNC_DAYS)))
  tbl <- if (isTRUE(bw_sens_has_c)) {
    bw_sens_df |>
      transmute(
        Bandwidth = label,
        `sigma (km)` = ifelse(is.finite(sigma_km), sprintf("%.3g", sigma_km), "NA"),
        `C mean saved` = ifelse(is.finite(C_mean_saved), sprintf("%.1f", C_mean_saved), "NA"),
        `C SD` = ifelse(is.finite(C_sd_saved), sprintf("%.1f", C_sd_saved), "NA"),
        `D mean saved` = ifelse(is.finite(D_mean_saved), sprintf("%.1f", D_mean_saved), "NA"),
        `D SD` = ifelse(is.finite(D_sd_saved), sprintf("%.1f", D_sd_saved), "NA"),
        Relabels = ifelse(is.finite(n_relabel), as.character(n_relabel), "NA")
      )
  } else {
    bw_sens_df |>
      transmute(
        Bandwidth = label,
        `sigma (km)` = ifelse(is.finite(sigma_km), sprintf("%.3g", sigma_km), "NA"),
        `Mean saved` = ifelse(is.finite(D_mean_saved), sprintf("%.1f", D_mean_saved), "NA"),
        `SD saved` = ifelse(is.finite(D_sd_saved), sprintf("%.1f", D_sd_saved), "NA"),
        `ATE sims` = ifelse(is.finite(D_n_sims), as.character(D_n_sims), "NA"),
        Relabels = ifelse(is.finite(n_relabel), as.character(n_relabel), "NA")
      )
  }
  knitr::kable(tbl, align = "lcccccc", caption = cap, row.names = FALSE) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

```{r fig-kde-bw-sensitivity}
#| label: fig-kde-bw-sensitivity
#| fig-cap: "Fit C/D all-or-nothing ATE versus isotropic KDE bandwidth."
if (!is.null(bw_sens_plot_df) &&
    requireNamespace("ggplot2", quietly = TRUE) &&
    sum(is.finite(bw_sens_plot_df$mean_saved) & is.finite(bw_sens_plot_df$sigma_km)) >= 2L) {
  p_bw <- ggplot2::ggplot(
    bw_sens_plot_df[is.finite(bw_sens_plot_df$mean_saved) & is.finite(bw_sens_plot_df$sigma_km), , drop = FALSE],
    ggplot2::aes(x = sigma_km, y = mean_saved, colour = Fit)
  ) +
    ggplot2::geom_line(linewidth = 0.8) +
    ggplot2::geom_point(size = 2.4) +
    ggplot2::geom_errorbar(
      ggplot2::aes(
        ymin = mean_saved - sd_saved,
        ymax = mean_saved + sd_saved
      ),
      width = 0.2, alpha = 0.7,
      na.rm = TRUE
    ) +
    ggplot2::scale_colour_manual(
      values = c("C: Naive biv+KDE" = "#fdb462", "D: SEM biv+KDE" = "#80b1d3")
    ) +
    ggplot2::labs(
      title = "Fit C/D ATE vs KDE bandwidth",
      x = "Isotropic bandwidth (km)",
      y = "Mean saved events (AoN)",
      colour = NULL
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(legend.position = "top")
  print(p_bw)
}
```

## Fixed Parameters

```{r tbl-fixed}
#| label: tbl-fixed
fixed_cfg <- cfg$FIXED_STRUCTURAL
if (is.null(fixed_cfg) || length(fixed_cfg) == 0) {
  fixed_df <- data.frame(
    Parameter = "None",
    Value = "Estimated (not fixed)",
    Role = "All ETAS parameters are estimated from data",
    stringsAsFactors = FALSE
  )
  cap <- "Fixed structural parameters: none for this run."
} else {
  role_map <- c(
    c = "Omori onset delay",
    p = "Omori decay exponent",
    D = "Spatial scale (magnitude-independent)",
    gamma = "Fixed at 0 (no mag-spatial scaling)",
    q = "Spatial decay exponent"
  )
  fixed_df <- data.frame(
    Parameter = names(fixed_cfg),
    Value = as.numeric(unlist(fixed_cfg)),
    Role = unname(role_map[names(fixed_cfg)]),
    stringsAsFactors = FALSE
  )
  if (!("gamma" %in% fixed_df$Parameter)) {
    g0 <- suppressWarnings(as.numeric(cfg$GAMMA_FIXED %||% 0))
    fixed_df <- rbind(
      fixed_df,
      data.frame(
        Parameter = "gamma",
        Value = if (is.finite(g0)) g0 else 0,
        Role = role_map[["gamma"]],
        stringsAsFactors = FALSE
      )
    )
  }
  cap <- "Fixed structural parameters (gamma=0 in all models)."
}
kable(fixed_df, align = "lcl", caption = cap) |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

```{r tbl-fixed-control-snapshots}
#| label: tbl-fixed-control-snapshots
cs <- results$control_snapshot_fits
if (is.null(cs) || length(cs) == 0) {
  kable(
    data.frame(
      Snapshot = "Control-only breakpoint fits",
      Status = "Not available in this RDS; rerun analysis with updated script.",
      stringsAsFactors = FALSE
    ),
    align = "ll",
    caption = "Control-only ETAS snapshots at three breakpoints (diagnostic only)."
  ) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
} else {
  ord <- c("pre50", "tstar", "end")
  keep <- ord[ord %in% names(cs)]
  if (length(keep) < 1L) keep <- names(cs)
  rows <- lapply(keep, function(k) {
    z <- cs[[k]]
    p <- z$params
    data.frame(
      Snapshot = z$label %||% k,
      `Window end (days)` = if (is.null(z$window_end)) NA_real_ else as.numeric(z$window_end),
      `Events used` = if (is.null(z$n_events)) NA_integer_ else as.integer(z$n_events),
      Status = z$status %||% NA_character_,
      mu = if (is.null(p$mu)) NA_real_ else as.numeric(p$mu),
      A = if (is.null(p$A)) NA_real_ else as.numeric(p$A),
      alpha_m = if (is.null(p$alpha_m)) NA_real_ else as.numeric(p$alpha_m),
      c = if (is.null(p$c)) NA_real_ else as.numeric(p$c),
      p = if (is.null(p$p)) NA_real_ else as.numeric(p$p),
      D = if (is.null(p$D)) NA_real_ else as.numeric(p$D),
      gamma = if (is.null(p$gamma)) NA_real_ else as.numeric(p$gamma),
      q = if (is.null(p$q)) NA_real_ else as.numeric(p$q),
      loglik = if (is.null(z$loglik)) NA_real_ else as.numeric(z$loglik),
      check.names = FALSE
    )
  })
  snap_df <- do.call(rbind, rows)
  kable(
    snap_df,
    digits = 4,
    align = "lcccrrrrrrrrr",
    caption = paste0(
      "Control-only ETAS fits at three breakpoints (all points/regions treated as control). ",
      "Diagnostic for parameter drift over time; SEM initialization machinery remains unchanged."
    )
  ) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                  font_size = 9, latex_options = c("scale_down", "hold_position"))
}
```

## SEM Configuration

Pilot tuning is optional and only targets the Model D SEM relabelling controls.
When enabled, the code runs a parallel grid over:

- `change_factor`
- `change_factor_min_mult`
- `change_factor_max_mult`
- `selection_temperature`

using a short inner SEM run (default 100 iterations), then ranks candidates by:
high final SEM score and low tail accepted/average flips. The chosen settings are
then used for the main primary-tessellation SEM fits in the run.

```{r sem-final-config}
#| label: sem-final-config
fmt <- function(x, default = "N/A") {
  if (length(x) < 1L || is.null(x) || all(is.na(x))) return(default)
  as.character(x[[1]])
}
run_sem_inner_props <- if (!is.null(cfg$SEM_INNER_PROPS)) cfg$SEM_INNER_PROPS else cfg$SEM_N_LABELLINGS
final_cfg_tbl <- data.frame(
  Setting = c(
    "Outer iterations",
    "Inner iterations",
    "SEM inner proposals",
    "SEM labellings",
    "Warm start fixed step",
    "Selection method",
    "Selection temperature",
    "Change factor",
    "Change-factor min multiplier",
    "Change-factor max multiplier",
    "Pilot enabled",
    "Pilot inner iterations",
    "Pilot cores",
    "Pilot max combos"
  ),
  Value = c(
    fmt(cfg$SEM_N_ITER),
    fmt(cfg$SEM_INNER_ITER),
    fmt(run_sem_inner_props),
    fmt(cfg$SEM_N_LABELLINGS),
    fmt(cfg$SEM_WARMSTART_FIXED),
    fmt(cfg$SEM_OPTIM_METHOD),
    fmt(cfg$SEM_SELECTION_TEMPERATURE),
    fmt(cfg$SEM_CHANGE_FACTOR),
    fmt(cfg$SEM_CHANGE_FACTOR_MIN_MULT),
    fmt(cfg$SEM_CHANGE_FACTOR_MAX_MULT),
    fmt(cfg$RUN_SEM_PILOT),
    fmt(cfg$SEM_PILOT_INNER_ITER),
    fmt(cfg$SEM_PILOT_CORES),
    fmt(cfg$SEM_PILOT_MAX_COMBOS)
  ),
  stringsAsFactors = FALSE
)
kable(final_cfg_tbl, align = "lc", caption = "Final SEM settings used for this run.") |>
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

# Results — Tessellation Partition

```{r extract}
#| label: extract
get_fit_by_letter <- function(letter, legacy_key = NULL) {
  if (!is.null(results$fits_named) && !is.null(results$fits_named[[letter]])) {
    return(results$fits_named[[letter]])
  }
  if (!is.null(legacy_key) && !is.null(results[[legacy_key]])) return(results[[legacy_key]])
  NULL
}

fit_A_obj <- get_fit_by_letter("A", "fitB")
fit_B_obj <- get_fit_by_letter("B", "fitD")
# Primary all_free C/D; legacy top-level fitE/fitF alias the same content.
fit_C_obj <- get_fit_by_letter("C", "fitE")
fit_D_obj <- get_fit_by_letter("D", "fitF")
fit_E_obj <- get_fit_by_letter("E")
fit_F_obj <- get_fit_by_letter("F")
# Univ+KDE G/H (legacy fitI/fitJ); univ homog I/J (legacy fitK/fitL).
fit_G_obj <- get_fit_by_letter("G", "fitI")
fit_H_obj <- get_fit_by_letter("H", "fitJ")
fit_I_obj <- get_fit_by_letter("I", "fitK")
fit_J_obj <- get_fit_by_letter("J", "fitL")
# Optional K/L fallbacks after normalize (prefer clean A–J above).
fit_K_obj <- get_fit_by_letter("K", "fitK")
if (is.null(fit_K_obj)) fit_K_obj <- fit_I_obj
fit_L_obj <- get_fit_by_letter("L", "fitL")
if (is.null(fit_L_obj)) fit_L_obj <- fit_J_obj

A_par <- if (!is.null(fit_A_obj)) fit_A_obj$params else NULL
B_par <- if (!is.null(fit_B_obj)) fit_B_obj$params else NULL
C_par <- if (!is.null(fit_C_obj)) fit_C_obj$params else NULL
D_par <- if (!is.null(fit_D_obj)) fit_D_obj$params else NULL
E_par <- if (!is.null(fit_E_obj)) fit_E_obj$params else NULL
F_par <- if (!is.null(fit_F_obj)) fit_F_obj$params else NULL
G_par <- if (!is.null(fit_G_obj)) fit_G_obj$params else NULL
H_par <- if (!is.null(fit_H_obj)) fit_H_obj$params else NULL
I_par <- if (!is.null(fit_I_obj)) fit_I_obj$params else NULL
J_par <- if (!is.null(fit_J_obj)) fit_J_obj$params else NULL

scalar_num_or_na <- function(x) {
  v <- suppressWarnings(as.numeric(x))
  if (length(v) < 1L) return(NA_real_)
  v <- v[[1]]
  if (is.finite(v)) v else NA_real_
}

bc <- function(par, p) {
  if (is.null(par)) return(NA_real_)
  if (is.list(par) && !is.null(par$control)) {
    return(scalar_num_or_na(par$control[[p]]))
  }
  if (p == "mu") {
    v <- scalar_num_or_na(par[["mu_0"]])
    if (is.finite(v)) return(v)
    v <- scalar_num_or_na(par[["mu"]])
    if (is.finite(v)) return(v)
  }
  v <- scalar_num_or_na(par[[paste0(p, "_00")]])
  if (is.finite(v)) return(v)
  scalar_num_or_na(par[[p]])
}
bt <- function(par, p) {
  if (is.null(par)) return(NA_real_)
  if (is.list(par) && !is.null(par$treated)) {
    return(scalar_num_or_na(par$treated[[p]]))
  }
  if (p == "mu") {
    v <- scalar_num_or_na(par[["mu_1"]])
    if (is.finite(v)) return(v)
    v <- scalar_num_or_na(par[["mu"]])
    if (is.finite(v)) return(v)
  }
  v <- scalar_num_or_na(par[[paste0(p, "_11")]])
  if (is.finite(v)) return(v)
  scalar_num_or_na(par[[p]])
}
safe_bc <- function(par, p) if (is.null(par)) NA_real_ else bc(par, p)
safe_bt <- function(par, p) if (is.null(par)) NA_real_ else bt(par, p)

etas_trigger_means <- function(par, beta_gr = NULL, t_trunc_days = NULL) {
  if (is.null(par)) {
    return(list(mean_time = NA_real_, mean_time_trunc = NA_real_, mean_distance = NA_real_, mean_scale = NA_real_))
  }
  c0 <- scalar_num_or_na(par[["c"]])
  p0 <- scalar_num_or_na(par[["p"]])
  D0 <- scalar_num_or_na(par[["D"]])
  g0 <- scalar_num_or_na(par[["gamma"]])
  q0 <- scalar_num_or_na(par[["q"]])
  ttr <- scalar_num_or_na(t_trunc_days)
  b0_vec <- suppressWarnings(as.numeric(unlist(beta_gr, use.names = FALSE)))
  b0 <- if (length(b0_vec) < 1L) NA_real_ else scalar_num_or_na(b0_vec[[1]])
  if (!is.finite(b0) || b0 <= 0) b0 <- 2.3

  mean_time <- if (is.finite(c0) && is.finite(p0) && !is.na(c0) && !is.na(p0) && p0 > 2) {
    c0 / (p0 - 2)
  } else {
    Inf
  }
  mean_time_trunc <- NA_real_
  if (is.finite(c0) && is.finite(p0) && !is.na(c0) && !is.na(p0) &&
      c0 > 0 && p0 > 1 && is.finite(ttr) && !is.na(ttr) && ttr > 0) {
    u <- 1 + (ttr / c0)
    if (p0 > 1 && p0 != 2) {
      i0 <- (c0 / (p0 - 1)) * (1 - u^(1 - p0))
      i1 <- c0^2 * ((u^(2 - p0) - 1) / (2 - p0) - (u^(1 - p0) - 1) / (1 - p0))
      if (is.finite(i0) && i0 > 0 && is.finite(i1)) {
        mean_time_trunc <- i1 / i0
      }
    } else if (identical(round(p0, 12), 2.0)) {
      # Closed form limit at p = 2.
      i0 <- c0 * (1 - 1 / u)
      i1 <- c0^2 * (log(u) + (1 / u) - 1)
      if (is.finite(i0) && i0 > 0 && is.finite(i1)) {
        mean_time_trunc <- i1 / i0
      }
    }
    if (!is.finite(mean_time_trunc) || is.na(mean_time_trunc)) {
      num <- tryCatch(
        integrate(function(tt) tt * (1 + tt / c0)^(-p0), lower = 0, upper = ttr)$value,
        error = function(e) NA_real_
      )
      den <- tryCatch(
        integrate(function(tt) (1 + tt / c0)^(-p0), lower = 0, upper = ttr)$value,
        error = function(e) NA_real_
      )
      if (is.finite(num) && is.finite(den) && den > 0) {
        mean_time_trunc <- num / den
      }
    }
  }

  mean_scale <- if (is.finite(D0) && is.finite(g0) && !is.na(D0) && !is.na(g0)) {
    D0 * exp(g0 / b0)
  } else {
    NA_real_
  }

  mean_distance <- if (is.finite(mean_scale) && !is.na(mean_scale) &&
                       is.finite(q0) && !is.na(q0) && q0 > 1.5) {
    sqrt(mean_scale) * (q0 - 1) * sqrt(pi) * gamma(q0 - 1.5) / gamma(q0)
  } else if (is.finite(q0) && !is.na(q0) && q0 <= 1.5) {
    Inf
  } else {
    NA_real_
  }

  list(mean_time = mean_time, mean_time_trunc = mean_time_trunc, mean_distance = mean_distance, mean_scale = mean_scale)
}

model_defs <- list(
  list(id = "A", fit = "fits_named$A", label = "A: Naive", long = "A: Naive Biv", par = A_par),
  list(id = "B", fit = "fits_named$B", label = "B: SEM", long = "B: SEM Biv", par = B_par),
  list(id = "C", fit = "fits_named$C", label = "C: Naive+KDE (all-free)", long = "C: Naive+KDE (all-free)", par = C_par),
  list(id = "D", fit = "fits_named$D", label = "D: SEM+KDE (all-free)", long = "D: SEM+KDE (all-free)", par = D_par),
  list(id = "E", fit = "fits_named$E", label = "E: Naive+KDE (struct-fixed)", long = "E: Naive+KDE (struct-fixed)", par = E_par),
  list(id = "F", fit = "fits_named$F", label = "F: SEM+KDE (struct-fixed)", long = "F: SEM+KDE (struct-fixed)", par = F_par),
  list(id = "G", fit = "fits_named$G", label = "G: Naive univariate+KDE", long = "G: Naive Univ+KDE", par = G_par),
  list(id = "H", fit = "fits_named$H", label = "H: SEM univariate+KDE", long = "H: SEM Univ+KDE", par = H_par),
  list(id = "I", fit = "fits_named$I", label = "I: Naive univariate (no KDE)", long = "I: Naive Univ (no KDE)", par = I_par),
  list(id = "J", fit = "fits_named$J", label = "J: SEM univariate (no KDE)", long = "J: SEM Univ (no KDE)", par = J_par)
)
param_models <- Filter(function(m) !is.null(m$par), model_defs)

fit_fill_colors <- c(
  "A: Naive Bivariate" = "#fc8d62",
  "B: SEM Bivariate" = "#8da0cb",
  "C: Naive Biv+KDE (all-free)" = "#fdb462",
  "D: SEM Biv+KDE (all-free)" = "#80b1d3",
  "E: Naive Biv+KDE (struct-fixed)" = "#fb9a99",
  "F: SEM Biv+KDE (struct-fixed)" = "#b3de69",
  "G: Naive Univ+KDE" = "#66c2a5",
  "H: SEM Univ+KDE" = "#ffd92f",
  "I: Naive Univ (no KDE)" = "#e78ac3",
  "J: SEM Univ (no KDE)" = "#a6d854"
)
fit_line_colors <- c(
  "A: Naive Bivariate" = "#d95f02",
  "B: SEM Bivariate" = "#7570b3",
  "C: Naive Biv+KDE (all-free)" = "#e6550d",
  "D: SEM Biv+KDE (all-free)" = "#3182bd",
  "E: Naive Biv+KDE (struct-fixed)" = "#e31a1c",
  "F: SEM Biv+KDE (struct-fixed)" = "#33a02c",
  "G: Naive Univ+KDE" = "#1b9e77",
  "H: SEM Univ+KDE" = "#b8860b",
  "I: Naive Univ (no KDE)" = "#c51b7d",
  "J: SEM Univ (no KDE)" = "#4d9221"
)

get_sem_metric_trace <- function(fit_obj) {
  if (is.null(fit_obj)) return(NULL)
  sem_obj <- if (!is.null(fit_obj$fit) && !is.null(fit_obj$fit$adaptive)) {
    fit_obj$fit
  } else if (!is.null(fit_obj$sem) && !is.null(fit_obj$sem$adaptive)) {
    fit_obj$sem
  } else {
    NULL
  }
  if (is.null(sem_obj) || is.null(sem_obj$adaptive)) return(NULL)
  ad <- sem_obj$adaptive
  if (!is.null(ad$metrics) && length(ad$metrics) > 0) {
    return(as.numeric(ad$metrics))
  }
  # Backward compatibility: reconstruct selected metric per iter as max over proposal metrics.
  if (!is.null(ad$all_metrics) && length(ad$all_metrics) > 0) {
    vals <- vapply(ad$all_metrics, function(v) {
      if (is.null(v) || length(v) == 0) return(NA_real_)
      max(as.numeric(v), na.rm = TRUE)
    }, numeric(1))
    vals[is.infinite(vals)] <- NA_real_
    return(vals)
  }
  NULL
}

# Drop the left prefix of scale-breaking SEM jumps (C-lock / MAP teleport)
# so later variation is visible. Keep original iteration indices.
sem_first_kept_iter <- function(y, frac = 0.10, range_keep = 0.25,
                                min_keep = 8L, early_frac = 0.2) {
  y <- as.numeric(y)
  n <- length(y)
  if (n < 3L) return(1L)
  finite <- is.finite(y)
  if (sum(finite) < 3L) return(1L)
  span <- diff(range(y[finite], na.rm = TRUE))
  final <- y[max(which(finite))]
  thresh <- if (is.finite(span) && span > 0) frac * span else Inf
  min_keep <- min(as.integer(min_keep), n)
  max_drop <- n - min_keep
  i <- 1L
  early_n <- max(2L, min(n - min_keep, ceiling(early_frac * n)))
  if (is.finite(thresh) && early_n >= 1L) {
    early_idx <- seq_len(early_n)
    early_idx <- early_idx[is.finite(y[early_idx])]
    if (length(early_idx)) {
      worst <- early_idx[which.max(abs(y[early_idx] - final))]
      if (abs(y[worst] - final) > thresh) i <- max(i, as.integer(worst) + 1L)
    }
  }
  i <- min(i, max_drop + 1L)
  while (i <= max_drop) {
    if (!is.finite(y[i])) { i <- i + 1L; next }
    if (is.finite(thresh) && abs(y[i] - final) > thresh) { i <- i + 1L; next }
    break
  }
  while (i <= max_drop) {
    rest <- y[i:n]
    rest <- rest[is.finite(rest)]
    rest2 <- y[(i + 1L):n]
    rest2 <- rest2[is.finite(rest2)]
    if (length(rest) < 2L || length(rest2) < 1L) break
    r1 <- diff(range(rest))
    r2 <- diff(range(rest2))
    if (is.finite(r1) && r1 > 0 && is.finite(r2) && (r2 / r1) < range_keep) {
      i <- i + 1L
      next
    }
    break
  }
  as.integer(i)
}

drop_sem_scale_prefix <- function(df, value_col, iter_col = "Iteration",
                                  group_cols = NULL) {
  if (is.null(df) || !is.data.frame(df) || nrow(df) < 1L) return(df)
  if (!value_col %in% names(df) || !iter_col %in% names(df)) return(df)
  if (is.null(group_cols)) {
    group_cols <- intersect(c("Model", "Metric", "entry"), names(df))
  }
  group_cols <- intersect(group_cols, names(df))
  split_keys <- if (length(group_cols)) {
    interaction(df[group_cols], drop = TRUE, lex.order = TRUE)
  } else {
    rep("all", nrow(df))
  }
  parts <- split(df, split_keys, drop = TRUE)
  out <- lapply(parts, function(d) {
    d <- d[order(as.numeric(d[[iter_col]])), , drop = FALSE]
    keep_from <- sem_first_kept_iter(d[[value_col]])
    if (keep_from > 1L && keep_from <= nrow(d)) {
      d <- d[seq.int(keep_from, nrow(d)), , drop = FALSE]
    }
    d
  })
  do.call(rbind, out)
}

get_sem_proposal_trace <- function(fit_obj) {
  if (is.null(fit_obj)) return(NULL)
  sem_obj <- if (!is.null(fit_obj$fit) && !is.null(fit_obj$fit$adaptive)) {
    fit_obj$fit
  } else if (!is.null(fit_obj$sem) && !is.null(fit_obj$sem$adaptive)) {
    fit_obj$sem
  } else {
    NULL
  }
  if (is.null(sem_obj) || is.null(sem_obj$adaptive$all_metrics)) return(NULL)
  am <- sem_obj$adaptive$all_metrics
  rows <- lapply(seq_along(am), function(i) {
    v <- as.numeric(am[[i]])
    if (!length(v)) return(NULL)
    data.frame(Iteration = i, Proposal = seq_along(v), LogLik = v)
  })
  rows <- Filter(Negate(is.null), rows)
  if (!length(rows)) return(NULL)
  do.call(rbind, rows)
}
```

## Marginal Parameters

```{r tbl-params}
#| label: tbl-params
beta_proxy <- suppressWarnings(as.numeric(unlist(cfg$BETA_GR, use.names = FALSE)))
beta_proxy <- beta_proxy[is.finite(beta_proxy)]
beta_proxy <- if (length(beta_proxy) > 0) beta_proxy[[1]] else 2.3

eta_exact <- function(mu, A, alpha_m, beta_gr = beta_proxy) {
  if (!all(is.finite(c(mu, A, alpha_m, beta_gr))) || beta_gr <= 0) return(NA_real_)
  if (alpha_m >= beta_gr) return(Inf)
  A * beta_gr / (beta_gr - alpha_m)
}

param_rows <- do.call(rbind, lapply(param_models, function(m) {
  mu_c <- bc(m$par, "mu"); A_c <- bc(m$par, "A"); a_c <- bc(m$par, "alpha_m")
  mu_t <- bt(m$par, "mu"); A_t <- bt(m$par, "A"); a_t <- bt(m$par, "alpha_m")
  eta_c <- eta_exact(mu_c, A_c, a_c)
  eta_t <- eta_exact(mu_t, A_t, a_t)
  rbind(
    data.frame(
      Fit = m$long, Process = "Control",
      mu = mu_c, A = A_c, alpha_m = a_c,
      branch_proxy = eta_c,
      mu_branch_proxy = if (is.finite(mu_c) && is.finite(eta_c)) mu_c * eta_c else NA_real_,
      stringsAsFactors = FALSE
    ),
    data.frame(
      Fit = m$long, Process = "Treated",
      mu = mu_t, A = A_t, alpha_m = a_t,
      branch_proxy = eta_t,
      mu_branch_proxy = if (is.finite(mu_t) && is.finite(eta_t)) mu_t * eta_t else NA_real_,
      stringsAsFactors = FALSE
    )
  )
}))

param_rows$Fit <- factor(param_rows$Fit, levels = vapply(param_models, function(m) m$long, character(1)))
param_rows <- param_rows[order(param_rows$Fit, param_rows$Process), , drop = FALSE]
rownames(param_rows) <- NULL
fit_end_rows <- which(param_rows$Process == "Treated")

tbl_k <- kable(
  param_rows,
  digits = 4,
  align = "llrrrrr",
  col.names = c("Fit", "Process", "mu", "A", "alpha_m", "Branching ratio", "mu * branching ratio"),
  caption = paste0(
    "Marginal ETAS parameters and exact branching ratios by fit. ",
    "Double separators mark fit boundaries. ",
    "The branching ratio is A*beta/(beta-alpha), with beta=",
    signif(beta_proxy, 4), "."
  ),
  booktabs = TRUE,
  linesep = "",
  hline_after = fit_end_rows
)

tbl_k |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE,
                font_size = 8, latex_options = c("scale_down", "hold_position")) |>
  row_spec(fit_end_rows, extra_css = "border-bottom: 3px double #444;")
```

## Cross-excitation

```{r tbl-cross}
#| label: tbl-cross
cn <- c("A_01","alpha_m_01","A_10","alpha_m_10")
cl <- c("$A_{01}$ (treat$\\to$ctrl)","$\\alpha_{m,01}$",
        "$A_{10}$ (ctrl$\\to$treat)","$\\alpha_{m,10}$")
cross_tbl <- data.frame(Parameter = cl)
safe_cross <- function(par, nm) {
  if (is.null(par)) return(NA_real_)
  if (is.list(par) && !is.null(par[[nm]])) return(scalar_num_or_na(par[[nm]]))
  scalar_num_or_na(par[[nm]])
}
for (m in param_models) {
  cross_tbl[[m$id]] <- sapply(cn, function(nm) safe_cross(m$par, nm))
}
rownames(cross_tbl) <- NULL
kable(cross_tbl, digits = 4, escape = FALSE,
      align = paste0("l", paste(rep("c", length(param_models)), collapse = "")),
      col.names = c("", vapply(param_models, function(m) m$label, character(1))),
      caption = "Cross-excitation parameters for available models.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Parameter Comparison

```{r fig-params}
#| label: fig-params
#| fig-cap: "Marginal parameter estimates across bivariate fits."
#| fig-height: 5
param_rows <- lapply(param_models, function(m) {
  rbind(
    data.frame(Fit = m$long, Process = "Control",
               mu = bc(m$par, "mu"), A = bc(m$par, "A"), alpha_m = bc(m$par, "alpha_m")),
    data.frame(Fit = m$long, Process = "Treated",
               mu = bt(m$par, "mu"), A = bt(m$par, "A"), alpha_m = bt(m$par, "alpha_m"))
  )
})
param_long <- do.call(rbind, param_rows)
pm <- pivot_longer(param_long, cols=c(mu,A,alpha_m), names_to="Parameter", values_to="Value")
pm <- pm[is.finite(pm$Value), , drop = FALSE]
pm$Parameter <- factor(pm$Parameter, levels=c("mu","A","alpha_m"), labels=c("mu","A","alpha[m]"))
pm$Fit <- factor(pm$Fit, levels = vapply(param_models, function(m) m$long, character(1)))
ggplot(pm, aes(x=Fit, y=Value, colour=Process)) +
  geom_point(size=3, position=position_dodge(0.4)) +
  facet_wrap(~Parameter, scales="free_y", labeller=label_parsed) +
  scale_colour_manual(values=c(Control="#377eb8",Treated="#e41a1c")) +
  labs(y="Estimate", x=NULL) +
  theme_minimal(base_size=13) +
  theme(axis.text.x=element_text(angle=30,hjust=1), strip.text=element_text(face="bold"), legend.position="top")
```

## SEM Convergence

```{r fig-flips}
#| label: fig-flips
#| fig-cap: "Accepted SEM confusion-matrix entry counts by inner iteration (county SEM fits). The initial scale-breaking MAP/C-lock jump is omitted so later 1-count changes are visible; each entry has its own y-scale."
#| fig-height: 8
extract_selected_confusion_trace <- function(sem_obj, model_label) {
  ad <- sem_obj$fit$adaptive
  if (is.null(ad) || is.null(ad$class_results) || is.null(ad$all_metrics)) return(NULL)
  if (length(ad$class_results) < 1 || length(ad$all_metrics) < 1) return(NULL)

  offset <- 0L
  rows <- list()
  for (i in seq_along(ad$all_metrics)) {
    metric_i <- ad$all_metrics[[i]]
    n_i <- length(metric_i)
    if (n_i < 1L) next
    best_j <- which.max(as.numeric(metric_i))
    class_idx <- offset + best_j
    if (class_idx > length(ad$class_results)) break
    cm <- ad$class_results[[class_idx]]
    if (is.null(cm)) {
      offset <- offset + n_i
      next
    }
    rows[[length(rows) + 1L]] <- data.frame(
      iteration = i,
      Model = model_label,
      control_as_control = as.numeric(if (!is.null(cm$control_as_control)) cm$control_as_control else 0),
      control_as_treated = as.numeric(if (!is.null(cm$control_as_treated)) cm$control_as_treated else 0),
      treated_as_control = as.numeric(if (!is.null(cm$treated_as_control)) cm$treated_as_control else 0),
      treated_as_treated = as.numeric(if (!is.null(cm$treated_as_treated)) cm$treated_as_treated else 0)
    )
    offset <- offset + n_i
  }
  if (length(rows) < 1L) return(NULL)
  do.call(rbind, rows)
}

trace_rows <- list()
sem_fit_convergence <- list(
  list(obj = if (!is.null(fit_B_obj)) fit_B_obj$fit else NULL, lab = "B: SEM Biv"),
  list(obj = if (!is.null(fit_D_obj)) fit_D_obj$fit else NULL, lab = "D: SEM Biv+KDE (all-free)"),
  list(obj = if (!is.null(fit_F_obj)) fit_F_obj$fit else NULL, lab = "F: SEM Biv+KDE (struct-fixed)"),
  list(obj = if (!is.null(fit_H_obj)) fit_H_obj$fit else NULL, lab = "H: SEM Univ+KDE"),
  list(obj = if (!is.null(fit_J_obj)) fit_J_obj$fit else NULL, lab = "J: SEM Univ (no KDE)")
)
for (nm in sem_fit_convergence) {
  if (is.null(nm$obj) || is.null(nm$obj$adaptive)) next
  trace_df <- extract_selected_confusion_trace(
    list(fit = nm$obj),
    nm$lab
  )
  if (!is.null(trace_df) && nrow(trace_df) > 0) {
    trace_rows[[length(trace_rows) + 1L]] <- trace_df
  } 
}

if (length(trace_rows) > 0) {
  trace_df <- do.call(rbind, trace_rows)
  trace_long <- tidyr::pivot_longer(
    trace_df,
    cols = c(control_as_control, control_as_treated, treated_as_control, treated_as_treated),
    names_to = "entry",
    values_to = "count"
  )
  trace_long$entry <- factor(
    trace_long$entry,
    levels = c("control_as_control", "control_as_treated", "treated_as_control", "treated_as_treated"),
    labels = c("Control -> Control", "Control -> Treated", "Treated -> Control", "Treated -> Treated")
  )
  trace_long <- drop_sem_scale_prefix(trace_long, value_col = "count", iter_col = "iteration")
}
if (length(trace_rows) > 0 && !is.null(trace_long) && nrow(trace_long) > 0) {
  ggplot(trace_long, aes(x = iteration, y = count, colour = entry)) +
    geom_line(linewidth = 0.9) +
    facet_grid(entry ~ Model, scales = "free_y") +
    scale_color_manual(name = "Confusion entry", values = c(
      "Control -> Control" = "#1f78b4",
      "Control -> Treated" = "#e31a1c",
      "Treated -> Control" = "#ff7f00",
      "Treated -> Treated" = "#33a02c"
    )) +
    labs(x = "Iteration", y = "Count") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none", strip.text.y = element_text(size = 8))
} else {
  plot.new()
  text(0.5, 0.5, "No SEM confusion-matrix trace found for county SEM fits (B/D/F/H/J).")
}
```

```{r fig-sem-adaptive-loglik}
#| label: fig-sem-adaptive-loglik
#| fig-cap: "Adaptive SEM selected-proposal score (solid) after omitting the initial scale-breaking MAP/C-lock jump. Grey points are other proposals in the same zoom window. Running best is dashed."
#| fig-height: 4.5
sem_metric_rows <- list()
sem_prop_rows <- list()
for (info in list(
  list(obj = fit_B_obj, lab = "B: SEM Biv"),
  list(obj = fit_D_obj, lab = "D: SEM Biv+KDE (all-free)"),
  list(obj = fit_F_obj, lab = "F: SEM Biv+KDE (struct-fixed)"),
  list(obj = fit_H_obj, lab = "H: SEM Univ+KDE"),
  list(obj = fit_J_obj, lab = "J: SEM Univ (no KDE)")
)) {
  tr <- get_sem_metric_trace(info$obj)
  if (!is.null(tr) && length(tr) > 0) {
    sem_metric_rows[[length(sem_metric_rows) + 1]] <- data.frame(
      Iteration = seq_along(tr),
      LogLik = as.numeric(tr),
      Model = info$lab
    )
  }
  pr <- get_sem_proposal_trace(info$obj)
  if (!is.null(pr) && nrow(pr) > 0) {
    pr$Model <- info$lab
    sem_prop_rows[[length(sem_prop_rows) + 1]] <- pr
  }
}
if (length(sem_metric_rows) > 0) {
  sem_metric_df <- do.call(rbind, sem_metric_rows)
  sem_metric_df <- sem_metric_df[is.finite(sem_metric_df$LogLik), , drop = FALSE]
  sem_metric_df <- drop_sem_scale_prefix(sem_metric_df, value_col = "LogLik")
}
if (length(sem_metric_rows) > 0 && !is.null(sem_metric_df) && nrow(sem_metric_df) > 0) {
  sem_metric_df <- sem_metric_df |>
    dplyr::group_by(.data$Model) |>
    dplyr::arrange(.data$Iteration, .by_group = TRUE) |>
    dplyr::mutate(RunningBest = cummax(.data$LogLik)) |>
    dplyr::ungroup()
  sem_metric_last <- sem_metric_df |>
    dplyr::group_by(.data$Model) |>
    dplyr::filter(.data$Iteration == max(.data$Iteration, na.rm = TRUE)) |>
    dplyr::slice_tail(n = 1) |>
    dplyr::ungroup() |>
    dplyr::mutate(lbl = sprintf("final = %.2f", .data$LogLik))
  sem_prop_zoom <- NULL
  if (length(sem_prop_rows) > 0) {
    sem_prop_df <- do.call(rbind, sem_prop_rows)
    kept <- unique(sem_metric_df[, c("Model", "Iteration")])
    sem_prop_df <- merge(sem_prop_df, kept, by = c("Model", "Iteration"))
    lims <- sem_metric_df |>
      dplyr::group_by(.data$Model) |>
      dplyr::summarise(
        ymin = min(.data$LogLik, na.rm = TRUE),
        ymax = max(.data$LogLik, na.rm = TRUE),
        .groups = "drop"
      ) |>
      dplyr::mutate(
        pad = pmax(0.08 * (.data$ymax - .data$ymin), 0.2),
        ymin = .data$ymin - .data$pad,
        ymax = .data$ymax + .data$pad
      )
    sem_prop_df <- merge(sem_prop_df, lims, by = "Model")
    sem_prop_zoom <- sem_prop_df[
      is.finite(sem_prop_df$LogLik) &
        sem_prop_df$LogLik >= sem_prop_df$ymin &
        sem_prop_df$LogLik <= sem_prop_df$ymax,
      , drop = FALSE
    ]
  }
  p_ll <- ggplot(sem_metric_df, aes(x = Iteration, y = LogLik, colour = Model))
  if (!is.null(sem_prop_zoom) && nrow(sem_prop_zoom) > 0) {
    p_ll <- p_ll + geom_point(
      data = sem_prop_zoom,
      aes(x = Iteration, y = LogLik),
      inherit.aes = FALSE,
      alpha = 0.18, size = 0.55, colour = "grey30"
    )
  }
  p_ll +
    geom_line(linewidth = 0.9) +
    geom_line(aes(y = RunningBest), linewidth = 0.7, linetype = "dashed", alpha = 0.7) +
    geom_text(
      data = sem_metric_last,
      aes(x = Inf, y = -Inf, label = lbl),
      inherit.aes = FALSE,
      hjust = 1.05, vjust = -0.4, size = 3.2, colour = "black"
    ) +
    scale_colour_manual(values = c(
      "B: SEM Biv" = "#6a3d9a",
      "D: SEM Biv+KDE (all-free)" = "#e31a1c",
      "F: SEM Biv+KDE (struct-fixed)" = "#1f78b4",
      "H: SEM Univ+KDE" = "#ff7f00",
      "J: SEM Univ (no KDE)" = "#b8860b"
    )) +
    labs(x = "Adaptive SEM iteration", y = "Selected SEM score (solid) / running best (dashed)") +
    facet_wrap(~Model, scales = "free_y") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
} else {
  plot.new()
  text(0.5, 0.5, "No adaptive SEM log-likelihood trace found in this results file.")
}
```

```{r fig-sem-cumulative-flips}
#| label: fig-sem-cumulative-flips
#| fig-cap: "Per-iteration SEM label flips after omitting the initial MAP Hamming spike. Average and accepted flips have separate y-scales so later 0–2 flip noise is visible."
#| fig-height: 6
flip_rows <- list()
for (info in list(
  list(obj = fit_B_obj, lab = "B: SEM Biv"),
  list(obj = fit_D_obj, lab = "D: SEM Biv+KDE (all-free)"),
  list(obj = fit_F_obj, lab = "F: SEM Biv+KDE (struct-fixed)"),
  list(obj = fit_H_obj, lab = "H: SEM Univ+KDE"),
  list(obj = fit_J_obj, lab = "J: SEM Univ (no KDE)")
)) {
  ad <- if (!is.null(info$obj)) info$obj$fit$adaptive else NULL
  if (is.null(ad) || is.null(ad$average_flips) || is.null(ad$max_metric_flips)) next
  n <- min(length(ad$average_flips), length(ad$max_metric_flips))
  if (n < 1) next
  tmp <- data.frame(
    Iteration = seq_len(n),
    Model = info$lab,
    AverageFlips = as.numeric(ad$average_flips[seq_len(n)]),
    AcceptedFlips = as.numeric(ad$max_metric_flips[seq_len(n)])
  )
  flip_rows[[length(flip_rows) + 1]] <- tmp
}
if (length(flip_rows) > 0) {
  flip_df <- do.call(rbind, flip_rows)
  flip_df <- flip_df |>
    tidyr::pivot_longer(
      cols = c(AverageFlips, AcceptedFlips),
      names_to = "Metric",
      values_to = "Flips"
    )
  flip_df$Metric <- factor(
    flip_df$Metric,
    levels = c("AverageFlips", "AcceptedFlips"),
    labels = c("Average flips", "Accepted flips")
  )
  flip_df <- drop_sem_scale_prefix(flip_df, value_col = "Flips")
}
if (length(flip_rows) > 0 && !is.null(flip_df) && nrow(flip_df) > 0) {
  ggplot(flip_df, aes(x = Iteration, y = Flips, colour = Metric)) +
    geom_line(linewidth = 0.9) +
    facet_grid(Metric ~ Model, scales = "free_y") +
    scale_colour_manual(values = c(
      "Average flips" = "#08519c",
      "Accepted flips" = "#a50f15"
    )) +
    labs(x = "Adaptive SEM iteration", y = "Flips per iteration") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
} else {
  plot.new()
  text(0.5, 0.5, "No SEM flip traces found in this results file.")
}
```

```{r tbl-final-loglik-compare}
#| label: tbl-final-loglik-compare
#| tbl-cap: "Final selected objective values from adaptive relabelling traces (higher is better)."
final_ll_rows <- list()

for (info in list(
  list(obj = fit_B_obj, lab = "B: SEM Biv", method = "SEM"),
  list(obj = fit_D_obj, lab = "D: SEM Biv+KDE (all-free)", method = "SEM"),
  list(obj = fit_F_obj, lab = "F: SEM Biv+KDE (struct-fixed)", method = "SEM"),
  list(obj = fit_H_obj, lab = "H: SEM Univ+KDE", method = "SEM"),
  list(obj = fit_J_obj, lab = "J: SEM Univ (no KDE)", method = "SEM")
)) {
  tr <- get_sem_metric_trace(info$obj)
  tr <- as.numeric(tr)
  tr <- tr[is.finite(tr)]
  if (length(tr) > 0) {
    final_ll_rows[[length(final_ll_rows) + 1]] <- data.frame(
      Model = info$lab,
      Method = info$method,
      Iterations = length(tr),
      Final_Objective = tail(tr, 1),
      stringsAsFactors = FALSE
    )
  }
}

if (length(final_ll_rows) > 0) {
  final_ll_tbl <- do.call(rbind, final_ll_rows)
  final_ll_tbl <- final_ll_tbl |>
    dplyr::arrange(.data$Method, .data$Model)
  final_ll_tbl$Final_Objective <- round(final_ll_tbl$Final_Objective, 2)
  kable(final_ll_tbl, align = "llrc") |>
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = FALSE)
}
```

## SEM Relabelling Summary

```{r tbl-cm-summary}
#| label: tbl-cm-summary
build_cm <- function(sem_obj) {
  if (is.null(sem_obj)) return(NULL)
  lab <- sem_obj$adaptive$adaptive_labelling
  post <- lab[lab$t >= 0, ]
  if (is.null(post$location_process) || is.null(post$inferred_process)) return(NULL)
  tab <- table(Location = post$location_process, Inferred = post$inferred_process)
  as.data.frame.matrix(tab)
}

sem_fit_obj <- function(obj) {
  if (is.null(obj)) return(NULL)
  if (!is.null(obj$fit) && !is.null(obj$fit$adaptive)) return(obj$fit)
  if (!is.null(obj$sem) && !is.null(obj$sem$adaptive)) return(obj$sem)
  if (!is.null(obj$adaptive)) return(obj)
  obj$fit %||% obj$sem %||% NULL
}

cm_specs <- list(
  list(cm = build_cm(sem_fit_obj(fit_B_obj)), lab = "B: SEM Bivariate (homogeneous)"),
  list(cm = build_cm(sem_fit_obj(fit_D_obj)), lab = "D: SEM Bivariate+KDE (all-free)"),
  list(cm = build_cm(sem_fit_obj(fit_F_obj)), lab = "F: SEM Bivariate+KDE (struct-fixed)"),
  list(cm = build_cm(sem_fit_obj(fit_H_obj)), lab = "H: SEM Univariate+KDE"),
  list(cm = build_cm(sem_fit_obj(fit_J_obj)), lab = "J: SEM Univariate (homogeneous)")
)

sum_rows <- list()
for (info in cm_specs) {
  cm <- info$cm
  if (is.null(cm)) next
  n_total <- sum(cm)
  ctrl_to_treat <- if ("treated" %in% colnames(cm) && "control" %in% rownames(cm)) {
    cm["control", "treated"]
  } else {
    0
  }
  treat_to_ctrl <- if ("control" %in% colnames(cm) && "treated" %in% rownames(cm)) {
    cm["treated", "control"]
  } else {
    0
  }
  n_flipped <- ctrl_to_treat + treat_to_ctrl
  sum_rows[[length(sum_rows) + 1L]] <- data.frame(
    Model = info$lab,
    `Post events` = n_total,
    `Ctrl to Treat` = ctrl_to_treat,
    `Treat to Ctrl` = treat_to_ctrl,
    `Total relabelled` = n_flipped,
    `Pct relabelled` = paste0(round(n_flipped / n_total * 100, 1), "%"),
    check.names = FALSE
  )
}
if (length(sum_rows) > 0) {
  kable(do.call(rbind, sum_rows),
        caption = "Summary of SEM relabelling for all county SEM fits (B, D, F, H, J).",
        align = "lccccc") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("No SEM relabelling summaries available.\n")
}
```

### Triggering Distance Diagnostics (All Fits)

```{r tbl-trigger-distance-all}
#| label: tbl-trigger-distance-all
diag_rows <- list()
for (fit_info in model_defs) {
  if (is.null(fit_info$par)) next
  m <- etas_trigger_means(fit_info$par, cfg$BETA_GR, cfg$SEM_T_TRUNC_DAYS)
  for (proc in c("Control", "Treated")) {
    diag_rows[[length(diag_rows) + 1]] <- data.frame(
      Model = fit_info$long,
      Process = proc,
      `Mean trigger time (days)` = if (is.finite(m$mean_time)) round(m$mean_time, 3) else Inf,
      `Mean trigger distance (km)` = if (is.finite(m$mean_distance)) round(m$mean_distance, 3) else Inf,
      check.names = FALSE
    )
    if (is.finite(m$mean_time_trunc) && !is.na(m$mean_time_trunc)) {
      diag_rows[[length(diag_rows) + 1]] <- data.frame(
        Model = fit_info$long,
        Process = paste0(proc, " (t<=t_trunc)"),
        `Mean trigger time (days)` = round(m$mean_time_trunc, 3),
        `Mean trigger distance (km)` = if (is.finite(m$mean_distance)) round(m$mean_distance, 3) else Inf,
        check.names = FALSE
      )
    }
  }
}
if (length(diag_rows) > 0) {
  diag_df <- do.call(rbind, diag_rows)
  kable(diag_df, align = "llcc",
        caption = paste0(
          "Model-implied ETAS triggering means across all available fits. ",
          "Rows marked '(t<=t_trunc)' report temporal means under the SEM truncation window. ",
          "Infinite values indicate heavy-tail regimes where the corresponding mean does not exist ",
          "(time: p <= 2, space: q <= 1.5)."
        )) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
}
```

# All-or-Nothing Treatment Effect

## Primary Tessellation

```{r tbl-ate-helpers}
#| label: tbl-ate-helpers
#| include: false
n_cty <- suppressWarnings(as.integer(unlist(cfg$n_tiles, use.names = FALSE)))
if (length(n_cty) < 1L || !is.finite(n_cty[[1]]) || is.na(n_cty[[1]]) || n_cty[[1]] < 1L) {
  n_cty <- suppressWarnings(as.integer(unlist(cfg$n_counties, use.names = FALSE)))
}
if (length(n_cty) < 1L || !is.finite(n_cty[[1]]) || is.na(n_cty[[1]]) || n_cty[[1]] < 1L) {
  n_cty <- suppressWarnings(as.integer(unlist(results$counties$n_counties, use.names = FALSE)))
}
if (length(n_cty) < 1L || !is.finite(n_cty[[1]]) || is.na(n_cty[[1]]) || n_cty[[1]] < 1L) {
  n_cty <- 1L
}
n_cty <- as.integer(n_cty[[1]])
ate_days <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 365
obs_post_n <- NA_integer_
if (!is.null(results$pp_data) && !is.null(results$pp_data$pp_post)) {
  pp_post_df <- coerce_pp_df(results$pp_data$pp_post)
  if (nrow(pp_post_df) > 0 && "t" %in% names(pp_post_df)) {
    t_num <- suppressWarnings(as.numeric(pp_post_df$t))
    if (all(!is.na(t_num))) {
      obs_post_n <- as.integer(sum(t_num >= 0 & t_num <= ate_days))
    } else {
      obs_post_n <- as.integer(nrow(pp_post_df))
    }
  } else {
    obs_post_n <- as.integer(nrow(pp_post_df))
  }
}

get_total_effect_sim <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  if (!is.finite(n_tiles) || is.na(n_tiles) || n_tiles < 1) n_tiles <- 1
  sim_df <- ate_obj$all_nothing_sim
  if ("total_saved" %in% names(sim_df)) return(sim_df$total_saved)
  if ("total_effect" %in% names(sim_df)) return(sim_df$total_effect)
  sim_df$ATE * n_tiles
}

get_expected_totals <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  sim_df <- ate_obj$all_nothing_sim
  if (all(c("c_total", "t_total") %in% names(sim_df))) {
    return(list(c_total = sim_df$c_total, t_total = sim_df$t_total))
  }
  list(c_total = sim_df$c_mean * n_tiles, t_total = sim_df$t_mean * n_tiles)
}

paper_ok_counts_row <- function(method_label, ate_obj, estimand_label = NA_character_) {
  expected <- tryCatch(get_expected_totals(ate_obj, n_tiles = n_cty), error = function(e) NULL)
  if (is.null(expected)) return(NULL)
  c_mean <- mean(expected$c_total, na.rm = TRUE)
  t_mean <- mean(expected$t_total, na.rm = TRUE)
  if (!is.finite(c_mean) || !is.finite(t_mean)) return(NULL)
  data.frame(
    Method = method_label,
    Estimand = estimand_label,
    `E[N(control)]` = round(c_mean, 0),
    `E[N(right)]` = round(t_mean, 0),
    `Delta = control - right` = round(c_mean - t_mean, 0),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}
pick_ate_contrast_early <- function(ate_obj, contrast = c("all_or_nothing", "observed")) {
  contrast <- match.arg(contrast)
  if (is.null(ate_obj)) return(NULL)
  bc <- ate_obj$by_contrast
  if (is.list(bc) && !is.null(bc[[contrast]])) return(bc[[contrast]])
  if (!is.null(ate_obj$contrast) && identical(as.character(ate_obj$contrast)[1], contrast)) {
    return(ate_obj)
  }
  if (identical(contrast, "all_or_nothing") && !is.null(ate_obj$all_nothing_sim)) {
    return(ate_obj)
  }
  NULL
}
```

### Paper-style counts: univariate Naive vs SEM (G–J)

Compact analogue of manuscript Table~13 (`tab:ok_counts`) for univariate fits,
with both all-or-nothing and observed-vs-control estimands when available.

```{r tbl-ok-counts-univ}
#| label: tbl-ok-counts-univ
univ_fit_specs <- list(
  list(lab = "Naive (G, univariate+KDE)", ate = if (!is.null(fit_G_obj)) fit_G_obj$ate else NULL),
  list(lab = "SEM (H, univariate+KDE)", ate = if (!is.null(fit_H_obj)) fit_H_obj$ate else NULL),
  list(lab = "Naive (I, univariate, no KDE)", ate = if (!is.null(fit_I_obj)) fit_I_obj$ate else NULL),
  list(lab = "SEM (J, univariate, no KDE)", ate = if (!is.null(fit_J_obj)) fit_J_obj$ate else NULL)
)
univ_count_rows <- Filter(Negate(is.null), unlist(lapply(univ_fit_specs, function(sp) {
  list(
    paper_ok_counts_row(sp$lab, pick_ate_contrast_early(sp$ate, "all_or_nothing"), "All-or-nothing"),
    paper_ok_counts_row(sp$lab, pick_ate_contrast_early(sp$ate, "observed"), "Observed vs control")
  )
}), recursive = FALSE))
if (length(univ_count_rows) > 0L) {
  univ_counts <- do.call(rbind, univ_count_rows)
  kable(
    univ_counts,
    align = "llrrr",
    caption = paste0(
      "Expected post-treatment earthquake counts for univariate fits I–L under both estimands ",
      "(All-or-nothing: right = treated everywhere; Observed: right = AOI regime; ",
      ate_days, "-day window). Point estimates only; bootstrap uncertainty omitted."
    )
  ) |>
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = FALSE)
} else {
  cat("Univariate I–L ATE simulations are unavailable in this results file.\n")
}
```

### Full model comparison (all available fits)

```{r tbl-ate-biv}
#| label: tbl-ate-biv

biv_rows <- list(
  list(lab = "A: Naive Bivariate", ate = if (!is.null(fit_A_obj)) fit_A_obj$ate else NULL, louis = NULL),
  list(lab = "B: SEM Bivariate",   ate = if (!is.null(fit_B_obj)) fit_B_obj$ate else NULL, louis = if (!is.null(fit_B_obj)) fit_B_obj$louis else NULL),
  list(lab = "C: Naive Biv+KDE (all-free)",   ate = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL, louis = NULL),
  list(lab = "D: SEM Biv+KDE (all-free)",     ate = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL, louis = if (!is.null(fit_D_obj)) fit_D_obj$louis else NULL),
  list(lab = "E: Naive Biv+KDE (struct-fixed)", ate = if (!is.null(fit_E_obj)) fit_E_obj$ate else NULL, louis = NULL),
  list(lab = "F: SEM Biv+KDE (struct-fixed)",   ate = if (!is.null(fit_F_obj)) fit_F_obj$ate else NULL, louis = if (!is.null(fit_F_obj)) fit_F_obj$louis else NULL))
if (!is.null(fit_G_obj) && !is.null(fit_G_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "G: Naive Univ+KDE", ate = fit_G_obj$ate, louis = NULL)
}
if (!is.null(fit_H_obj) && !is.null(fit_H_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "H: SEM Univ+KDE", ate = fit_H_obj$ate, louis = NULL)
}
if (!is.null(fit_I_obj) && !is.null(fit_I_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "I: Naive Univ (no KDE)", ate = fit_I_obj$ate, louis = NULL)
}
if (!is.null(fit_J_obj) && !is.null(fit_J_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "J: SEM Univ (no KDE)", ate = fit_J_obj$ate, louis = NULL)
}

boot_obj <- results$bootstrap_ate
# Primary bootstrap pair is C/D (all_free). Accept legacy E/F target tags.
boot_primary_ids <- c("C", "D")
if (!is.null(boot_obj) && !is.null(boot_obj$config) && !is.null(boot_obj$config$targets)) {
  tgt <- unique(toupper(trimws(as.character(boot_obj$config$targets))))
  if (length(tgt) >= 2 && all(c("C", "D") %in% tgt)) {
    boot_primary_ids <- c("C", "D")
  } else if (length(tgt) >= 2 && all(c("E", "F") %in% tgt)) {
    boot_primary_ids <- c("C", "D")
  }
}
model_label_by_id <- function(id) {
  switch(
    toupper(id),
    "C" = "C: Naive Biv+KDE (all-free)",
    "D" = "D: SEM Biv+KDE (all-free)",
    "E" = "E: Naive Biv+KDE (struct-fixed)",
    "F" = "F: SEM Biv+KDE (struct-fixed)",
    paste0(id, ": model")
  )
}
get_boot_df <- function(id) {
  if (is.null(boot_obj)) return(NULL)
  stable_rows <- function(df) {
    df <- as.data.frame(df)
    if ("is_stable" %in% names(df)) {
      df <- df[!is.na(df$is_stable) & df$is_stable, , drop = FALSE]
    }
    df
  }
  key <- paste0("fit_", toupper(id))
  if (!is.null(boot_obj[[key]]) &&
      !is.null(boot_obj[[key]]$replicate_summary) &&
      nrow(boot_obj[[key]]$replicate_summary) > 0) {
    return(stable_rows(boot_obj[[key]]$replicate_summary))
  }
  # Cross-alias: legacy primary stored under fit_E/fit_F maps to C/D.
  alias <- switch(toupper(id), "C" = "fit_E", "D" = "fit_F", NA_character_)
  if (is.character(alias) && !is.na(alias) &&
      !is.null(boot_obj[[alias]]) &&
      !is.null(boot_obj[[alias]]$replicate_summary) &&
      nrow(boot_obj[[alias]]$replicate_summary) > 0) {
    return(stable_rows(boot_obj[[alias]]$replicate_summary))
  }
  NULL
}
boot_naive_id <- boot_primary_ids[[1]]
boot_sem_id <- boot_primary_ids[[2]]
boot_naive_df <- get_boot_df(boot_naive_id)
boot_sem_df <- get_boot_df(boot_sem_id)
boot_naive_label <- model_label_by_id(boot_naive_id)
boot_sem_label <- model_label_by_id(boot_sem_id)
boot_pair_label <- paste0(boot_naive_id, "/", boot_sem_id)

# Explosive refits are removed in the analysis pipeline before these summaries.
# Retain every finite ATE from the stable remainder; do not apply an ATE-tail
# filter, because a large but stable estimate is part of parameter uncertainty.
retain_boot_vals <- function(x) {
  x <- suppressWarnings(as.numeric(x))
  ok <- is.finite(x)
  list(
    vals = x[ok],
    n_in = length(x),
    n_out = as.integer(sum(!ok)),
    dropped = x[!ok]
  )
}

get_boot_summary <- function(model_label) {
  vals <- NULL
  if (model_label == boot_naive_label && !is.null(boot_naive_df)) vals <- boot_naive_df$ate_total_mean
  if (model_label == boot_sem_label && !is.null(boot_sem_df)) vals <- boot_sem_df$ate_total_mean
  # Compatibility fallback if table labels differ from selected pair labels.
  if (is.null(vals) && model_label == "C: Naive Biv+KDE (all-free)" && !is.null(get_boot_df("C"))) vals <- get_boot_df("C")$ate_total_mean
  if (is.null(vals) && model_label == "D: SEM Biv+KDE (all-free)" && !is.null(get_boot_df("D"))) vals <- get_boot_df("D")$ate_total_mean
  if (is.null(vals)) return(NULL)
  vals <- retain_boot_vals(vals)$vals
  if (length(vals) < 2) return(NULL)
  list(
    sd = stats::sd(vals, na.rm = TRUE),
    lo = as.numeric(stats::quantile(vals, probs = 0.025, na.rm = TRUE)),
    hi = as.numeric(stats::quantile(vals, probs = 0.975, na.rm = TRUE)),
    n = length(vals),
    vals = vals
  )
}

boot_ci_efron_bc <- function(boot_vals, theta_hat, conf = 0.95) {
  x <- suppressWarnings(as.numeric(boot_vals))
  x <- x[is.finite(x)]
  theta_hat <- suppressWarnings(as.numeric(theta_hat))
  if (length(x) < 5L || !is.finite(theta_hat) || is.na(theta_hat)) {
    return(NULL)
  }
  alpha <- (1 - conf) / 2
  q_pct <- stats::quantile(x, probs = c(alpha, 1 - alpha), na.rm = TRUE, names = FALSE)
  p_less <- mean(x < theta_hat, na.rm = TRUE)
  # Guard against p=0 or p=1 leading to infinite z0.
  eps <- 1 / (length(x) + 1)
  p_less <- min(max(p_less, eps), 1 - eps)
  z0 <- stats::qnorm(p_less)
  z_lo <- stats::qnorm(alpha)
  z_hi <- stats::qnorm(1 - alpha)
  a_lo <- stats::pnorm(2 * z0 + z_lo)
  a_hi <- stats::pnorm(2 * z0 + z_hi)
  a_lo <- min(max(a_lo, eps), 1 - eps)
  a_hi <- min(max(a_hi, eps), 1 - eps)
  if (a_lo > a_hi) {
    tmp <- a_lo; a_lo <- a_hi; a_hi <- tmp
  }
  # If BC collapses to a numerically degenerate interval, use percentile fallback.
  if (!is.finite(a_lo) || !is.finite(a_hi) || is.na(a_lo) || is.na(a_hi) ||
      (a_hi - a_lo) < (1 / length(x))) {
    return(list(
      lo = as.numeric(q_pct[[1]]),
      hi = as.numeric(q_pct[[2]]),
      z0 = z0,
      p_less = p_less,
      alpha_lo = alpha,
      alpha_hi = 1 - alpha,
      method = "percentile_fallback"
    ))
  }
  lo <- as.numeric(stats::quantile(x, probs = a_lo, na.rm = TRUE, names = FALSE))
  hi <- as.numeric(stats::quantile(x, probs = a_hi, na.rm = TRUE, names = FALSE))
  list(lo = lo, hi = hi, z0 = z0, p_less = p_less, alpha_lo = a_lo, alpha_hi = a_hi, method = "bc")
}

boot_ci_quantile_length <- function(boot_vals, theta_hat, conf = 0.95, center_stat = c("median", "mean")) {
  center_stat <- match.arg(center_stat)
  x <- suppressWarnings(as.numeric(boot_vals))
  x <- x[is.finite(x)]
  theta_hat <- suppressWarnings(as.numeric(theta_hat))
  if (length(x) < 5L || !is.finite(theta_hat) || is.na(theta_hat)) return(NULL)
  alpha <- (1 - conf) / 2
  q_lo <- as.numeric(stats::quantile(x, probs = alpha, na.rm = TRUE, names = FALSE))
  q_hi <- as.numeric(stats::quantile(x, probs = 1 - alpha, na.rm = TRUE, names = FALSE))
  center <- if (identical(center_stat, "median")) stats::median(x, na.rm = TRUE) else mean(x, na.rm = TRUE)
  if (!is.finite(q_lo) || !is.finite(q_hi) || !is.finite(center)) return(NULL)
  left_len <- center - q_lo
  right_len <- q_hi - center
  if (!is.finite(left_len) || !is.finite(right_len)) return(NULL)
  list(
    lo = theta_hat - left_len,
    hi = theta_hat + right_len,
    q_lo = q_lo,
    q_hi = q_hi,
    center = center
  )
}

format_partition_label <- function(part_id) {
  if (is.na(part_id) || is.null(part_id)) return(part_id)
  if (part_id == "county") return("County")
  if (part_id == "aoi_region") return("AOI region")
  if (part_id == "grid_coarse") return("Grid (coarse)")
  m <- regexec("^grid_([0-9.]+)R$", part_id)
  g <- regmatches(part_id, m)[[1]]
  if (length(g) == 2) {
    mult <- sub("\\.0$", "", g[2])
    return(paste0("Grid (", mult, "x triggering range)"))
  }
  part_id
}

# Prefer contrast-specific payload from $by_contrast when present (new dual-estimand runs).
pick_ate_contrast <- function(ate_obj, contrast = c("all_or_nothing", "observed")) {
  contrast <- match.arg(contrast)
  if (is.null(ate_obj)) return(NULL)
  bc <- ate_obj$by_contrast
  if (is.list(bc) && !is.null(bc[[contrast]])) return(bc[[contrast]])
  if (!is.null(ate_obj$contrast) && identical(as.character(ate_obj$contrast)[1], contrast)) {
    return(ate_obj)
  }
  # Legacy single-contrast objects: only reuse for all_or_nothing if sims are present.
  if (identical(contrast, "all_or_nothing") && !is.null(ate_obj$all_nothing_sim)) {
    return(ate_obj)
  }
  NULL
}

estimand_specs <- list(
  list(
    id = "all_or_nothing",
    label = "All-or-nothing",
    right_lab = "E[N(treated everywhere)]"
  ),
  list(
    id = "observed",
    label = "Observed vs control",
    right_lab = "E[N(actual treatment regime)]"
  )
)

ate_row_for <- function(r, est) {
  ate_use <- pick_ate_contrast(r$ate, est$id)
  if (is.null(ate_use)) return(NULL)
  expected <- tryCatch(get_expected_totals(ate_use, n_tiles = n_cty), error = function(e) NULL)
  sim_totals <- tryCatch(get_total_effect_sim(ate_use, n_tiles = n_cty), error = function(e) NULL)
  c_mean <- if (!is.null(expected) && !is.null(expected$c_total) && length(expected$c_total) > 0) {
    mean(expected$c_total, na.rm = TRUE)
  } else { NA_real_ }
  t_mean <- if (!is.null(expected) && !is.null(expected$t_total) && length(expected$t_total) > 0) {
    mean(expected$t_total, na.rm = TRUE)
  } else { NA_real_ }
  c_sd <- if (!is.null(expected) && !is.null(expected$c_total) && length(expected$c_total) > 1L) {
    sd(expected$c_total, na.rm = TRUE)
  } else { NA_real_ }
  t_sd <- if (!is.null(expected) && !is.null(expected$t_total) && length(expected$t_total) > 1L) {
    sd(expected$t_total, na.rm = TRUE)
  } else { NA_real_ }
  saved_diff_mean <- if (is.finite(c_mean) && is.finite(t_mean)) c_mean - t_mean else NA_real_
  sim_sd_total  <- if (!is.null(sim_totals) && length(sim_totals) > 1L) sd(sim_totals, na.rm = TRUE) else NA_real_
  base_row <- data.frame(
    Model = r$lab,
    Estimand = est$label,
    `Observed N(actual regime)` = obs_post_n,
    `E[N(control everywhere)]` = round(c_mean, 0),
    `E[N(right-hand regime)]` = round(t_mean, 0),
    `SD(control everywhere)` = round(c_sd, 0),
    `SD(right-hand regime)` = round(t_sd, 0),
    `Saved = control - right` = round(saved_diff_mean, 0),
    `SD(saved)` = round(sim_sd_total, 0),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  if (using_pre_bootstrap) return(base_row)
  # Bootstrap uncertainty only for primary all-free pair (C/D) on the run's primary contrast.
  boot_models_for_table <- c("C: Naive Biv+KDE (all-free)", "D: SEM Biv+KDE (all-free)")
  primary_contrast <- tolower(as.character(cfg$ATE_CONTRAST %||% "all_or_nothing")[1])
  boot_sum <- if (r$lab %in% boot_models_for_table && identical(est$id, primary_contrast)) {
    get_boot_summary(r$lab)
  } else {
    NULL
  }
  boot_sd <- if (!is.null(boot_sum)) round(boot_sum$sd, 0) else NA
  boot_ci_q_len <- if (!is.null(boot_sum) && is.finite(saved_diff_mean)) {
    ci <- boot_ci_quantile_length(boot_sum$vals, saved_diff_mean, conf = 0.95, center_stat = "median")
    if (!is.null(ci) && is.finite(ci$lo) && is.finite(ci$hi)) {
      sprintf("[%d, %d]", round(ci$lo, 0), round(ci$hi, 0))
    } else {
      "NA"
    }
  } else {
    "NA"
  }
  base_row$`Bootstrap SD (ATE)` <- boot_sd
  base_row$`95% quantile-length CI` <- boot_ci_q_len
  base_row
}

ate_rows <- unlist(lapply(biv_rows, function(r) {
  lapply(estimand_specs, function(est) ate_row_for(r, est))
}), recursive = FALSE)
ate_rows <- Filter(Negate(is.null), ate_rows)
ate_biv <- if (length(ate_rows) > 0L) do.call(rbind, ate_rows) else NULL
if (!is.null(ate_biv) && nrow(ate_biv) > 0) {
  core_cols <- c(
    "Model",
    "Estimand",
    "Observed N(actual regime)",
    "E[N(control everywhere)]",
    "E[N(right-hand regime)]",
    "SD(control everywhere)",
    "SD(right-hand regime)",
    "Saved = control - right",
    "SD(saved)"
  )
  if (using_pre_bootstrap) {
    ate_biv <- ate_biv[, core_cols, drop = FALSE]
  } else {
    full_cols <- c(core_cols, "Bootstrap SD (ATE)", "95% quantile-length CI")
    ate_biv <- ate_biv[, full_cols, drop = FALSE]
  }
  kable(ate_biv,
        col.names = if (using_pre_bootstrap) {
          c(
            "Model", "Estimand",
            paste0("Observed N(actual regime) / ", ate_days, "d"),
            "E[N(control everywhere)]",
            "E[N(right-hand regime)]",
            "SD(control everywhere)", "SD(right-hand regime)",
            paste0("Saved = control - right / ", n_cty, " tiles / ", ate_days, "d"),
            "SD(saved)"
          )
        } else {
          c(
            "Model", "Estimand",
            paste0("Observed N(actual regime) / ", ate_days, "d"),
            "E[N(control everywhere)]",
            "E[N(right-hand regime)]",
            "SD(control everywhere)", "SD(right-hand regime)",
            paste0("Saved = control - right / ", n_cty, " tiles / ", ate_days, "d"),
            "SD(saved)", "Bootstrap SD (ATE)", "95% quantile-length CI"
          )
        },
        caption = paste0(
          "Expected counts over ", ate_days, " days for both estimands: ",
          "All-or-nothing (right-hand = treated everywhere) and Observed vs control ",
          "(right-hand = actual AOI regime). Saved = E[N(control everywhere)] - E[N(right-hand)]. ",
          if (using_pre_bootstrap) {
            "Point estimates only (pre-bootstrap checkpoint)."
          } else {
            paste0(
              "Bootstrap SD/CI (when present) apply to the run's primary contrast only; ",
              "quantile-length 95% intervals recenter bootstrap lengths around the point ATE."
            )
          }
        ),
        align = if (using_pre_bootstrap) "llccccccc" else "llccccccccc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE)
} else {
  cat(sprintf("ATE table unavailable in this checkpoint stage (%s). Run until pre-bootstrap or full stage for Table 6.1.\n",
              if (!is.null(results$checkpoint) && !is.null(results$checkpoint$stage)) results$checkpoint$stage else "unknown"))
}
```

### Combined parameter comparison (all fits A–J)

```{r tbl-params-all-fits}
#| label: tbl-params-all-fits
# Unified A–J comparison only (pair-wise tables removed).
# Bivariate fits use shared structural (c,p,D,q); univariate fits map A_k -> A_kk,
# leave cross channels blank, and keep margin-specific structural terms.
to_named_num <- function(x) {
  if (is.null(x)) return(setNames(numeric(0), character(0)))
  if (is.list(x)) x <- unlist(x, use.names = TRUE)
  v <- suppressWarnings(as.numeric(x))
  nms <- names(x)
  if (is.null(nms)) nms <- rep("", length(v))
  names(v) <- nms
  v
}
beta_gr_tbl <- suppressWarnings(as.numeric(unlist(cfg$BETA_GR, use.names = FALSE)))
beta_gr_tbl <- beta_gr_tbl[is.finite(beta_gr_tbl) & beta_gr_tbl > 0]
beta_gr_tbl <- if (length(beta_gr_tbl)) beta_gr_tbl[[1]] else 2.3
etas_branch_matrix <- function(par, beta_gr) {
  getp <- function(nm) {
    v <- suppressWarnings(as.numeric(par[[nm]]))
    if (!length(v) || !is.finite(v[[1]])) return(NA_real_)
    v[[1]]
  }
  eta_ij <- function(A, alpha) {
    if (!is.finite(A) || !is.finite(alpha) || !is.finite(beta_gr) || beta_gr <= 0) return(NA_real_)
    gap <- beta_gr - alpha
    if (!is.finite(gap) || gap <= 1e-8) return(Inf)
    A * beta_gr / gap
  }
  matrix(
    c(
      eta_ij(getp("A_00"), getp("alpha_m_00")),
      eta_ij(getp("A_01"), getp("alpha_m_01")),
      eta_ij(getp("A_10"), getp("alpha_m_10")),
      eta_ij(getp("A_11"), getp("alpha_m_11"))
    ),
    nrow = 2, byrow = TRUE,
    dimnames = list(c("from_0", "from_1"), c("to_0", "to_1"))
  )
}
spectral_radius_etas <- function(par, beta_gr) {
  M <- etas_branch_matrix(par, beta_gr)
  if (!all(is.finite(M))) return(Inf)
  max(Re(eigen(M, only.values = TRUE)$values))
}
univ_eta <- function(par, beta_gr) {
  if (is.null(par)) return(NA_real_)
  pv <- to_named_num(par)
  A <- suppressWarnings(as.numeric(pv[["A"]]))
  alpha <- suppressWarnings(as.numeric(pv[["alpha_m"]]))
  if (!length(A) || !length(alpha) || !is.finite(A[[1]]) || !is.finite(alpha[[1]]) ||
      !is.finite(beta_gr) || beta_gr <= 0) {
    return(NA_real_)
  }
  gap <- beta_gr - alpha[[1]]
  if (!is.finite(gap) || gap <= 1e-8) return(Inf)
  A[[1]] * beta_gr / gap
}
extract_univ_margin <- function(fit_obj, margin) {
  if (is.null(fit_obj) || is.null(fit_obj$params)) return(NULL)
  p <- fit_obj$params
  if (!is.null(p[[margin]])) return(p[[margin]])
  NULL
}
fmt_sigfig_all <- function(x, digits = 3L) {
  if (is.null(x) || length(x) < 1L || is.na(x)) return("—")
  if (!is.finite(x)) return("Inf")
  v <- signif(as.numeric(x), digits)
  if (!is.finite(v) || v == 0) return(format(v, scientific = FALSE, trim = TRUE))
  order <- floor(log10(abs(v)))
  dp <- max(0L, as.integer(digits) - order - 1L)
  sprintf(paste0("%.", dp, "f"), v)
}
pull_num <- function(par, nm) {
  if (is.null(par)) return(NA_real_)
  v <- to_named_num(par)
  suppressWarnings(as.numeric(v[[nm]]))
}
biv_col <- function(fit_obj) {
  p <- to_named_num(if (!is.null(fit_obj)) fit_obj$params else NULL)
  if (!length(p)) return(NULL)
  list(
    kind = "biv",
    p = p,
    rho = spectral_radius_etas(p, beta_gr_tbl)
  )
}
univ_col <- function(fit_obj) {
  ctrl <- extract_univ_margin(fit_obj, "control")
  treat <- extract_univ_margin(fit_obj, "treated")
  if (is.null(ctrl) && is.null(treat)) return(NULL)
  eta0 <- univ_eta(ctrl, beta_gr_tbl)
  eta1 <- univ_eta(treat, beta_gr_tbl)
  rho <- max(c(eta0, eta1), na.rm = TRUE)
  if (!is.finite(rho)) rho <- Inf
  list(kind = "univ", ctrl = ctrl, treat = treat, rho = rho)
}
fit_cols <- list(
  A = biv_col(fit_A_obj), B = biv_col(fit_B_obj),
  C = biv_col(fit_C_obj), D = biv_col(fit_D_obj),
  E = biv_col(fit_E_obj), F = biv_col(fit_F_obj),
  G = univ_col(fit_G_obj), H = univ_col(fit_H_obj),
  I = univ_col(fit_I_obj), J = univ_col(fit_J_obj)
)
param_rows_all <- c(
  "mu_0", "mu_1",
  "A_00", "alpha_m_00", "A_11", "alpha_m_11",
  "A_01", "alpha_m_01", "A_10", "alpha_m_10",
  "c_0", "p_0", "D_0", "q_0",
  "c_1", "p_1", "D_1", "q_1",
  "rho", "explosive?"
)
val_for <- function(col, nm) {
  if (is.null(col)) return(NA_real_)
  if (identical(nm, "rho")) return(col$rho)
  if (identical(nm, "explosive?")) {
    return(if (!is.finite(col$rho)) Inf else if (col$rho >= 1) 1 else 0)
  }
  if (identical(col$kind, "biv")) {
    base <- sub("_[01]$", "", nm)
    if (grepl("^[cpDq]_", nm) || nm %in% c("c", "p", "D", "q")) {
      # Shared structural under bivariate law.
      return(pull_num(col$p, base))
    }
    return(pull_num(col$p, nm))
  }
  # univariate
  if (grepl("_0$", nm)) {
    return(pull_num(col$ctrl, sub("_0$", "", nm)))
  }
  if (grepl("_1$", nm)) {
    return(pull_num(col$treat, sub("_1$", "", nm)))
  }
  if (grepl("_(00|11)$", nm)) {
    margin <- if (grepl("_00$", nm)) col$ctrl else col$treat
    return(pull_num(margin, sub("_[01]{2}$", "", nm)))
  }
  # Cross channels absent under independent margins.
  if (grepl("_(01|10)$", nm)) return(NA_real_)
  NA_real_
}
fmt_cell <- function(col, nm) {
  if (identical(nm, "explosive?")) {
    if (is.null(col)) return("—")
    if (!is.finite(col$rho)) return("yes (Inf)")
    if (col$rho >= 1) "yes" else "no"
  } else {
    fmt_sigfig_all(val_for(col, nm), 3L)
  }
}
present_letters <- names(fit_cols)[vapply(fit_cols, function(z) !is.null(z), logical(1))]
if (length(present_letters) > 0L) {
  all_mat <- sapply(present_letters, function(let) {
    vapply(param_rows_all, function(nm) fmt_cell(fit_cols[[let]], nm), character(1))
  }, simplify = "matrix")
  if (is.null(dim(all_mat))) {
    all_mat <- matrix(all_mat, ncol = 1, dimnames = list(param_rows_all, present_letters))
  }
  all_tbl <- data.frame(Parameter = param_rows_all, all_mat, check.names = FALSE)
  kable(
    all_tbl,
    caption = paste0(
      "Combined ETAS parameter comparison across available fits A–J ",
      "(gamma fixed at 0). Bivariate columns share structural (c,p,D,q) across margins; ",
      "univariate columns leave cross-excitation (A_01/A_10) blank and report margin-specific ",
      "structural terms. rho is the offspring spectral radius (bivariate) or max margin branching ",
      "(univariate); explosive means rho >= 1 (beta=", signif(beta_gr_tbl, 3), ")."
    ),
    align = paste0("l", paste(rep("c", length(present_letters)), collapse = ""))
  ) |>
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = FALSE) |>
    scroll_box(width = "100%", height = "640px")
} else {
  cat("Combined A–J parameter table unavailable in this results checkpoint.\n")
}
```

```{r fig-ate-biv}
#| label: fig-ate-biv
#| fig-cap: "Simulation distribution of the all-or-nothing total saved over the post-treatment horizon, facetted by model family with Naive vs SEM overlays."
#| fig-height: 4
biv_sim <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  data.frame(Model = r$lab, TotalATE = get_total_effect_sim(r$ate, n_tiles = n_cty))
}))
if (!is.null(biv_sim) && nrow(biv_sim) > 0) {
  assign_role <- function(model_label) {
    if (grepl("^A:|^C:|^E:", model_label)) return("Naive")
    if (grepl("^B:|^D:|^F:", model_label)) return("SEM")
    NA_character_
  }
  assign_pair <- function(model_label) {
    if (grepl("^A:|^B:", model_label)) return("Homogeneous (A/B)")
    if (grepl("^C:|^D:", model_label)) return("KDE all-free (C/D)")
    if (grepl("^E:|^F:", model_label)) return("KDE struct-fixed (E/F)")
    NA_character_
  }
  biv_sim$Role <- vapply(as.character(biv_sim$Model), assign_role, character(1))
  biv_sim$Pair <- vapply(as.character(biv_sim$Model), assign_pair, character(1))
  biv_sim <- biv_sim[!is.na(biv_sim$Role) & !is.na(biv_sim$Pair), , drop = FALSE]
  pair_ok <- vapply(split(biv_sim$Role, biv_sim$Pair), function(x) {
    all(c("Naive", "SEM") %in% unique(x))
  }, logical(1))
  biv_sim <- biv_sim[biv_sim$Pair %in% names(pair_ok)[pair_ok], , drop = FALSE]
  if (nrow(biv_sim) < 1L) {
    cat("Bivariate A–F ATE simulations are not in this results file.\n")
  } else {
    biv_sim$Role <- factor(biv_sim$Role, levels = c("Naive", "SEM"))
    biv_sim$Pair <- factor(
      biv_sim$Pair,
      levels = c("Homogeneous (A/B)", "KDE all-free (C/D)", "KDE struct-fixed (E/F)")
    )
    biv_means <- aggregate(TotalATE ~ Pair + Role, biv_sim, mean)
    ggplot(biv_sim, aes(x = TotalATE, fill = Role)) +
      geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
      geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
      geom_vline(data = biv_means, aes(xintercept = TotalATE, colour = Role),
                 linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
      facet_wrap(~Pair, scales = "free") +
      scale_fill_manual(values = c(Naive = "#fdb462", SEM = "#80b1d3")) +
      scale_colour_manual(values = c(Naive = "#e6550d", SEM = "#3182bd")) +
      labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes saved /" ~ .(ate_days) ~ "days)"),
           y = "Simulation count",
           title = "All-or-Nothing Total Saved: Naive vs SEM by Model Family") +
      theme_minimal(base_size = 13) +
      theme(legend.position = "top",
            plot.title = element_text(hjust = 0.5, face = "bold"))
  }
} else {
  cat("Bivariate A–F ATE simulations are not in this results file.\n")
}
```

```{r ate-scenario-compare-note, results='asis'}
#| echo: false
if (using_pre_bootstrap || is.null(boot_obj)) {
  cat("\n### Four-scenario ATE comparison\n\n")
  cat("Bootstrap scenario panels are omitted for this pre-bootstrap / no-bootstrap render.\n\n")
} else {
  cat("\n### Four-scenario ATE comparison\n\n")
  cat("The **2×2 panel below** overlays Naive (C) and SEM (D) bias-corrected bootstrap ")
  cat("histograms for univariate vs bivariate × all-or-nothing vs observed. ")
  cat("When present, **512-rep** scenario RDS (`*_boot512`) are preferred; otherwise the ")
  cat("largest available bootstrap *n* per scenario is used. ")
  cat("Detailed companion HTML: ")
  scen_html <- file.path(dirname(result_file), "oklahoma_ate_scenarios.html")
  scen_csv <- file.path(dirname(result_file), "ate_scenarios", "scenario_summary.csv")
  if (file.exists(scen_html)) {
    cat(sprintf("[`oklahoma_ate_scenarios.html`](%s)", basename(scen_html)))
  } else {
    cat("`oklahoma_ate_scenarios.html` (render with `render_ate_scenario_compare.R`)")
  }
  cat(".\n\n")
  if (file.exists(scen_csv)) {
    scen_sum <- utils::read.csv(scen_csv, stringsAsFactors = FALSE)
    if (nrow(scen_sum) > 0) {
      knitr::kable(
        scen_sum[, intersect(c("label", "method", "point_E", "point_F",
                               "boot_bc_mean_E", "boot_bc_sd_E",
                               "boot_bc_mean_F", "boot_bc_sd_F"), names(scen_sum)),
                 drop = FALSE],
        digits = 1,
        caption = "ATE scenario summary (bias-corrected bootstrap)."
      ) |>
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
    }
  }
}
```

```{r fig-ate-scenarios-4panel}
#| label: fig-ate-scenarios-4panel
#| fig-cap: "Four-scenario bias-corrected bootstrap comparison (univariate vs bivariate × all-or-nothing vs observed). Each panel overlays Naive (C) and SEM (D), using all finite ATEs retained after failed or explosive refits are excluded. Prefers 512-attempt scenario RDS when available."
#| fig-height: 7.2
#| eval: !expr '!using_pre_bootstrap && !is.null(boot_obj)'
scen_specs <- list(
  list(id = "univ_aon", law = "Univariate", contrast = "All-or-nothing",
       files = c("for_paper_univ_aon_boot512.rds", "for_paper_univ_aon.rds",
                 file.path("ate_scenarios", "univ_aon_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "univ_aon", "for_paper.rds"))),
  list(id = "univ_obs", law = "Univariate", contrast = "Observed vs control",
       files = c("for_paper_univ_obs_boot512.rds", "for_paper_univ_obs.rds",
                 file.path("ate_scenarios", "univ_obs_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "univ_obs", "for_paper.rds"))),
  list(id = "biv_aon", law = "Bivariate", contrast = "All-or-nothing",
       files = c("for_paper_biv_aon_boot512.rds", "for_paper.rds",
                 "for_paper_biv_aon_boot256.rds", "for_paper_biv_aon.rds",
                 file.path("ate_scenarios", "biv_aon_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "biv_aon", "for_paper.rds"))),
  list(id = "biv_obs", law = "Bivariate", contrast = "Observed vs control",
       files = c("for_paper_biv_obs_boot512.rds", "for_paper_biv_obs.rds",
                 "for_paper_biv_obs_boot256.rds",
                 file.path("ate_scenarios", "biv_obs_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "biv_obs", "for_paper.rds"),
                 file.path("ate_scenarios", "biv_obs_boot256", "for_paper.rds")))
)
scen_primary_boot <- function(boot) {
  if (is.null(boot)) return(list(naive = NULL, sem = NULL))
  list(
    naive = if (!is.null(boot$fit_C)) boot$fit_C else boot$fit_E,
    sem = if (!is.null(boot$fit_D)) boot$fit_D else boot$fit_F
  )
}
scen_primary_fits <- function(fn) {
  if (is.null(fn)) return(list(naive = NULL, sem = NULL))
  list(
    naive = if (!is.null(fn$C)) fn$C else fn$E,
    sem = if (!is.null(fn$D)) fn$D else fn$F
  )
}
resolve_scen_file <- function(rel_files) {
  cands <- unique(unlist(lapply(c(OUT_DIR, candidate_out_dirs), function(d) {
    file.path(d, rel_files)
  }), use.names = FALSE))
  # Prefer the largest bootstrap n among existing candidates.
  existing <- cands[file.exists(cands)]
  if (!length(existing)) return(NA_character_)
  n_boot <- vapply(existing, function(f) {
    r <- tryCatch(normalize_oklahoma_fit_letters(readRDS(f)), error = function(e) NULL)
    blk <- scen_primary_boot(r$bootstrap_ate)$sem
    if (is.null(blk) || is.null(blk$replicate_summary)) return(0L)
    as.integer(nrow(blk$replicate_summary))
  }, integer(1))
  existing[[which.max(n_boot)]]
}
recenter_boot_scen <- function(boot_vals, target_mean) {
  bv <- suppressWarnings(as.numeric(boot_vals))
  if (!is.finite(target_mean) || is.na(target_mean)) return(bv)
  bmean <- mean(bv, na.rm = TRUE)
  if (!is.finite(bmean) || is.na(bmean)) return(bv)
  bv + (target_mean - bmean)
}
point_mean_scen <- function(fit_obj) {
  if (is.null(fit_obj) || is.null(fit_obj$ate)) return(NA_real_)
  vals <- get_total_effect_sim(fit_obj$ate, n_tiles = n_cty)
  vals <- suppressWarnings(as.numeric(vals))
  vals <- vals[is.finite(vals)]
  if (!length(vals)) return(NA_real_)
  mean(vals, na.rm = TRUE)
}
scen_boot_rows <- list()
scen_n_note <- character(0)
for (sp in scen_specs) {
  fp <- resolve_scen_file(sp$files)
  if (is.na(fp)) next
  rr <- tryCatch(normalize_oklahoma_fit_letters(readRDS(fp)), error = function(e) NULL)
  boot_pair <- scen_primary_boot(rr$bootstrap_ate)
  if (is.null(rr) || is.null(boot_pair$naive) || is.null(boot_pair$sem)) next
  be_df <- as.data.frame(boot_pair$naive$replicate_summary)
  bf_df <- as.data.frame(boot_pair$sem$replicate_summary)
  if ("is_stable" %in% names(be_df)) {
    be_df <- be_df[!is.na(be_df$is_stable) & be_df$is_stable, , drop = FALSE]
  }
  if ("is_stable" %in% names(bf_df)) {
    bf_df <- bf_df[!is.na(bf_df$is_stable) & bf_df$is_stable, , drop = FALSE]
  }
  be_raw <- be_df$ate_total_mean
  bf_raw <- bf_df$ate_total_mean
  fit_pair <- scen_primary_fits(rr$fits_named)
  pe <- point_mean_scen(fit_pair$naive)
  pf <- point_mean_scen(fit_pair$sem)
  be <- retain_boot_vals(be_raw)
  bf <- retain_boot_vals(bf_raw)
  attempted_C <- boot_pair$naive$n_attempted %||%
    rr$bootstrap_ate$config$reps %||% be$n_in
  attempted_D <- boot_pair$sem$n_attempted %||%
    rr$bootstrap_ate$config$reps %||% bf$n_in
  scen_n_note <- c(
    scen_n_note,
    sprintf(
      "%s retained C=%d/%d, D=%d/%d",
      sp$id, length(be$vals), attempted_C, length(bf$vals), attempted_D
    )
  )
  scen_boot_rows[[length(scen_boot_rows) + 1L]] <- rbind(
    data.frame(
      Law = sp$law, Contrast = sp$contrast, Model = "Naive (C)",
      TotalATE = recenter_boot_scen(be$vals, pe), stringsAsFactors = FALSE
    ),
    data.frame(
      Law = sp$law, Contrast = sp$contrast, Model = "SEM (D)",
      TotalATE = recenter_boot_scen(bf$vals, pf), stringsAsFactors = FALSE
    )
  )
}
if (length(scen_boot_rows) > 0) {
  scen_boot_df <- do.call(rbind, scen_boot_rows)
  scen_boot_df <- scen_boot_df[is.finite(scen_boot_df$TotalATE), , drop = FALSE]
  scen_boot_df$Law <- factor(scen_boot_df$Law, levels = c("Univariate", "Bivariate"))
  scen_boot_df$Contrast <- factor(scen_boot_df$Contrast, levels = c("All-or-nothing", "Observed vs control"))
  scen_boot_df$Model <- factor(scen_boot_df$Model, levels = c("Naive (C)", "SEM (D)"))
  scen_means <- aggregate(TotalATE ~ Law + Contrast + Model, scen_boot_df, mean)
  ggplot(scen_boot_df, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_vline(data = scen_means, aes(xintercept = TotalATE, colour = Model),
               linewidth = 0.9, show.legend = FALSE) +
    facet_grid(Law ~ Contrast, scales = "free_y") +
    scale_fill_manual(values = c("Naive (C)" = "#fdb462", "SEM (D)" = "#80b1d3")) +
    scale_colour_manual(values = c("Naive (C)" = "#e6550d", "SEM (D)" = "#3182bd")) +
    coord_cartesian(xlim = c(-1000, 1000)) +
    labs(
      x = bquote("Bias-corrected bootstrap " ~ hat(Delta) ~ "(total saved /" ~ .(ate_days) ~ "days)"),
      y = "Bootstrap replicate count",
      fill = "Model",
      title = "Four-scenario ATE bootstrap comparison",
      subtitle = paste(scen_n_note, collapse = " · ")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 9),
      strip.text = element_text(face = "bold")
    )
} else {
  plot.new()
  text(0.5, 0.5, "Four-scenario RDS files not found under oklahoma output.")
}
```

```{r fig-ate-bootstrap-cd}
#| label: fig-ate-bootstrap-cd
#| fig-cap: "Bias-corrected bootstrap replicate distribution of total saved for the primary KDE pair C/D, using all finite ATEs retained after failed or explosive refits are excluded. Shape/spread preserved, recentered to the point ATE."
#| fig-height: 4
#| eval: !expr '!using_pre_bootstrap && !is.null(boot_naive_df) && !is.null(boot_sem_df)'
# Labels / centers follow results$config$ATE_METHOD for this RDS.
if (!is.null(boot_naive_df) && !is.null(boot_sem_df) &&
    nrow(boot_naive_df) > 0 && nrow(boot_sem_df) > 0 &&
    "ate_total_mean" %in% names(boot_naive_df) && "ate_total_mean" %in% names(boot_sem_df)) {
  get_fit_obj_by_id <- function(id) {
    switch(
      toupper(id),
      "C" = fit_C_obj,
      "D" = fit_D_obj,
      "E" = fit_E_obj,
      "F" = fit_F_obj,
      NULL
    )
  }
  obs_ate_mean <- function(id) {
    fo <- get_fit_obj_by_id(id)
    if (is.null(fo) || is.null(fo$ate)) return(NA_real_)
    vals <- get_total_effect_sim(fo$ate, n_tiles = n_cty)
    vals <- suppressWarnings(as.numeric(vals))
    vals <- vals[is.finite(vals)]
    if (length(vals) < 1L) return(NA_real_)
    mean(vals, na.rm = TRUE)
  }
  recenter_boot <- function(boot_vals, target_mean) {
    bv <- suppressWarnings(as.numeric(boot_vals))
    if (!is.finite(target_mean) || is.na(target_mean)) return(bv)
    bmean <- mean(bv, na.rm = TRUE)
    if (!is.finite(bmean) || is.na(bmean)) return(bv)
    bv + (target_mean - bmean)
  }

  naive_obs_mean <- obs_ate_mean(boot_naive_id)
  sem_obs_mean <- obs_ate_mean(boot_sem_id)
  naive_filt <- retain_boot_vals(boot_naive_df$ate_total_mean)
  sem_filt <- retain_boot_vals(boot_sem_df$ate_total_mean)
  naive_bc_vals <- recenter_boot(naive_filt$vals, naive_obs_mean)
  sem_bc_vals <- recenter_boot(sem_filt$vals, sem_obs_mean)

  boot_cd <- rbind(
    data.frame(Model = boot_naive_label, TotalATE = naive_bc_vals),
    data.frame(Model = boot_sem_label, TotalATE = sem_bc_vals)
  )
  boot_cd <- boot_cd[is.finite(boot_cd$TotalATE), , drop = FALSE]
  if (nrow(boot_cd) > 0) {
    boot_means <- aggregate(TotalATE ~ Model, boot_cd, mean)
    n_drop <- naive_filt$n_out + sem_filt$n_out
    ggplot(boot_cd, aes(x = TotalATE, fill = Model)) +
      geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
      geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
      geom_vline(data = boot_means, aes(xintercept = TotalATE, colour = Model),
                 linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
      scale_fill_manual(values = setNames(c("#fdb462", "#80b1d3"), c(boot_naive_label, boot_sem_label))) +
      scale_colour_manual(values = setNames(c("#e6550d", "#3182bd"), c(boot_naive_label, boot_sem_label))) +
      coord_cartesian(xlim = c(-1000, 1000)) +
      labs(x = bquote(hat(Delta)[AoN] ~ "(bias-corrected bootstrap replicate mean total saved /" ~ .(ate_days) ~ "days)"),
           y = "Bootstrap replicate count",
           title = paste0("Bias-corrected bootstrap AoN total-saved distribution (", boot_pair_label, ")"),
           subtitle = sprintf(
             "Centers: %s=%.1f, %s=%.1f; retained stable finite ATEs %d/%d (%s) and %d/%d (%s); %d non-finite ATE(s) omitted",
             boot_naive_id, naive_obs_mean, boot_sem_id, sem_obs_mean,
             length(naive_filt$vals), naive_filt$n_in, boot_naive_id,
             length(sem_filt$vals), sem_filt$n_in, boot_sem_id,
             n_drop
           )) +
      theme_minimal(base_size = 13) +
      theme(legend.position = "top",
            plot.title = element_text(hjust = 0.5, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, size = 10))
  } else {
    plot.new()
    text(0.5, 0.5, sprintf("Bias-corrected bootstrap %s replicate means are present but non-finite.", boot_pair_label))
  }
} else {
  plot.new()
  text(0.5, 0.5, sprintf("Bias-corrected bootstrap %s results are unavailable in this results checkpoint.", boot_pair_label))
}
```

```{r fig-ate-bootstrap-biv-obs}
#| label: fig-ate-bootstrap-biv-obs
#| fig-cap: "Bias-corrected bootstrap replicate distribution for the bivariate observed-vs-control-everywhere contrast (C/D), using all finite ATEs retained after failed or explosive refits are excluded. Same display conventions as the all-or-nothing bootstrap figure."
#| fig-height: 4
#| eval: !expr '!using_pre_bootstrap && !is.null(boot_obj)'
# Companion scenario RDS written by launch_ate_scenarios_nesi.sh / bootstrap-only jobs.
biv_obs_candidates <- unique(c(
  file.path(OUT_DIR, "for_paper_biv_obs.rds"),
  file.path(OUT_DIR, "ate_scenarios", "biv_obs", "for_paper.rds"),
  unlist(lapply(candidate_out_dirs, function(d) {
    c(
      file.path(d, "for_paper_biv_obs.rds"),
      file.path(d, "ate_scenarios", "biv_obs", "for_paper.rds")
    )
  }), use.names = FALSE)
))
biv_obs_file <- biv_obs_candidates[file.exists(biv_obs_candidates)][1]
if (!using_pre_bootstrap && !is.na(biv_obs_file) && nzchar(biv_obs_file)) {
  res_obs <- normalize_oklahoma_fit_letters(readRDS(biv_obs_file))
  boot_obs <- res_obs$bootstrap_ate
  fit_C_obs <- if (!is.null(res_obs$fits_named)) res_obs$fits_named$C else NULL
  fit_D_obs <- if (!is.null(res_obs$fits_named)) res_obs$fits_named$D else NULL
  boot_C_blk <- if (!is.null(boot_obs$fit_C)) boot_obs$fit_C else boot_obs$fit_E
  boot_D_blk <- if (!is.null(boot_obs$fit_D)) boot_obs$fit_D else boot_obs$fit_F
  boot_C_obs <- if (!is.null(boot_C_blk)) boot_C_blk$replicate_summary else NULL
  boot_D_obs <- if (!is.null(boot_D_blk)) boot_D_blk$replicate_summary else NULL
  if (!is.null(boot_C_obs) && "is_stable" %in% names(boot_C_obs)) {
    boot_C_obs <- boot_C_obs[
      !is.na(boot_C_obs$is_stable) & boot_C_obs$is_stable, , drop = FALSE
    ]
  }
  if (!is.null(boot_D_obs) && "is_stable" %in% names(boot_D_obs)) {
    boot_D_obs <- boot_D_obs[
      !is.na(boot_D_obs$is_stable) & boot_D_obs$is_stable, , drop = FALSE
    ]
  }
  if (!is.null(boot_C_obs) && !is.null(boot_D_obs) &&
      nrow(boot_C_obs) > 0 && nrow(boot_D_obs) > 0 &&
      "ate_total_mean" %in% names(boot_C_obs) && "ate_total_mean" %in% names(boot_D_obs)) {
    obs_ate_mean_from_fit <- function(fit_obj) {
      if (is.null(fit_obj) || is.null(fit_obj$ate)) return(NA_real_)
      vals <- get_total_effect_sim(fit_obj$ate, n_tiles = n_cty)
      vals <- suppressWarnings(as.numeric(vals))
      vals <- vals[is.finite(vals)]
      if (length(vals) < 1L) return(NA_real_)
      mean(vals, na.rm = TRUE)
    }
    recenter_boot_obs <- function(boot_vals, target_mean) {
      bv <- suppressWarnings(as.numeric(boot_vals))
      if (!is.finite(target_mean) || is.na(target_mean)) return(bv)
      bmean <- mean(bv, na.rm = TRUE)
      if (!is.finite(bmean) || is.na(bmean)) return(bv)
      bv + (target_mean - bmean)
    }
    lab_C <- "C: Naive Biv+KDE (all-free)"
    lab_D <- "D: SEM Biv+KDE (all-free)"
    mean_C_obs <- obs_ate_mean_from_fit(fit_C_obs)
    mean_D_obs <- obs_ate_mean_from_fit(fit_D_obs)
    filt_C_obs <- retain_boot_vals(boot_C_obs$ate_total_mean)
    filt_D_obs <- retain_boot_vals(boot_D_obs$ate_total_mean)
    boot_cd_obs <- rbind(
      data.frame(Model = lab_C, TotalATE = recenter_boot_obs(filt_C_obs$vals, mean_C_obs)),
      data.frame(Model = lab_D, TotalATE = recenter_boot_obs(filt_D_obs$vals, mean_D_obs))
    )
    boot_cd_obs <- boot_cd_obs[is.finite(boot_cd_obs$TotalATE), , drop = FALSE]
    if (nrow(boot_cd_obs) > 0) {
      boot_means_obs <- aggregate(TotalATE ~ Model, boot_cd_obs, mean)
      n_drop_obs <- filt_C_obs$n_out + filt_D_obs$n_out
      ate_days_obs <- suppressWarnings(as.numeric(res_obs$config$ATE_WINDOW_DAYS))
      if (!is.finite(ate_days_obs)) ate_days_obs <- ate_days
      ggplot(boot_cd_obs, aes(x = TotalATE, fill = Model)) +
        geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
        geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
        geom_vline(data = boot_means_obs, aes(xintercept = TotalATE, colour = Model),
                   linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
        scale_fill_manual(values = setNames(c("#fdb462", "#80b1d3"), c(lab_C, lab_D))) +
        scale_colour_manual(values = setNames(c("#e6550d", "#3182bd"), c(lab_C, lab_D))) +
        coord_cartesian(xlim = c(-1000, 1000)) +
        labs(
          x = bquote(hat(Delta)[obs] ~ "(bias-corrected bootstrap replicate mean total saved /" ~ .(ate_days_obs) ~ "days)"),
          y = "Bootstrap replicate count",
          title = "Bias-corrected bootstrap observed-vs-control distribution (C/D, bivariate)",
          subtitle = sprintf(
            "Contrast: bivariate/observed; centers C=%.1f, D=%.1f; retained stable finite ATEs %d/%d (C) and %d/%d (D); omitted %d non-finite; source %s",
            mean_C_obs, mean_D_obs,
            length(filt_C_obs$vals), filt_C_obs$n_in,
            length(filt_D_obs$vals), filt_D_obs$n_in,
            n_drop_obs,
            basename(biv_obs_file)
          )
        ) +
        theme_minimal(base_size = 13) +
        theme(
          legend.position = "top",
          plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, size = 10)
        )
    } else {
      plot.new()
      text(0.5, 0.5, "Bivariate observed bootstrap replicate means are present but non-finite.")
    }
  } else {
    plot.new()
    text(0.5, 0.5, sprintf("Bivariate observed bootstrap C/D summaries missing in %s.", basename(biv_obs_file)))
  }
} else {
  plot.new()
  text(0.5, 0.5, "Bivariate observed scenario RDS not found (expected for_paper_biv_obs.rds).")
}
```

# Partition Sensitivity

The analysis is repeated under alternative spatial partitions to assess
sensitivity to spatial resolution:

- **Grid partitions** with cell diameters equal to 1, 2, and 5
  times the estimated triggering range
  (`r fmt_scaled_num(if (!is.null(results$kde_info)) results$kde_info$trigger_range_km else NULL)` km).
- **AOI region** partition: the AOI polygon defines one treated region,
  everything else is control (2 tiles total).

For each partition, both the naive bivariate + KDE (C-type, all-free)
and SEM bivariate + KDE (D-type, all-free) models are fitted. The maps
below use the same pre/post overlay as the tessellation point-pattern figure.

The log-likelihood table below scores every tessellation's fitted
parameters on the **same AOI geometry** (OCC polygon = treated, complement =
control, same KDE field). **Fit C LL** uses that tessellation's naive C
parameters with AOI-boundary labels (C's native complete-data labelling).
The D columns keep that tessellation's SEM-D parameters and change only
the process labels: the SEM's last proposed labelling versus the
AOI-boundary location labels. A higher last-labelling likelihood means
the sampler's proposed labels outscore the geographic AOI assignment
under that tessellation's fitted D model. Pre-treatment events remain
control in every column.

A separate parameter table lists the fitted bivariate ETAS coefficients
for C and D on each partition (background, productivity, cross-excitation,
shared structural terms, and branching). Use it to see whether a tessellation
stays C-like (\(\mu_1\gg\mu_0\), moderate \(\rho\)) or walks toward a
low-background, near-critical treated Hawkes.

```{r prep-partition-overlays}
#| label: prep-partition-overlays
#| include: false
geom_path <- file.path(SCRIPT_DIR, "oklahoma_geometry.R")
if (file.exists(geom_path) && !exists("oklahoma_sf_to_owin_km", mode = "function")) {
  source(geom_path, local = FALSE)
}
support_path <- file.path(SCRIPT_DIR, "oklahoma_report_support.R")
if (file.exists(support_path) && !exists("oklahoma_report_plot_partition_pp", mode = "function")) {
  source(support_path, local = FALSE)
}

partition_overlay_specs <- list()
partition_overlay_caps <- character(0)
pr_overlay <- results$partition_results
if (!is.null(pr_overlay) && length(pr_overlay) >= 1L &&
    exists("oklahoma_sf_to_owin_km", mode = "function") &&
    exists("oklahoma_report_plot_partition_pp", mode = "function") &&
    exists("counties_km") && exists("aoi_sf") && exists("counties_sf")) {
  xs <- c(pp_pre$x, pp_post$x)
  ys <- c(pp_pre$y, pp_post$y)
  xs <- xs[is.finite(xs)]
  ys <- ys[is.finite(ys)]
  if (length(xs) > 0L && length(ys) > 0L) {
    xlims_part <- range(xs, na.rm = TRUE)
    ylims_part <- range(ys, na.rm = TRUE)
  } else {
    bb_map <- sf::st_bbox(counties_km)
    xlims_part <- as.numeric(bb_map[c("xmin", "xmax")])
    ylims_part <- as.numeric(bb_map[c("ymin", "ymax")])
  }
  theme_part <- ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(hjust = 0.5, face = "bold"),
      legend.position = "none"
    )
  fill_vals <- c(control = "#deebf7", treated = "#fc9272")

  ok_boundary <- sf::st_union(counties_sf)
  ok_boundary <- sf::st_make_valid(ok_boundary)
  win_km_part <- oklahoma_sf_to_owin_km(ok_boundary)
  aoi_union_part <- sf::st_union(sf::st_make_valid(aoi_sf))
  aoi_owin_part <- tryCatch(
    oklahoma_sf_to_owin_km(aoi_union_part),
    error = function(e) NULL
  )
  aoi_km <- aoi_sf
  sf::st_geometry(aoi_km) <- sf::st_geometry(aoi_sf) / 1000
  sf::st_crs(aoi_km) <- NA

  trigger_km <- suppressWarnings(as.numeric(results$kde_info$trigger_range_km))
  if (!is.finite(trigger_km) || trigger_km <= 0) {
    trigger_km <- suppressWarnings(as.numeric(cfg$spatial_scale_D))
  }
  grid_mults <- unique(c(
    1.0, 2.0, 5.0,
    suppressWarnings(as.numeric(unlist(cfg$grid_multipliers)))
  ))
  grid_mults <- grid_mults[is.finite(grid_mults)]
  test_mode <- isTRUE(cfg$TEST_MODE)
  grid_max_lookup <- if (isTRUE(test_mode)) {
    c(`1` = 2000L, `2` = 1600L, `5` = 1200L)
  } else {
    c(`1` = 5000L, `2` = 3000L, `5` = 1500L)
  }
  grid_max_tiles <- unname(grid_max_lookup[as.character(as.integer(grid_mults))])
  grid_max_tiles[is.na(grid_max_tiles)] <- if (isTRUE(test_mode)) 1600L else 3000L

  wanted <- unique(c(primary_partition_id, "county", names(pr_overlay)))
  wanted <- wanted[wanted %in% c("county", names(pr_overlay), "aoi_region", primary_partition_id) |
                     grepl("^grid_", wanted)]
  order_ids <- unique(c(
    primary_partition_id,
    "county",
    sort(wanted[grepl("^grid_", wanted)]),
    "aoi_region"
  ))
  order_ids <- intersect(order_ids, wanted)

  for (pid in order_ids) {
    lab <- format_partition_label(pid)
    spec <- NULL
    if (identical(pid, "county")) {
      spec <- list(
        kind = "sf",
        tiles = counties_km,
        post = {
          out <- pp_post
          process_col <- dplyr::case_when(
            "location_process" %in% names(out) ~ "location_process",
            "inferred_process" %in% names(out) ~ "inferred_process",
            "process" %in% names(out) ~ "process",
            TRUE ~ NA_character_
          )
          if (is.na(process_col)) {
            out$Process <- factor(rep("control", nrow(out)), levels = c("control", "treated"))
          } else {
            out$Process <- factor(as.character(out[[process_col]]), levels = c("control", "treated"))
          }
          out
        }
      )
    } else if (grepl("^grid_", pid) && !is.null(aoi_owin_part)) {
      if (identical(pid, "grid_coarse")) {
        gp <- oklahoma_report_grid_partition(
          diameter = NA_real_,
          win = win_km_part,
          aoi_owin = aoi_owin_part,
          label = pid,
          max_tiles = 2000L
        )
      } else {
        if (!is.finite(trigger_km)) next
        mult <- suppressWarnings(as.numeric(sub("^grid_([0-9.]+)R$", "\\1", pid)))
        if (!is.finite(mult)) next
        mi <- match(mult, grid_mults)
        if (is.na(mi)) mi <- which.min(abs(grid_mults - mult))
        gp <- oklahoma_report_grid_partition(
          diameter = mult * trigger_km,
          win = win_km_part,
          aoi_owin = aoi_owin_part,
          label = pid,
          max_tiles = grid_max_tiles[[max(1L, as.integer(mi))]]
        )
      }
      post <- pp_post
      post$Process <- factor(
        oklahoma_assign_partition_process(post$x, post$y, gp$partition, gp$processes),
        levels = c("control", "treated")
      )
      spec <- list(
        kind = "rect",
        tiles = oklahoma_report_tess_rects(gp$partition, gp$processes),
        post = post
      )
    } else if (identical(pid, "aoi_region") && !is.null(aoi_owin_part)) {
      aoi_treat <- tryCatch(
        sf::st_make_valid(sf::st_intersection(sf::st_union(aoi_km), sf::st_union(counties_km))),
        error = function(e) sf::st_union(aoi_km)
      )
      aoi_ctrl <- tryCatch(
        sf::st_make_valid(sf::st_difference(sf::st_union(counties_km), aoi_treat)),
        error = function(e) NULL
      )
      tiles <- rbind(
        sf::st_sf(treatment = "treated", geometry = sf::st_geometry(aoi_treat))
      )
      if (!is.null(aoi_ctrl) && !sf::st_is_empty(aoi_ctrl)) {
        tiles <- rbind(
          tiles,
          sf::st_sf(treatment = "control", geometry = sf::st_geometry(aoi_ctrl))
        )
      }
      post <- pp_post
      post$Process <- factor(
        ifelse(spatstat.geom::inside.owin(post$x, post$y, aoi_owin_part), "treated", "control"),
        levels = c("control", "treated")
      )
      spec <- list(kind = "sf", tiles = tiles, post = post)
    }
    if (is.null(spec)) next
    spec$id <- pid
    spec$label <- lab
    spec$xlims <- xlims_part
    spec$ylims <- ylims_part
    spec$theme <- theme_part
    spec$fill_vals <- fill_vals
    partition_overlay_specs[[length(partition_overlay_specs) + 1L]] <- spec
    partition_overlay_caps <- c(
      partition_overlay_caps,
      sprintf(
        "Pre-treatment (left) and post-treatment (right) events overlaid on the %s partition.",
        lab
      )
    )
  }
}
```

```{r fig-partition-pp-all}
#| label: fig-partition-pp-all
#| fig-height: 5
#| fig-cap: !expr if (length(partition_overlay_caps) > 0L) partition_overlay_caps else "Partition event overlays."
if (length(partition_overlay_specs) < 1L) {
  cat("Partition geometries could not be rebuilt for event overlays in this results file.\n")
} else {
  for (spec in partition_overlay_specs) {
    if (identical(spec$kind, "rect")) {
      base_map <- ggplot2::ggplot() +
        ggplot2::geom_rect(
          data = spec$tiles,
          ggplot2::aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = treatment),
          colour = "grey70", linewidth = 0.05
        ) +
        ggplot2::scale_fill_manual(values = spec$fill_vals, guide = "none") +
        ggplot2::coord_fixed(
          xlim = spec$xlims, ylim = spec$ylims, expand = FALSE, ratio = 1
        ) +
        spec$theme
    } else {
      base_map <- ggplot2::ggplot() +
        ggplot2::geom_sf(
          data = spec$tiles, ggplot2::aes(fill = treatment),
          colour = "grey60", linewidth = 0.25
        ) +
        ggplot2::scale_fill_manual(values = spec$fill_vals, guide = "none") +
        ggplot2::coord_sf(
          xlim = spec$xlims, ylim = spec$ylims, datum = NA, expand = FALSE
        ) +
        spec$theme
    }
    oklahoma_report_plot_partition_pp(base_map, pp_pre, spec$post)
  }
}
```

```{r tbl-partition-sensitivity}
#| label: tbl-partition-sensitivity
pr <- results$partition_results
ap <- results$ate_partitions
if (is.null(pr)) pr <- list()
primary_info <- list(
  label = primary_partition_id,
  E_params = if (!is.null(fit_C_obj)) fit_C_obj$params else NULL,
  F_params = if (!is.null(fit_D_obj)) fit_D_obj$params else NULL,
  semF = sem_fit_obj(fit_D_obj),
  n_tiles = n_cty,
  n_treated = suppressWarnings(as.integer(unlist(cfg$n_treated, use.names = FALSE)))[1],
  n_relabel = NA_integer_
)
if (!primary_partition_id %in% names(pr)) {
  pr <- c(setNames(list(primary_info), primary_partition_id), pr)
}

part_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (identical(info$label, primary_partition_id)) {
    list(ate_E = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL,
         ate_F = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL)
  } else {
    ap[[pname]]
  }

  for (fit_type in list(
    list(lab = "C: Naive Biv+KDE (all-free)", par = info$E_params, ate = ate_info$ate_E, sem_obj = NULL),
    list(lab = "D: SEM Biv+KDE (all-free)",   par = info$F_params, ate = ate_info$ate_F, sem_obj = info$semF)
  )) {
    expected <- if (!is.null(fit_type$ate)) {
      get_expected_totals(fit_type$ate, n_tiles = info$n_tiles)
    } else { NULL }
    sim_totals <- if (!is.null(fit_type$ate)) {
      get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles)
    } else { NULL }
    c_mean <- if (!is.null(expected) && !is.null(expected$c_total)) {
      mean(expected$c_total, na.rm = TRUE)
    } else { NA_real_ }
    t_mean <- if (!is.null(expected) && !is.null(expected$t_total)) {
      mean(expected$t_total, na.rm = TRUE)
    } else { NA_real_ }
    t_all_mean <- if (!is.null(fit_type$ate$analytic) && !is.null(fit_type$ate$analytic$eta_treat)) {
      as.numeric(fit_type$ate$analytic$eta_treat) * info$n_tiles
    } else { NA_real_ }
    saved_diff_mean <- if (is.finite(c_mean) && is.finite(t_mean)) c_mean - t_mean else NA_real_
    sim_sd_total <- if (!is.null(sim_totals) && length(sim_totals) > 1L) {
      sd(sim_totals, na.rm = TRUE)
    } else { NA_real_ }
    spill_total <- if (!is.null(fit_type$ate$total_spillover)) {
      fit_type$ate$total_spillover
    } else if (!is.null(fit_type$ate$ATE_spillover)) {
      fit_type$ate$ATE_spillover * info$n_tiles
    } else { NA_real_ }
    n_post <- NA_integer_
    if (!is.null(info$n_post_ctrl) && !is.null(info$n_post_treat) &&
        is.finite(info$n_post_ctrl) && is.finite(info$n_post_treat)) {
      n_post <- as.integer(info$n_post_ctrl) + as.integer(info$n_post_treat)
    } else if (identical(info$label, primary_partition_id) && exists("pp_post") && !is.null(pp_post)) {
      n_post <- as.integer(nrow(pp_post))
    }
    n_relabel <- if (!is.null(fit_type$sem_obj) &&
                     !is.null(fit_type$sem_obj$adaptive$adaptive_labelling)) {
      lp <- fit_type$sem_obj$adaptive$adaptive_labelling
      lp <- lp[lp$t >= 0, , drop = FALSE]
      sum(lp$location_process != lp$inferred_process, na.rm = TRUE)
    } else if (identical(fit_type$lab, "D: SEM Biv+KDE (all-free)") &&
               !is.null(info$n_relabel) && is.finite(info$n_relabel)) {
      as.integer(info$n_relabel)
    } else {
      NA_integer_
    }
    relabel_txt <- if (!grepl("^D:", fit_type$lab)) {
      "—"
    } else if (is.finite(n_relabel) && is.finite(n_post) && n_post > 0L) {
      sprintf("%d / %d", as.integer(n_relabel), as.integer(n_post))
    } else if (is.finite(n_relabel)) {
      as.character(as.integer(n_relabel))
    } else {
      "NA"
    }
    rnd <- function(x, digits = 0L) {
      v <- suppressWarnings(as.numeric(x)[1L])
      if (!is.finite(v)) NA_real_ else round(v, digits)
    }
    pget <- function(par, nm) {
      if (is.null(par)) return(NA_real_)
      v <- tryCatch(par[[nm]], error = function(e) NULL)
      suppressWarnings(as.numeric(v)[1L])
    }

    part_rows[[length(part_rows) + 1]] <- data.frame(
      PartitionID = info$label,
      Partition = format_partition_label(info$label),
      Model = fit_type$lab,
      Tiles = info$n_tiles,
      Treated = info$n_treated,
      `Relabelled / post` = relabel_txt,
      `E[N(control everywhere)]` = rnd(c_mean, 0),
      `E[N(actual treatment regime)]` = rnd(t_mean, 0),
      `E[N(treated everywhere)]` = rnd(t_all_mean, 0),
      `Saved = control - actual` = rnd(saved_diff_mean, 0),
      `SD(saved)` = rnd(sim_sd_total, 0),
      `Spillover` = rnd(spill_total, 0),
      mu_0 = rnd(pget(fit_type$par, "mu_0"), 4),
      mu_1 = rnd(pget(fit_type$par, "mu_1"), 4),
      A_0 = rnd(pget(fit_type$par, "A_00"), 4),
      A_1 = rnd(pget(fit_type$par, "A_11"), 4),
      alpha_m_0 = rnd(pget(fit_type$par, "alpha_m_00"), 4),
      alpha_m_1 = rnd(pget(fit_type$par, "alpha_m_11"), 4),
      check.names = FALSE)
  }
}

if (length(part_rows) > 0) {
  part_df <- do.call(rbind, part_rows)
  # Keep the primary tessellation first, then remaining partitions.
  part_ids <- unique(as.character(part_df$PartitionID))
  part_df$PartitionID <- factor(part_df$PartitionID,
                                levels = c(primary_partition_id,
                                           sort(setdiff(part_ids, primary_partition_id))))
  part_df <- part_df[order(part_df$PartitionID, part_df$Model), ]
  part_df$Partition <- factor(as.character(part_df$Partition),
                              levels = unique(as.character(part_df$Partition)))
  part_df_print <- subset(part_df, select = -PartitionID)
  rownames(part_df_print) <- NULL
  kable(part_df_print, align = "llccccccccccccccc", row.names = FALSE,
        caption = paste0("Partition comparison over ", ate_days, " days using the same expected-count and ATE metrics as Section 6.1 (plus marginal ETAS parameters). SEM relabels are post-treatment events whose inferred process differs from the partition location label; naive Fit C does not relabel.")) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                  font_size = 9, latex_options = c("scale_down", "hold_position"))
}
```

```{r tbl-partition-params}
#| label: tbl-partition-params
#| tbl-cap: "Bivariate ETAS parameters for C (location labels) and D (SEM) on each partition. Branching ratios are A*beta/(beta-alpha); rho is the offspring spectral radius. Shared structural terms (c, p, D, q) are one value per bivariate fit (gamma = 0)."
pnum <- function(par, nm) {
  if (is.null(par)) return(NA_real_)
  if (is.list(par) && nm %in% names(par)) {
    v <- suppressWarnings(as.numeric(par[[nm]]))
    return(if (length(v) && is.finite(v[[1]])) v[[1]] else NA_real_)
  }
  pv <- tryCatch(unlist(par), error = function(e) NULL)
  if (is.null(pv) || is.null(names(pv)) || !nm %in% names(pv)) return(NA_real_)
  v <- suppressWarnings(as.numeric(pv[[nm]]))
  if (!length(v) || !is.finite(v[[1]])) return(NA_real_)
  v[[1]]
}
beta_part <- suppressWarnings(as.numeric(unlist(cfg$BETA_GR, use.names = FALSE)))
beta_part <- beta_part[is.finite(beta_part) & beta_part > 0]
beta_part <- if (length(beta_part)) beta_part[[1]] else 2.3
eta_part <- function(A, alpha, beta_gr = beta_part) {
  if (!is.finite(A) || !is.finite(alpha) || !is.finite(beta_gr) || beta_gr <= 0) {
    return(NA_real_)
  }
  gap <- beta_gr - alpha
  if (!is.finite(gap) || gap <= 1e-8) return(Inf)
  A * beta_gr / gap
}
rho_part <- function(par) {
  e00 <- eta_part(pnum(par, "A_00"), pnum(par, "alpha_m_00"))
  e01 <- eta_part(pnum(par, "A_01"), pnum(par, "alpha_m_01"))
  e10 <- eta_part(pnum(par, "A_10"), pnum(par, "alpha_m_10"))
  e11 <- eta_part(pnum(par, "A_11"), pnum(par, "alpha_m_11"))
  M <- matrix(c(e00, e01, e10, e11), nrow = 2, byrow = TRUE)
  if (!all(is.finite(M))) return(Inf)
  max(Re(eigen(M, only.values = TRUE)$values))
}
fmt_part <- function(x, digits = 4L) {
  if (length(x) < 1L || is.null(x) || is.na(x)) return("—")
  if (!is.finite(x)) return("Inf")
  sprintf(paste0("%.", as.integer(digits), "f"), as.numeric(x))
}
fmt_tiny <- function(x) {
  if (length(x) < 1L || is.null(x) || is.na(x)) return("—")
  if (!is.finite(x)) return("Inf")
  if (abs(x) > 0 && abs(x) < 1e-3) return(format(signif(x, 3), scientific = TRUE))
  sprintf("%.4f", x)
}

part_param_rows <- list()
if (exists("pr") && length(pr) > 0L) {
  part_ids_p <- unique(as.character(unlist(lapply(pr, function(z) z$label))))
  part_ids_p <- c(primary_partition_id, sort(setdiff(part_ids_p, primary_partition_id)))
  part_ids_p <- intersect(part_ids_p, unique(c(names(pr), primary_partition_id)))
  for (pid in part_ids_p) {
    info <- pr[[pid]]
    if (is.null(info)) next
    for (ft in list(
      list(lab = "C", par = info$E_params),
      list(lab = "D", par = info$F_params)
    )) {
      par <- ft$par
      if (is.null(par)) next
      e00 <- eta_part(pnum(par, "A_00"), pnum(par, "alpha_m_00"))
      e11 <- eta_part(pnum(par, "A_11"), pnum(par, "alpha_m_11"))
      part_param_rows[[length(part_param_rows) + 1L]] <- data.frame(
        Partition = format_partition_label(pid),
        Model = ft$lab,
        mu_0 = fmt_part(pnum(par, "mu_0")),
        mu_1 = fmt_part(pnum(par, "mu_1")),
        A_00 = fmt_part(pnum(par, "A_00")),
        A_11 = fmt_part(pnum(par, "A_11")),
        alpha_00 = fmt_part(pnum(par, "alpha_m_00")),
        alpha_11 = fmt_part(pnum(par, "alpha_m_11")),
        A_01 = fmt_tiny(pnum(par, "A_01")),
        A_10 = fmt_tiny(pnum(par, "A_10")),
        c = fmt_tiny(pnum(par, "c")),
        p = fmt_part(pnum(par, "p")),
        D = fmt_part(pnum(par, "D")),
        q = fmt_part(pnum(par, "q")),
        eta_00 = fmt_part(e00, 3L),
        eta_11 = fmt_part(e11, 3L),
        rho = fmt_part(rho_part(par), 3L),
        check.names = FALSE,
        stringsAsFactors = FALSE
      )
    }
  }
}
if (length(part_param_rows) > 0L) {
  part_param_df <- do.call(rbind, part_param_rows)
  rownames(part_param_df) <- NULL
  kable(
    part_param_df,
    align = "llcccccccccccccccc",
    row.names = FALSE,
    col.names = c(
      "Partition", "Model",
      "$\\mu_0$", "$\\mu_1$",
      "$A_{00}$", "$A_{11}$",
      "$\\alpha_{00}$", "$\\alpha_{11}$",
      "$A_{01}$", "$A_{10}$",
      "$c$", "$p$", "$D$", "$q$",
      "$\\eta_{00}$", "$\\eta_{11}$", "$\\rho$"
    ),
    escape = FALSE
  ) |>
    kable_styling(
      bootstrap_options = c("striped", "hover", "bordered"),
      full_width = FALSE, font_size = 8,
      latex_options = c("scale_down", "hold_position")
    ) |>
    scroll_box(width = "100%")
} else {
  cat("Partition C/D parameters are not stored in this results file.\n")
}
```

```{r tbl-partition-sem-aoi-loglik}
#| label: tbl-partition-sem-aoi-loglik
#| tbl-cap: "Fit C and SEM D log-likelihoods on the AOI geometry. Fit C uses that tessellation's naive C parameters with AOI-boundary labels. The D columns use that tessellation's SEM-D parameters with last proposed labelling vs AOI-boundary labels. All columns share the AOI control/treated regions and the same KDE field. Higher is better. Delta = D last labelling minus D AOI-boundary labels."
support_path <- file.path(SCRIPT_DIR, "oklahoma_report_support.R")
if (file.exists(support_path) &&
    !exists("oklahoma_report_partition_sem_aoi_loglik", mode = "function")) {
  source(support_path, local = FALSE)
}
pp_pre_holdout_ll <- NULL
if (!is.null(results$pp_data) && !is.null(results$pp_data$pp_pre_holdout)) {
  pp_pre_holdout_ll <- coerce_pp_df(results$pp_data$pp_pre_holdout)
}
primary_ll_info <- list(
  label = primary_partition_id,
  E_params = if (!is.null(fit_C_obj)) fit_C_obj$params else NULL,
  F_params = if (!is.null(fit_D_obj)) fit_D_obj$params else NULL,
  semF = if (!is.null(fit_D_obj)) {
    if (!is.null(fit_D_obj$fit) && !is.null(fit_D_obj$fit$adaptive)) {
      fit_D_obj$fit
    } else {
      fit_D_obj$sem
    }
  } else {
    NULL
  },
  pp_post_sem = NULL
)
if (!is.null(pr[[primary_partition_id]]) &&
    !is.null(pr[[primary_partition_id]]$pp_post_sem)) {
  primary_ll_info$pp_post_sem <- pr[[primary_partition_id]]$pp_post_sem
}
sem_aoi_ll_df <- if (exists("oklahoma_report_partition_sem_aoi_loglik", mode = "function") &&
                     exists("win_km_part") && exists("aoi_owin_part") &&
                     !is.null(win_km_part) && !is.null(aoi_owin_part) &&
                     !is.null(pp_pre_holdout_ll)) {
  oklahoma_report_partition_sem_aoi_loglik(
    partition_results = pr,
    primary_id = primary_partition_id,
    primary_info = primary_ll_info,
    pp_pre = pp_pre,
    pp_post = pp_post,
    pp_pre_holdout = pp_pre_holdout_ll,
    win_km = win_km_part,
    aoi_owin = aoi_owin_part,
    kde_info = results$kde_info,
    cfg = cfg,
    format_label = format_partition_label
  )
} else {
  NULL
}
if (!is.null(sem_aoi_ll_df) && nrow(sem_aoi_ll_df) > 0L) {
  ll_print <- sem_aoi_ll_df
  ll_print$Fit_C_LL <- round(ll_print$Fit_C_LL, 2)
  ll_print$Last_labelling_LL <- round(ll_print$Last_labelling_LL, 2)
  ll_print$AOI_boundary_LL <- round(ll_print$AOI_boundary_LL, 2)
  ll_print$Delta_last_minus_AOI <- round(ll_print$Delta_last_minus_AOI, 2)
  names(ll_print) <- c(
    "PartitionID",
    "Partition",
    "Fit C LL",
    "D last labelling LL",
    "D AOI-boundary LL",
    "Delta (D last - D AOI)",
    "Post labels last != AOI",
    "N post"
  )
  ll_print <- subset(ll_print, select = -PartitionID)
  rownames(ll_print) <- NULL
  kable(ll_print, align = "lrrrrrr", row.names = FALSE) |>
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"),
                  full_width = FALSE, font_size = 10)
} else {
  cat("C/D AOI-vs-last-labelling log-likelihoods could not be rebuilt from this results file (need KDE holdout, AOI geometry, and Fit C or Fit D parameters).\n")
}
```

```{r fig-partition-params}
#| label: fig-partition-params
#| fig-cap: "Marginal ETAS parameters (mu, A, alpha_m) across partition schemes. The primary tessellation is highlighted."
#| fig-height: 6
if (exists("part_df") && nrow(part_df) > 0) {
  plot_df <- part_df
  plot_df$is_county <- as.character(plot_df$PartitionID) == primary_partition_id

  param_long <- rbind(
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "control", mu = plot_df$mu_0, A = plot_df$A_0, alpha_m = plot_df$alpha_m_0),
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "treated", mu = plot_df$mu_1, A = plot_df$A_1, alpha_m = plot_df$alpha_m_1)
  )
  param_long <- pivot_longer(param_long, cols = c(mu, A, alpha_m),
                             names_to = "Parameter", values_to = "Value")
  param_long$Process <- factor(param_long$Process, levels = c("control", "treated"))

  param_long$Parameter <- factor(
    param_long$Parameter,
    levels = c("mu", "A", "alpha_m"),
    labels = c(expression(mu), expression(A), expression(alpha[m]))
  )

  ggplot(param_long, aes(x = Partition, y = Value, colour = Model, group = Model)) +
    geom_line(alpha = 0.45) +
    geom_point(aes(size = is_county), alpha = 0.9) +
    scale_size_manual(values = c(`FALSE` = 2.1, `TRUE` = 3.6), guide = "none") +
    scale_colour_manual(values = c("C: Naive Biv+KDE (all-free)" = "#fdb462", "D: SEM Biv+KDE (all-free)" = "#80b1d3")) +
    facet_grid(Parameter ~ Process, scales = "free_y", labeller = label_parsed) +
    labs(x = "Partition scheme", y = "Estimate", colour = "Model") +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      axis.text.x = element_text(angle = 30, hjust = 1),
      strip.text = element_text(face = "bold")
    )
}
```

```{r fig-partition-ate}
#| label: fig-partition-ate
#| fig-cap: "Partition-sensitivity total-saved distributions (same units and x-axis scale as Section 6.1)."
if (exists("biv_sim") && !is.null(biv_sim) && nrow(biv_sim) > 0) {
  biv_xlim <- range(biv_sim$TotalATE, na.rm = TRUE)
} else {
  biv_xlim <- NULL
}

part_sim_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (identical(info$label, primary_partition_id)) {
    list(ate_E = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL,
         ate_F = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL)
  } else {
    ap[[pname]]
  }
  for (fit_type in list(list(lab = "C: Naive Biv+KDE (all-free)", ate = ate_info$ate_E),
                        list(lab = "D: SEM Biv+KDE (all-free)",   ate = ate_info$ate_F))) {
    if (!is.null(fit_type$ate) && !is.null(fit_type$ate$all_nothing_sim)) {
      part_sim_rows[[length(part_sim_rows) + 1]] <- data.frame(
        Partition = info$label,
        Model = fit_type$lab,
        TotalATE = get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles)
      )
    }
  }
}

if (length(part_sim_rows) > 0) {
  part_sim <- do.call(rbind, part_sim_rows)
  part_sim$PartitionID <- part_sim$Partition
  part_sim$Partition <- vapply(part_sim$Partition, format_partition_label, character(1))
  part_sim_ids <- unique(as.character(part_sim$PartitionID))
  part_sim$PartitionID <- factor(part_sim$PartitionID,
                                 levels = c(primary_partition_id,
                                            sort(setdiff(part_sim_ids, primary_partition_id))))
  part_order <- unique(as.character(part_sim[order(part_sim$PartitionID), "Partition"]))
  part_sim$Partition <- factor(part_sim$Partition, levels = part_order)
  p_part <- ggplot(part_sim, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 25, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    facet_wrap(~Partition, scales = "fixed") +
    scale_fill_manual(values = c("C: Naive Biv+KDE (all-free)" = "#fdb462", "D: SEM Biv+KDE (all-free)" = "#80b1d3")) +
    labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes saved /" ~ .(ate_days) ~ "days)"),
         y = "Simulation count") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          strip.text = element_text(face = "bold"))
  if (!is.null(biv_xlim) && all(is.finite(biv_xlim))) {
    p_part <- p_part + coord_cartesian(xlim = biv_xlim)
  }
  print(p_part)
}
```

# SEM initialization variability (Fit D)

Each replicate below starts from **that tessellation's C MLE** $\theta$ and an independent sequential Bernoulli labelling of post-treatment events, then polishes for the configured SEM inner steps. This is initialization variability, not the parametric bootstrap.

```{r fig-fitvar-d-ate}
#| label: fig-fitvar-d-ate
#| fig-cap: "Histogram of Fit D all-or-nothing ATE across independent sequential-Bernoulli initializations (same C MLE)."
fitvar_helper <- file.path(SCRIPT_DIR, "oklahoma_fitvar_report.R")
if (file.exists(fitvar_helper) && !exists("oklahoma_fitvar_d_ates", mode = "function")) {
  source(fitvar_helper, local = FALSE)
}
if (!exists("oklahoma_fitvar_d_ates", mode = "function")) {
  oklahoma_fitvar_d_ates <- function(results) numeric(0)
  oklahoma_fitvar_d_summary <- function(results) {
    data.frame(rep = integer(), success = logical(),
               mc_total_saved_mean = numeric(), stringsAsFactors = FALSE)
  }
}
fitvar_d_df <- oklahoma_fitvar_d_summary(results)
fitvar_d_ates <- oklahoma_fitvar_d_ates(results)
fitvar_cfg <- results$fit_variability$config
d_point_ate <- if (!is.null(fit_D_obj) && !is.null(fit_D_obj$ate)) {
  vals <- get_total_effect_sim(fit_D_obj$ate, n_tiles = n_cty)
  vals <- suppressWarnings(as.numeric(vals))
  vals <- vals[is.finite(vals)]
  if (length(vals) > 0L) mean(vals) else NA_real_
} else {
  NA_real_
}
if (length(fitvar_d_ates) > 0L) {
  n_ok <- length(fitvar_d_ates)
  n_att <- if (!is.null(fitvar_cfg$reps)) as.integer(fitvar_cfg$reps) else n_ok
  init_lab <- if (!is.null(fitvar_cfg$label_init)) {
    as.character(fitvar_cfg$label_init)
  } else if (!is.null(cfg$SEM_LABEL_INIT)) {
    as.character(cfg$SEM_LABEL_INIT)
  } else {
    "sequential_bernoulli"
  }
  inner_n <- if (!is.null(fitvar_cfg$sem_inner_iter)) {
    as.integer(fitvar_cfg$sem_inner_iter)
  } else if (!is.null(cfg$SEM_INNER_ITER)) {
    as.integer(cfg$SEM_INNER_ITER)
  } else {
    NA_integer_
  }
  cat(sprintf(
    "Successful D inits: %d / %d. Mean ATE = %.1f (SD %.1f; range [%.1f, %.1f]). SEM inner = %s; label init = %s.\n",
    n_ok, n_att, mean(fitvar_d_ates), stats::sd(fitvar_d_ates),
    min(fitvar_d_ates), max(fitvar_d_ates),
    if (is.finite(inner_n)) as.character(inner_n) else "NA",
    init_lab
  ))
  hist_df <- data.frame(TotalATE = fitvar_d_ates)
  p_fitvar <- ggplot(hist_df, aes(x = TotalATE)) +
    geom_histogram(bins = min(20L, max(8L, n_ok %/% 3L)),
                   fill = "#80b1d3", colour = "white", alpha = 0.85) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    labs(
      x = bquote(hat(Delta)[AoN] ~ "(total earthquakes saved /" ~ .(ate_days) ~ "days)"),
      y = "Initialization count",
      subtitle = sprintf("n = %d successful D initializations", n_ok)
    ) +
    theme_minimal(base_size = 12)
  if (is.finite(d_point_ate)) {
    p_fitvar <- p_fitvar +
      geom_vline(xintercept = d_point_ate, colour = "#3182bd", linewidth = 0.9) +
      annotate(
        "text", x = d_point_ate, y = Inf, vjust = 1.4, hjust = -0.05,
        colour = "#3182bd", size = 3.2,
        label = sprintf("primary D = %.0f", d_point_ate)
      )
  }
  print(p_fitvar)
  if (nrow(fitvar_d_df) > 0L) {
    sm <- data.frame(
      Setting = c("Successful D inits", "Mean ATE", "SD ATE", "Min ATE", "Max ATE",
                  "Median ATE", "Median init Hamming vs location",
                  "Primary D ATE (vline)"),
      Value = c(
        sprintf("%d / %d", n_ok, n_att),
        sprintf("%.1f", mean(fitvar_d_ates)),
        sprintf("%.1f", stats::sd(fitvar_d_ates)),
        sprintf("%.1f", min(fitvar_d_ates)),
        sprintf("%.1f", max(fitvar_d_ates)),
        sprintf("%.1f", stats::median(fitvar_d_ates)),
        {
          ham <- fitvar_d_df$n_label_init_flips[fitvar_d_df$success %in% TRUE]
          ham <- ham[is.finite(ham)]
          if (length(ham) > 0L) sprintf("%.0f", stats::median(ham)) else "NA"
        },
        if (is.finite(d_point_ate)) sprintf("%.1f", d_point_ate) else "NA"
      ),
      stringsAsFactors = FALSE
    )
    knitr::kable(sm, align = "lc",
                 caption = "Fit D sequential-Bernoulli initialization variability.") |>
      kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)
  }
} else {
  cat("Fit-variability D ATEs are unavailable in this results file (stage not run, or every replicate failed).\n")
}
```

# Interpretation

```{r summaries, include=FALSE}
n_cty_interp <- if (!is.null(cfg$n_tiles)) cfg$n_tiles else cfg$n_counties
ate_days_interp <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 100

safe_mu_ratio <- function(par) {
  if (is.null(par)) return(NA_real_)
  denom <- safe_bc(par, "mu")
  numer <- safe_bt(par, "mu")
  if (!is.finite(denom) || is.na(denom) || abs(denom) < .Machine$double.eps) return(NA_real_)
  numer / denom
}
mu_ratio_C <- safe_mu_ratio(C_par)
mu_ratio_D <- safe_mu_ratio(D_par)
mu_ratio_E <- safe_mu_ratio(E_par)
mu_ratio_F <- safe_mu_ratio(F_par)
mu_ratio_G <- safe_mu_ratio(G_par)
mu_ratio_H <- safe_mu_ratio(H_par)

total_C_1yr <- if (!is.null(fit_C_obj) && !is.null(fit_C_obj$ate)) round(mean(get_total_effect_sim(fit_C_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_D_1yr <- if (!is.null(fit_D_obj) && !is.null(fit_D_obj$ate)) round(mean(get_total_effect_sim(fit_D_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_E_1yr <- if (!is.null(fit_E_obj) && !is.null(fit_E_obj$ate)) round(mean(get_total_effect_sim(fit_E_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_F_1yr <- if (!is.null(fit_F_obj) && !is.null(fit_F_obj$ate)) round(mean(get_total_effect_sim(fit_F_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_G_1yr <- if (!is.null(fit_G_obj) && !is.null(fit_G_obj$ate)) round(mean(get_total_effect_sim(fit_G_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_H_1yr <- if (!is.null(fit_H_obj) && !is.null(fit_H_obj$ate)) round(mean(get_total_effect_sim(fit_H_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_J_1yr <- if (!is.null(fit_J_obj) && !is.null(fit_J_obj$ate)) round(mean(get_total_effect_sim(fit_J_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA

effect_direction <- function(x, tol = 10) {
  if (is.na(x)) return("unavailable")
  if (x > tol) return("worse under treatment")
  if (x < -tol) return("beneficial under treatment")
  "close to no detectable difference"
}
```

**Inhomogeneous background rate focus.** Fits C and D (primary all-free)
use a spatially inhomogeneous background estimated from the first half of
control pre-treatment data. This provides an independent estimate of
the baseline seismicity pattern without using post-treatment data.

**Background contrast (inhomogeneous variants).** The treated-to-control
background ratio $\mu_1/\mu_0$ is:
`r paste(sprintf("C=%.1fx, D=%.1fx, E=%.1fx, F=%.1fx, G=%.1fx, H=%.1fx",
                 mu_ratio_C, mu_ratio_D, mu_ratio_E, mu_ratio_F, mu_ratio_G, mu_ratio_H), collapse = "")`.

**All-or-nothing total saved (`r ate_days_interp`-day horizon).**

```{r interp-effects, results='asis'}
effect_lines <- c(
  paste0("- **Fit C (naive+KDE, all-free):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_C_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit D (SEM+KDE, all-free):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_D_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit E (naive+KDE, struct-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_E_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit F (SEM+KDE, struct-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_F_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  {
    sem_totals <- c(total_D_1yr, total_F_1yr)
    sem_totals <- sem_totals[is.finite(sem_totals)]
    if (length(sem_totals) > 0L) {
      paste0("- **Interpretation:** Across KDE bivariate SEM fits (D/F), effects range from **",
             effect_direction(min(sem_totals)),
             "** to **",
             effect_direction(max(sem_totals)),
             "**.")
    } else {
      "- **Interpretation:** KDE SEM variant effects are unavailable in this results file."
    }
  }
)
if (!is.na(total_H_1yr)) {
  effect_lines <- c(effect_lines, paste0("- **Fit H (SEM univariate+KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_H_1yr, "** total earthquakes saved / ", ate_days_interp, " days"))
}
if (!is.na(total_J_1yr)) {
  effect_lines <- c(effect_lines, paste0("- **Fit J (SEM univariate, no KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_J_1yr, "** total earthquakes saved / ", ate_days_interp, " days"))
}
cat(paste(effect_lines, collapse = "\n"), "\n")
```

**Partition sensitivity.** The alternative partition schemes (grids at
different scales and the AOI-as-single-region partition) show how the
treatment effect estimate varies with spatial resolution. The C/D
log-likelihood table scores each tessellation's fitted parameters on the
same AOI geometry. Fit C is the naive complete-data likelihood on
AOI-boundary labels; a D last-labelling gain over D's AOI-boundary
labels is a labelling comparison rather than a change of spatial
compensator.

# Configuration

```{r tbl-config}
#| label: tbl-config
scalar_chr <- function(x, default = "N/A") {
  if (is.null(x)) return(default)
  xv <- unlist(x, use.names = FALSE)
  if (length(xv) < 1L) return(default)
  x1 <- xv[[1]]
  if (is.na(x1)) return(default)
  as.character(x1)
}
ate_win <- scalar_chr(cfg$ATE_WINDOW_DAYS)
sem_inner_props <- scalar_chr(cfg$SEM_INNER_PROPS %||% cfg$SEM_N_LABELLINGS)
grid_vals <- suppressWarnings(as.numeric(unlist(cfg$grid_diameters, use.names = FALSE)))
grid_vals <- grid_vals[is.finite(grid_vals)]
grid_str <- if (length(grid_vals) > 0L) {
  paste(round(grid_vals, 2), "km", collapse = ", ")
} else { "N/A" }
cdf <- data.frame(
  Setting = c("Magnitude threshold","GR \u03b2","Primary tessellation",
              "Primary tiles (total / treated)",
              "Counties (total / treated)",
              "SEM outer iters","SEM inner iters","SEM inner proposals","SEM labellings",
              "SEM change factor","SEM max relabel step frac","SEM label init","SEM map teleport once","SEM forced refit flip frac","Fit-variability reps / cores","Fitting window (days)",
              "SEM t_trunc (days)","SEM t_trunc source","SEM t_trunc rel level",
              "ATE simulation window (days)","Simulation replicates",
              "Spatial scale D (km)", "Estimated triggering range (km)", "Grid diameters",
              "Test mode"),
  Value = c(scalar_chr(cfg$ETAS_M0), scalar_chr(cfg$BETA_GR),
            scalar_chr(cfg$PRIMARY_PARTITION),
            paste0(scalar_chr(cfg$n_tiles), " / ", scalar_chr(cfg$n_treated)),
            paste0(scalar_chr(cfg$n_counties), " / ",
                   scalar_chr(cfg$n_treated_counties %||% results$counties$n_treated)),
            scalar_chr(cfg$SEM_N_ITER), scalar_chr(cfg$SEM_INNER_ITER), sem_inner_props, scalar_chr(cfg$SEM_N_LABELLINGS),
            scalar_chr(cfg$SEM_CHANGE_FACTOR),
            scalar_chr(cfg$SEM_MAX_RELABEL_STEP_FRAC),
            scalar_chr(cfg$SEM_LABEL_INIT),
            scalar_chr(cfg$SEM_MAP_TELEPORT_ONCE),
            scalar_chr(cfg$SEM_FORCE_PARAM_UPDATE_FLIP_FRAC),
            paste0(scalar_chr(cfg$FIT_VARIABILITY_REPS), " / ", scalar_chr(cfg$FIT_VARIABILITY_CORES)),
            if (!is.null(cfg$windowT_post)) paste(unlist(cfg$windowT_post, use.names = FALSE), collapse=" \u2013 ") else "N/A",
            fmt_t_trunc_cfg(cfg$SEM_T_TRUNC_DAYS),
            scalar_chr(cfg$SEM_T_TRUNC_SOURCE),
            scalar_chr(cfg$SEM_T_TRUNC_REL),
            ate_win, scalar_chr(cfg$ATE_N_SIMS),
            fmt_scaled_num(cfg$spatial_scale_D),
            fmt_scaled_num(if (!is.null(results$kde_info)) results$kde_info$trigger_range_km else NULL),
            grid_str,
            scalar_chr(cfg$TEST_MODE)),
  stringsAsFactors = FALSE
)
kable(cdf, align="lc", caption="Analysis configuration.") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)
```

# References

- Ogata, Y. (1988). Statistical models for earthquake occurrences.
  *JASA*, 83(401), 9--27.
- Zhuang, J., Ogata, Y. & Vere-Jones, D. (2002). Stochastic
  declustering of space-time earthquake occurrences. *JASA*, 97(458),
  369--380.
- Helmstetter, A. & Sornette, D. (2003). Importance of direct and
  indirect triggered seismicity in the ETAS model of seismicity.
  *Geophysical Research Letters*, 30(11), 1576.
- Langenbruch, C. & Zoback, M. D. (2016). How will induced seismicity
  in Oklahoma respond to decreased saltwater injection rates? *Science
  Advances*, 2(11), e1601542.

---
*Report generated by PPDisentangle.*
