---
title: "Oklahoma Induced Seismicity: Causal ETAS Analysis"
subtitle: "Bivariate ETAS with Partition Sensitivity"
author: "PPDisentangle"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 7
    fig-dpi: 150
    self-contained: true
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    fig-width: 7
    fig-height: 5
    fig-dpi: 300
    keep-tex: true
    documentclass: article
    geometry:
      - margin=1in
    echo: false
    include-in-header:
      text: |
        \usepackage{longtable}[=v4.13]
execute:
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: false
library(PPDisentangle)
library(spatstat)
library(sf)
library(tigris)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(fig.align = "center", out.width = "100%")

SCRIPT_DIR <- getwd()
DATA_DIR <- file.path(SCRIPT_DIR,
                       "oklahoma_induced_seismicity_data_regional20150318")
candidate_out_dirs <- unique(c(
  file.path(SCRIPT_DIR, "output"),
  file.path(dirname(dirname(SCRIPT_DIR)), "output")
))
result_patterns <- c("^oklahoma_results(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_bootstrap(_[0-9]+)?\\.rds$")
candidate_result_files <- unlist(lapply(candidate_out_dirs, function(d) {
  if (!dir.exists(d)) return(character(0))
  unique(unlist(lapply(result_patterns, function(pat) {
    list.files(d, pattern = pat, full.names = TRUE)
  })))
}), use.names = FALSE)
existing_result_files <- candidate_result_files[file.exists(candidate_result_files)]
if (length(existing_result_files) == 0) {
  stop("No Oklahoma results RDS file found in expected output directories.")
}
result_file <- existing_result_files[which.max(file.info(existing_result_files)$mtime)]
OUT_DIR <- dirname(result_file)
candidate_plot_dirs <- unique(file.path(candidate_out_dirs, "plots"))

results <- readRDS(result_file)
cfg <- results$config
cty <- results$counties
using_pre_bootstrap <- isTRUE(identical(results$metadata$stage, "pre_bootstrap")) ||
  grepl("pre_bootstrap", basename(result_file), fixed = TRUE)

resolve_plot_file <- function(filename) {
  candidates <- file.path(candidate_plot_dirs, filename)
  existing <- candidates[file.exists(candidates)]
  if (length(existing) == 0) return(NA_character_)
  existing[which.max(file.info(existing)$mtime)]
}

render_timestamp <- format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")

has_fit <- function(fit_name) {
  !is.null(results[[fit_name]])
}
decode_enabled <- isTRUE(cfg$RUN_DECODE) || has_fit("fitG") || has_fit("fitH")

coerce_pp_df <- function(obj) {
  if (is.null(obj)) return(data.frame())

  # Handle spatstat point-pattern objects.
  if (inherits(obj, "ppp")) {
    out <- data.frame(x = obj$x, y = obj$y)
    if (!is.null(obj$marks)) {
      marks_df <- as.data.frame(obj$marks)
      out <- cbind(out, marks_df, stringsAsFactors = FALSE)
    }
    return(out)
  }

  # Standard data-frame-like objects.
  if (is.data.frame(obj)) return(as.data.frame(obj))
  if (is.matrix(obj)) return(as.data.frame(obj))

  # Common list structure with coordinate/time vectors.
  if (is.list(obj) && all(c("x", "y") %in% names(obj))) {
    n <- min(length(obj$x), length(obj$y))
    out <- data.frame(
      x = as.numeric(obj$x)[seq_len(n)],
      y = as.numeric(obj$y)[seq_len(n)]
    )
    if ("t" %in% names(obj)) out$t <- as.numeric(obj$t)[seq_len(min(n, length(obj$t)))]
    for (nm in setdiff(names(obj), c("x", "y", "t"))) {
      v <- obj[[nm]]
      if (length(v) >= n && is.atomic(v)) out[[nm]] <- v[seq_len(n)]
    }
    return(out)
  }

  # Last fallback.
  out <- tryCatch(as.data.frame(obj), error = function(e) NULL)
  if (is.null(out)) {
    stop("Could not coerce point-pattern object to data frame for plotting.")
  }
  out
}
```

# Introduction

**Report rendered at:** `r render_timestamp`

**Results source:** `r basename(result_file)` `r if (using_pre_bootstrap) "(pre-bootstrap checkpoint)" else "(full results)"`

On 18 March 2015 the Oklahoma Corporation Commission (OCC) issued
directive **AOI_20150318**, requiring operators within a defined Area of
Interest (AOI) in central Oklahoma to reduce wastewater disposal volumes.
This report estimates the **all-or-nothing causal effect** of the
directive on earthquake rates using the **PPDisentangle** framework.

## Estimand

The target quantity is the **all-or-nothing total treatment effect**:
the expected difference in total earthquake counts between
a world in which *every* county receives the directive
($\mathbf{Z}=(1,\dots,1)$) and a world in which *no* county receives it
($\mathbf{Z}=(0,\dots,0)$):

$$
\Delta_{\text{AoN}}
= \bigl\{
   \mathbb{E}\!\bigl[N(\mathbf{1})\bigr]
  -\mathbb{E}\!\bigl[N(\mathbf{0})\bigr]
\bigr\},
$$

where $N(\mathbf{z})$ denotes the total event count under treatment
assignment $\mathbf{z}$.

## Models

The analysis fits four bivariate ETAS models on the **county partition**:

| Label | Model | Background | Labels |
|:---:|:---|:---|:---|
| **B** | Naive bivariate ETAS | Homogeneous | Location-based |
| **D** | SEM bivariate ETAS | Homogeneous | SEM-corrected |
| **E** | Naive bivariate ETAS | KDE (non-parametric) | Location-based |
| **F** | SEM bivariate ETAS | KDE (non-parametric) | SEM-corrected |

Additionally, fits B and D are repeated under alternative **partitioning
schemes** to assess sensitivity to spatial resolution.

# Study Design

```{r design}
#| label: design
meta <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
```

**Treatment.** OCC directive AOI_20150318: mandatory reduction in
wastewater disposal volumes within the AOI.

**Tessellation.** The `r cty$n_counties` Oklahoma counties (US Census
Bureau via `tigris`). Counties whose centroid falls inside the AOI are
treated.

**Treated counties** (`r cty$n_treated`):
`r paste(cty$treated_names, collapse = ", ")`.

**Temporal windows.**

- Pre-treatment: `r meta$design$pre_start_utc` to `r meta$design$t_star_utc`
- Post-treatment: `r meta$design$t_star_utc` to `r meta$design$post_end_utc`

**Magnitude threshold.** $m_0 =$ `r cfg$ETAS_M0`.

**Event counts.**
- Pre-treatment total: `r if (!is.null(cfg$n_pre_total)) cfg$n_pre_total else cfg$n_pre`
- Pre-treatment holdout (first 50%): `r if (!is.null(cfg$n_pre_holdout)) cfg$n_pre_holdout else NA`
- Pre-treatment used for estimation/carryover (second 50%): `r cfg$n_pre`
- Post-treatment: `r cfg$n_post`
- Total used in SEM estimation (used pre + post): `r cfg$n_pre + cfg$n_post`

# Data

```{r load-data}
#| label: load-data
ev_all <- fread(file.path(DATA_DIR, "events_all.csv"))
t_star_utc <- as.POSIXct(meta$design$t_star_utc, tz = "UTC")
ev_all[, time_utc := as.POSIXct(time_utc, tz = "UTC")]
ev_all[, t_days := as.numeric(difftime(time_utc, t_star_utc, units = "days"))]
ev_all <- ev_all[mag >= cfg$ETAS_M0]
pp_pre  <- results$pp_data$pp_pre
pp_post <- results$pp_data$pp_post
pp_pre <- coerce_pp_df(pp_pre)
pp_post <- coerce_pp_df(pp_post)
if (!("t" %in% names(pp_pre))) pp_pre$t <- numeric(nrow(pp_pre))
if (!("t" %in% names(pp_post))) pp_post$t <- numeric(nrow(pp_post))
```

## Magnitude Distribution

```{r fig-mag}
#| label: fig-mag
#| fig-cap: "Magnitude distribution of events above the completeness threshold."
ggplot(ev_all, aes(x = mag)) +
  geom_histogram(binwidth = 0.1, fill = "#4292c6", color = "white",
                 boundary = cfg$ETAS_M0) +
  geom_vline(xintercept = cfg$ETAS_M0, linetype = "dashed", color = "red") +
  annotate("text", x = cfg$ETAS_M0 + 0.15, y = Inf, vjust = 2,
           label = paste0("m0 = ", cfg$ETAS_M0), color = "red", size = 4) +
  labs(x = "Magnitude", y = "Count") +
  theme_minimal(base_size = 13)
```

## Temporal Profile

```{r fig-temporal}
#| label: fig-temporal
#| fig-cap: "Cumulative event count; dashed line marks the OCC directive."
ev_all[, cum_n := seq_len(.N)]
ggplot(ev_all, aes(x = t_days, y = cum_n)) +
  geom_step(color = "#2171b5", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("text", x = 5, y = max(ev_all$cum_n) * 0.1,
           label = "OCC directive\n(18 Mar 2015)", color = "red", hjust = 0, size = 3.5) +
  labs(x = "Days since treatment", y = "Cumulative events") +
  theme_minimal(base_size = 13)
```

## County Partition

```{r fig-partition}
#| label: fig-partition
#| fig-cap: "Oklahoma county-level partition. Red counties are treated (centroid inside AOI)."
options(tigris_use_cache = TRUE)
counties_sf <- counties(state = "OK", cb = TRUE, year = 2022)
counties_sf <- st_transform(counties_sf, 5070)
counties_sf <- st_make_valid(counties_sf)

aoi_sf <- st_read(file.path(DATA_DIR, "occ_aoi_layer_2.geojson"), quiet = TRUE)
aoi_sf <- st_transform(aoi_sf, 5070) |> st_make_valid()
aoi_union <- st_union(aoi_sf)

county_centroids <- st_centroid(counties_sf)
inside_aoi <- lengths(st_within(county_centroids, aoi_union)) > 0
counties_sf$treatment <- ifelse(inside_aoi, "treated", "control")

ggplot() +
  geom_sf(data = counties_sf, aes(fill = treatment),
          color = "grey40", linewidth = 0.4) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    name = "Assignment") +
  geom_sf(data = aoi_sf, fill = NA, color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Oklahoma County Partition",
       subtitle = paste0("Treated: ", paste(cty$treated_names, collapse = ", "))) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 9))

counties_km <- counties_sf
st_geometry(counties_km) <- st_geometry(counties_sf) / 1000
st_crs(counties_km) <- NA
```

## Point Patterns

```{r fig-pp}
#| label: fig-pp
#| fig-cap: "Pre-treatment (left) and post-treatment (right) events overlaid on the county partition."
#| fig-height: 5
xlims <- range(c(pp_pre$x, pp_post$x))
ylims <- range(c(pp_pre$y, pp_post$y))

base_map <- ggplot() +
  geom_sf(data = counties_km, aes(fill = treatment),
          colour = "grey60", linewidth = 0.25) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    guide = "none") +
  coord_sf(xlim = xlims, ylim = ylims, datum = NA, expand = FALSE) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = "none")

p1 <- base_map +
  geom_point(data = pp_pre, aes(x = x, y = y, alpha = t),
             colour = "#2166ac", size = 0.8) +
  scale_alpha_continuous(range = c(0.15, 0.8), guide = "none") +
  labs(title = "Pre-treatment", x = "X (km)", y = "Y (km)")

pp_post_plot <- pp_post
process_col <- dplyr::case_when(
  "location_process" %in% names(pp_post_plot) ~ "location_process",
  "inferred_process" %in% names(pp_post_plot) ~ "inferred_process",
  "process" %in% names(pp_post_plot) ~ "process",
  TRUE ~ NA_character_
)
if (is.na(process_col)) {
  pp_post_plot$Process <- factor("control", levels = c("control", "treated"))
} else {
  pp_post_plot$Process <- factor(as.character(pp_post_plot[[process_col]]),
                                 levels = c("control", "treated"))
}
p2 <- base_map +
  geom_point(data = pp_post_plot,
             aes(x = x, y = y, colour = Process, alpha = t),
             size = 0.8) +
  scale_colour_manual(values = c(control = "#2166ac", treated = "#b2182b"),
                      guide = "none") +
  scale_alpha_continuous(range = c(0.3, 1), guide = "none") +
  labs(title = "Post-treatment", x = "X (km)", y = "Y (km)")

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

# Model

## ETAS Conditional Intensity

$$
\lambda(t,x,y) = \frac{\mu}{|S|} + \sum_{j:\,t_j<t} A\,e^{\alpha_m(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

with Omori--Utsu temporal kernel $g$ and power-law spatial kernel $f$.

The kernels used in the fits are:

$$
g(\Delta t) = \frac{p-1}{c}\left(1+\frac{\Delta t}{c}\right)^{-p}, \qquad \Delta t > 0
$$

and

$$
f(\Delta x,\Delta y \mid m_j) =
\frac{q-1}{\pi\,d(m_j)}\,
\left(1+\frac{r^2}{d(m_j)}\right)^{-q},
\quad r^2 = \Delta x^2 + \Delta y^2,
\quad d(m_j)=D\exp\{\gamma(m_j-m_0)\}.
$$

## Bivariate Extension

$$
\lambda_k(t,x,y) = \frac{\mu_k}{|S_k|}\,W_k(x,y) + \sum_{l\in\{0,1\}}\sum_{\substack{j:\,t_j<t\\Z_j=l}} A_{kl}\,e^{\alpha_{m,kl}(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

Off-diagonal terms $A_{01}$ (treated$\to$control) and $A_{10}$
(control$\to$treated) capture cross-excitation. For the KDE background
models (E, F), $W_k(x,y)$ is the normalized kernel density estimate from
control pre-treatment data; for the homogeneous models (B, D),
$W_k \equiv 1$.

## Non-Parametric Background Rate

```{r tbl-kde}
#| label: tbl-kde
kde <- results$kde_info
kde_holdout <- if (!is.null(kde$n_pre_holdout)) kde$n_pre_holdout else NA
kde_used <- if (!is.null(kde$n_pre_used)) kde$n_pre_used else NA
kde_df <- data.frame(
  Setting = c("KDE training events (held-out control pre)",
              "Held-out pre-treatment events (all processes)",
              "Pre-treatment events used for estimation",
              "Bandwidth (km)", "Estimated triggering range (km)", "Method"),
  Value = c(kde$n_training, kde_holdout, kde_used,
            round(kde$bw_sigma, 2), round(kde$trigger_range_km, 2),
            "Diggle cross-validation"))
kable(kde_df, align = "lc",
      caption = "KDE background rate estimation details.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

Only the **second 50%** of pre-treatment events are used for
estimation/carryover in both homogeneous and inhomogeneous SEM fits.
The **first 50%** is held out for fairness; in the inhomogeneous case,
its control events are used to estimate the KDE background via
`bw.diggle`.

```{r fig-kde-bg}
#| label: fig-kde-bg
#| fig-cap: "Estimated inhomogeneous background rate from held-out control pre-treatment events."
bg_plot_file <- resolve_plot_file("background_rate_kde.png")
if (file.exists(bg_plot_file)) {
  knitr::include_graphics(bg_plot_file)
}
```

## Fixed Parameters

```{r tbl-fixed}
#| label: tbl-fixed
fixed_cfg <- cfg$FIXED_STRUCTURAL
if (is.null(fixed_cfg) || length(fixed_cfg) == 0) {
  fixed_df <- data.frame(
    Parameter = "None",
    Value = "Estimated (not fixed)",
    Role = "All ETAS parameters are estimated from data",
    stringsAsFactors = FALSE
  )
  cap <- "Fixed structural parameters: none for this run."
} else {
  role_map <- c(
    c = "Omori onset delay",
    p = "Omori decay exponent",
    D = "Spatial scale",
    gamma = "Mag-spatial scaling",
    q = "Spatial decay exponent"
  )
  fixed_df <- data.frame(
    Parameter = names(fixed_cfg),
    Value = as.numeric(unlist(fixed_cfg)),
    Role = unname(role_map[names(fixed_cfg)]),
    stringsAsFactors = FALSE
  )
  cap <- "Fixed structural parameters."
}
kable(fixed_df, align = "lcl", caption = cap) |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## SEM Configuration

- **Outer iterations:** `r cfg$SEM_N_ITER`
- **Inner iterations:** `r cfg$SEM_INNER_ITER`
- **SEM inner proposals (`SEM_INNER_PROPS`):** `r if (!is.null(cfg$SEM_INNER_PROPS)) cfg$SEM_INNER_PROPS else cfg$SEM_N_LABELLINGS`
- **SEM labellings (`SEM_N_LABELLINGS`):** `r cfg$SEM_N_LABELLINGS`
- **Change factor:** `r cfg$SEM_CHANGE_FACTOR`

# Results — County Partition

```{r extract}
#| label: extract
B_par <- results$fitB$params
D_par <- results$fitD$params
E_par <- results$fitE$params
F_par <- results$fitF$params
G_par <- if (!is.null(results$fitG)) results$fitG$params else NULL
H_par <- if (!is.null(results$fitH)) results$fitH$params else NULL

bc <- function(par, p) if (p=="mu") par[["mu_0"]] else par[[paste0(p,"_00")]]
bt <- function(par, p) if (p=="mu") par[["mu_1"]] else par[[paste0(p,"_11")]]
safe_bc <- function(par, p) if (is.null(par)) NA_real_ else bc(par, p)
safe_bt <- function(par, p) if (is.null(par)) NA_real_ else bt(par, p)

model_defs <- list(
  list(id = "B", fit = "fitB", label = "B: Naive", long = "B: Naive Biv", par = B_par),
  list(id = "D", fit = "fitD", label = "D: SEM", long = "D: SEM Biv", par = D_par),
  list(id = "E", fit = "fitE", label = "E: Naive+KDE", long = "E: Naive+KDE", par = E_par),
  list(id = "F", fit = "fitF", label = "F: SEM+KDE", long = "F: SEM+KDE", par = F_par),
  list(id = "G", fit = "fitG", label = "G: Decode", long = "G: Decode Biv", par = G_par),
  list(id = "H", fit = "fitH", label = "H: Decode+KDE", long = "H: Decode+KDE", par = H_par)
)
param_models <- Filter(function(m) !is.null(m$par), model_defs)

fit_fill_colors <- c(
  "B: Naive Bivariate" = "#fc8d62",
  "D: SEM Bivariate" = "#8da0cb",
  "E: Naive Biv+KDE" = "#e78ac3",
  "F: SEM Biv+KDE" = "#a6d854",
  "G: Decode Bivariate" = "#66c2a5",
  "H: Decode Biv+KDE" = "#ffd92f"
)
fit_line_colors <- c(
  "B: Naive Bivariate" = "#d95f02",
  "D: SEM Bivariate" = "#7570b3",
  "E: Naive Biv+KDE" = "#c51b7d",
  "F: SEM Biv+KDE" = "#4d9221",
  "G: Decode Bivariate" = "#1b9e77",
  "H: Decode Biv+KDE" = "#b8860b"
)

get_sem_metric_trace <- function(fit_obj) {
  if (is.null(fit_obj) || is.null(fit_obj$sem) || is.null(fit_obj$sem$adaptive)) return(NULL)
  ad <- fit_obj$sem$adaptive
  if (!is.null(ad$metrics) && length(ad$metrics) > 0) {
    return(as.numeric(ad$metrics))
  }
  # Backward compatibility: reconstruct selected metric per iter as max over proposal metrics.
  if (!is.null(ad$all_metrics) && length(ad$all_metrics) > 0) {
    vals <- vapply(ad$all_metrics, function(v) {
      if (is.null(v) || length(v) == 0) return(NA_real_)
      max(as.numeric(v), na.rm = TRUE)
    }, numeric(1))
    vals[is.infinite(vals)] <- NA_real_
    return(vals)
  }
  NULL
}
```

## Marginal Parameters

```{r tbl-params}
#| label: tbl-params
pr <- c("mu","A","alpha_m")
pl <- c("$\\mu$","$A$","$\\alpha_m$")
tbl <- data.frame(Parameter = pl)
for (m in param_models) {
  tbl[[paste0(m$id, "_c")]] <- sapply(pr, function(p) bc(m$par, p))
  tbl[[paste0(m$id, "_t")]] <- sapply(pr, function(p) bt(m$par, p))
}
rownames(tbl) <- NULL
align_vec <- paste0("l", paste(rep("cc", length(param_models)), collapse = ""))
col_names <- c("", unlist(lapply(param_models, function(m) c(paste0(m$id, ":ctrl"), paste0(m$id, ":treat")))))
header <- c(" " = 1)
for (m in param_models) header[m$long] <- 2
kable(tbl, digits = 4, escape = FALSE, align = align_vec,
      col.names = col_names,
      caption = "Marginal ETAS parameters across available bivariate fits.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                font_size = 8, latex_options = c("scale_down", "hold_position")) |>
  add_header_above(header)
```

## Cross-excitation

```{r tbl-cross}
#| label: tbl-cross
cn <- c("A_01","alpha_m_01","A_10","alpha_m_10")
cl <- c("$A_{01}$ (treat$\\to$ctrl)","$\\alpha_{m,01}$",
        "$A_{10}$ (ctrl$\\to$treat)","$\\alpha_{m,10}$")
cross_tbl <- data.frame(Parameter = cl)
for (m in param_models) {
  cross_tbl[[m$id]] <- sapply(cn, function(nm) m$par[[nm]])
}
rownames(cross_tbl) <- NULL
kable(cross_tbl, digits = 4, escape = FALSE,
      align = paste0("l", paste(rep("c", length(param_models)), collapse = "")),
      col.names = c("", vapply(param_models, function(m) m$label, character(1))),
      caption = "Cross-excitation parameters for available models.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Parameter Comparison

```{r fig-params}
#| label: fig-params
#| fig-cap: "Marginal parameter estimates across bivariate fits."
#| fig-height: 5
param_rows <- lapply(param_models, function(m) {
  rbind(
    data.frame(Fit = m$long, Process = "Control",
               mu = bc(m$par, "mu"), A = bc(m$par, "A"), alpha_m = bc(m$par, "alpha_m")),
    data.frame(Fit = m$long, Process = "Treated",
               mu = bt(m$par, "mu"), A = bt(m$par, "A"), alpha_m = bt(m$par, "alpha_m"))
  )
})
param_long <- do.call(rbind, param_rows)
pm <- pivot_longer(param_long, cols=c(mu,A,alpha_m), names_to="Parameter", values_to="Value")
pm <- pm[is.finite(pm$Value), , drop = FALSE]
pm$Parameter <- factor(pm$Parameter, levels=c("mu","A","alpha_m"), labels=c("mu","A","alpha[m]"))
pm$Fit <- factor(pm$Fit, levels = vapply(param_models, function(m) m$long, character(1)))
ggplot(pm, aes(x=Fit, y=Value, colour=Process)) +
  geom_point(size=3, position=position_dodge(0.4)) +
  facet_wrap(~Parameter, scales="free_y", labeller=label_parsed) +
  scale_colour_manual(values=c(Control="#377eb8",Treated="#e41a1c")) +
  labs(y="Estimate", x=NULL) +
  theme_minimal(base_size=13) +
  theme(axis.text.x=element_text(angle=30,hjust=1), strip.text=element_text(face="bold"), legend.position="top")
```

## SEM Convergence

```{r fig-flips}
#| label: fig-flips
#| fig-cap: "Label flips per relabelling iteration."
#| fig-height: 4
build_flip_df <- function(avg_flips, max_flips) {
  if (is.null(avg_flips) || is.null(max_flips)) return(NULL)
  n <- min(length(avg_flips), length(max_flips))
  if (n == 0) return(NULL)
  data.frame(
    iteration = seq_len(n),
    average_flips = as.numeric(avg_flips[seq_len(n)]),
    max_metric_flips = as.numeric(max_flips[seq_len(n)])
  )
}

plot_flip_df <- function(df, ttl) {
  if (is.null(df) || nrow(df) == 0) return(NULL)
  ggplot(df, aes(x = iteration)) +
    geom_line(aes(y = average_flips, color = "Average flips")) +
    geom_line(aes(y = max_metric_flips, color = "Max metric flips")) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    labs(title = ttl, x = "Iteration", y = "Number of flips") +
    theme_minimal() +
    scale_color_manual(name = "Metric", values = c("Average flips" = "blue", "Max metric flips" = "red"))
}

plots <- list()
for (nm in list(
  list(kind = "sem", obj = results$fitD$sem, lab = "D: SEM Biv"),
  list(kind = "sem", obj = results$fitF$sem, lab = "F: SEM Biv+KDE"),
  list(kind = "decode", obj = results$fitG$decode, lab = "G: Decode Biv"),
  list(kind = "decode", obj = results$fitH$decode, lab = "H: Decode Biv+KDE")
)) {
  if (nm$kind == "sem" && !is.null(nm$obj) && !is.null(nm$obj$adaptive)) {
    df <- build_flip_df(nm$obj$adaptive$average_flips, nm$obj$adaptive$max_metric_flips)
    p <- plot_flip_df(df, nm$lab)
    if (!is.null(p)) plots[[length(plots)+1]] <- p
  }
  if (nm$kind == "decode" && !is.null(nm$obj)) {
    if (!is.null(nm$obj$trace)) {
      df <- build_flip_df(nm$obj$trace$average_flips, nm$obj$trace$max_metric_flips)
    } else {
      # Backward compatibility for older result files without decode traces.
      nf <- if (!is.null(nm$obj$n_flips)) as.numeric(nm$obj$n_flips) else 0
      df <- build_flip_df(c(nf, nf), c(nf, 0))
    }
    p <- plot_flip_df(df, nm$lab)
    if (!is.null(p)) plots[[length(plots)+1]] <- p
  }
}
if (length(plots) >= 2) {
  ncol_use <- if (length(plots) >= 4) 2 else length(plots)
  do.call(gridExtra::grid.arrange, c(plots, ncol = ncol_use))
} else if (length(plots) == 1) {
  print(plots[[1]])
}
```

```{r fig-sem-adaptive-loglik}
#| label: fig-sem-adaptive-loglik
#| fig-cap: "Adaptive SEM selected-label log-likelihood progression (county SEM fits)."
#| fig-height: 4
sem_metric_rows <- list()
sem_trace_D <- get_sem_metric_trace(results$fitD)
sem_trace_F <- get_sem_metric_trace(results$fitF)
if (!is.null(sem_trace_D) && length(sem_trace_D) > 0) {
  sem_metric_rows[[length(sem_metric_rows) + 1]] <- data.frame(
    Iteration = seq_along(sem_trace_D),
    LogLik = as.numeric(sem_trace_D),
    Model = "D: SEM Biv"
  )
}
if (!is.null(sem_trace_F) && length(sem_trace_F) > 0) {
  sem_metric_rows[[length(sem_metric_rows) + 1]] <- data.frame(
    Iteration = seq_along(sem_trace_F),
    LogLik = as.numeric(sem_trace_F),
    Model = "F: SEM Biv+KDE"
  )
}
if (length(sem_metric_rows) > 0) {
  sem_metric_df <- do.call(rbind, sem_metric_rows)
  sem_metric_df <- sem_metric_df[is.finite(sem_metric_df$LogLik), , drop = FALSE]
  ggplot(sem_metric_df, aes(x = Iteration, y = LogLik, colour = Model)) +
    geom_line(linewidth = 0.9) +
    scale_colour_manual(values = c("D: SEM Biv" = "#7570b3",
                                   "F: SEM Biv+KDE" = "#4d9221")) +
    labs(x = "Adaptive SEM iteration", y = "Selected log-likelihood") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top")
} else {
  plot.new()
  text(0.5, 0.5, "No adaptive SEM log-likelihood trace found in this results file.")
}
```

```{r fig-selected-loglik}
#| label: fig-selected-loglik
#| fig-cap: "Decode selected log-likelihood progression across relabelling iterations (when Decode models are present)."
#| fig-height: 5
metric_rows <- list()
if (!is.null(results$fitG$decode) && !is.null(results$fitG$decode$trace$objective)) {
  metric_rows[[length(metric_rows) + 1]] <- data.frame(
    Iteration = seq_along(results$fitG$decode$trace$objective),
    Objective = as.numeric(results$fitG$decode$trace$objective),
    Model = "G: Decode Biv"
  )
}
if (!is.null(results$fitH$decode) && !is.null(results$fitH$decode$trace$objective)) {
  metric_rows[[length(metric_rows) + 1]] <- data.frame(
    Iteration = seq_along(results$fitH$decode$trace$objective),
    Objective = as.numeric(results$fitH$decode$trace$objective),
    Model = "H: Decode Biv+KDE"
  )
}
if (length(metric_rows) > 0) {
  metric_df <- do.call(rbind, metric_rows)
  metric_df <- metric_df[is.finite(metric_df$Objective), , drop = FALSE]
  ggplot(metric_df, aes(x = Iteration, y = Objective, colour = Model)) +
    geom_line(linewidth = 0.8, alpha = 0.9) +
    scale_colour_manual(values = c(
      "G: Decode Biv" = "#1b9e77",
      "H: Decode Biv+KDE" = "#b8860b"
    )) +
    labs(x = "Iteration", y = "Selected objective (log-likelihood scale)") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top")
}
```

```{r fig-decode-flips}
#| label: fig-decode-flips
#| fig-cap: "Decode relabelling totals (county models)."
decode_counts <- data.frame(
  Model = c("G: Decode Bivariate", "H: Decode Biv+KDE"),
  Flips = c(
    if (!is.null(results$fitG$decode$n_flips)) results$fitG$decode$n_flips else NA_real_,
    if (!is.null(results$fitH$decode$n_flips)) results$fitH$decode$n_flips else NA_real_
  )
)
decode_counts <- decode_counts[is.finite(decode_counts$Flips), , drop = FALSE]
if (nrow(decode_counts) > 0) {
  ggplot(decode_counts, aes(x = Model, y = Flips, fill = Model)) +
    geom_col(width = 0.6, alpha = 0.9) +
    scale_fill_manual(values = c("G: Decode Bivariate" = "#66c2a5",
                                 "H: Decode Biv+KDE" = "#ffd92f")) +
    labs(x = NULL, y = "Number of relabelled points") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 20, hjust = 1))
}
```

## SEM Relabelling: Confusion Matrices

```{r tbl-confusion}
#| label: tbl-confusion
build_cm <- function(sem_obj, label) {
  if (is.null(sem_obj)) return(NULL)
  lab <- sem_obj$adaptive$adaptive_labelling
  post <- lab[lab$t >= 0, ]
  if (is.null(post$location_process) || is.null(post$inferred_process)) return(NULL)
  tab <- table(Location = post$location_process, Inferred = post$inferred_process)
  as.data.frame.matrix(tab)
}

build_cm_from_labelling <- function(lab) {
  if (is.null(lab)) return(NULL)
  post <- lab[lab$t >= 0, ]
  if (is.null(post$location_process) || is.null(post$inferred_process)) return(NULL)
  tab <- table(Location = post$location_process, Inferred = post$inferred_process)
  as.data.frame.matrix(tab)
}

cm_D <- build_cm(results$fitD$sem, "D")
cm_F <- build_cm(results$fitF$sem, "F")
cm_G <- build_cm_from_labelling(if (!is.null(results$fitG$decode)) results$fitG$decode$labelling else NULL)
cm_H <- build_cm_from_labelling(if (!is.null(results$fitH$decode)) results$fitH$decode$labelling else NULL)
```

### Fit D — SEM Bivariate ETAS

```{r tbl-cm-D}
#| label: tbl-cm-D
if (!is.null(cm_D)) {
  n_post_D <- sum(cm_D)
  cm_D_pct <- round(cm_D / n_post_D * 100, 1)
  cm_display_D <- cm_D
  for (rn in rownames(cm_D)) for (cn in colnames(cm_D)) {
    cm_display_D[rn, cn] <- paste0(cm_D[rn, cn], "  (", cm_D_pct[rn, cn], "%)")
  }
  kable(cm_display_D,
        caption = paste0("Fit D: Location vs SEM-inferred label (n = ", n_post_D,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_D)))
}
```

### Fit F — SEM Bivariate + KDE Background

```{r tbl-cm-F}
#| label: tbl-cm-F
if (!is.null(cm_F)) {
  n_post_F <- sum(cm_F)
  cm_F_pct <- round(cm_F / n_post_F * 100, 1)
  cm_display_F <- cm_F
  for (rn in rownames(cm_F)) for (cn in colnames(cm_F)) {
    cm_display_F[rn, cn] <- paste0(cm_F[rn, cn], "  (", cm_F_pct[rn, cn], "%)")
  }
  kable(cm_display_F,
        caption = paste0("Fit F: Location vs SEM-inferred label (n = ", n_post_F,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_F)))
}
```

### Fit G — Decode Bivariate ETAS

```{r tbl-cm-G}
#| label: tbl-cm-G
if (!is.null(cm_G)) {
  n_post_G <- sum(cm_G)
  cm_G_pct <- round(cm_G / n_post_G * 100, 1)
  cm_display_G <- cm_G
  for (rn in rownames(cm_G)) for (cn in colnames(cm_G)) {
    cm_display_G[rn, cn] <- paste0(cm_G[rn, cn], "  (", cm_G_pct[rn, cn], "%)")
  }
  kable(cm_display_G,
        caption = paste0("Fit G: Location vs Decode-inferred label (n = ", n_post_G,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_G)))
}
```

### Fit H — Decode Bivariate + KDE Background

```{r tbl-cm-H}
#| label: tbl-cm-H
if (!is.null(cm_H)) {
  n_post_H <- sum(cm_H)
  cm_H_pct <- round(cm_H / n_post_H * 100, 1)
  cm_display_H <- cm_H
  for (rn in rownames(cm_H)) for (cn in colnames(cm_H)) {
    cm_display_H[rn, cn] <- paste0(cm_H[rn, cn], "  (", cm_H_pct[rn, cn], "%)")
  }
  kable(cm_display_H,
        caption = paste0("Fit H: Location vs Decode-inferred label (n = ", n_post_H,
                         " post-treatment events)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_H)))
}
```

### Relabelling Summary

```{r tbl-cm-summary}
#| label: tbl-cm-summary
sum_rows <- list()
for (info in list(list(cm = cm_D, lab = "D: SEM Bivariate"),
                  list(cm = cm_F, lab = "F: SEM Bivariate+KDE"),
                  list(cm = cm_G, lab = "G: Decode Bivariate"),
                  list(cm = cm_H, lab = "H: Decode Biv+KDE"))) {
  cm <- info$cm
  if (is.null(cm)) next
  n_total <- sum(cm)
  ctrl_to_treat <- if ("treated" %in% colnames(cm) && "control" %in% rownames(cm)) cm["control","treated"] else 0
  treat_to_ctrl <- if ("control" %in% colnames(cm) && "treated" %in% rownames(cm)) cm["treated","control"] else 0
  n_flipped <- ctrl_to_treat + treat_to_ctrl
  sum_rows[[length(sum_rows)+1]] <- data.frame(
    Model = info$lab,
    `Post events` = n_total,
    `Ctrl to Treat` = ctrl_to_treat,
    `Treat to Ctrl` = treat_to_ctrl,
    `Total relabelled` = n_flipped,
    `Pct relabelled` = paste0(round(n_flipped / n_total * 100, 1), "%"),
    check.names = FALSE)
}
if (length(sum_rows) > 0) {
  kable(do.call(rbind, sum_rows),
        caption = "Summary of SEM relabelling across models.",
        align = "lccccc") |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
}
```

# All-or-Nothing Treatment Effect

## County Partition

```{r tbl-ate-biv}
#| label: tbl-ate-biv
n_cty <- cfg$n_counties
ate_days <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 365

biv_rows <- list(
  list(lab = "B: Naive Bivariate", ate = results$fitB$ate, louis = NULL),
  list(lab = "D: SEM Bivariate",   ate = results$fitD$ate, louis = results$fitD$louis),
  list(lab = "E: Naive Biv+KDE",   ate = results$fitE$ate, louis = NULL),
  list(lab = "F: SEM Biv+KDE",     ate = results$fitF$ate, louis = results$fitF$louis))
if (!is.null(results$fitG) && !is.null(results$fitG$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "G: Decode Bivariate", ate = results$fitG$ate, louis = NULL)
}
if (!is.null(results$fitH) && !is.null(results$fitH$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "H: Decode Biv+KDE", ate = results$fitH$ate, louis = NULL)
}

boot_obj <- results$bootstrap_ate
boot_E_df <- NULL
boot_F_df <- NULL
if (!is.null(boot_obj) &&
    !is.null(boot_obj$fit_E) &&
    !is.null(boot_obj$fit_E$replicate_summary) &&
    nrow(boot_obj$fit_E$replicate_summary) > 0) {
  boot_E_df <- boot_obj$fit_E$replicate_summary
}
if (!is.null(boot_obj) &&
    !is.null(boot_obj$fit_F) &&
    !is.null(boot_obj$fit_F$replicate_summary) &&
    nrow(boot_obj$fit_F$replicate_summary) > 0) {
  boot_F_df <- boot_obj$fit_F$replicate_summary
}

get_boot_summary <- function(model_label) {
  vals <- NULL
  if (model_label == "E: Naive Biv+KDE" && !is.null(boot_E_df)) vals <- boot_E_df$ate_total_mean
  if (model_label == "F: SEM Biv+KDE" && !is.null(boot_F_df)) vals <- boot_F_df$ate_total_mean
  if (is.null(vals)) return(NULL)
  vals <- vals[is.finite(vals)]
  if (length(vals) < 2) return(NULL)
  list(
    sd = stats::sd(vals, na.rm = TRUE),
    lo = as.numeric(stats::quantile(vals, probs = 0.025, na.rm = TRUE)),
    hi = as.numeric(stats::quantile(vals, probs = 0.975, na.rm = TRUE)),
    n = length(vals)
  )
}

get_total_effect_sim <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  sim_df <- ate_obj$all_nothing_sim
  if ("total_effect" %in% names(sim_df)) return(sim_df$total_effect)
  # Backward compatibility with old result files.
  sim_df$ATE * n_tiles
}

get_expected_totals <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  sim_df <- ate_obj$all_nothing_sim
  if (all(c("c_total", "t_total") %in% names(sim_df))) {
    return(list(c_total = sim_df$c_total, t_total = sim_df$t_total))
  }
  # Backward compatibility with old result files.
  list(c_total = sim_df$c_mean * n_tiles, t_total = sim_df$t_mean * n_tiles)
}

format_partition_label <- function(part_id) {
  if (is.na(part_id) || is.null(part_id)) return(part_id)
  if (part_id == "county") return("County")
  if (part_id == "aoi_region") return("AOI region")
  m <- regexec("^grid_([0-9.]+)R$", part_id)
  g <- regmatches(part_id, m)[[1]]
  if (length(g) == 2) {
    mult <- sub("\\.0$", "", g[2])
    return(paste0("Grid (", mult, "x triggering range)"))
  }
  part_id
}

ate_biv <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  expected <- get_expected_totals(r$ate, n_tiles = n_cty)
  sim_totals <- get_total_effect_sim(r$ate, n_tiles = n_cty)
  sim_total <- mean(sim_totals, na.rm = TRUE)
  c_mean <- mean(expected$c_total, na.rm = TRUE)
  t_mean <- mean(expected$t_total, na.rm = TRUE)
  sim_sd_total  <- sd(sim_totals, na.rm = TRUE)
  base_row <- data.frame(
    Model = r$lab,
    `E[N(0)]` = round(c_mean, 0),
    `E[N(1)]` = round(t_mean, 0),
    `Total Effect` = round(sim_total, 0),
    `SD` = round(sim_sd_total, 0),
    `Spillover` = round(ifelse(!is.null(r$ate$total_spillover), r$ate$total_spillover, r$ate$ATE_spillover * n_cty), 0),
    check.names = FALSE
  )
  if (using_pre_bootstrap) return(base_row)
  boot_sum <- get_boot_summary(r$lab)
  boot_sd <- if (!is.null(boot_sum)) round(boot_sum$sd, 0) else NA
  boot_ci <- if (!is.null(boot_sum)) {
    sprintf("[%d, %d]", round(boot_sum$lo, 0), round(boot_sum$hi, 0))
  } else {
    "NA"
  }
  base_row$`Bootstrap SD` <- boot_sd
  base_row$`Bootstrap 95% CI` <- boot_ci
  base_row
}))
kable(ate_biv,
      col.names = if (using_pre_bootstrap) {
        c("Model", "E[N(0)]", "E[N(1)]",
          paste0("Total Effect / ", n_cty, " counties / ", ate_days, "d"),
          "SD", paste0("Spillover / ", ate_days, "d"))
      } else {
        c("Model", "E[N(0)]", "E[N(1)]",
          paste0("Total Effect / ", n_cty, " counties / ", ate_days, "d"),
          "SD", "Bootstrap SD", "Bootstrap 95% CI",
          paste0("Spillover / ", ate_days, "d"))
      },
      caption = if (using_pre_bootstrap) {
        paste0("All-or-nothing total effect over ", ate_days, " days (pre-bootstrap checkpoint).")
      } else {
        paste0("All-or-nothing total effect over ", ate_days, " days.")
      },
      align = if (using_pre_bootstrap) "lccccc" else "lccccccc") |>
  kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE)
if (using_pre_bootstrap) {
  cat("\nBootstrap results are not shown because the newest available source is a pre-bootstrap checkpoint.\n")
} else if (!is.null(boot_E_df) || !is.null(boot_F_df)) {
  labs <- c()
  if (!is.null(boot_E_df)) labs <- c(labs, "E (Naive Biv+KDE)")
  if (!is.null(boot_F_df)) labs <- c(labs, "F (SEM Biv+KDE)")
  cat(sprintf("\nBootstrap uncertainty is available for: %s.\n", paste(labs, collapse = ", ")))
}
```

```{r fig-ate-biv}
#| label: fig-ate-biv
#| fig-cap: "Simulation distribution of the all-or-nothing total effect over the post-treatment horizon."
#| fig-height: 4
biv_sim <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  data.frame(Model = r$lab, TotalATE = get_total_effect_sim(r$ate, n_tiles = n_cty))
}))
if (!is.null(biv_sim) && nrow(biv_sim) > 0) {
  biv_means <- aggregate(TotalATE ~ Model, biv_sim, mean)
  model_levels <- unique(as.character(biv_sim$Model))
  fill_vals <- fit_fill_colors[names(fit_fill_colors) %in% model_levels]
  line_vals <- fit_line_colors[names(fit_line_colors) %in% model_levels]
  ggplot(biv_sim, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_vline(data = biv_means, aes(xintercept = TotalATE, colour = Model),
               linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
    scale_fill_manual(values = fill_vals) +
    scale_colour_manual(values = line_vals) +
    labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes /" ~ .(ate_days) ~ "days)"),
         y = "Simulation count",
         title = "All-or-Nothing Total Effect: Bivariate Models") +
    theme_minimal(base_size = 13) +
    theme(legend.position = "top",
          plot.title = element_text(hjust = 0.5, face = "bold"))
}
```

# Partition Sensitivity

The analysis is repeated under alternative spatial partitions to assess
sensitivity to spatial resolution:

- **Grid partitions** with cell diameters equal to 1, 2, and 5
  times the estimated triggering range
  (`r round(results$kde_info$trigger_range_km, 2)` km).
- **AOI region** partition: the AOI polygon defines one treated region,
  everything else is control (2 tiles total).

For each partition, both the naive bivariate + KDE (E-type) and
SEM bivariate + KDE (F-type) models are fitted.

```{r fig-partition-all}
#| label: fig-partition-all
#| fig-cap: "Partition maps for all partition schemes evaluated in the run."
partition_ids <- names(results$partition_results)
partition_plot_files <- character(0)
partition_plot_labels <- character(0)
for (pid in partition_ids) {
  pfile <- resolve_plot_file(paste0("partition_map_", pid, ".png"))
  if (file.exists(pfile)) {
    partition_plot_files <- c(partition_plot_files, pfile)
    partition_plot_labels <- c(partition_plot_labels, format_partition_label(pid))
  }
}
if (length(partition_plot_files) > 0) {
  knitr::kable(
    data.frame(`Panels (top to bottom)` = partition_plot_labels),
    align = "l"
  )
  knitr::include_graphics(partition_plot_files)
}
```

```{r tbl-partition-sensitivity}
#| label: tbl-partition-sensitivity
pr <- results$partition_results
ap <- results$ate_partitions

part_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (info$label == "county") {
    list(ate_E = results$fitE$ate, ate_F = results$fitF$ate)
  } else {
    ap[[pname]]
  }

  for (fit_type in list(
    list(lab = "Naive Biv+KDE", par = info$E_params, ate = ate_info$ate_E, sem_obj = NULL),
    list(lab = "SEM Biv+KDE",   par = info$F_params, ate = ate_info$ate_F, sem_obj = info$semF)
  )) {
    expected <- if (!is.null(fit_type$ate)) {
      get_expected_totals(fit_type$ate, n_tiles = info$n_tiles)
    } else { NULL }
    sim_total <- if (!is.null(fit_type$ate)) {
      round(mean(get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles), na.rm = TRUE), 0)
    } else { NA }
    n_relabel <- if (!is.null(fit_type$sem_obj) &&
                     !is.null(fit_type$sem_obj$adaptive$adaptive_labelling)) {
      lp <- fit_type$sem_obj$adaptive$adaptive_labelling
      lp <- lp[lp$t >= 0, , drop = FALSE]
      sum(lp$location_process != lp$inferred_process, na.rm = TRUE)
    } else { 0 }
    c_mean <- if (!is.null(expected)) round(mean(expected$c_total, na.rm = TRUE), 0) else NA
    t_mean <- if (!is.null(expected)) round(mean(expected$t_total, na.rm = TRUE), 0) else NA

    part_rows[[length(part_rows) + 1]] <- data.frame(
      PartitionID = info$label,
      Partition = format_partition_label(info$label),
      Model = fit_type$lab,
      Tiles = info$n_tiles,
      Treated = info$n_treated,
      `Relabelled points` = n_relabel,
      `E[N(0)]` = c_mean,
      `E[N(1)]` = t_mean,
      mu_0 = round(fit_type$par[["mu_0"]], 4),
      mu_1 = round(fit_type$par[["mu_1"]], 4),
      A_0 = round(fit_type$par[["A_00"]], 4),
      A_1 = round(fit_type$par[["A_11"]], 4),
      alpha_m_0 = round(fit_type$par[["alpha_m_00"]], 4),
      alpha_m_1 = round(fit_type$par[["alpha_m_11"]], 4),
      `Total Effect` = sim_total,
      check.names = FALSE)
  }
}

if (length(part_rows) > 0) {
  part_df <- do.call(rbind, part_rows)
  # Keep county first, then all remaining partitions alphabetically by ID.
  part_df$PartitionID <- factor(part_df$PartitionID,
                                levels = c("county",
                                           sort(setdiff(unique(as.character(part_df$PartitionID)),
                                                        "county"))))
  part_df <- part_df[order(part_df$PartitionID, part_df$Model), ]
  part_df$Partition <- factor(as.character(part_df$Partition),
                              levels = unique(as.character(part_df$Partition)))
  part_df_print <- subset(part_df, select = -PartitionID)
  kable(part_df_print, align = "llccccccccccccc",
        caption = paste0("Partition comparison over ", ate_days, " days: E[N(0)], E[N(1)], marginal ETAS parameters, and total effect.")) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                  font_size = 9, latex_options = c("scale_down", "hold_position"))
}
```

```{r fig-partition-params}
#| label: fig-partition-params
#| fig-cap: "Marginal ETAS parameters (mu, A, alpha_m) across partition schemes. County is highlighted to compare against grid-based partitions."
#| fig-height: 6
if (exists("part_df") && nrow(part_df) > 0) {
  plot_df <- part_df
  plot_df$is_county <- as.character(plot_df$Partition) == "county"

  param_long <- rbind(
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "control", mu = plot_df$mu_0, A = plot_df$A_0, alpha_m = plot_df$alpha_m_0),
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "treated", mu = plot_df$mu_1, A = plot_df$A_1, alpha_m = plot_df$alpha_m_1)
  )
  param_long <- pivot_longer(param_long, cols = c(mu, A, alpha_m),
                             names_to = "Parameter", values_to = "Value")
  param_long$Process <- factor(param_long$Process, levels = c("control", "treated"))

  param_long$Parameter <- factor(
    param_long$Parameter,
    levels = c("mu", "A", "alpha_m"),
    labels = c(expression(mu), expression(A), expression(alpha[m]))
  )

  ggplot(param_long, aes(x = Partition, y = Value, colour = Model, group = Model)) +
    geom_line(alpha = 0.45) +
    geom_point(aes(size = is_county), alpha = 0.9) +
    scale_size_manual(values = c(`FALSE` = 2.1, `TRUE` = 3.6), guide = "none") +
    scale_colour_manual(values = c("Naive Biv+KDE" = "#e78ac3", "SEM Biv+KDE" = "#a6d854")) +
    facet_grid(Parameter ~ Process, scales = "free_y", labeller = label_parsed) +
    labs(x = "Partition scheme", y = "Estimate", colour = "Model") +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      axis.text.x = element_text(angle = 30, hjust = 1),
      strip.text = element_text(face = "bold")
    )
}
```

```{r fig-partition-ate}
#| label: fig-partition-ate
#| fig-cap: "Partition-sensitivity total-effect distributions (same units and x-axis scale as Section 6.1)."
if (exists("biv_sim") && !is.null(biv_sim) && nrow(biv_sim) > 0) {
  biv_xlim <- range(biv_sim$TotalATE, na.rm = TRUE)
} else {
  biv_xlim <- NULL
}

part_sim_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (info$label == "county") {
    list(ate_E = results$fitE$ate, ate_F = results$fitF$ate)
  } else {
    ap[[pname]]
  }
  for (fit_type in list(list(lab = "Naive Biv+KDE", ate = ate_info$ate_E),
                        list(lab = "SEM Biv+KDE",   ate = ate_info$ate_F))) {
    if (!is.null(fit_type$ate) && !is.null(fit_type$ate$all_nothing_sim)) {
      part_sim_rows[[length(part_sim_rows) + 1]] <- data.frame(
        Partition = info$label,
        Model = fit_type$lab,
        TotalATE = get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles)
      )
    }
  }
}

if (length(part_sim_rows) > 0) {
  part_sim <- do.call(rbind, part_sim_rows)
  part_sim$PartitionID <- part_sim$Partition
  part_sim$Partition <- vapply(part_sim$Partition, format_partition_label, character(1))
  part_sim$PartitionID <- factor(part_sim$PartitionID,
                                 levels = c("county",
                                            sort(setdiff(unique(as.character(part_sim$PartitionID)),
                                                         "county"))))
  part_order <- unique(as.character(part_sim[order(part_sim$PartitionID), "Partition"]))
  part_sim$Partition <- factor(part_sim$Partition, levels = part_order)
  p_part <- ggplot(part_sim, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 25, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    facet_wrap(~Partition, scales = "fixed") +
    scale_fill_manual(values = c("Naive Biv+KDE" = "#e78ac3", "SEM Biv+KDE" = "#a6d854")) +
    labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes /" ~ .(ate_days) ~ "days)"),
         y = "Simulation count") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          strip.text = element_text(face = "bold"))
  if (!is.null(biv_xlim) && all(is.finite(biv_xlim))) {
    p_part <- p_part + coord_cartesian(xlim = biv_xlim)
  }
  print(p_part)
}
```

## KDE Bandwidth Sensitivity (County Inhomogeneous E/F)

County-level inhomogeneous fits are re-run across KDE bandwidths:
Diggle, Diggle$\\times$2, Diggle$\\times$5, and Diggle$\\times$10.

```{r tbl-kde-bandwidth-sensitivity}
#| label: tbl-kde-bandwidth-sensitivity
bws <- results$kde_bandwidth_sensitivity
if (!is.null(bws) && length(bws) > 0) {
  bw_rows <- list()
  bw_order <- c("diggle", "digglex2", "digglex5", "digglex10")
  bw_label <- function(x) {
    switch(as.character(x),
           diggle = "Diggle",
           digglex2 = "Diggle x2",
           digglex5 = "Diggle x5",
           digglex10 = "Diggle x10",
           x)
  }
  for (nm in names(bws)) {
    info <- bws[[nm]]
    if (is.null(info)) next
    for (fit_type in list(
      list(model = "Naive Biv+KDE", par = info$E_params, ate = info$ate_E, relab = 0),
      list(model = "SEM Biv+KDE",   par = info$F_params, ate = info$ate_F,
           relab = if (!is.null(info$n_relabel)) info$n_relabel else NA)
    )) {
      if (is.null(fit_type$par)) next
      expected <- if (!is.null(fit_type$ate)) get_expected_totals(fit_type$ate, n_tiles = n_cty) else NULL
      sim_total <- if (!is.null(fit_type$ate)) {
        round(mean(get_total_effect_sim(fit_type$ate, n_tiles = n_cty), na.rm = TRUE), 0)
      } else { NA }
      c_mean <- if (!is.null(expected)) round(mean(expected$c_total, na.rm = TRUE), 0) else NA
      t_mean <- if (!is.null(expected)) round(mean(expected$t_total, na.rm = TRUE), 0) else NA
      bw_rows[[length(bw_rows) + 1]] <- data.frame(
        BandwidthID = nm,
        Bandwidth = bw_label(nm),
        Model = fit_type$model,
        Tiles = n_cty,
        Treated = cfg$n_treated,
        `Relabelled points` = fit_type$relab,
        `E[N(0)]` = c_mean,
        `E[N(1)]` = t_mean,
        mu_0 = round(fit_type$par[["mu_0"]], 4),
        mu_1 = round(fit_type$par[["mu_1"]], 4),
        A_0 = round(fit_type$par[["A_00"]], 4),
        A_1 = round(fit_type$par[["A_11"]], 4),
        alpha_m_0 = round(fit_type$par[["alpha_m_00"]], 4),
        alpha_m_1 = round(fit_type$par[["alpha_m_11"]], 4),
        `Total Effect` = sim_total,
        check.names = FALSE
      )
    }
  }
  if (length(bw_rows) > 0) {
    bw_df <- do.call(rbind, bw_rows)
    bw_df$BandwidthID <- factor(bw_df$BandwidthID, levels = bw_order)
    bw_df <- bw_df[order(bw_df$BandwidthID, bw_df$Model), ]
    bw_df$Bandwidth <- factor(as.character(bw_df$Bandwidth),
                              levels = unique(as.character(bw_df$Bandwidth)))
    bw_print <- subset(bw_df, select = -BandwidthID)
    kable(bw_print, align = "llccccccccccccc",
          caption = paste0("County inhomogeneous KDE bandwidth sensitivity over ", ate_days,
                           " days: E[N(0)], E[N(1)], marginal ETAS parameters, and total effect.")) |>
      kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                    font_size = 9, latex_options = c("scale_down", "hold_position"))
  }
}
```

# Interpretation

```{r summaries, include=FALSE}
n_cty_interp <- cfg$n_counties
ate_days_interp <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 100

mu_ratio_E <- bt(E_par,"mu") / bc(E_par,"mu")
mu_ratio_F <- bt(F_par,"mu") / bc(F_par,"mu")

total_E_1yr <- if (!is.null(results$fitE$ate)) round(mean(get_total_effect_sim(results$fitE$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_F_1yr <- if (!is.null(results$fitF$ate)) round(mean(get_total_effect_sim(results$fitF$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_H_1yr <- if (!is.null(results$fitH) && !is.null(results$fitH$ate)) round(mean(get_total_effect_sim(results$fitH$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA

effect_direction <- function(x, tol = 10) {
  if (is.na(x)) return("unavailable")
  if (x > tol) return("worse under treatment")
  if (x < -tol) return("beneficial under treatment")
  "close to no detectable difference"
}
```

**Inhomogeneous background rate focus.** Fits E and F use a
spatially inhomogeneous background estimated from the first half of
control pre-treatment data. This provides an independent estimate of
the baseline seismicity pattern without using post-treatment data.

**Background contrast (inhomogeneous only).** In Fit E (naive labels),
the treated-region background rate $\mu_1$ is `r round(mu_ratio_E, 1)`$\times$
higher than $\mu_0$. After SEM relabelling (Fit F), this ratio is
`r round(mu_ratio_F, 1)`$\times$.

**All-or-nothing total effect (`r ate_days_interp`-day horizon).**

```{r interp-effects, results='asis'}
effect_lines <- c(
  paste0("- **Fit E (naive+KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_E_1yr, "** total earthquakes / ", ate_days_interp, " days"),
  paste0("- **Fit F (SEM+KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_F_1yr, "** total earthquakes / ", ate_days_interp, " days"),
  paste0("- **Interpretation:** In the inhomogeneous SEM model (F), treatment looks **",
         effect_direction(total_F_1yr), "**; in the inhomogeneous naive model (E), treatment looks **",
         effect_direction(total_E_1yr), "**.")
)
if (!is.na(total_H_1yr)) {
  effect_lines <- c(effect_lines, paste0("- **Fit H (Decode+KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_H_1yr, "** total earthquakes / ", ate_days_interp, " days"))
}
cat(paste(effect_lines, collapse = "\n"), "\n")
```

**Partition sensitivity.** The alternative partition schemes (grids at
different scales and the AOI-as-single-region partition) show how the
treatment effect estimate varies with spatial resolution. Coarser
partitions aggregate more spatial heterogeneity, while finer grids
may better capture localized effects but have fewer events per cell.

# Configuration

```{r tbl-config}
#| label: tbl-config
ate_win <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else "N/A"
sem_inner_props <- if (!is.null(cfg$SEM_INNER_PROPS)) cfg$SEM_INNER_PROPS else cfg$SEM_N_LABELLINGS
grid_str <- if (!is.null(cfg$grid_diameters)) {
  paste(round(cfg$grid_diameters, 2), "km", collapse = ", ")
} else { "N/A" }
cdf <- data.frame(
  Setting = c("Magnitude threshold","GR \u03b2","Counties (total / treated)",
              "SEM outer iters","SEM inner iters","SEM inner proposals","SEM labellings",
              "SEM change factor","Decode enabled","Decode iterations","Fitting window (days)",
              "ATE simulation window (days)","Simulation replicates",
              "Spatial scale D (km)", "Estimated triggering range (km)", "Grid diameters",
              "Test mode"),
  Value = c(cfg$ETAS_M0, cfg$BETA_GR,
            paste0(cfg$n_counties," / ",cfg$n_treated),
            cfg$SEM_N_ITER, cfg$SEM_INNER_ITER, sem_inner_props, cfg$SEM_N_LABELLINGS,
            cfg$SEM_CHANGE_FACTOR,
            if (!is.null(cfg$RUN_DECODE)) cfg$RUN_DECODE else decode_enabled,
            if (!is.null(cfg$DECODE_ITER)) cfg$DECODE_ITER else "N/A",
            paste(cfg$windowT_post, collapse=" \u2013 "),
            ate_win, cfg$ATE_N_SIMS,
            if (!is.null(cfg$spatial_scale_D)) cfg$spatial_scale_D else "N/A",
            if (!is.null(results$kde_info$trigger_range_km)) round(results$kde_info$trigger_range_km, 2) else "N/A",
            grid_str,
            cfg$TEST_MODE))
kable(cdf, align="lc", caption="Analysis configuration.") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)
```

# References

- Ogata, Y. (1988). Statistical models for earthquake occurrences.
  *JASA*, 83(401), 9--27.
- Zhuang, J., Ogata, Y. & Vere-Jones, D. (2002). Stochastic
  declustering of space-time earthquake occurrences. *JASA*, 97(458),
  369--380.
- Helmstetter, A. & Sornette, D. (2003). Importance of direct and
  indirect triggered seismicity in the ETAS model of seismicity.
  *Geophysical Research Letters*, 30(11), 1576.
- Langenbruch, C. & Zoback, M. D. (2016). How will induced seismicity
  in Oklahoma respond to decreased saltwater injection rates? *Science
  Advances*, 2(11), e1601542.

---
*Report generated by PPDisentangle.*
