---
title: "Oklahoma Induced Seismicity: Causal ETAS Analysis"
subtitle: "Four-Way Comparison Using County-Level Tessellation"
author: "PPDisentangle"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 7
    fig-dpi: 150
    self-contained: true
execute:
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: false
library(PPDisentangle)
library(spatstat)
library(sf)
library(tigris)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(fig.align = "center", out.width = "100%")

SCRIPT_DIR <- getwd()
DATA_DIR <- file.path(SCRIPT_DIR,
                       "oklahoma_induced_seismicity_data_regional20150318")
OUT_DIR  <- file.path(SCRIPT_DIR, "output")
if (!file.exists(file.path(OUT_DIR, "oklahoma_results.rds")))
  OUT_DIR <- file.path(dirname(dirname(SCRIPT_DIR)), "output")

results <- readRDS(file.path(OUT_DIR, "oklahoma_results.rds"))
cfg <- results$config
cty <- results$counties
```

# Introduction

On 18 March 2015 the Oklahoma Corporation Commission (OCC) issued
directive **AOI_20150318**, requiring operators within a defined Area of
Interest (AOI) in central Oklahoma to reduce wastewater disposal volumes.
This report quantifies the causal effect of the directive on earthquake
rates and clustering using the **PPDisentangle** framework.

The analysis uses **Oklahoma county boundaries** as the spatial
tessellation (`r cty$n_counties` counties, `r cty$n_treated` treated)
and fits four progressively richer models:

| Label | Model | Labels | Cross-excitation |
|:---:|:---|:---|:---|
| **A** | Naive independent ETAS | Location-based | None |
| **B** | Naive bivariate ETAS | Location-based | Estimated |
| **C** | SEM independent ETAS | SEM-corrected | None |
| **D** | SEM bivariate ETAS | SEM-corrected | Estimated |

Each model produces an **Average Treatment Effect (ATE)** — the expected
per-county change in earthquake counts from applying the directive.

# Study Design

```{r design}
#| label: design
meta <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
```

**Treatment.** OCC directive AOI_20150318: mandatory reduction in
wastewater disposal volumes within the AOI.

**Tessellation.** The 77 Oklahoma counties (US Census Bureau via
`tigris`). Counties whose centroid falls inside the AOI are treated.

**Treated counties** (`r cty$n_treated`):
`r paste(cty$treated_names, collapse = ", ")`.

**Temporal windows.**

- Pre-treatment: `r meta$design$pre_start_utc` to `r meta$design$t_star_utc`
- Post-treatment: `r meta$design$t_star_utc` to `r meta$design$post_end_utc`
  (ends before next directive AOI_20150624)

**Magnitude threshold.** $m_0 =$ `r cfg$ETAS_M0`.

**Event counts.** `r cfg$n_pre` pre-treatment,
`r cfg$n_post` post-treatment
(`r cfg$n_pre + cfg$n_post` total within Oklahoma).


# Data

```{r load-data}
#| label: load-data
ev_all <- fread(file.path(DATA_DIR, "events_all.csv"))
t_star_utc <- as.POSIXct(meta$design$t_star_utc, tz = "UTC")
ev_all[, time_utc := as.POSIXct(time_utc, tz = "UTC")]
ev_all[, t_days := as.numeric(difftime(time_utc, t_star_utc, units = "days"))]
ev_all <- ev_all[mag >= cfg$ETAS_M0]
pp_pre  <- results$pp_data$pp_pre
pp_post <- results$pp_data$pp_post
```

## Magnitude Distribution

```{r fig-mag}
#| label: fig-mag
#| fig-cap: "Magnitude distribution of events above the completeness threshold."
ggplot(ev_all, aes(x = mag)) +
  geom_histogram(binwidth = 0.1, fill = "#4292c6", color = "white",
                 boundary = cfg$ETAS_M0) +
  geom_vline(xintercept = cfg$ETAS_M0, linetype = "dashed", color = "red") +
  annotate("text", x = cfg$ETAS_M0 + 0.15, y = Inf, vjust = 2,
           label = paste0("m0 = ", cfg$ETAS_M0), color = "red", size = 4) +
  labs(x = "Magnitude", y = "Count") +
  theme_minimal(base_size = 13)
```

## Temporal Profile

```{r fig-temporal}
#| label: fig-temporal
#| fig-cap: "Cumulative event count; dashed line marks the OCC directive."
ev_all[, cum_n := seq_len(.N)]
ggplot(ev_all, aes(x = t_days, y = cum_n)) +
  geom_step(color = "#2171b5", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("text", x = 5, y = max(ev_all$cum_n) * 0.1,
           label = "OCC directive\n(18 Mar 2015)", color = "red", hjust = 0, size = 3.5) +
  labs(x = "Days since treatment", y = "Cumulative events") +
  theme_minimal(base_size = 13)
```

## County Partition

```{r fig-partition}
#| label: fig-partition
#| fig-cap: "Oklahoma county-level partition. Red counties are treated (centroid inside AOI)."
options(tigris_use_cache = TRUE)
counties_sf <- counties(state = "OK", cb = TRUE, year = 2022)
counties_sf <- st_transform(counties_sf, 5070)
counties_sf <- st_make_valid(counties_sf)

aoi_sf <- st_read(file.path(DATA_DIR, "occ_aoi_layer_2.geojson"), quiet = TRUE)
aoi_sf <- st_transform(aoi_sf, 5070) |> st_make_valid()
aoi_union <- st_union(aoi_sf)

county_centroids <- st_centroid(counties_sf)
inside_aoi <- lengths(st_within(county_centroids, aoi_union)) > 0
counties_sf$treatment <- ifelse(inside_aoi, "treated", "control")

ggplot() +
  geom_sf(data = counties_sf, aes(fill = treatment),
          color = "grey40", linewidth = 0.4) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    name = "Assignment") +
  geom_sf(data = aoi_sf, fill = NA, color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Oklahoma County Partition",
       subtitle = paste0("Treated: ", paste(cty$treated_names, collapse = ", "))) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 9))
```

## Point Patterns

```{r fig-pp}
#| label: fig-pp
#| fig-cap: "Pre-treatment (left) and post-treatment events coloured by location label (right)."
#| fig-height: 5
p1 <- ggplot(pp_pre, aes(x = x, y = y, alpha = t)) +
  geom_point(color = "#377eb8", fill = "#377eb8", size = 1.2, shape = 21) +
  scale_alpha_continuous(range = c(0.15, 0.8), guide = "none") +
  labs(title = "Pre-treatment", x = "X (km)", y = "Y (km)") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

pp_post_plot <- pp_post
pp_post_plot$Process <- factor(pp_post_plot$location_process, levels = c("control","treated"))
p2 <- ggplot(pp_post_plot, aes(x = x, y = y, fill = Process, alpha = t)) +
  geom_point(size = 1.2, shape = 21, stroke = 0.2) +
  scale_fill_manual(values = c(control = "#377eb8", treated = "#e41a1c")) +
  scale_alpha_continuous(range = c(0.3, 1), guide = "none") +
  labs(title = "Post-treatment", x = "X (km)", y = "Y (km)") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

# Model

## ETAS Conditional Intensity

$$
\lambda(t,x,y) = \frac{\mu}{|S|} + \sum_{j:\,t_j<t} A\,e^{\alpha_m(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

with Omori–Utsu temporal kernel $g$ and power-law spatial kernel $f$.

## Bivariate Extension

$$
\lambda_k(t,x,y) = \frac{\mu_k}{|S_k|} + \sum_{l\in\{0,1\}}\sum_{\substack{j:\,t_j<t\\Z_j=l}} A_{kl}\,e^{\alpha_{m,kl}(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

Off-diagonal terms $A_{01}$ (treated→control) and $A_{10}$
(control→treated) capture cross-excitation.

## Fixed Parameters

```{r tbl-fixed}
#| label: tbl-fixed
fixed_df <- data.frame(
  Parameter = c("c", "p", "D", "γ", "q"),
  Value = c(0.05, 1.2, 5.0, 0.5, 1.5),
  Role = c("Omori onset delay", "Omori decay exponent",
           "Spatial scale", "Mag-spatial scaling", "Spatial decay exponent"))
kable(fixed_df, align = "lcl",
      caption = "Fixed structural parameters.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## SEM Configuration

- **Outer iterations:** `r cfg$SEM_N_ITER`
- **Inner iterations:** `r cfg$SEM_INNER_ITER`
- **Labelling proposals:** `r cfg$SEM_N_LABELLINGS`
- **Change factor:** `r cfg$SEM_CHANGE_FACTOR`

# Results

```{r extract}
#| label: extract
A_ctrl  <- unlist(results$fitA$ctrl);  A_treat <- unlist(results$fitA$treat)
B_par   <- results$fitB$params
C_ctrl  <- unlist(results$fitC$ctrl);  C_treat <- unlist(results$fitC$treat)
D_par   <- results$fitD$params

bc <- function(par, p) if (p=="mu") par[["mu_0"]] else par[[paste0(p,"_00")]]
bt <- function(par, p) if (p=="mu") par[["mu_1"]] else par[[paste0(p,"_11")]]
```

## Marginal Parameters

```{r tbl-params}
#| label: tbl-params
pr <- c("mu","A","alpha_m")
pl <- c("$\\mu$","$A$","$\\alpha_m$")
tbl <- data.frame(
  Parameter = pl,
  A_c = sapply(pr, \(p) A_ctrl[[p]]),  A_t = sapply(pr, \(p) A_treat[[p]]),
  B_c = sapply(pr, \(p) bc(B_par,p)),  B_t = sapply(pr, \(p) bt(B_par,p)),
  C_c = sapply(pr, \(p) C_ctrl[[p]]),  C_t = sapply(pr, \(p) C_treat[[p]]),
  D_c = sapply(pr, \(p) bc(D_par,p)),  D_t = sapply(pr, \(p) bt(D_par,p)))
rownames(tbl) <- NULL
kable(tbl, digits = 4, escape = FALSE, align = "lcccccccc",
      col.names = c("","A:ctrl","A:treat","B:ctrl","B:treat",
                     "C:ctrl","C:treat","D:ctrl","D:treat"),
      caption = "Marginal ETAS parameters across four fits.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = TRUE) |>
  add_header_above(c(" "=1,"Naive Indep"=2,"Naive Biv"=2,"SEM Indep"=2,"SEM Biv"=2))
```

## Cross-excitation

```{r tbl-cross}
#| label: tbl-cross
cn <- c("A_01","alpha_m_01","A_10","alpha_m_10")
cl <- c("$A_{01}$ (treat$\\to$ctrl)","$\\alpha_{m,01}$",
        "$A_{10}$ (ctrl$\\to$treat)","$\\alpha_{m,10}$")
cross_tbl <- data.frame(Parameter = cl,
  B = sapply(cn, \(nm) B_par[[nm]]),
  D = sapply(cn, \(nm) D_par[[nm]]))
rownames(cross_tbl) <- NULL
kable(cross_tbl, digits = 4, escape = FALSE, align = "lcc",
      col.names = c("","B: Naive Biv","D: SEM Biv"),
      caption = "Cross-excitation parameters.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Parameter Comparison

```{r fig-params}
#| label: fig-params
#| fig-cap: "Marginal parameter estimates across all four fits."
#| fig-height: 5
param_long <- rbind(
  data.frame(Fit="A: Naive Indep",Process="Control",mu=A_ctrl[["mu"]],A=A_ctrl[["A"]],alpha_m=A_ctrl[["alpha_m"]]),
  data.frame(Fit="A: Naive Indep",Process="Treated",mu=A_treat[["mu"]],A=A_treat[["A"]],alpha_m=A_treat[["alpha_m"]]),
  data.frame(Fit="B: Naive Biv",Process="Control",mu=bc(B_par,"mu"),A=bc(B_par,"A"),alpha_m=bc(B_par,"alpha_m")),
  data.frame(Fit="B: Naive Biv",Process="Treated",mu=bt(B_par,"mu"),A=bt(B_par,"A"),alpha_m=bt(B_par,"alpha_m")),
  data.frame(Fit="C: SEM Indep",Process="Control",mu=C_ctrl[["mu"]],A=C_ctrl[["A"]],alpha_m=C_ctrl[["alpha_m"]]),
  data.frame(Fit="C: SEM Indep",Process="Treated",mu=C_treat[["mu"]],A=C_treat[["A"]],alpha_m=C_treat[["alpha_m"]]),
  data.frame(Fit="D: SEM Biv",Process="Control",mu=bc(D_par,"mu"),A=bc(D_par,"A"),alpha_m=bc(D_par,"alpha_m")),
  data.frame(Fit="D: SEM Biv",Process="Treated",mu=bt(D_par,"mu"),A=bt(D_par,"A"),alpha_m=bt(D_par,"alpha_m")))
pm <- pivot_longer(param_long, cols=c(mu,A,alpha_m), names_to="Parameter", values_to="Value")
pm$Parameter <- factor(pm$Parameter, levels=c("mu","A","alpha_m"), labels=c("mu","A","alpha[m]"))
pm$Fit <- factor(pm$Fit, levels=c("A: Naive Indep","B: Naive Biv","C: SEM Indep","D: SEM Biv"))
ggplot(pm, aes(x=Fit, y=Value, colour=Process)) +
  geom_point(size=3, position=position_dodge(0.4)) +
  facet_wrap(~Parameter, scales="free_y", labeller=label_parsed) +
  scale_colour_manual(values=c(Control="#377eb8",Treated="#e41a1c")) +
  labs(y="Estimate", x=NULL) +
  theme_minimal(base_size=13) +
  theme(axis.text.x=element_text(angle=30,hjust=1), strip.text=element_text(face="bold"), legend.position="top")
```

## SEM Convergence

```{r fig-flips}
#| label: fig-flips
#| fig-cap: "Label flips per SEM iteration."
#| fig-height: 4
plots <- list()
for (nm in list(list(res=results$fitC$sem,lab="C: SEM Indep"),
                list(res=results$fitD$sem,lab="D: SEM Biv"))) {
  if (!is.null(nm$res)) plots[[length(plots)+1]] <- plot_flips(nm$res) + ggtitle(nm$lab)
}
if (length(plots)==2) {
  gridExtra::grid.arrange(plots[[1]],plots[[2]],ncol=2)
} else if (length(plots)==1) {
  print(plots[[1]])
}
```

## SEM Relabelling: Confusion Matrices

The SEM corrects the naive location-based labels by relabelling events
whose clustering signature is more consistent with the other process.
The tables below compare the **location-based label** (which county the
event physically occurred in) against the **SEM-inferred label** (the
process the SEM assigned). Off-diagonal entries represent events the SEM
believes were triggered by parents in the other region — i.e. spatial
spillover.

```{r tbl-confusion}
#| label: tbl-confusion
build_cm <- function(sem_obj, label) {
  if (is.null(sem_obj)) return(NULL)
  lab <- sem_obj$adaptive$adaptive_labelling
  post <- lab[lab$t >= 0, ]
  if (is.null(post$location_process) || is.null(post$inferred_process)) return(NULL)
  tab <- table(Location = post$location_process, Inferred = post$inferred_process)
  as.data.frame.matrix(tab)
}

cm_C <- build_cm(results$fitC$sem, "C")
cm_D <- build_cm(results$fitD$sem, "D")
```

### Fit C — SEM Independent ETAS

The independent model can only relabel events from one process to the
other; it has no cross-excitation terms, so relabelling corrects for
events that the model thinks are background from the wrong region.

```{r tbl-cm-C}
#| label: tbl-cm-C
if (!is.null(cm_C)) {
  n_post_C <- sum(cm_C)
  cm_C_pct <- round(cm_C / n_post_C * 100, 1)
  cm_display <- cm_C
  for (rn in rownames(cm_C)) for (cn in colnames(cm_C)) {
    cm_display[rn, cn] <- paste0(cm_C[rn, cn], "  (", cm_C_pct[rn, cn], "%)")
  }
  kable(cm_display,
        caption = paste0("Fit C: Location vs SEM-inferred label (n = ", n_post_C,
                         " post-treatment events). Counts (% of total)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_C)))
}
```

```{r fig-cm-C}
#| label: fig-cm-C
#| fig-cap: "Fit C — Post-treatment events coloured by SEM-inferred label."
#| fig-height: 5
if (!is.null(results$fitC$sem)) {
  lab_C <- results$fitC$sem$adaptive$adaptive_labelling
  post_C <- lab_C[lab_C$t >= 0, ]
  post_C$Relabelled <- post_C$location_process != post_C$inferred_process
  post_C$Label <- paste0(post_C$location_process, " \u2192 ", post_C$inferred_process)
  post_C$Label <- factor(post_C$Label)
  ggplot(post_C, aes(x = x, y = y, colour = inferred_process, shape = Relabelled)) +
    geom_point(size = 1.5, alpha = 0.7) +
    scale_colour_manual(values = c(control = "#377eb8", treated = "#e41a1c"),
                        name = "SEM label") +
    scale_shape_manual(values = c(`FALSE` = 16, `TRUE` = 4),
                       name = "Relabelled?") +
    labs(title = "Fit C: SEM Independent — Inferred Labels",
         x = "X (km)", y = "Y (km)") +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = "right")
}
```

### Fit D — SEM Bivariate ETAS

The bivariate model has cross-excitation, so its relabelling accounts
for the branching structure: an event in a control county that was
likely triggered by a parent in a treated county is relabelled as
"treated."

```{r tbl-cm-D}
#| label: tbl-cm-D
if (!is.null(cm_D)) {
  n_post_D <- sum(cm_D)
  cm_D_pct <- round(cm_D / n_post_D * 100, 1)
  cm_display_D <- cm_D
  for (rn in rownames(cm_D)) for (cn in colnames(cm_D)) {
    cm_display_D[rn, cn] <- paste0(cm_D[rn, cn], "  (", cm_D_pct[rn, cn], "%)")
  }
  kable(cm_display_D,
        caption = paste0("Fit D: Location vs SEM-inferred label (n = ", n_post_D,
                         " post-treatment events). Counts (% of total)."),
        align = "cc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
    add_header_above(c("Location \u2193 / Inferred \u2192" = 1, " " = ncol(cm_D)))
}
```

```{r fig-cm-D}
#| label: fig-cm-D
#| fig-cap: "Fit D — Post-treatment events coloured by SEM-inferred label. Crosses mark relabelled events."
#| fig-height: 5
if (!is.null(results$fitD$sem)) {
  lab_D <- results$fitD$sem$adaptive$adaptive_labelling
  post_D <- lab_D[lab_D$t >= 0, ]
  post_D$Relabelled <- post_D$location_process != post_D$inferred_process
  ggplot(post_D, aes(x = x, y = y, colour = inferred_process, shape = Relabelled)) +
    geom_point(size = 1.5, alpha = 0.7) +
    scale_colour_manual(values = c(control = "#377eb8", treated = "#e41a1c"),
                        name = "SEM label") +
    scale_shape_manual(values = c(`FALSE` = 16, `TRUE` = 4),
                       name = "Relabelled?") +
    labs(title = "Fit D: SEM Bivariate — Inferred Labels",
         x = "X (km)", y = "Y (km)") +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = "right")
}
```

### Comparison

```{r tbl-cm-summary}
#| label: tbl-cm-summary
sum_rows <- list()
for (info in list(list(cm = cm_C, lab = "C: SEM Independent"),
                  list(cm = cm_D, lab = "D: SEM Bivariate"))) {
  cm <- info$cm
  if (is.null(cm)) next
  n_total <- sum(cm)
  ctrl_to_treat <- if ("treated" %in% colnames(cm) && "control" %in% rownames(cm)) cm["control","treated"] else 0
  treat_to_ctrl <- if ("control" %in% colnames(cm) && "treated" %in% rownames(cm)) cm["treated","control"] else 0
  n_flipped <- ctrl_to_treat + treat_to_ctrl
  sum_rows[[length(sum_rows)+1]] <- data.frame(
    Model = info$lab,
    `Post events` = n_total,
    `Ctrl to Treat` = ctrl_to_treat,
    `Treat to Ctrl` = treat_to_ctrl,
    `Total relabelled` = n_flipped,
    `Pct relabelled` = paste0(round(n_flipped / n_total * 100, 1), "%"),
    check.names = FALSE)
}
if (length(sum_rows) > 0) {
  kable(do.call(rbind, sum_rows),
        caption = "Summary of SEM relabelling across models.",
        align = "lccccc") |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
}
```

# Average Treatment Effect

## Branching Ratio and Criticality

The ETAS branching ratio $\eta = A\,\beta_{\text{GR}}/(\beta_{\text{GR}} - \alpha_m)$
gives the expected number of direct offspring per event, integrated over the
Gutenberg–Richter magnitude distribution. When $\eta \geq 1$ the process is
formally supercritical: the infinite-horizon expected count diverges.

This is **not a modelling artefact**. ETAS fits to real seismicity
routinely produce $\eta \geq 1$ (Helmstetter & Sornette, 2003).
The finite observation window, spatial boundaries, and magnitude bounds
prevent actual divergence — the process generates finite realisations and
the simulation-based ATE is well-defined.

Because the closed-form formula
$\mathbb{E}[N] = \mu T/(1-\eta)$ is inapplicable when $\eta \geq 1$,
we report: (i) the **background-rate ATE** ($\Delta\mu\cdot T/n$),
which isolates the treatment effect on immigrant seismicity, and
(ii) the **simulation-based ATE**, which fully accounts for cascade
amplification within the finite observation window.

```{r tbl-branching}
#| label: tbl-branching
ate_rows <- list(
  list(lab = "B: Naive Bivariate",   ate = results$fitB$ate),
  list(lab = "D: SEM Bivariate",     ate = results$fitD$ate))

br_df <- do.call(rbind, lapply(ate_rows, function(r) {
  if (is.null(r$ate) || is.null(r$ate$analytic)) {
    return(data.frame(Model = r$lab, `eta ctrl` = NA, `eta treat` = NA,
                      `BG Rate ATE` = NA, check.names = FALSE))
  }
  a <- r$ate$analytic
  data.frame(Model = r$lab,
    `eta ctrl`    = round(a$eta_ctrl, 3),
    `eta treat`   = round(a$eta_treat, 3),
    `BG Rate ATE` = round(a$ATE_bg, 4),
    check.names = FALSE)
}))
kable(br_df, align = "lccc",
      col.names = c("Model", "$\\eta_0$ (ctrl)", "$\\eta_1$ (treat)", "BG Rate ATE"),
      escape = FALSE,
      caption = "Branching ratios and background-rate ATE for bivariate models.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Naive Bivariate vs SEM Bivariate: 1-Year ATE

The central comparison is between **Fit B** (naive location-based labels)
and **Fit D** (SEM-corrected labels), both using the bivariate ETAS model
with cross-excitation. The ATE is estimated by simulating complete
realisations directly over a **1-year (365-day) horizon** under the
fitted parameters.

```{r tbl-ate-biv}
#| label: tbl-ate-biv
n_cty <- cfg$n_counties
ate_days <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 365

biv_rows <- list(
  list(lab = "B: Naive Bivariate", ate = results$fitB$ate, louis = NULL),
  list(lab = "D: SEM Bivariate",   ate = results$fitD$ate, louis = results$fitD$louis))

ate_biv <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  sim_ate <- mean(r$ate$all_nothing_sim$ATE, na.rm = TRUE)
  sim_sd  <- sd(r$ate$all_nothing_sim$ATE, na.rm = TRUE)
  total   <- sim_ate * n_cty
  data.frame(
    Model = r$lab,
    `ATE/county` = round(sim_ate, 1),
    `SD` = round(sim_sd, 1),
    `Total` = round(total, 0),
    Spillover = round(r$ate$ATE_spillover, 2),
    check.names = FALSE)
}))
kable(ate_biv,
      col.names = c("Model", "ATE / county / yr", "SD", paste0("Total (", n_cty, " counties)"), "Spillover"),
      caption = paste0("1-year ATE: expected earthquakes per county per year (simulated over ", ate_days, " days)."),
      align = "lcccc") |>
  kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE)
```

```{r fig-ate-biv}
#| label: fig-ate-biv
#| fig-cap: "1-year all-or-nothing ATE: Naive Bivariate (B) vs SEM Bivariate (D)."
#| fig-height: 4
biv_sim <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  data.frame(Model = r$lab, ATE = r$ate$all_nothing_sim$ATE)
}))
if (!is.null(biv_sim) && nrow(biv_sim) > 0) {
  biv_means <- aggregate(ATE ~ Model, biv_sim, mean)
  ggplot(biv_sim, aes(x = ATE, fill = Model)) +
    geom_histogram(bins = 30, alpha = 0.65, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_vline(data = biv_means, aes(xintercept = ATE, colour = Model),
               linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
    scale_fill_manual(values = c("B: Naive Bivariate" = "#fc8d62",
                                 "D: SEM Bivariate" = "#8da0cb")) +
    scale_colour_manual(values = c("B: Naive Bivariate" = "#d95f02",
                                   "D: SEM Bivariate" = "#7570b3")) +
    labs(x = "Per-county ATE (earthquakes / year)", y = "Simulation count",
         title = "1-Year All-or-Nothing ATE: Naive vs SEM Bivariate") +
    theme_minimal(base_size = 13) +
    theme(legend.position = "top",
          plot.title = element_text(hjust = 0.5, face = "bold"))
}
```

```{r tbl-ate-biv-total}
#| label: tbl-ate-biv-total
total_df <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  sim_ate <- mean(r$ate$all_nothing_sim$ATE, na.rm = TRUE)
  data.frame(
    Model = r$lab,
    `Per county per year` = round(sim_ate, 1),
    `All counties per year` = round(sim_ate * n_cty, 0),
    `Interpretation` = ifelse(sim_ate > 0,
      paste0("+", round(abs(sim_ate * n_cty)), " more earthquakes/yr under treatment"),
      paste0(round(abs(sim_ate * n_cty)), " fewer earthquakes/yr under treatment")),
    check.names = FALSE)
}))
kable(total_df, align = "lccc",
      caption = "1-year treatment effect on earthquake counts.") |>
  kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE) |>
  column_spec(4, bold = TRUE)
```

## All Four Models (Reference)

```{r tbl-ate-all}
#| label: tbl-ate-all
all_rows <- list(
  list(lab = "A: Naive Independent", ate = results$fitA$ate),
  list(lab = "B: Naive Bivariate",   ate = results$fitB$ate),
  list(lab = "C: SEM Independent",   ate = results$fitC$ate),
  list(lab = "D: SEM Bivariate",     ate = results$fitD$ate))

ate_all <- do.call(rbind, lapply(all_rows, function(r) {
  if (is.null(r$ate)) return(data.frame(Model=r$lab, `ATE/county/yr`=NA, `Total/yr`=NA, Spillover=NA, check.names=FALSE))
  sim_ate <- mean(r$ate$all_nothing_sim$ATE, na.rm=TRUE)
  data.frame(
    Model = r$lab,
    `ATE/county/yr` = round(sim_ate, 1),
    `Total/yr` = round(sim_ate * n_cty, 0),
    Spillover = round(r$ate$ATE_spillover, 2),
    check.names = FALSE)
}))
kable(ate_all, col.names = c("Model","ATE / county / yr", paste0("Total / ", n_cty, " counties / yr"), "Spillover"),
      caption = "1-year ATE across all four models.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Louis-Method Standard Errors (SEM Fits)

For the SEM-based fits, uncertainty in the ATE arises from two
sources: (1) parameter estimation error, and (2) label uncertainty —
different labellings assign different events to each process.
The **Louis (1982) identity** decomposes the observed information as

$$
I_{\text{obs}} = I_{\text{complete}} - I_{\text{missing}}
$$

where $I_{\text{missing}}$ is estimated from the weighted variance of the
per-labelling ATE across the importance-sampled proposals.

```{r tbl-louis}
#| label: tbl-louis
louis_rows <- list()
for (info in list(list(louis = results$fitC$louis, lab = "C: SEM Independent"),
                  list(louis = results$fitD$louis, lab = "D: SEM Bivariate"))) {
  l <- info$louis
  if (is.null(l)) {
    louis_rows[[length(louis_rows)+1]] <- data.frame(
      Model = info$lab, ATE = NA, SE = NA, `CI lower` = NA, `CI upper` = NA,
      ESS = NA, check.names = FALSE)
  } else {
    louis_rows[[length(louis_rows)+1]] <- data.frame(
      Model = info$lab,
      ATE = round(l$ate_mean, 2),
      SE = round(l$ate_se, 4),
      `CI lower` = round(l$ci_lower, 2),
      `CI upper` = round(l$ci_upper, 2),
      ESS = round(l$ess, 1),
      check.names = FALSE)
  }
}
kable(do.call(rbind, louis_rows),
      col.names = c("Model","ATE","SE","95% CI lower","95% CI upper","ESS"),
      caption = "Louis-method standard errors for SEM ATE estimates.",
      align = "lccccc") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

The small SEs reflect that label uncertainty affects only a handful of
events across proposals (the SEM converges tightly). The ESS of ~3
indicates weight concentration on a few dominant labellings — a common
feature of importance sampling in high-dimensional label spaces.

# Interpretation

```{r summaries, include=FALSE}
n_cty_interp <- cfg$n_counties

mu_ratio_D <- bt(D_par,"mu") / bc(D_par,"mu")
mu_ratio_B <- bt(B_par,"mu") / bc(B_par,"mu")

ate_B_1yr <- if (!is.null(results$fitB$ate)) round(mean(results$fitB$ate$all_nothing_sim$ATE, na.rm=TRUE), 1) else NA
ate_D_1yr <- if (!is.null(results$fitD$ate)) round(mean(results$fitD$ate$all_nothing_sim$ATE, na.rm=TRUE), 1) else NA
total_B_1yr <- round(ate_B_1yr * n_cty_interp)
total_D_1yr <- round(ate_D_1yr * n_cty_interp)

louis_D_se <- if (!is.null(results$fitD$louis)) round(results$fitD$louis$ate_se, 2) else NA
louis_D_ci <- if (!is.null(results$fitD$louis)) paste0("[", round(results$fitD$louis$ci_lower, 1), ", ", round(results$fitD$louis$ci_upper, 1), "]") else NA
```

The core comparison is **Fit B** (naive bivariate) vs **Fit D** (SEM
bivariate), both of which model cross-excitation between regions. The
ATE is simulated directly over a **1-year (365-day) horizon**.

**Background rate.** In Fit B (naive labels), the treated-region
background rate $\mu_1$ is `r round(mu_ratio_B, 1)`$\times$ higher
than $\mu_0$. After SEM relabelling (Fit D), this ratio becomes
`r round(mu_ratio_D, 1)`$\times$ — the SEM shifts some events out of
the treated label, partially redistributing background intensity.

**Cross-excitation.** Both bivariate models find $A_{01} > A_{10}$:
treated-region earthquakes trigger aftershocks in control counties more
than vice versa. The SEM changes the estimated cross-excitation
slightly ($A_{01}$: `r round(B_par[["A_01"]],4)` →
`r round(D_par[["A_01"]],4)`).

**1-Year ATE.**

- **Fit B (naive):** `r ate_B_1yr` events per county per year
  (**`r total_B_1yr` total** across `r n_cty_interp` counties)
- **Fit D (SEM):** `r ate_D_1yr` events per county per year
  (**`r total_D_1yr` total** across `r n_cty_interp` counties)

The naive bivariate model (B) estimates that treatment *increases*
seismicity — an artefact of mislabelling spillover events. The SEM
bivariate model (D) corrects for this and finds a near-zero or
slightly negative ATE, consistent with the directive having a modest
protective effect that is largely offset by cross-regional triggering.

**Branching ratio.** All fitted ETAS models have
$\eta = A\beta_{\text{GR}}/(\beta_{\text{GR}} - \alpha_m) \geq 1$. This
is typical for ETAS fits to real seismicity (Helmstetter & Sornette, 2003)
and does not indicate model failure — the finite observation window and
spatial boundaries bound the actual event count. It means the closed-form
$\mu T/(1-\eta)$ formula is inapplicable, and simulation is required.

**ATE uncertainty (Louis method).** The SEM bivariate ATE has a
Louis-method SE of **`r louis_D_se`** (95% CI: **`r louis_D_ci`**
events/county/year), reflecting tight convergence of the SEM labelling
step.

# Configuration

```{r tbl-config}
#| label: tbl-config
ate_win <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else "N/A"
cdf <- data.frame(
  Setting = c("Magnitude threshold","GR β","Counties (total / treated)",
              "SEM outer iters","SEM inner iters","SEM proposals",
              "SEM change factor","Fitting window (days)",
              "ATE simulation window (days)","ATE simulations","Test mode"),
  Value = c(cfg$ETAS_M0, cfg$BETA_GR,
            paste0(cfg$n_counties," / ",cfg$n_treated),
            cfg$SEM_N_ITER, cfg$SEM_INNER_ITER, cfg$SEM_N_LABELLINGS,
            cfg$SEM_CHANGE_FACTOR,
            paste(cfg$windowT_post, collapse=" – "),
            ate_win, cfg$ATE_N_SIMS, cfg$TEST_MODE))
kable(cdf, align="lc", caption="Analysis configuration.") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)
```

# References

- Ogata, Y. (1988). Statistical models for earthquake occurrences.
  *JASA*, 83(401), 9–27.
- Zhuang, J., Ogata, Y. & Vere-Jones, D. (2002). Stochastic
  declustering of space-time earthquake occurrences. *JASA*, 97(458),
  369–380.
- Helmstetter, A. & Sornette, D. (2003). Importance of direct and
  indirect triggered seismicity in the ETAS model of seismicity.
  *Geophysical Research Letters*, 30(11), 1576.
- Louis, T. A. (1982). Finding the observed information matrix when
  using the EM algorithm. *JRSS-B*, 44(2), 226–233.
- Langenbruch, C. & Zoback, M. D. (2016). How will induced seismicity
  in Oklahoma respond to decreased saltwater injection rates? *Science
  Advances*, 2(11), e1601542.

---
*Report generated by PPDisentangle.*
