---
title: "Oklahoma Induced Seismicity: Causal ETAS Analysis"
subtitle: "Four-Way Comparison of Independent and Bivariate ETAS Under Naive and SEM Labelling"
author: "PPDisentangle"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: hide
    theme: cosmo
    self_contained: true
    fig_width: 10
    fig_height: 7
---

```{r setup, include=FALSE}
library(PPDisentangle)
library(spatstat)
library(sf)
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(tidyr)

knitr::opts_chunk$set(
  fig.align = "center",
  out.width = "100%",
  warning = FALSE,
  message = FALSE,
  fig.retina = 2
)

SCRIPT_DIR <- getwd()
DATA_DIR <- file.path(SCRIPT_DIR,
                       "oklahoma_induced_seismicity_data_regional20150318")
OUT_DIR  <- file.path(SCRIPT_DIR, "output")
if (!file.exists(file.path(OUT_DIR, "oklahoma_results.rds"))) {
  OUT_DIR <- file.path(dirname(dirname(SCRIPT_DIR)), "output")
}

results <- readRDS(file.path(OUT_DIR, "oklahoma_results.rds"))
cfg <- results$config
```

# Introduction

On 18 March 2015 the Oklahoma Corporation Commission (OCC) issued
directive **AOI_20150318**, requiring operators within a defined Area of
Interest in central Oklahoma to reduce wastewater disposal volumes.  The
directive targeted the most seismically active zone linked to
injection-induced seismicity.  This report quantifies the causal effect
of the directive on earthquake rates and clustering using the
**PPDisentangle** framework for causal inference with point processes.

The analysis fits four nested models to the same data, progressively
relaxing assumptions about spatial labelling and cross-process
interaction:

| Label | Model | Labels | Cross-excitation |
|:---:|:---|:---|:---|
| **A** | Naive independent ETAS | Location-based | None |
| **B** | Naive bivariate ETAS | Location-based | Estimated |
| **C** | SEM independent ETAS | SEM-corrected | None |
| **D** | SEM bivariate ETAS | SEM-corrected | Estimated |


# Study Design

```{r design-params}
meta <- jsonlite::fromJSON(readLines(
  file.path(DATA_DIR, "metadata.json")))
```

**Treatment.** The OCC directive required significant reductions in
wastewater disposal volumes within the designated AOI in central
Oklahoma.

**Outcome.** Earthquake rate and spatio-temporal clustering, measured via
the Epidemic-Type Aftershock Sequence (ETAS) model.

**Spatial partition.** A `r meta$design$buffer_km` km buffer around the
AOI is tessellated into `r meta$design$grid_cellsize_km` km square grid
cells.  Cells whose centroid falls inside the AOI are designated
*treated*; all others are *control*.  In total there are
`r meta$counts$cells` cells, of which `r meta$counts$treated_cells` are
treated.

**Temporal windows.**

- Pre-treatment: `r meta$design$pre_start_utc` to
  `r meta$design$t_star_utc`
- Post-treatment: `r meta$design$t_star_utc` to
  `r meta$design$post_end_utc`

The post window ends before the next regional directive (AOI_20150624)
takes effect, avoiding contamination from subsequent interventions.

**Magnitude threshold.** $m_0$ = `r cfg$ETAS_M0` (completeness
threshold for the USGS catalog in this region).

**Event counts.** After filtering to $m \geq m_0$:
`r cfg$n_pre` pre-treatment events and `r cfg$n_post` post-treatment
events (`r cfg$n_pre + cfg$n_post` total).


# Data

```{r load-data}
ev_all <- fread(file.path(DATA_DIR, "events_all.csv"))
cells  <- fread(file.path(DATA_DIR, "cells.csv"))

t_star_utc <- as.POSIXct(meta$design$t_star_utc, tz = "UTC")
ev_all[, time_utc := as.POSIXct(time_utc, tz = "UTC")]
ev_all[, t_days := as.numeric(difftime(time_utc, t_star_utc, units = "days"))]
ev_all[, x_km := x_m / 1000]
ev_all[, y_km := y_m / 1000]
ev_all <- ev_all[mag >= cfg$ETAS_M0]

pp_pre  <- results$pp_data$pp_pre
pp_post <- results$pp_data$pp_post
```

## Magnitude Distribution

```{r fig-mag-hist, fig.cap="Magnitude distribution of all events above the completeness threshold."}
ggplot(ev_all, aes(x = mag)) +
  geom_histogram(binwidth = 0.1, fill = "#4292c6", color = "white",
                 boundary = cfg$ETAS_M0) +
  geom_vline(xintercept = cfg$ETAS_M0, linetype = "dashed", color = "red") +
  annotate("text", x = cfg$ETAS_M0 + 0.15, y = Inf, vjust = 2,
           label = paste0("m0 = ", cfg$ETAS_M0), color = "red", size = 4) +
  labs(x = "Magnitude", y = "Count",
       title = "Earthquake Magnitude Distribution") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

## Temporal Profile

```{r fig-temporal, fig.cap="Cumulative event count over time; vertical line marks the OCC directive."}
ev_all[, cum_n := seq_len(.N)]
ggplot(ev_all, aes(x = t_days, y = cum_n)) +
  geom_step(color = "#2171b5", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red",
             linewidth = 0.8) +
  annotate("text", x = 5, y = max(ev_all$cum_n) * 0.1,
           label = "OCC directive\n(18 Mar 2015)", color = "red",
           hjust = 0, size = 3.5) +
  labs(x = "Days since treatment", y = "Cumulative events",
       title = "Cumulative Earthquake Count") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

## Spatial Partition

```{r partition-setup}
grid_sf <- st_read(file.path(DATA_DIR, "grid_cells.gpkg"),
                   layer = "cells", quiet = TRUE)
grid_sf <- st_transform(grid_sf, 5070)
grid_sf <- st_make_valid(grid_sf)

bb <- st_bbox(st_union(grid_sf))
win_km <- owin(xrange = c(bb["xmin"], bb["xmax"]) / 1000,
               yrange = c(bb["ymin"], bb["ymax"]) / 1000)
partition <- quadrats(win_km,
  nx = round(diff(win_km$xrange) / 20),
  ny = round(diff(win_km$yrange) / 20))

partition_processes <- rep("control", partition$n)
for (i in seq_len(nrow(cells))) {
  if (cells$Z[i] == 1L) {
    cx <- cells$centroid_x_m[i] / 1000
    cy <- cells$centroid_y_m[i] / 1000
    ti <- as.integer(tileindex(cx, cy, partition))
    if (!is.na(ti) && ti >= 1 && ti <= partition$n)
      partition_processes[ti] <- "treated"
  }
}
```

```{r fig-partition, fig.cap="Spatial partition with treatment assignment. Points show post-treatment earthquakes."}
tess_sf <- tess_to_sf(partition)
tess_sf$treatment <- partition_processes

ggplot() +
  geom_sf(data = tess_sf, aes(fill = treatment),
          color = "grey50", linewidth = 0.3) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    name = "Assignment") +
  geom_point(data = pp_post, aes(x = x, y = y),
             size = 0.5, alpha = 0.5) +
  labs(title = "Oklahoma Spatial Partition",
       subtitle = "OCC directive AOI_20150318 (treated = red)",
       x = "X (km)", y = "Y (km)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))
```


# Point Pattern Plots

## Pre-treatment

```{r fig-pp-pre, fig.cap="Pre-treatment earthquake locations. Opacity encodes time (more opaque = more recent)."}
ggplot(pp_pre, aes(x = x, y = y, alpha = t)) +
  geom_point(color = "#377eb8", size = 1.5, shape = 21,
             fill = "#377eb8") +
  scale_alpha_continuous(name = "Time (days)", range = c(0.15, 0.8)) +
  labs(title = "Pre-treatment Events",
       x = "X (km)", y = "Y (km)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## Post-treatment (Location Labels)

```{r fig-pp-post-loc, fig.cap="Post-treatment events coloured by location-based process label."}
pp_post_plot <- pp_post
pp_post_plot$Process <- factor(pp_post_plot$location_process,
                               levels = c("control", "treated"))
ggplot(pp_post_plot, aes(x = x, y = y, fill = Process, alpha = t)) +
  geom_point(size = 1.8, shape = 21, stroke = 0.3) +
  scale_fill_manual(values = c(control = "#377eb8", treated = "#e41a1c")) +
  scale_alpha_continuous(name = "Time (days)", range = c(0.3, 1)) +
  labs(title = "Post-treatment Events (Location Labels)",
       x = "X (km)", y = "Y (km)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## Post-treatment (SEM-inferred Labels)

```{r fig-pp-post-sem, fig.cap="Post-treatment events coloured by SEM bivariate-inferred process label (Fit D). Events whose label changed from the location assignment are candidates for cross-regional triggering."}
sem_lab <- results$fitD$sem$adaptive$adaptive_labelling
if (!is.null(sem_lab)) {
  sem_post <- sem_lab[sem_lab$t >= 0, ]
  sem_post$Process <- factor(sem_post$inferred_process,
                             levels = c("control", "treated"))

  print(
    ggplot(sem_post, aes(x = x, y = y, fill = Process, alpha = t)) +
      geom_point(size = 1.8, shape = 21, stroke = 0.3) +
      scale_fill_manual(values = c(control = "#377eb8",
                                   treated = "#e41a1c")) +
      scale_alpha_continuous(name = "Time (days)", range = c(0.3, 1)) +
      labs(title = "Post-treatment Events (SEM Bivariate Labels)",
           x = "X (km)", y = "Y (km)") +
      theme_minimal(base_size = 13) +
      theme(plot.title = element_text(hjust = 0.5, face = "bold"))
  )

  n_flipped <- sum(sem_post$inferred_process != sem_post$location_process)
  pct_flipped <- round(100 * n_flipped / nrow(sem_post), 1)
} else {
  n_flipped <- NA
  pct_flipped <- NA
  cat("SEM bivariate result not available.")
}
```

`r if (!is.na(n_flipped)) paste0("The SEM relabelled **", n_flipped, "** of ", nrow(sem_post), " post-treatment events (", pct_flipped, "%) relative to the location-based assignment.")`


# Model

## ETAS Specification

The spatio-temporal ETAS conditional intensity for a single process is:

$$
\lambda(t, x, y) = \frac{\mu}{|S|} + \sum_{j:\, t_j < t} \kappa(m_j)\, g(t - t_j)\, f(x - x_j,\, y - y_j \mid m_j)
$$

where:

- $\mu / |S|$ is the homogeneous background rate
- $\kappa(m) = A \exp\bigl(\alpha_m (m - m_0)\bigr)$ is the magnitude-dependent productivity
- $g(\Delta t) = \frac{p-1}{c}\left(1 + \frac{\Delta t}{c}\right)^{-p}$ is the Omori--Utsu temporal kernel
- $f(\Delta x, \Delta y \mid m) \propto \bigl(1 + r^2 / (D e^{\gamma(m-m_0)})\bigr)^{-q}$ is the spatial kernel with $r^2 = \Delta x^2 + \Delta y^2$

## Bivariate Extension

The bivariate ETAS model introduces a $2 \times 2$ triggering structure.
For process $k \in \{0\text{ (ctrl)},\, 1\text{ (treat)}\}$:

$$
\lambda_k(t, x, y) = \frac{\mu_k}{|S_k|} + \sum_{l \in \{0,1\}} \sum_{\substack{j:\, t_j < t \\ Z_j = l}} A_{kl}\, e^{\alpha_{m,kl}(m_j - m_0)}\, g(t - t_j)\, f(x - x_j, y - y_j \mid m_j)
$$

The diagonal terms $(A_{00}, A_{11})$ capture self-excitation within each
process; the off-diagonal terms $(A_{01}, A_{10})$ capture
**cross-excitation** — an earthquake in the treated region triggering
aftershocks in the control region, and vice versa.  Setting
$A_{01} = A_{10} = 0$ recovers two independent processes.

The 15 bivariate parameters are:

| Parameters | Role |
|:---|:---|
| $\mu_0,\, \mu_1$ | Background rates |
| $A_{00},\, \alpha_{m,00}$ | Control self-excitation |
| $A_{11},\, \alpha_{m,11}$ | Treated self-excitation |
| $A_{01},\, \alpha_{m,01}$ | Treated $\to$ control spillover |
| $A_{10},\, \alpha_{m,10}$ | Control $\to$ treated spillover |
| $c,\, p,\, D,\, \gamma,\, q$ | Shared structural (fixed) |

## Estimation Strategy

**Naive fits (A, B)** use maximum likelihood on events assigned to
processes by their spatial location (events in treated tiles are
"treated," all others "control").  This ignores that an aftershock
triggered by a treated-region parent may land in the control region.

**SEM fits (C, D)** use the Stochastic Expectation-Maximisation
algorithm to jointly infer latent process labels and model parameters:

1. **Inner loop** (adaptive labelling): proposes relabellings of
   post-treatment events using simulated discrepancy, accepting changes
   that improve the likelihood.
2. **Outer loop**: re-estimates parameters using importance-weighted
   proposals from the current label distribution.

The SEM runs for `r cfg$SEM_N_ITER` outer iterations with
`r cfg$SEM_N_LABELLINGS` labelling proposals per iteration.

## Fixed Structural Parameters

The temporal and spatial decay parameters are held fixed across all four
fits to reduce dimensionality and improve identifiability:

```{r tbl-fixed}
fixed_df <- data.frame(
  Parameter = c("c", "p", "D", "γ", "q"),
  Value = c(0.05, 1.2, 5.0, 0.5, 1.5),
  Role = c("Omori-Utsu onset delay",
           "Omori-Utsu decay exponent",
           "Spatial decay scale",
           "Magnitude-spatial scaling",
           "Spatial decay exponent")
)
kable(fixed_df, align = "lcl",
      caption = "Fixed structural parameters shared by all models.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE)
```


# Results

```{r extract-params}
A_ctrl  <- unlist(results$fitA$ctrl)
A_treat <- unlist(results$fitA$treat)
B_par   <- results$fitB$params
C_ctrl  <- unlist(results$fitC$ctrl)
C_treat <- unlist(results$fitC$treat)
D_par   <- results$fitD$params

biv_ctrl <- function(par, p) {
  if (p == "mu") par[["mu_0"]] else par[[paste0(p, "_00")]]
}
biv_treat <- function(par, p) {
  if (p == "mu") par[["mu_1"]] else par[[paste0(p, "_11")]]
}
```

## Marginal Parameters

```{r tbl-params}
param_rows <- c("mu", "A", "alpha_m")
param_labels <- c("$\\mu$", "$A$", "$\\alpha_m$")

tbl <- data.frame(
  Parameter = param_labels,
  A_ctrl  = sapply(param_rows, function(p) A_ctrl[[p]]),
  A_treat = sapply(param_rows, function(p) A_treat[[p]]),
  B_ctrl  = sapply(param_rows, function(p) biv_ctrl(B_par, p)),
  B_treat = sapply(param_rows, function(p) biv_treat(B_par, p)),
  C_ctrl  = sapply(param_rows, function(p) C_ctrl[[p]]),
  C_treat = sapply(param_rows, function(p) C_treat[[p]]),
  D_ctrl  = sapply(param_rows, function(p) biv_ctrl(D_par, p)),
  D_treat = sapply(param_rows, function(p) biv_treat(D_par, p))
)
rownames(tbl) <- NULL

kable(tbl, digits = 4, escape = FALSE, align = "lcccccccc",
      col.names = c("", "A: ctrl", "A: treat",
                     "B: ctrl", "B: treat",
                     "C: ctrl", "C: treat",
                     "D: ctrl", "D: treat"),
      caption = "Estimated marginal ETAS parameters across the four fits.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = TRUE) %>%
  add_header_above(c(" " = 1,
                      "Naive Independent" = 2,
                      "Naive Bivariate" = 2,
                      "SEM Independent" = 2,
                      "SEM Bivariate" = 2))
```

## Cross-excitation Parameters

```{r tbl-cross}
cross_names <- c("A_01", "alpha_m_01", "A_10", "alpha_m_10")
cross_labels <- c("$A_{01}$ (treat$\\to$ctrl)", "$\\alpha_{m,01}$",
                   "$A_{10}$ (ctrl$\\to$treat)", "$\\alpha_{m,10}$")

cross_tbl <- data.frame(
  Parameter = cross_labels,
  B = sapply(cross_names, function(nm) B_par[[nm]]),
  D = sapply(cross_names, function(nm) D_par[[nm]])
)
rownames(cross_tbl) <- NULL

kable(cross_tbl, digits = 4, escape = FALSE, align = "lcc",
      col.names = c("", "B: Naive Bivariate", "D: SEM Bivariate"),
      caption = "Cross-excitation parameters (zero by construction in independent models A and C).") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE)
```

## Parameter Comparison Plot

```{r fig-param-comparison, fig.cap="Comparison of the three free marginal parameters across all four fits.", fig.height=5}
param_long <- rbind(
  data.frame(Fit = "A: Naive Indep", Process = "Control",
             mu = A_ctrl[["mu"]], A = A_ctrl[["A"]],
             alpha_m = A_ctrl[["alpha_m"]]),
  data.frame(Fit = "A: Naive Indep", Process = "Treated",
             mu = A_treat[["mu"]], A = A_treat[["A"]],
             alpha_m = A_treat[["alpha_m"]]),
  data.frame(Fit = "B: Naive Biv", Process = "Control",
             mu = biv_ctrl(B_par, "mu"), A = biv_ctrl(B_par, "A"),
             alpha_m = biv_ctrl(B_par, "alpha_m")),
  data.frame(Fit = "B: Naive Biv", Process = "Treated",
             mu = biv_treat(B_par, "mu"), A = biv_treat(B_par, "A"),
             alpha_m = biv_treat(B_par, "alpha_m")),
  data.frame(Fit = "C: SEM Indep", Process = "Control",
             mu = C_ctrl[["mu"]], A = C_ctrl[["A"]],
             alpha_m = C_ctrl[["alpha_m"]]),
  data.frame(Fit = "C: SEM Indep", Process = "Treated",
             mu = C_treat[["mu"]], A = C_treat[["A"]],
             alpha_m = C_treat[["alpha_m"]]),
  data.frame(Fit = "D: SEM Biv", Process = "Control",
             mu = biv_ctrl(D_par, "mu"), A = biv_ctrl(D_par, "A"),
             alpha_m = biv_ctrl(D_par, "alpha_m")),
  data.frame(Fit = "D: SEM Biv", Process = "Treated",
             mu = biv_treat(D_par, "mu"), A = biv_treat(D_par, "A"),
             alpha_m = biv_treat(D_par, "alpha_m"))
)

param_melt <- pivot_longer(param_long, cols = c(mu, A, alpha_m),
                           names_to = "Parameter", values_to = "Value")
param_melt$Parameter <- factor(param_melt$Parameter,
  levels = c("mu", "A", "alpha_m"),
  labels = c("mu", "A", "alpha[m]"))
param_melt$Fit <- factor(param_melt$Fit,
  levels = c("A: Naive Indep", "B: Naive Biv",
             "C: SEM Indep", "D: SEM Biv"))

ggplot(param_melt, aes(x = Fit, y = Value, colour = Process)) +
  geom_point(size = 3, position = position_dodge(width = 0.4)) +
  facet_wrap(~ Parameter, scales = "free_y", labeller = label_parsed) +
  scale_colour_manual(values = c(Control = "#377eb8", Treated = "#e41a1c")) +
  labs(y = "Estimate", x = NULL) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        strip.text = element_text(face = "bold"),
        legend.position = "top")
```

## SEM Convergence

```{r fig-sem-flips, fig.cap="Label flips per SEM outer iteration for the independent and bivariate models.", fig.height=4}
plots_made <- list()
for (nm in list(list(res = results$fitC$sem, label = "C: SEM Independent"),
                list(res = results$fitD$sem, label = "D: SEM Bivariate"))) {
  if (!is.null(nm$res)) {
    p <- plot_flips(nm$res) + ggtitle(nm$label)
    plots_made[[length(plots_made) + 1]] <- p
  }
}
if (length(plots_made) == 2) {
  gridExtra::grid.arrange(plots_made[[1]], plots_made[[2]], ncol = 2)
} else if (length(plots_made) == 1) {
  print(plots_made[[1]])
} else {
  cat("No SEM results available for convergence plots.")
}
```


# Interpretation

```{r compute-summaries, include=FALSE}
mu_ratio_A <- A_treat[["mu"]] / A_ctrl[["mu"]]
mu_ratio_D <- biv_treat(D_par, "mu") / biv_ctrl(D_par, "mu")
A_ratio_A  <- A_ctrl[["A"]] / A_treat[["A"]]
A_ratio_D  <- biv_ctrl(D_par, "A") / biv_treat(D_par, "A")
```

## Background Rate

The background rate $\mu$ is the most direct measure of the treatment
effect.  Across all four fits, the treated region has a dramatically
higher background rate than the control region:

- **Fit A** (naive independent): $\mu_{\text{treat}} / \mu_{\text{ctrl}}$ = `r round(mu_ratio_A, 1)`
- **Fit D** (SEM bivariate): $\mu_{\text{treat}} / \mu_{\text{ctrl}}$ = `r round(mu_ratio_D, 1)`

The treated region's background rate is on the order of
**`r round(mu_ratio_D)`$\times$** higher than the control region's,
reflecting the intense injection-driven seismicity concentrated in the
AOI.

The bivariate models (B, D) estimate even lower control-region background
rates than the independent models (A, C), because some of the apparent
control-region background can be explained by cross-excitation from
treated-region parents.

## Self-excitation

The control region exhibits markedly higher self-excitation productivity
($A \approx$ `r round(A_ctrl[["A"]], 2)`) than the treated region
($A \approx$ `r round(A_treat[["A"]], 2)`).  This may seem
counterintuitive: the treated region has more earthquakes, yet lower
clustering.  The explanation is that the treated region's seismicity is
dominated by *background* events (injection-driven), whereas the control
region's smaller population of events is more *clustered* (natural
aftershock sequences).

## Cross-excitation

The bivariate models reveal asymmetric cross-excitation:

- **Treated $\to$ Control** ($A_{01}$): `r round(D_par[["A_01"]], 4)` —
  earthquakes in the treated region trigger a small but non-zero rate of
  aftershocks in the control region.
- **Control $\to$ Treated** ($A_{10}$): `r round(D_par[["A_10"]], 4)` —
  negligible spillover from control to treated.

This asymmetry is consistent with the physics: the high-rate treated
region produces many more potential parents, and some of their aftershocks
spill spatially into the control region.  The near-zero $A_{10}$ means
the control region does not meaningfully trigger events in the treated
region.

## Effect of Label Correction

Comparing naive (A, B) to SEM-corrected (C, D) estimates:

- The SEM slightly reduces the treated background rate and adjusts the
  self-excitation, as some events previously assigned to "treated" by
  location are recognised as aftershocks of control-region parents
  (and vice versa).
- The cross-excitation parameters increase under the SEM (e.g.,
  $A_{01}$: `r round(B_par[["A_01"]], 4)` $\to$ `r round(D_par[["A_01"]], 4)`),
  suggesting that naive location labelling underestimates spillover.


# Configuration

```{r tbl-config}
config_df <- data.frame(
  Setting = c("Magnitude threshold (m₀)",
              "Gutenberg-Richter β",
              "SEM outer iterations",
              "SEM inner iterations",
              "SEM labelling proposals",
              "SEM change factor",
              "Post-treatment window (days)",
              "Test mode"),
  Value = c(cfg$ETAS_M0, cfg$BETA_GR,
            cfg$SEM_N_ITER, cfg$SEM_INNER_ITER,
            cfg$SEM_N_LABELLINGS,
            cfg$SEM_CHANGE_FACTOR,
            paste(cfg$windowT_post, collapse = " – "),
            cfg$TEST_MODE)
)
kable(config_df, align = "lc",
      caption = "Analysis configuration.") %>%
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE)
```


# References

- Ogata, Y. (1988). Statistical models for earthquake occurrences and
  residual analysis for point processes. *JASA*, 83(401), 9–27.
- Zhuang, J., Ogata, Y., & Vere-Jones, D. (2002). Stochastic
  declustering of space-time earthquake occurrences. *JASA*, 97(458),
  369–380.
- Langenbruch, C. & Zoback, M. D. (2016). How will induced seismicity
  in Oklahoma respond to decreased saltwater injection rates? *Science
  Advances*, 2(11), e1601542.

---

*Report generated by PPDisentangle. Results loaded from
`output/oklahoma_results.rds`.*
