---
title: "Oklahoma Induced Seismicity: Causal ETAS Analysis"
subtitle: "Point estimates (pre-bootstrap) · univariate G–J counts + bivariate C/D primary"
author: "PPDisentangle"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 7
    fig-dpi: 150
    self-contained: true
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    fig-width: 7
    fig-height: 5
    fig-dpi: 300
    keep-tex: true
    documentclass: article
    geometry:
      - margin=1in
    echo: false
    include-in-header:
      text: |
        \usepackage{longtable}[=v4.13]
execute:
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: false
SCRIPT_DIR <- getwd()
REPO_ROOT <- if (basename(SCRIPT_DIR) == "oklahoma" &&
                  basename(dirname(SCRIPT_DIR)) == "inst") {
  normalizePath(file.path(SCRIPT_DIR, "..", ".."), winslash = "/", mustWork = FALSE)
} else {
  normalizePath(SCRIPT_DIR, winslash = "/", mustWork = FALSE)
}
# Prefer the local package source so report rendering matches the repo
# (installed PPDisentangle can lag behind and miss helpers like pp_output_path).
if (file.exists(file.path(REPO_ROOT, "DESCRIPTION")) &&
    requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(REPO_ROOT, quiet = TRUE, export_all = FALSE, helpers = FALSE, attach_testthat = FALSE)
} else if (requireNamespace("PPDisentangle", quietly = TRUE)) {
  library(PPDisentangle)
} else {
  stop("PPDisentangle is not available via local source or installed package.")
}
library(spatstat)
library(sf)
library(tigris)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(fig.align = "center", out.width = "100%")
`%||%` <- function(a, b) if (!is.null(a)) a else b

DATA_DIR <- file.path(SCRIPT_DIR,
                       "oklahoma_induced_seismicity_data_regional20150318")
candidate_out_dirs <- unique(c(
  PPDisentangle::pp_output_path("oklahoma", repo_root = REPO_ROOT)
))
explicit_result_file <- trimws(Sys.getenv("OK_RESULTS_FILE", ""))
if (nzchar(explicit_result_file) && !file.exists(explicit_result_file)) {
  stop("OK_RESULTS_FILE was provided but does not exist: ", explicit_result_file)
}
# Prefer the canonical publication snapshot when present. Scenario job RDS
# files (univ/obs/biv) can be newer but are not the default main-report source.
canonical_for_paper <- unlist(lapply(candidate_out_dirs, function(d) {
  fp <- file.path(d, "for_paper.rds")
  if (file.exists(fp)) fp else character(0)
}), use.names = FALSE)
result_patterns <- c("^oklahoma_results(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_job([0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_bootstrap(_[0-9]+)?\\.rds$",
                     "^oklahoma_results_pre_bootstrap_job([0-9]+)?\\.rds$",
                     "^for_paper\\.rds$")
candidate_result_files <- unlist(lapply(candidate_out_dirs, function(d) {
  if (!dir.exists(d)) return(character(0))
  unique(unlist(lapply(result_patterns, function(pat) {
    list.files(d, pattern = pat, full.names = TRUE)
  })))
}), use.names = FALSE)
existing_result_files <- candidate_result_files[file.exists(candidate_result_files)]
if (length(existing_result_files) == 0 && length(canonical_for_paper) == 0) {
  stop("No Oklahoma results RDS file found in expected output directories.")
}
result_file <- if (nzchar(explicit_result_file)) {
  normalizePath(explicit_result_file, winslash = "/", mustWork = TRUE)
} else if (length(canonical_for_paper) > 0) {
  normalizePath(canonical_for_paper[[1]], winslash = "/", mustWork = TRUE)
} else {
  result_meta <- data.frame(
    file = existing_result_files,
    base = basename(existing_result_files),
    mtime = as.POSIXct(file.info(existing_result_files)$mtime, origin = "1970-01-01"),
    stringsAsFactors = FALSE
  )
  result_meta$stage <- ifelse(
    grepl("pre_bootstrap", result_meta$base, fixed = TRUE), "pre_bootstrap", "full"
  )
  result_meta$stage_rank <- match(result_meta$stage, c("pre_bootstrap", "full"))
  job_id <- sub(".*_job([0-9]+).*", "\\1", result_meta$base)
  job_id[!grepl("_job[0-9]+", result_meta$base)] <- NA_character_
  result_meta$run_key <- ifelse(!is.na(job_id), paste0("job:", job_id), "untagged")
  run_latest_idx <- tapply(seq_len(nrow(result_meta)), result_meta$run_key, function(idx) {
    idx[which.max(result_meta$mtime[idx])]
  })
  run_latest_rows <- result_meta[as.integer(run_latest_idx), , drop = FALSE]
  newest_run_key <- run_latest_rows$run_key[which.max(run_latest_rows$mtime)]
  run_pool <- result_meta[result_meta$run_key == newest_run_key, , drop = FALSE]
  best_stage <- max(run_pool$stage_rank, na.rm = TRUE)
  stage_pool <- run_pool[run_pool$stage_rank == best_stage, , drop = FALSE]
  stage_pool$file[which.max(stage_pool$mtime)]
}
OUT_DIR <- dirname(result_file)
candidate_plot_dirs <- unique(file.path(candidate_out_dirs, "plots"))

results <- readRDS(result_file)

# Remap legacy A–L lettering (E/F = all_free primary, duplicate C/D slot, K/L univ homog)
# onto the current A–J scheme (C/D = all_free primary; no K/L).
normalize_oklahoma_fit_letters <- function(res) {
  fn <- res$fits_named
  if (is.null(fn) || !is.list(fn)) return(res)
  e_txt <- if (!is.null(fn$E)) {
    paste(c(fn$E$method, fn$E$label, fn$E$id), collapse = " ")
  } else {
    ""
  }
  e_is_all_free <- grepl("all[_ -]?free", e_txt, ignore.case = TRUE)
  old_scheme <- !is.null(fn$L) || (isTRUE(e_is_all_free) && !is.null(fn$C) && !is.null(fn$E))
  if (!isTRUE(old_scheme)) return(res)
  old <- fn
  remap_one <- function(obj, letter) {
    if (is.null(obj)) return(NULL)
    obj$letter <- letter
    obj
  }
  res$fits_named <- list(
    A = remap_one(old$A, "A"),
    B = remap_one(old$B, "B"),
    C = remap_one(old$E, "C"),
    D = remap_one(old$F, "D"),
    E = remap_one(old$G, "E"),
    F = remap_one(old$H, "F"),
    G = remap_one(old$I, "G"),
    H = remap_one(old$J, "H"),
    I = remap_one(old$K, "I"),
    J = remap_one(old$L, "J")
  )
  boot <- res$bootstrap_ate
  if (is.list(boot)) {
    if (!is.null(boot$fit_E) && is.null(boot$fit_C)) boot$fit_C <- boot$fit_E
    if (!is.null(boot$fit_F) && is.null(boot$fit_D)) boot$fit_D <- boot$fit_F
    res$bootstrap_ate <- boot
  }
  if (!is.null(res$fit_name_map)) {
    res$fit_name_map <- list(
      A = "fits_named$A", B = "fits_named$B",
      C = "fits_named$C", D = "fits_named$D",
      E = "fits_named$E", F = "fits_named$F",
      G = "fits_named$G", H = "fits_named$H",
      I = "fits_named$I", J = "fits_named$J"
    )
  }
  res
}
results <- normalize_oklahoma_fit_letters(results)

meta_for_cfg <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
cfg <- if (!is.null(results$config)) results$config else list()
if (is.null(cfg$ETAS_M0) || length(cfg$ETAS_M0) == 0) {
  cfg$ETAS_M0 <- meta_for_cfg$design$min_mag
}
if (is.null(cfg$FIXED_STRUCTURAL)) cfg$FIXED_STRUCTURAL <- list()

expected_main_inner <- suppressWarnings(as.integer(trimws(Sys.getenv("OK_SEM_INNER_ITER", ""))))
expected_sens_inner <- suppressWarnings(as.integer(trimws(Sys.getenv("OK_SENS_SEM_INNER_ITER", ""))))
expected_boot_inner <- suppressWarnings(as.integer(trimws(Sys.getenv("OK_BOOT_SEM_INNER_ITER", ""))))
if (is.finite(expected_main_inner) && !is.na(expected_main_inner) &&
    !identical(as.integer(cfg$SEM_INNER_ITER), as.integer(expected_main_inner))) {
  stop(sprintf(
    "Report/results mismatch: cfg$SEM_INNER_ITER=%s but OK_SEM_INNER_ITER=%s (file: %s)",
    as.character(cfg$SEM_INNER_ITER), as.character(expected_main_inner), result_file
  ))
}
if (is.finite(expected_sens_inner) && !is.na(expected_sens_inner) &&
    !identical(as.integer(cfg$SENS_SEM_INNER_ITER), as.integer(expected_sens_inner))) {
  stop(sprintf(
    "Report/results mismatch: cfg$SENS_SEM_INNER_ITER=%s but OK_SENS_SEM_INNER_ITER=%s (file: %s)",
    as.character(cfg$SENS_SEM_INNER_ITER), as.character(expected_sens_inner), result_file
  ))
}
if (is.finite(expected_boot_inner) && !is.na(expected_boot_inner) &&
    !identical(as.integer(cfg$BOOT_SEM_INNER_ITER), as.integer(expected_boot_inner))) {
  stop(sprintf(
    "Report/results mismatch: cfg$BOOT_SEM_INNER_ITER=%s but OK_BOOT_SEM_INNER_ITER=%s (file: %s)",
    as.character(cfg$BOOT_SEM_INNER_ITER), as.character(expected_boot_inner), result_file
  ))
}

# Older pre-bootstrap RDS files omitted pp_data/counties/full config; rebuild from CSV.
report_rebuilt <- NULL
if (is.null(results$pp_data) || is.null(results$counties)) {
  support_path <- file.path(SCRIPT_DIR, "oklahoma_report_support.R")
  if (!file.exists(support_path)) {
    stop("Missing ", support_path, " — cannot rebuild pp_data for this checkpoint.")
  }
  source(support_path, local = TRUE)
  report_rebuilt <- oklahoma_report_rebuild_pp_and_counties(DATA_DIR, etas_m0 = cfg$ETAS_M0)
  results$pp_data <- report_rebuilt$pp_data
  results$counties <- report_rebuilt$counties
}
cty <- results$counties
# Back-fill config keys expected by the narrative / tables (narrow legacy checkpoints).
if (is.null(cfg$n_counties) && !is.null(cty$n_counties)) cfg$n_counties <- cty$n_counties
if (is.null(cfg$n_treated) && !is.null(cty$n_treated)) cfg$n_treated <- cty$n_treated
pd0 <- results$pp_data
if (!is.null(pd0)) {
  if (is.null(cfg$n_pre) || (length(cfg$n_pre) == 1L && is.na(cfg$n_pre))) {
    cfg$n_pre <- nrow(as.data.frame(pd0$pp_pre))
  }
  if (is.null(cfg$n_post) || (length(cfg$n_post) == 1L && is.na(cfg$n_post))) {
    cfg$n_post <- nrow(as.data.frame(pd0$pp_post))
  }
  if ((is.null(cfg$n_pre_holdout) || (length(cfg$n_pre_holdout) == 1L && is.na(cfg$n_pre_holdout))) &&
      !is.null(pd0$pp_pre_holdout)) {
    cfg$n_pre_holdout <- nrow(as.data.frame(pd0$pp_pre_holdout))
  }
  if (is.null(cfg$n_pre_total) || (length(cfg$n_pre_total) == 1L && is.na(cfg$n_pre_total))) {
    cfg$n_pre_total <- (cfg$n_pre_holdout %||% 0L) + (cfg$n_pre %||% 0L)
  }
}
# Default ATE horizon if missing (days post-treatment in design)
if (is.null(cfg$ATE_WINDOW_DAYS) || (length(cfg$ATE_WINDOW_DAYS) == 1L && is.na(cfg$ATE_WINDOW_DAYS))) {
  t_star <- as.POSIXct(meta_for_cfg$design$t_star_utc, tz = "UTC")
  post_end <- as.POSIXct(meta_for_cfg$design$post_end_utc, tz = "UTC")
  cfg$ATE_WINDOW_DAYS <- as.integer(ceiling(as.numeric(difftime(post_end, t_star, units = "days"))))
}

bootstrap_absent <- is.null(results$bootstrap_ate) ||
  (is.list(results$bootstrap_ate) &&
     is.null(results$bootstrap_ate$fit_C) &&
     is.null(results$bootstrap_ate$fit_D))
using_pre_bootstrap <- isTRUE(identical(results$metadata$stage, "pre_bootstrap")) ||
  isTRUE(identical(results$metadata$stage, "pre_sensitivity")) ||
  (!is.null(results$checkpoint) && isTRUE(identical(results$checkpoint$stage, "pre_bootstrap"))) ||
  (!is.null(results$checkpoint) && isTRUE(identical(results$checkpoint$stage, "pre_sensitivity"))) ||
  grepl("pre_sensitivity", basename(result_file), fixed = TRUE) ||
  grepl("pre_bootstrap", basename(result_file), fixed = TRUE) ||
  isTRUE(bootstrap_absent)

path_rel_to <- function(path, start) {
  path <- normalizePath(path, winslash = "/", mustWork = FALSE)
  start <- normalizePath(start, winslash = "/", mustWork = FALSE)
  path_seg <- strsplit(path, "/", fixed = TRUE)[[1L]]
  start_seg <- strsplit(start, "/", fixed = TRUE)[[1L]]
  path_seg <- path_seg[nzchar(path_seg)]
  start_seg <- start_seg[nzchar(start_seg)]
  n <- min(length(path_seg), length(start_seg))
  i <- 1L
  while (i <= n && identical(path_seg[[i]], start_seg[[i]])) i <- i + 1L
  rel_parts <- c(rep("..", length(start_seg) - i + 1L), path_seg[i:length(path_seg)])
  if (length(rel_parts) < 1L) "." else paste(rel_parts, collapse = "/")
}

resolve_plot_file <- function(filename) {
  plot_dirs <- unique(c(file.path(OUT_DIR, "plots"), candidate_plot_dirs))
  candidates <- file.path(plot_dirs, filename)
  existing <- candidates[file.exists(candidates)]
  if (length(existing) > 0) {
    abs_path <- existing[which.max(file.info(existing)$mtime)]
    return(path_rel_to(abs_path, SCRIPT_DIR))
  }
  stem <- sub("\\.[^.]+$", "", filename)
  ext <- sub("^.*(\\.[^.]+)$", "\\1", filename)
  ext <- if (identical(ext, filename)) "" else ext
  tagged_pat <- paste0("^", gsub("([.|()\\^{}+$*?\\[\\]\\\\])", "\\\\\\1", stem), "_job[0-9]+", gsub("([.|()\\^{}+$*?\\[\\]\\\\])", "\\\\\\1", ext), "$")
  tagged <- unlist(lapply(plot_dirs, function(d) {
    if (!dir.exists(d)) return(character(0))
    list.files(d, pattern = tagged_pat, full.names = TRUE)
  }), use.names = FALSE)
  tagged <- tagged[file.exists(tagged)]
  if (length(tagged) == 0) return(NA_character_)
  path_rel_to(tagged[which.max(file.info(tagged)$mtime)], SCRIPT_DIR)
}

render_timestamp <- format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")

# KDE/config scalars may be NULL, list-wrapped, or non-numeric after saveRDS / merges.
fmt_scaled_num <- function(x, digits = 2) {
  if (is.null(x)) return("N/A")
  xv <- suppressWarnings(as.numeric(unlist(x, use.names = FALSE)))
  xv <- xv[is.finite(xv)]
  if (length(xv) < 1L) return("N/A")
  as.character(round(xv[[1]], digits))
}

has_fit <- function(fit_name) {
  !is.null(results[[fit_name]])
}

coerce_pp_df <- function(obj) {
  if (is.null(obj)) return(data.frame())

  # Handle spatstat point-pattern objects.
  if (inherits(obj, "ppp")) {
    out <- data.frame(x = obj$x, y = obj$y)
    if (!is.null(obj$marks)) {
      marks_df <- as.data.frame(obj$marks)
      out <- cbind(out, marks_df, stringsAsFactors = FALSE)
    }
    return(out)
  }

  # Standard data-frame-like objects.
  if (is.data.frame(obj)) return(as.data.frame(obj))
  if (is.matrix(obj)) return(as.data.frame(obj))

  # Common list structure with coordinate/time vectors.
  if (is.list(obj) && all(c("x", "y") %in% names(obj))) {
    n <- min(length(obj$x), length(obj$y))
    out <- data.frame(
      x = as.numeric(obj$x)[seq_len(n)],
      y = as.numeric(obj$y)[seq_len(n)]
    )
    if ("t" %in% names(obj)) out$t <- as.numeric(obj$t)[seq_len(min(n, length(obj$t)))]
    for (nm in setdiff(names(obj), c("x", "y", "t"))) {
      v <- obj[[nm]]
      if (length(v) >= n && is.atomic(v)) out[[nm]] <- v[seq_len(n)]
    }
    return(out)
  }

  # Last fallback.
  out <- tryCatch(as.data.frame(obj), error = function(e) NULL)
  if (is.null(out)) {
    stop("Could not coerce point-pattern object to data frame for plotting.")
  }
  out
}
```

# Introduction

**Report rendered at:** `r render_timestamp`

**Results source:** `r basename(result_file)` `r if (using_pre_bootstrap) "(pre-bootstrap checkpoint)" else "(full results)"`

On 18 March 2015 the Oklahoma Corporation Commission (OCC) issued
directive **AOI_20150318**, requiring operators within a defined Area of
Interest (AOI) in central Oklahoma to reduce wastewater disposal volumes.
This report estimates the **all-or-nothing causal effect** of the
directive on earthquake rates using the **PPDisentangle** framework.
Counterfactual evaluations use the **full bivariate** ETAS law
(including cross-excitation).

## Estimand

The target quantity is the **all-or-nothing total treatment effect**
under the fitted bivariate law: the expected difference in total
earthquake counts between a world in which *every* county receives the
directive ($\mathbf{Z}=(1,\dots,1)$) and a world in which *no* county
receives it ($\mathbf{Z}=(0,\dots,0)$):

$$
\Delta_{\text{AoN}}
= \bigl\{
   \mathbb{E}\!\bigl[N(\mathbf{1})\bigr]
  -\mathbb{E}\!\bigl[N(\mathbf{0})\bigr]
\bigr\},
$$

where $N(\mathbf{z})$ denotes the total event count under treatment
assignment $\mathbf{z}$.

## Models

Primary county-partition fits (publication pair):

| Label | Model | Background | Labels |
|:---:|:---|:---|:---|
| **C** | Naive bivariate ETAS | KDE (non-parametric) | Location-based |
| **D** | SEM bivariate ETAS | KDE (non-parametric) | SEM-corrected |

All fits fix $\gamma\equiv 0$ (magnitude-independent spatial scale $d(m)=D$).
**A/B** homogeneous bivariate (all-free except $\gamma$);
**C/D** KDE all-free (primary);
**E/F** KDE with structural $(c,p,D,q)$ fixed from first-half control pre-treatment;
**G/H** univariate+KDE; **I/J** univariate homogeneous (no KDE).
Partition-sensitivity refits use the same C/D (all-free) specifications on
alternative tessellations.

# Study Design

```{r design}
#| label: design
meta <- jsonlite::fromJSON(readLines(file.path(DATA_DIR, "metadata.json")))
```

**Treatment.** OCC directive AOI_20150318: mandatory reduction in
wastewater disposal volumes within the AOI.

**Tessellation.** The `r cty$n_counties` Oklahoma counties (US Census
Bureau via `tigris`). Counties whose centroid falls inside the AOI are
treated.

**Treated counties** (`r cty$n_treated`):
`r paste(cty$treated_names, collapse = ", ")`.

**Temporal windows.**

- Pre-treatment: `r meta$design$pre_start_utc` to `r meta$design$t_star_utc`
- Post-treatment: `r meta$design$t_star_utc` to `r meta$design$post_end_utc`

**Magnitude threshold.** $m_0 =$ `r cfg$ETAS_M0`.

**Event counts.**
- Pre-treatment total: `r if (!is.null(cfg$n_pre_total)) cfg$n_pre_total else cfg$n_pre`
- Pre-treatment holdout (first 50%): `r if (!is.null(cfg$n_pre_holdout)) cfg$n_pre_holdout else NA`
- Pre-treatment used for estimation/carryover (second 50%): `r cfg$n_pre`
- Post-treatment: `r cfg$n_post`
- Total used in SEM estimation (used pre + post): `r cfg$n_pre + cfg$n_post`

# Data

```{r load-data}
#| label: load-data
ev_all <- fread(file.path(DATA_DIR, "events_all.csv"))
t_star_utc <- as.POSIXct(meta$design$t_star_utc, tz = "UTC")
ev_all[, time_utc := as.POSIXct(time_utc, tz = "UTC")]
ev_all[, t_days := as.numeric(difftime(time_utc, t_star_utc, units = "days"))]
ev_all <- ev_all[mag >= cfg$ETAS_M0]
pp_pre  <- results$pp_data$pp_pre
pp_post <- results$pp_data$pp_post
pp_pre <- coerce_pp_df(pp_pre)
pp_post <- coerce_pp_df(pp_post)
if (!("t" %in% names(pp_pre))) pp_pre$t <- numeric(nrow(pp_pre))
if (!("t" %in% names(pp_post))) pp_post$t <- numeric(nrow(pp_post))
```

## Magnitude Distribution

```{r fig-mag}
#| label: fig-mag
#| fig-cap: "Magnitude distribution of events above the completeness threshold."
ggplot(ev_all, aes(x = mag)) +
  geom_histogram(binwidth = 0.1, fill = "#4292c6", color = "white",
                 boundary = cfg$ETAS_M0) +
  geom_vline(xintercept = cfg$ETAS_M0, linetype = "dashed", color = "red") +
  annotate("text", x = cfg$ETAS_M0 + 0.15, y = Inf, vjust = 2,
           label = paste0("m0 = ", cfg$ETAS_M0), color = "red", size = 4) +
  labs(x = "Magnitude", y = "Count") +
  theme_minimal(base_size = 13)
```

## Temporal Profile

```{r fig-temporal}
#| label: fig-temporal
#| fig-cap: "Cumulative event count; dashed line marks the OCC directive."
ev_all[, cum_n := seq_len(.N)]
ggplot(ev_all, aes(x = t_days, y = cum_n)) +
  geom_step(color = "#2171b5", linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("text", x = 5, y = max(ev_all$cum_n) * 0.1,
           label = "OCC directive\n(18 Mar 2015)", color = "red", hjust = 0, size = 3.5) +
  labs(x = "Days since treatment", y = "Cumulative events") +
  theme_minimal(base_size = 13)
```

## County Partition

```{r fig-partition}
#| label: fig-partition
#| fig-cap: "Oklahoma county-level partition. Red counties are treated (centroid inside AOI)."
options(tigris_use_cache = TRUE)
counties_sf <- counties(state = "OK", cb = TRUE, year = 2022)
counties_sf <- st_transform(counties_sf, 5070)
counties_sf <- st_make_valid(counties_sf)

aoi_sf <- st_read(file.path(DATA_DIR, "occ_aoi_layer_2.geojson"), quiet = TRUE)
aoi_sf <- st_transform(aoi_sf, 5070) |> st_make_valid()
aoi_union <- st_union(aoi_sf)

county_centroids <- st_centroid(counties_sf)
inside_aoi <- lengths(st_within(county_centroids, aoi_union)) > 0
counties_sf$treatment <- ifelse(inside_aoi, "treated", "control")

ggplot() +
  geom_sf(data = counties_sf, aes(fill = treatment),
          color = "grey40", linewidth = 0.4) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    name = "Assignment") +
  geom_sf(data = aoi_sf, fill = NA, color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Oklahoma County Partition",
       subtitle = paste0("Treated: ", paste(cty$treated_names, collapse = ", "))) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 9))

counties_km <- counties_sf
st_geometry(counties_km) <- st_geometry(counties_sf) / 1000
st_crs(counties_km) <- NA
```

## Point Patterns

```{r fig-pp}
#| label: fig-pp
#| fig-cap: "Pre-treatment (left) and post-treatment (right) events overlaid on the county partition."
#| fig-height: 5
xs <- c(pp_pre$x, pp_post$x)
ys <- c(pp_pre$y, pp_post$y)
xs <- xs[is.finite(xs)]
ys <- ys[is.finite(ys)]
if (length(xs) == 0L || length(ys) == 0L) {
  bb_map <- sf::st_bbox(counties_km)
  xlims <- as.numeric(bb_map[c("xmin", "xmax")])
  ylims <- as.numeric(bb_map[c("ymin", "ymax")])
} else {
  xlims <- range(xs, na.rm = TRUE)
  ylims <- range(ys, na.rm = TRUE)
}

base_map <- ggplot() +
  geom_sf(data = counties_km, aes(fill = treatment),
          colour = "grey60", linewidth = 0.25) +
  scale_fill_manual(values = c(control = "#deebf7", treated = "#fc9272"),
                    guide = "none") +
  coord_sf(xlim = xlims, ylim = ylims, datum = NA, expand = FALSE) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = "none")

p1 <- base_map +
  geom_point(data = pp_pre, aes(x = x, y = y, alpha = t),
             colour = "#2166ac", size = 0.8) +
  scale_alpha_continuous(range = c(0.15, 0.8), guide = "none") +
  labs(title = "Pre-treatment", x = "X (km)", y = "Y (km)")

pp_post_plot <- pp_post
process_col <- dplyr::case_when(
  "location_process" %in% names(pp_post_plot) ~ "location_process",
  "inferred_process" %in% names(pp_post_plot) ~ "inferred_process",
  "process" %in% names(pp_post_plot) ~ "process",
  TRUE ~ NA_character_
)
nr_post <- nrow(pp_post_plot)
if (nr_post == 0L) {
  pp_post_plot$Process <- factor(character(0), levels = c("control", "treated"))
} else if (is.na(process_col)) {
  pp_post_plot$Process <- factor(rep("control", nr_post), levels = c("control", "treated"))
} else {
  pp_post_plot$Process <- factor(as.character(pp_post_plot[[process_col]]),
                                 levels = c("control", "treated"))
}
p2 <- base_map +
  geom_point(data = pp_post_plot,
             aes(x = x, y = y, colour = Process, alpha = t),
             size = 0.8) +
  scale_colour_manual(values = c(control = "#2166ac", treated = "#b2182b"),
                      guide = "none") +
  scale_alpha_continuous(range = c(0.3, 1), guide = "none") +
  labs(title = "Post-treatment", x = "X (km)", y = "Y (km)")

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

# Model

## ETAS Conditional Intensity

$$
\lambda(t,x,y) = \frac{\mu}{|S|} + \sum_{j:\,t_j<t} A\,e^{\alpha_m(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

with Omori--Utsu temporal kernel $g$ and power-law spatial kernel $f$.

The kernels used in the fits are:

$$
g(\Delta t) = \frac{p-1}{c}\left(1+\frac{\Delta t}{c}\right)^{-p}, \qquad \Delta t > 0
$$

and

$$
f(\Delta x,\Delta y \mid m_j) =
\frac{q-1}{\pi\,D}\,
\left(1+\frac{r^2}{D}\right)^{-q},
\quad r^2 = \Delta x^2 + \Delta y^2.
$$

The spatial scale is **magnitude-independent**: we fix $\gamma\equiv 0$ in
every fit (equivalently $d(m)=D$), so larger events are more productive only
through $\alpha_m>0$, not through a wider spatial kernel.

## Bivariate Extension

$$
\lambda_k(t,x,y) = \frac{\mu_k}{|S_k|}\,W_k(x,y) + \sum_{l\in\{0,1\}}\sum_{\substack{j:\,t_j<t\\Z_j=l}} A_{kl}\,e^{\alpha_{m,kl}(m_j-m_0)}\,g(t-t_j)\,f(x-x_j,y-y_j\mid m_j)
$$

Off-diagonal terms $A_{01}$ (treated$\to$control) and $A_{10}$
(control$\to$treated) capture cross-excitation. For the KDE background
models (C, D, G, H), $W_k(x,y)$ is the normalized kernel density estimate from
control pre-treatment data; for the homogeneous models (A, B, I, J),
$W_k \equiv 1$.

## Branching Matrix, Spectral Radius, and Explosiveness

The bivariate ETAS law admits a Poisson-cluster (branching) representation.
Averaging productivity over magnitudes under a Gutenberg–Richter law with
slope $\beta$ (here $\beta=`r signif(as.numeric(cfg$BETA_GR %||% 2.3), 3)`$)
gives the expected number of type-$k$ offspring of a type-$\ell$ parent,

$$
\eta_{k\ell}
=
A_{k\ell}\,\frac{\beta}{\beta-\alpha_{m,k\ell}},
\qquad
\alpha_{m,k\ell}<\beta
$$

(and $\eta_{k\ell}=\infty$ if $\alpha_{m,k\ell}\ge\beta$). These four
channel means form the **reproduction matrix**

$$
M
=
\begin{pmatrix}
\eta_{00} & \eta_{01} \\
\eta_{10} & \eta_{11}
\end{pmatrix}.
$$

Its **spectral radius** $\rho(M)=\max\{|\lambda|:\lambda\text{ eigenvalue of }M\}$
is the long-run growth factor of a typical cluster:

- $\rho(M)<1$ (**subcritical**): clusters die out almost surely; the expected
  total progeny of one immigrant is finite, equal to
  $\mathbf{1}^\top(I-M)^{-1}\mathbf{e}_\ell$ from a type-$\ell$ immigrant.
  Mean event counts over a finite horizon are well-defined, so all-or-nothing
  DAITE contrasts computed by forward simulation are meaningful.
- $\rho(M)=1$ (**critical**): expected cluster size diverges; finite-horizon
  counts remain random but have heavy tails.
- $\rho(M)>1$ (**explosive / supercritical**): clusters grow exponentially in
  expectation. Simulated catalogues routinely explode, and plug-in causal
  contrasts become numerically unstable and scientifically uninterpretable
  as finite expected savings.

In the independent (univariate) case there is no cross-excitation, so
$M=\mathrm{diag}(\eta_0,\eta_1)$ and $\rho(M)=\max(\eta_0,\eta_1)$: each
margin is checked separately. With cross-excitation, even two subcritical
self-excitation ratios can combine with large off-diagonal $\eta_{01},\eta_{10}$
to push $\rho(M)$ above one.

Fits in this analysis enforce a hard subcritical constraint
$\rho(M)<`r signif(as.numeric(cfg$ETAS_BRANCHING_MAX %||% 0.98), 3)`$
(and $\alpha_m>0$), projecting productivity when needed. Parameter tables
below report $\rho$ and an explosive flag for each fitted law.

## Likelihood Truncation in This Run

For ETAS likelihood evaluation in this project, temporal triggering is
truncated at a finite horizon `t_trunc` (days) when enabled. In code,
parent-child contributions with $\Delta t > t_{\text{trunc}}$ are dropped
from the triggering sum. The Omori part in the **log-intensity** is scaled by

$$
\text{temporal\_norm}
= 1 - \left(1 + \frac{t_{\text{trunc}}}{c}\right)^{-(p-1)},
$$

so $p$ directly controls how much triggering mass remains inside the truncation
window. Larger $p$ concentrates mass near zero lag (less tail), smaller $p$
(still $>1$) leaves heavier temporal tails.

The spatial exponent $q$ does **not** interact with temporal truncation directly.
It continues to govern spatial tail decay in
$f(r\mid m)\propto (1+r^2/d(m))^{-q}$, while truncation only limits temporal
history depth.

```{r tbl-truncation-params}
#| label: tbl-truncation-params
safe_num <- function(x) {
  v <- suppressWarnings(as.numeric(x))
  if (length(v) < 1L) return(NA_real_)
  v <- v[[1]]
  if (is.finite(v)) v else NA_real_
}
fs <- cfg$FIXED_STRUCTURAL
c_val <- if (!is.null(fs)) safe_num(fs[["c"]]) else NA_real_
p_val <- if (!is.null(fs)) safe_num(fs[["p"]]) else NA_real_
q_val <- if (!is.null(fs)) safe_num(fs[["q"]]) else NA_real_
ttr <- safe_num(cfg$SEM_T_TRUNC_DAYS)
rel_target <- safe_num(cfg$SEM_T_TRUNC_REL)
t_source <- if (!is.null(cfg$SEM_T_TRUNC_SOURCE)) as.character(cfg$SEM_T_TRUNC_SOURCE[[1]]) else "N/A"

temporal_mass_retained <- if (is.finite(ttr) && is.finite(c_val) && is.finite(p_val) && c_val > 0 && p_val > 1) {
  1 - (1 + ttr / c_val)^(-(p_val - 1))
} else NA_real_
tail_mass_dropped <- if (is.finite(temporal_mass_retained)) 1 - temporal_mass_retained else NA_real_
kernel_rel_at_ttr <- if (is.finite(ttr) && is.finite(c_val) && is.finite(p_val) && c_val > 0) {
  (1 + ttr / c_val)^(-p_val)
} else NA_real_

trunc_df <- data.frame(
  Setting = c(
    "SEM t_trunc (days)",
    "SEM t_trunc source",
    "Legacy auto-trunc rel level (unused)",
    "Structural c (Omori offset)",
    "Structural p (temporal exponent)",
    "Structural q (spatial exponent)",
    "Temporal mass retained up to t_trunc",
    "Temporal tail mass dropped beyond t_trunc",
    "Relative Omori kernel at t_trunc: ((t_trunc+c)/c)^(-p)"
  ),
  Value = c(
    if (is.finite(ttr)) sprintf("%.4f", ttr) else "N/A",
    t_source,
    if (is.finite(rel_target)) sprintf("%.4f", rel_target) else "N/A",
    if (is.finite(c_val)) sprintf("%.4f", c_val) else "N/A",
    if (is.finite(p_val)) sprintf("%.4f", p_val) else "N/A",
    if (is.finite(q_val)) sprintf("%.4f", q_val) else "N/A",
    if (is.finite(temporal_mass_retained)) sprintf("%.4f", temporal_mass_retained) else "N/A",
    if (is.finite(tail_mass_dropped)) sprintf("%.4f", tail_mass_dropped) else "N/A",
    if (is.finite(kernel_rel_at_ttr)) sprintf("%.6f", kernel_rel_at_ttr) else "N/A"
  ),
  stringsAsFactors = FALSE
)
kable(trunc_df, align = "lc",
      caption = "Temporal truncation settings and implied Omori-tail diagnostics used in likelihood/simulation.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

```{r prep-t-trunc-sensitivity}
#| label: prep-t-trunc-sensitivity
#| include: false
trunc_sens <- results$t_trunc_sensitivity
trunc_sens_df <- NULL
trunc_sens_main_ttr <- suppressWarnings(as.numeric(cfg$SEM_T_TRUNC_DAYS))
if (!is.null(trunc_sens) && length(trunc_sens) >= 1L) {
  rows <- lapply(trunc_sens, function(x) {
    if (is.null(x)) return(NULL)
    data.frame(
      t_trunc_days = suppressWarnings(as.numeric(x$t_trunc_days)),
      status = as.character(x$status %||% NA_character_),
      mean_saved = suppressWarnings(as.numeric(x$ate_mean_saved)),
      sd_saved = suppressWarnings(as.numeric(x$ate_sd_saved)),
      n_sims = suppressWarnings(as.integer(x$ate_n_sims)),
      n_relabel = suppressWarnings(as.integer(x$n_relabel)),
      elapsed_min = suppressWarnings(as.numeric(x$elapsed_sec)) / 60,
      stringsAsFactors = FALSE
    )
  })
  trunc_sens_df <- do.call(rbind, rows[!vapply(rows, is.null, logical(1))])
  if (!is.null(trunc_sens_df) && nrow(trunc_sens_df) >= 1L) {
    trunc_sens_df <- trunc_sens_df[order(trunc_sens_df$t_trunc_days), , drop = FALSE]
  } else {
    trunc_sens_df <- NULL
  }
}
```

```{r tbl-t-trunc-sensitivity}
#| label: tbl-t-trunc-sensitivity
if (is.null(trunc_sens_df)) {
  cat("No Fit D t_trunc sensitivity grid was saved in this results file.\n")
} else {
  # Plain-text caption: nested LaTeX in kable captions can drop HTML tables.
  cap <- "Fit D (SEM biv+KDE) all-or-nothing ATE vs temporal truncation horizon."
  if (is.finite(trunc_sens_main_ttr)) {
    cap <- paste0(cap, sprintf(" Main-run t_trunc=%.3f days.", trunc_sens_main_ttr))
  }
  knitr::kable(
    trunc_sens_df |>
      transmute(
        `t_trunc (days)` = sprintf("%.3g", t_trunc_days),
        Status = status,
        `Mean saved` = ifelse(is.finite(mean_saved), sprintf("%.1f", mean_saved), "NA"),
        `SD saved` = ifelse(is.finite(sd_saved), sprintf("%.1f", sd_saved), "NA"),
        `ATE sims` = ifelse(is.finite(n_sims), as.character(n_sims), "NA"),
        Relabels = ifelse(is.finite(n_relabel), as.character(n_relabel), "NA"),
        `Elapsed (min)` = ifelse(is.finite(elapsed_min), sprintf("%.1f", elapsed_min), "NA")
      ),
    align = "lcccccc",
    caption = cap,
    row.names = FALSE
  ) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

```{r fig-t-trunc-sensitivity}
#| label: fig-t-trunc-sensitivity
#| fig-cap: "Fit D all-or-nothing ATE versus temporal truncation horizon."
if (!is.null(trunc_sens_df) &&
    requireNamespace("ggplot2", quietly = TRUE) &&
    sum(is.finite(trunc_sens_df$mean_saved)) >= 2L) {
  p_trunc <- ggplot2::ggplot(
    trunc_sens_df[is.finite(trunc_sens_df$mean_saved), , drop = FALSE],
    ggplot2::aes(x = t_trunc_days, y = mean_saved)
  ) +
    ggplot2::geom_line(linewidth = 0.8, colour = "#1b9e77") +
    ggplot2::geom_point(size = 2.4, colour = "#1b9e77") +
    ggplot2::geom_errorbar(
      ggplot2::aes(
        ymin = mean_saved - sd_saved,
        ymax = mean_saved + sd_saved
      ),
      width = 0.4, colour = "#1b9e77", alpha = 0.7,
      na.rm = TRUE
    ) +
    ggplot2::labs(
      title = "Fit D ATE vs temporal truncation",
      x = expression(t[trunc]~"(days)"),
      y = "Mean saved events (AoN)"
    ) +
    ggplot2::theme_minimal()
  if (is.finite(trunc_sens_main_ttr)) {
    p_trunc <- p_trunc +
      ggplot2::geom_vline(
        xintercept = trunc_sens_main_ttr, linetype = "dashed", colour = "grey40"
      )
  }
  print(p_trunc)
}
```

## Non-Parametric Background Rate

```{r tbl-kde}
#| label: tbl-kde
kde <- results$kde_info
if (is.null(kde)) {
  kde_df <- data.frame(
    Setting = "KDE summary",
    Value = paste0("Not stored in this results file. ",
                   if (using_pre_bootstrap) {
                     "Re-run oklahoma_analysis.R through the pre-bootstrap checkpoint (current script saves kde_info), or use full oklahoma_results.rds."
                   } else {
                     "Use a full results RDS that includes kde_info."
                   }),
    stringsAsFactors = FALSE
  )
} else {
  kde_holdout <- if (!is.null(kde$n_pre_holdout)) kde$n_pre_holdout else NA
  kde_used <- if (!is.null(kde$n_pre_used)) kde$n_pre_used else NA
  kde_df <- data.frame(
    Setting = c("KDE training events (held-out control pre)",
                "Held-out pre-treatment events (all processes)",
                "Pre-treatment events used for estimation",
                "Bandwidth (km)", "Estimated triggering range (km)", "Method"),
    Value = c(kde$n_training, kde_holdout, kde_used,
              fmt_scaled_num(kde$bw_sigma), fmt_scaled_num(kde$trigger_range_km),
              "Diggle cross-validation"))
}
kable(kde_df, align = "lc",
      caption = "KDE background rate estimation details.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

Only the **second 50%** of pre-treatment events are used for
estimation/carryover in both homogeneous and inhomogeneous SEM fits.
The **first 50%** is held out for fairness; in the inhomogeneous case,
its control events are used to estimate the KDE background via
`bw.diggle`.

```{r fig-kde-bg}
#| label: fig-kde-bg
#| fig-cap: "Estimated inhomogeneous background rate (KDE) from held-out control pre-treatment events."
#| fig-height: 6
bg_plot <- NULL
if (!is.null(results$plots) && !is.null(results$plots$background_rate_kde)) {
  bg_plot <- results$plots$background_rate_kde
}
if (inherits(bg_plot, "gg") || inherits(bg_plot, "ggplot")) {
  # Analysis stores this ggplot on the results RDS; PNG export is optional.
  if (exists("counties_km") && inherits(counties_km, "sf")) {
    county_xy <- tryCatch({
      as.data.frame(sf::st_coordinates(sf::st_cast(counties_km, "MULTILINESTRING")))
    }, error = function(e) NULL)
    if (!is.null(county_xy) && all(c("X", "Y", "L1") %in% names(county_xy))) {
      bg_plot <- bg_plot +
        geom_path(
          data = county_xy,
          aes(x = X, y = Y, group = L1),
          inherit.aes = FALSE,
          colour = "grey25",
          linewidth = 0.2,
          alpha = 0.7
        )
    }
  }
  print(bg_plot)
} else {
  bg_plot_file <- resolve_plot_file("background_rate_kde.png")
  if (is.character(bg_plot_file) && length(bg_plot_file) == 1L &&
      nzchar(bg_plot_file) && file.exists(bg_plot_file)) {
    knitr::include_graphics(bg_plot_file)
  } else {
    plot.new()
    text(0.5, 0.5, "KDE background-rate plot unavailable in this results file.")
  }
}
```

## Fixed Parameters

```{r tbl-fixed}
#| label: tbl-fixed
fixed_cfg <- cfg$FIXED_STRUCTURAL
if (is.null(fixed_cfg) || length(fixed_cfg) == 0) {
  fixed_df <- data.frame(
    Parameter = "None",
    Value = "Estimated (not fixed)",
    Role = "All ETAS parameters are estimated from data",
    stringsAsFactors = FALSE
  )
  cap <- "Fixed structural parameters: none for this run."
} else {
  role_map <- c(
    c = "Omori onset delay",
    p = "Omori decay exponent",
    D = "Spatial scale (magnitude-independent)",
    gamma = "Fixed at 0 (no mag-spatial scaling)",
    q = "Spatial decay exponent"
  )
  fixed_df <- data.frame(
    Parameter = names(fixed_cfg),
    Value = as.numeric(unlist(fixed_cfg)),
    Role = unname(role_map[names(fixed_cfg)]),
    stringsAsFactors = FALSE
  )
  if (!("gamma" %in% fixed_df$Parameter)) {
    g0 <- suppressWarnings(as.numeric(cfg$GAMMA_FIXED %||% 0))
    fixed_df <- rbind(
      fixed_df,
      data.frame(
        Parameter = "gamma",
        Value = if (is.finite(g0)) g0 else 0,
        Role = role_map[["gamma"]],
        stringsAsFactors = FALSE
      )
    )
  }
  cap <- "Fixed structural parameters (gamma=0 in all models)."
}
kable(fixed_df, align = "lcl", caption = cap) |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

```{r tbl-fixed-control-snapshots}
#| label: tbl-fixed-control-snapshots
cs <- results$control_snapshot_fits
if (is.null(cs) || length(cs) == 0) {
  kable(
    data.frame(
      Snapshot = "Control-only breakpoint fits",
      Status = "Not available in this RDS; rerun analysis with updated script.",
      stringsAsFactors = FALSE
    ),
    align = "ll",
    caption = "Control-only ETAS snapshots at three breakpoints (diagnostic only)."
  ) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
} else {
  ord <- c("pre50", "tstar", "end")
  keep <- ord[ord %in% names(cs)]
  if (length(keep) < 1L) keep <- names(cs)
  rows <- lapply(keep, function(k) {
    z <- cs[[k]]
    p <- z$params
    data.frame(
      Snapshot = z$label %||% k,
      `Window end (days)` = if (is.null(z$window_end)) NA_real_ else as.numeric(z$window_end),
      `Events used` = if (is.null(z$n_events)) NA_integer_ else as.integer(z$n_events),
      Status = z$status %||% NA_character_,
      mu = if (is.null(p$mu)) NA_real_ else as.numeric(p$mu),
      A = if (is.null(p$A)) NA_real_ else as.numeric(p$A),
      alpha_m = if (is.null(p$alpha_m)) NA_real_ else as.numeric(p$alpha_m),
      c = if (is.null(p$c)) NA_real_ else as.numeric(p$c),
      p = if (is.null(p$p)) NA_real_ else as.numeric(p$p),
      D = if (is.null(p$D)) NA_real_ else as.numeric(p$D),
      gamma = if (is.null(p$gamma)) NA_real_ else as.numeric(p$gamma),
      q = if (is.null(p$q)) NA_real_ else as.numeric(p$q),
      loglik = if (is.null(z$loglik)) NA_real_ else as.numeric(z$loglik),
      check.names = FALSE
    )
  })
  snap_df <- do.call(rbind, rows)
  kable(
    snap_df,
    digits = 4,
    align = "lcccrrrrrrrrr",
    caption = paste0(
      "Control-only ETAS fits at three breakpoints (all points/regions treated as control). ",
      "Diagnostic for parameter drift over time; SEM initialization machinery remains unchanged."
    )
  ) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                  font_size = 9, latex_options = c("scale_down", "hold_position"))
}
```

## SEM Configuration

Pilot tuning is optional and only targets the Model D SEM relabelling controls.
When enabled, the code runs a parallel grid over:

- `change_factor`
- `change_factor_min_mult`
- `change_factor_max_mult`
- `selection_temperature`

using a short inner SEM run (default 100 iterations), then ranks candidates by:
high final SEM score and low tail accepted/average flips. The chosen settings are
then used for the main county SEM fits in the run.

```{r sem-final-config}
#| label: sem-final-config
fmt <- function(x, default = "N/A") {
  if (length(x) < 1L || is.null(x) || all(is.na(x))) return(default)
  as.character(x[[1]])
}
run_sem_inner_props <- if (!is.null(cfg$SEM_INNER_PROPS)) cfg$SEM_INNER_PROPS else cfg$SEM_N_LABELLINGS
final_cfg_tbl <- data.frame(
  Setting = c(
    "Outer iterations",
    "Inner iterations",
    "SEM inner proposals",
    "SEM labellings",
    "Warm start fixed step",
    "Selection method",
    "Selection temperature",
    "Change factor",
    "Change-factor min multiplier",
    "Change-factor max multiplier",
    "Pilot enabled",
    "Pilot inner iterations",
    "Pilot cores",
    "Pilot max combos"
  ),
  Value = c(
    fmt(cfg$SEM_N_ITER),
    fmt(cfg$SEM_INNER_ITER),
    fmt(run_sem_inner_props),
    fmt(cfg$SEM_N_LABELLINGS),
    fmt(cfg$SEM_WARMSTART_FIXED),
    fmt(cfg$SEM_OPTIM_METHOD),
    fmt(cfg$SEM_SELECTION_TEMPERATURE),
    fmt(cfg$SEM_CHANGE_FACTOR),
    fmt(cfg$SEM_CHANGE_FACTOR_MIN_MULT),
    fmt(cfg$SEM_CHANGE_FACTOR_MAX_MULT),
    fmt(cfg$RUN_SEM_PILOT),
    fmt(cfg$SEM_PILOT_INNER_ITER),
    fmt(cfg$SEM_PILOT_CORES),
    fmt(cfg$SEM_PILOT_MAX_COMBOS)
  ),
  stringsAsFactors = FALSE
)
kable(final_cfg_tbl, align = "lc", caption = "Final SEM settings used for this run.") |>
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

# Results — County Partition

```{r extract}
#| label: extract
get_fit_by_letter <- function(letter, legacy_key = NULL) {
  if (!is.null(results$fits_named) && !is.null(results$fits_named[[letter]])) {
    return(results$fits_named[[letter]])
  }
  if (!is.null(legacy_key) && !is.null(results[[legacy_key]])) return(results[[legacy_key]])
  NULL
}

fit_A_obj <- get_fit_by_letter("A", "fitB")
fit_B_obj <- get_fit_by_letter("B", "fitD")
# Primary all_free C/D; legacy top-level fitE/fitF alias the same content.
fit_C_obj <- get_fit_by_letter("C", "fitE")
fit_D_obj <- get_fit_by_letter("D", "fitF")
fit_E_obj <- get_fit_by_letter("E")
fit_F_obj <- get_fit_by_letter("F")
# Univ+KDE G/H (legacy fitI/fitJ); univ homog I/J (legacy fitK/fitL).
fit_G_obj <- get_fit_by_letter("G", "fitI")
fit_H_obj <- get_fit_by_letter("H", "fitJ")
fit_I_obj <- get_fit_by_letter("I", "fitK")
fit_J_obj <- get_fit_by_letter("J", "fitL")
# Optional K/L fallbacks after normalize (prefer clean A–J above).
fit_K_obj <- get_fit_by_letter("K", "fitK")
if (is.null(fit_K_obj)) fit_K_obj <- fit_I_obj
fit_L_obj <- get_fit_by_letter("L", "fitL")
if (is.null(fit_L_obj)) fit_L_obj <- fit_J_obj

A_par <- if (!is.null(fit_A_obj)) fit_A_obj$params else NULL
B_par <- if (!is.null(fit_B_obj)) fit_B_obj$params else NULL
C_par <- if (!is.null(fit_C_obj)) fit_C_obj$params else NULL
D_par <- if (!is.null(fit_D_obj)) fit_D_obj$params else NULL
E_par <- if (!is.null(fit_E_obj)) fit_E_obj$params else NULL
F_par <- if (!is.null(fit_F_obj)) fit_F_obj$params else NULL
G_par <- if (!is.null(fit_G_obj)) fit_G_obj$params else NULL
H_par <- if (!is.null(fit_H_obj)) fit_H_obj$params else NULL
I_par <- if (!is.null(fit_I_obj)) fit_I_obj$params else NULL
J_par <- if (!is.null(fit_J_obj)) fit_J_obj$params else NULL

scalar_num_or_na <- function(x) {
  v <- suppressWarnings(as.numeric(x))
  if (length(v) < 1L) return(NA_real_)
  v <- v[[1]]
  if (is.finite(v)) v else NA_real_
}

bc <- function(par, p) {
  if (is.null(par)) return(NA_real_)
  if (is.list(par) && !is.null(par$control)) {
    return(scalar_num_or_na(par$control[[p]]))
  }
  if (p == "mu") {
    v <- scalar_num_or_na(par[["mu_0"]])
    if (is.finite(v)) return(v)
    v <- scalar_num_or_na(par[["mu"]])
    if (is.finite(v)) return(v)
  }
  v <- scalar_num_or_na(par[[paste0(p, "_00")]])
  if (is.finite(v)) return(v)
  scalar_num_or_na(par[[p]])
}
bt <- function(par, p) {
  if (is.null(par)) return(NA_real_)
  if (is.list(par) && !is.null(par$treated)) {
    return(scalar_num_or_na(par$treated[[p]]))
  }
  if (p == "mu") {
    v <- scalar_num_or_na(par[["mu_1"]])
    if (is.finite(v)) return(v)
    v <- scalar_num_or_na(par[["mu"]])
    if (is.finite(v)) return(v)
  }
  v <- scalar_num_or_na(par[[paste0(p, "_11")]])
  if (is.finite(v)) return(v)
  scalar_num_or_na(par[[p]])
}
safe_bc <- function(par, p) if (is.null(par)) NA_real_ else bc(par, p)
safe_bt <- function(par, p) if (is.null(par)) NA_real_ else bt(par, p)

etas_trigger_means <- function(par, beta_gr = NULL, t_trunc_days = NULL) {
  if (is.null(par)) {
    return(list(mean_time = NA_real_, mean_time_trunc = NA_real_, mean_distance = NA_real_, mean_scale = NA_real_))
  }
  c0 <- scalar_num_or_na(par[["c"]])
  p0 <- scalar_num_or_na(par[["p"]])
  D0 <- scalar_num_or_na(par[["D"]])
  g0 <- scalar_num_or_na(par[["gamma"]])
  q0 <- scalar_num_or_na(par[["q"]])
  ttr <- scalar_num_or_na(t_trunc_days)
  b0_vec <- suppressWarnings(as.numeric(unlist(beta_gr, use.names = FALSE)))
  b0 <- if (length(b0_vec) < 1L) NA_real_ else scalar_num_or_na(b0_vec[[1]])
  if (!is.finite(b0) || b0 <= 0) b0 <- 2.3

  mean_time <- if (is.finite(c0) && is.finite(p0) && !is.na(c0) && !is.na(p0) && p0 > 2) {
    c0 / (p0 - 2)
  } else {
    Inf
  }
  mean_time_trunc <- NA_real_
  if (is.finite(c0) && is.finite(p0) && !is.na(c0) && !is.na(p0) &&
      c0 > 0 && p0 > 1 && is.finite(ttr) && !is.na(ttr) && ttr > 0) {
    u <- 1 + (ttr / c0)
    if (p0 > 1 && p0 != 2) {
      i0 <- (c0 / (p0 - 1)) * (1 - u^(1 - p0))
      i1 <- c0^2 * ((u^(2 - p0) - 1) / (2 - p0) - (u^(1 - p0) - 1) / (1 - p0))
      if (is.finite(i0) && i0 > 0 && is.finite(i1)) {
        mean_time_trunc <- i1 / i0
      }
    } else if (identical(round(p0, 12), 2.0)) {
      # Closed form limit at p = 2.
      i0 <- c0 * (1 - 1 / u)
      i1 <- c0^2 * (log(u) + (1 / u) - 1)
      if (is.finite(i0) && i0 > 0 && is.finite(i1)) {
        mean_time_trunc <- i1 / i0
      }
    }
    if (!is.finite(mean_time_trunc) || is.na(mean_time_trunc)) {
      num <- tryCatch(
        integrate(function(tt) tt * (1 + tt / c0)^(-p0), lower = 0, upper = ttr)$value,
        error = function(e) NA_real_
      )
      den <- tryCatch(
        integrate(function(tt) (1 + tt / c0)^(-p0), lower = 0, upper = ttr)$value,
        error = function(e) NA_real_
      )
      if (is.finite(num) && is.finite(den) && den > 0) {
        mean_time_trunc <- num / den
      }
    }
  }

  mean_scale <- if (is.finite(D0) && is.finite(g0) && !is.na(D0) && !is.na(g0)) {
    D0 * exp(g0 / b0)
  } else {
    NA_real_
  }

  mean_distance <- if (is.finite(mean_scale) && !is.na(mean_scale) &&
                       is.finite(q0) && !is.na(q0) && q0 > 1.5) {
    sqrt(mean_scale) * (q0 - 1) * sqrt(pi) * gamma(q0 - 1.5) / gamma(q0)
  } else if (is.finite(q0) && !is.na(q0) && q0 <= 1.5) {
    Inf
  } else {
    NA_real_
  }

  list(mean_time = mean_time, mean_time_trunc = mean_time_trunc, mean_distance = mean_distance, mean_scale = mean_scale)
}

model_defs <- list(
  list(id = "A", fit = "fits_named$A", label = "A: Naive", long = "A: Naive Biv", par = A_par),
  list(id = "B", fit = "fits_named$B", label = "B: SEM", long = "B: SEM Biv", par = B_par),
  list(id = "C", fit = "fits_named$C", label = "C: Naive+KDE (all-free)", long = "C: Naive+KDE (all-free)", par = C_par),
  list(id = "D", fit = "fits_named$D", label = "D: SEM+KDE (all-free)", long = "D: SEM+KDE (all-free)", par = D_par),
  list(id = "E", fit = "fits_named$E", label = "E: Naive+KDE (struct-fixed)", long = "E: Naive+KDE (struct-fixed)", par = E_par),
  list(id = "F", fit = "fits_named$F", label = "F: SEM+KDE (struct-fixed)", long = "F: SEM+KDE (struct-fixed)", par = F_par),
  list(id = "G", fit = "fits_named$G", label = "G: Naive univariate+KDE", long = "G: Naive Univ+KDE", par = G_par),
  list(id = "H", fit = "fits_named$H", label = "H: SEM univariate+KDE", long = "H: SEM Univ+KDE", par = H_par),
  list(id = "I", fit = "fits_named$I", label = "I: Naive univariate (no KDE)", long = "I: Naive Univ (no KDE)", par = I_par),
  list(id = "J", fit = "fits_named$J", label = "J: SEM univariate (no KDE)", long = "J: SEM Univ (no KDE)", par = J_par)
)
param_models <- Filter(function(m) !is.null(m$par), model_defs)

fit_fill_colors <- c(
  "A: Naive Bivariate" = "#fc8d62",
  "B: SEM Bivariate" = "#8da0cb",
  "C: Naive Biv+KDE (all-free)" = "#fdb462",
  "D: SEM Biv+KDE (all-free)" = "#80b1d3",
  "E: Naive Biv+KDE (struct-fixed)" = "#fb9a99",
  "F: SEM Biv+KDE (struct-fixed)" = "#b3de69",
  "G: Naive Univ+KDE" = "#66c2a5",
  "H: SEM Univ+KDE" = "#ffd92f",
  "I: Naive Univ (no KDE)" = "#e78ac3",
  "J: SEM Univ (no KDE)" = "#a6d854"
)
fit_line_colors <- c(
  "A: Naive Bivariate" = "#d95f02",
  "B: SEM Bivariate" = "#7570b3",
  "C: Naive Biv+KDE (all-free)" = "#e6550d",
  "D: SEM Biv+KDE (all-free)" = "#3182bd",
  "E: Naive Biv+KDE (struct-fixed)" = "#e31a1c",
  "F: SEM Biv+KDE (struct-fixed)" = "#33a02c",
  "G: Naive Univ+KDE" = "#1b9e77",
  "H: SEM Univ+KDE" = "#b8860b",
  "I: Naive Univ (no KDE)" = "#c51b7d",
  "J: SEM Univ (no KDE)" = "#4d9221"
)

get_sem_metric_trace <- function(fit_obj) {
  if (is.null(fit_obj)) return(NULL)
  sem_obj <- if (!is.null(fit_obj$fit) && !is.null(fit_obj$fit$adaptive)) {
    fit_obj$fit
  } else if (!is.null(fit_obj$sem) && !is.null(fit_obj$sem$adaptive)) {
    fit_obj$sem
  } else {
    NULL
  }
  if (is.null(sem_obj) || is.null(sem_obj$adaptive)) return(NULL)
  ad <- sem_obj$adaptive
  if (!is.null(ad$metrics) && length(ad$metrics) > 0) {
    return(as.numeric(ad$metrics))
  }
  # Backward compatibility: reconstruct selected metric per iter as max over proposal metrics.
  if (!is.null(ad$all_metrics) && length(ad$all_metrics) > 0) {
    vals <- vapply(ad$all_metrics, function(v) {
      if (is.null(v) || length(v) == 0) return(NA_real_)
      max(as.numeric(v), na.rm = TRUE)
    }, numeric(1))
    vals[is.infinite(vals)] <- NA_real_
    return(vals)
  }
  NULL
}
```

## Marginal Parameters

```{r tbl-params}
#| label: tbl-params
beta_proxy <- suppressWarnings(as.numeric(unlist(cfg$BETA_GR, use.names = FALSE)))
beta_proxy <- beta_proxy[is.finite(beta_proxy)]
beta_proxy <- if (length(beta_proxy) > 0) beta_proxy[[1]] else 2.3

eta_exact <- function(mu, A, alpha_m, beta_gr = beta_proxy) {
  if (!all(is.finite(c(mu, A, alpha_m, beta_gr))) || beta_gr <= 0) return(NA_real_)
  if (alpha_m >= beta_gr) return(Inf)
  A * beta_gr / (beta_gr - alpha_m)
}

param_rows <- do.call(rbind, lapply(param_models, function(m) {
  mu_c <- bc(m$par, "mu"); A_c <- bc(m$par, "A"); a_c <- bc(m$par, "alpha_m")
  mu_t <- bt(m$par, "mu"); A_t <- bt(m$par, "A"); a_t <- bt(m$par, "alpha_m")
  eta_c <- eta_exact(mu_c, A_c, a_c)
  eta_t <- eta_exact(mu_t, A_t, a_t)
  rbind(
    data.frame(
      Fit = m$long, Process = "Control",
      mu = mu_c, A = A_c, alpha_m = a_c,
      branch_proxy = eta_c,
      mu_branch_proxy = if (is.finite(mu_c) && is.finite(eta_c)) mu_c * eta_c else NA_real_,
      stringsAsFactors = FALSE
    ),
    data.frame(
      Fit = m$long, Process = "Treated",
      mu = mu_t, A = A_t, alpha_m = a_t,
      branch_proxy = eta_t,
      mu_branch_proxy = if (is.finite(mu_t) && is.finite(eta_t)) mu_t * eta_t else NA_real_,
      stringsAsFactors = FALSE
    )
  )
}))

param_rows$Fit <- factor(param_rows$Fit, levels = vapply(param_models, function(m) m$long, character(1)))
param_rows <- param_rows[order(param_rows$Fit, param_rows$Process), , drop = FALSE]
rownames(param_rows) <- NULL
fit_end_rows <- which(param_rows$Process == "Treated")

tbl_k <- kable(
  param_rows,
  digits = 4,
  align = "llrrrrr",
  col.names = c("Fit", "Process", "mu", "A", "alpha_m", "Branching ratio", "mu * branching ratio"),
  caption = paste0(
    "Marginal ETAS parameters and exact branching ratios by fit. ",
    "Double separators mark fit boundaries. ",
    "The branching ratio is A*beta/(beta-alpha), with beta=",
    signif(beta_proxy, 4), "."
  ),
  booktabs = TRUE,
  linesep = "",
  hline_after = fit_end_rows
)

tbl_k |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE,
                font_size = 8, latex_options = c("scale_down", "hold_position")) |>
  row_spec(fit_end_rows, extra_css = "border-bottom: 3px double #444;")
```

## Cross-excitation

```{r tbl-cross}
#| label: tbl-cross
cn <- c("A_01","alpha_m_01","A_10","alpha_m_10")
cl <- c("$A_{01}$ (treat$\\to$ctrl)","$\\alpha_{m,01}$",
        "$A_{10}$ (ctrl$\\to$treat)","$\\alpha_{m,10}$")
cross_tbl <- data.frame(Parameter = cl)
safe_cross <- function(par, nm) {
  if (is.null(par)) return(NA_real_)
  if (is.list(par) && !is.null(par[[nm]])) return(scalar_num_or_na(par[[nm]]))
  scalar_num_or_na(par[[nm]])
}
for (m in param_models) {
  cross_tbl[[m$id]] <- sapply(cn, function(nm) safe_cross(m$par, nm))
}
rownames(cross_tbl) <- NULL
kable(cross_tbl, digits = 4, escape = FALSE,
      align = paste0("l", paste(rep("c", length(param_models)), collapse = "")),
      col.names = c("", vapply(param_models, function(m) m$label, character(1))),
      caption = "Cross-excitation parameters for available models.") |>
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```

## Parameter Comparison

```{r fig-params}
#| label: fig-params
#| fig-cap: "Marginal parameter estimates across bivariate fits."
#| fig-height: 5
param_rows <- lapply(param_models, function(m) {
  rbind(
    data.frame(Fit = m$long, Process = "Control",
               mu = bc(m$par, "mu"), A = bc(m$par, "A"), alpha_m = bc(m$par, "alpha_m")),
    data.frame(Fit = m$long, Process = "Treated",
               mu = bt(m$par, "mu"), A = bt(m$par, "A"), alpha_m = bt(m$par, "alpha_m"))
  )
})
param_long <- do.call(rbind, param_rows)
pm <- pivot_longer(param_long, cols=c(mu,A,alpha_m), names_to="Parameter", values_to="Value")
pm <- pm[is.finite(pm$Value), , drop = FALSE]
pm$Parameter <- factor(pm$Parameter, levels=c("mu","A","alpha_m"), labels=c("mu","A","alpha[m]"))
pm$Fit <- factor(pm$Fit, levels = vapply(param_models, function(m) m$long, character(1)))
ggplot(pm, aes(x=Fit, y=Value, colour=Process)) +
  geom_point(size=3, position=position_dodge(0.4)) +
  facet_wrap(~Parameter, scales="free_y", labeller=label_parsed) +
  scale_colour_manual(values=c(Control="#377eb8",Treated="#e41a1c")) +
  labs(y="Estimate", x=NULL) +
  theme_minimal(base_size=13) +
  theme(axis.text.x=element_text(angle=30,hjust=1), strip.text=element_text(face="bold"), legend.position="top")
```

## SEM Convergence

```{r fig-flips}
#| label: fig-flips
#| fig-cap: "Accepted SEM confusion-matrix entry counts by inner iteration (county SEM fits)."
#| fig-height: 4
extract_selected_confusion_trace <- function(sem_obj, model_label) {
  ad <- sem_obj$fit$adaptive
  if (is.null(ad) || is.null(ad$class_results) || is.null(ad$all_metrics)) return(NULL)
  if (length(ad$class_results) < 1 || length(ad$all_metrics) < 1) return(NULL)

  offset <- 0L
  rows <- list()
  for (i in seq_along(ad$all_metrics)) {
    metric_i <- ad$all_metrics[[i]]
    n_i <- length(metric_i)
    if (n_i < 1L) next
    best_j <- which.max(as.numeric(metric_i))
    class_idx <- offset + best_j
    if (class_idx > length(ad$class_results)) break
    cm <- ad$class_results[[class_idx]]
    if (is.null(cm)) {
      offset <- offset + n_i
      next
    }
    rows[[length(rows) + 1L]] <- data.frame(
      iteration = i,
      Model = model_label,
      control_as_control = as.numeric(if (!is.null(cm$control_as_control)) cm$control_as_control else 0),
      control_as_treated = as.numeric(if (!is.null(cm$control_as_treated)) cm$control_as_treated else 0),
      treated_as_control = as.numeric(if (!is.null(cm$treated_as_control)) cm$treated_as_control else 0),
      treated_as_treated = as.numeric(if (!is.null(cm$treated_as_treated)) cm$treated_as_treated else 0)
    )
    offset <- offset + n_i
  }
  if (length(rows) < 1L) return(NULL)
  do.call(rbind, rows)
}

trace_rows <- list()
sem_fit_convergence <- list(
  list(obj = if (!is.null(fit_B_obj)) fit_B_obj$fit else NULL, lab = "B: SEM Biv"),
  list(obj = if (!is.null(fit_D_obj)) fit_D_obj$fit else NULL, lab = "D: SEM Biv+KDE (all-free)"),
  list(obj = if (!is.null(fit_F_obj)) fit_F_obj$fit else NULL, lab = "F: SEM Biv+KDE (struct-fixed)"),
  list(obj = if (!is.null(fit_H_obj)) fit_H_obj$fit else NULL, lab = "H: SEM Univ+KDE"),
  list(obj = if (!is.null(fit_J_obj)) fit_J_obj$fit else NULL, lab = "J: SEM Univ (no KDE)")
)
for (nm in sem_fit_convergence) {
  if (is.null(nm$obj) || is.null(nm$obj$adaptive)) next
  trace_df <- extract_selected_confusion_trace(
    list(fit = nm$obj),
    nm$lab
  )
  if (!is.null(trace_df) && nrow(trace_df) > 0) {
    trace_rows[[length(trace_rows) + 1L]] <- trace_df
  } 
}

if (length(trace_rows) > 0) {
  trace_df <- do.call(rbind, trace_rows)
  trace_long <- tidyr::pivot_longer(
    trace_df,
    cols = c(control_as_control, control_as_treated, treated_as_control, treated_as_treated),
    names_to = "entry",
    values_to = "count"
  )
  trace_long$entry <- factor(
    trace_long$entry,
    levels = c("control_as_control", "control_as_treated", "treated_as_control", "treated_as_treated"),
    labels = c("Control -> Control", "Control -> Treated", "Treated -> Control", "Treated -> Treated")
  )
  ggplot(trace_long, aes(x = iteration, y = count, colour = entry)) +
    geom_line(linewidth = 0.9) +
    facet_wrap(~ Model, scales = "free_y") +
    scale_color_manual(name = "Confusion entry", values = c(
      "Control -> Control" = "#1f78b4",
      "Control -> Treated" = "#e31a1c",
      "Treated -> Control" = "#ff7f00",
      "Treated -> Treated" = "#33a02c"
    )) +
    labs(x = "Iteration", y = "Count") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top")
} else {
  plot.new()
  text(0.5, 0.5, "No SEM confusion-matrix trace found for county SEM fits (B/D/F/H/J).")
}
```

```{r fig-sem-adaptive-loglik}
#| label: fig-sem-adaptive-loglik
#| fig-cap: "Adaptive SEM selected-proposal score progression (county SEM fits). This trace is stochastic and is not expected to be monotone."
#| fig-height: 4
sem_metric_rows <- list()
for (info in list(
  list(obj = fit_B_obj, lab = "B: SEM Biv"),
  list(obj = fit_D_obj, lab = "D: SEM Biv+KDE (all-free)"),
  list(obj = fit_F_obj, lab = "F: SEM Biv+KDE (struct-fixed)"),
  list(obj = fit_H_obj, lab = "H: SEM Univ+KDE"),
  list(obj = fit_J_obj, lab = "J: SEM Univ (no KDE)")
)) {
  tr <- get_sem_metric_trace(info$obj)
  if (!is.null(tr) && length(tr) > 0) {
    sem_metric_rows[[length(sem_metric_rows) + 1]] <- data.frame(
      Iteration = seq_along(tr),
      LogLik = as.numeric(tr),
      Model = info$lab
    )
  }
}
if (length(sem_metric_rows) > 0) {
  sem_metric_df <- do.call(rbind, sem_metric_rows)
  sem_metric_df <- sem_metric_df[is.finite(sem_metric_df$LogLik), , drop = FALSE]
  sem_metric_df <- sem_metric_df |>
    dplyr::group_by(.data$Model) |>
    dplyr::arrange(.data$Iteration, .by_group = TRUE) |>
    dplyr::mutate(RunningBest = cummax(.data$LogLik)) |>
    dplyr::ungroup()
  sem_metric_last <- sem_metric_df |>
    dplyr::group_by(.data$Model) |>
    dplyr::filter(.data$Iteration == max(.data$Iteration, na.rm = TRUE)) |>
    dplyr::slice_tail(n = 1) |>
    dplyr::ungroup() |>
    dplyr::mutate(lbl = sprintf("final = %.2f", .data$LogLik))
  ggplot(sem_metric_df, aes(x = Iteration, y = LogLik, colour = Model)) +
    geom_line(linewidth = 0.9) +
    geom_line(aes(y = RunningBest), linewidth = 0.7, linetype = "dashed", alpha = 0.7) +
    geom_text(
      data = sem_metric_last,
      aes(x = Inf, y = -Inf, label = lbl),
      inherit.aes = FALSE,
      hjust = 1.05, vjust = -0.4, size = 3.2, colour = "black"
    ) +
    scale_colour_manual(values = c(
      "B: SEM Biv" = "#6a3d9a",
      "D: SEM Biv+KDE (all-free)" = "#e31a1c",
      "F: SEM Biv+KDE (struct-fixed)" = "#1f78b4",
      "H: SEM Univ+KDE" = "#ff7f00",
      "J: SEM Univ (no KDE)" = "#b8860b"
    )) +
    labs(x = "Adaptive SEM iteration", y = "Selected SEM score (solid) / running best (dashed)") +
    facet_wrap(~Model, scales = "free_y") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
} else {
  plot.new()
  text(0.5, 0.5, "No adaptive SEM log-likelihood trace found in this results file.")
}
```

```{r fig-sem-cumulative-flips}
#| label: fig-sem-cumulative-flips
#| fig-cap: "Per-iteration SEM label flips by relabelling iteration for county SEM fits (faceted by model)."
#| fig-height: 5
flip_rows <- list()
for (info in list(
  list(obj = fit_B_obj, lab = "B: SEM Biv"),
  list(obj = fit_D_obj, lab = "D: SEM Biv+KDE (all-free)"),
  list(obj = fit_F_obj, lab = "F: SEM Biv+KDE (struct-fixed)"),
  list(obj = fit_H_obj, lab = "H: SEM Univ+KDE"),
  list(obj = fit_J_obj, lab = "J: SEM Univ (no KDE)")
)) {
  ad <- if (!is.null(info$obj)) info$obj$fit$adaptive else NULL
  if (is.null(ad) || is.null(ad$average_flips) || is.null(ad$max_metric_flips)) next
  n <- min(length(ad$average_flips), length(ad$max_metric_flips))
  if (n < 1) next
  tmp <- data.frame(
    Iteration = seq_len(n),
    Model = info$lab,
    AverageFlips = as.numeric(ad$average_flips[seq_len(n)]),
    AcceptedFlips = as.numeric(ad$max_metric_flips[seq_len(n)])
  )
  flip_rows[[length(flip_rows) + 1]] <- tmp
}
if (length(flip_rows) > 0) {
  flip_df <- do.call(rbind, flip_rows)
  flip_df <- flip_df |>
    tidyr::pivot_longer(
      cols = c(AverageFlips, AcceptedFlips),
      names_to = "Metric",
      values_to = "Flips"
    )
  flip_df$Metric <- factor(
    flip_df$Metric,
    levels = c("AverageFlips", "AcceptedFlips"),
    labels = c("Average flips", "Accepted flips")
  )
  ggplot(flip_df, aes(x = Iteration, y = Flips, colour = Metric)) +
    geom_line(linewidth = 0.9) +
    facet_wrap(~ Model, scales = "free_y") +
    scale_colour_manual(values = c(
      "Average flips" = "#08519c",
      "Accepted flips" = "#a50f15"
    )) +
    labs(x = "Adaptive SEM iteration", y = "Flips per iteration") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top")
} else {
  plot.new()
  text(0.5, 0.5, "No SEM flip traces found in this results file.")
}
```

```{r tbl-final-loglik-compare}
#| label: tbl-final-loglik-compare
#| tbl-cap: "Final selected objective values from adaptive relabelling traces (higher is better)."
final_ll_rows <- list()

for (info in list(
  list(obj = fit_B_obj, lab = "B: SEM Biv", method = "SEM"),
  list(obj = fit_D_obj, lab = "D: SEM Biv+KDE (all-free)", method = "SEM"),
  list(obj = fit_F_obj, lab = "F: SEM Biv+KDE (struct-fixed)", method = "SEM"),
  list(obj = fit_H_obj, lab = "H: SEM Univ+KDE", method = "SEM"),
  list(obj = fit_J_obj, lab = "J: SEM Univ (no KDE)", method = "SEM")
)) {
  tr <- get_sem_metric_trace(info$obj)
  tr <- as.numeric(tr)
  tr <- tr[is.finite(tr)]
  if (length(tr) > 0) {
    final_ll_rows[[length(final_ll_rows) + 1]] <- data.frame(
      Model = info$lab,
      Method = info$method,
      Iterations = length(tr),
      Final_Objective = tail(tr, 1),
      stringsAsFactors = FALSE
    )
  }
}

if (length(final_ll_rows) > 0) {
  final_ll_tbl <- do.call(rbind, final_ll_rows)
  final_ll_tbl <- final_ll_tbl |>
    dplyr::arrange(.data$Method, .data$Model)
  final_ll_tbl$Final_Objective <- round(final_ll_tbl$Final_Objective, 2)
  kable(final_ll_tbl, align = "llrc") |>
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = FALSE)
}
```

## SEM Relabelling Summary

```{r tbl-cm-summary}
#| label: tbl-cm-summary
build_cm <- function(sem_obj) {
  if (is.null(sem_obj)) return(NULL)
  lab <- sem_obj$adaptive$adaptive_labelling
  post <- lab[lab$t >= 0, ]
  if (is.null(post$location_process) || is.null(post$inferred_process)) return(NULL)
  tab <- table(Location = post$location_process, Inferred = post$inferred_process)
  as.data.frame.matrix(tab)
}

sem_fit_obj <- function(obj) {
  if (is.null(obj)) return(NULL)
  if (!is.null(obj$fit) && !is.null(obj$fit$adaptive)) return(obj$fit)
  if (!is.null(obj$sem) && !is.null(obj$sem$adaptive)) return(obj$sem)
  if (!is.null(obj$adaptive)) return(obj)
  obj$fit %||% obj$sem %||% NULL
}

cm_specs <- list(
  list(cm = build_cm(sem_fit_obj(fit_B_obj)), lab = "B: SEM Bivariate (homogeneous)"),
  list(cm = build_cm(sem_fit_obj(fit_D_obj)), lab = "D: SEM Bivariate+KDE (all-free)"),
  list(cm = build_cm(sem_fit_obj(fit_F_obj)), lab = "F: SEM Bivariate+KDE (struct-fixed)"),
  list(cm = build_cm(sem_fit_obj(fit_H_obj)), lab = "H: SEM Univariate+KDE"),
  list(cm = build_cm(sem_fit_obj(fit_J_obj)), lab = "J: SEM Univariate (homogeneous)")
)

sum_rows <- list()
for (info in cm_specs) {
  cm <- info$cm
  if (is.null(cm)) next
  n_total <- sum(cm)
  ctrl_to_treat <- if ("treated" %in% colnames(cm) && "control" %in% rownames(cm)) {
    cm["control", "treated"]
  } else {
    0
  }
  treat_to_ctrl <- if ("control" %in% colnames(cm) && "treated" %in% rownames(cm)) {
    cm["treated", "control"]
  } else {
    0
  }
  n_flipped <- ctrl_to_treat + treat_to_ctrl
  sum_rows[[length(sum_rows) + 1L]] <- data.frame(
    Model = info$lab,
    `Post events` = n_total,
    `Ctrl to Treat` = ctrl_to_treat,
    `Treat to Ctrl` = treat_to_ctrl,
    `Total relabelled` = n_flipped,
    `Pct relabelled` = paste0(round(n_flipped / n_total * 100, 1), "%"),
    check.names = FALSE
  )
}
if (length(sum_rows) > 0) {
  kable(do.call(rbind, sum_rows),
        caption = "Summary of SEM relabelling for all county SEM fits (B, D, F, H, J).",
        align = "lccccc") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("No SEM relabelling summaries available.\n")
}
```

### Triggering Distance Diagnostics (All Fits)

```{r tbl-trigger-distance-all}
#| label: tbl-trigger-distance-all
diag_rows <- list()
for (fit_info in model_defs) {
  if (is.null(fit_info$par)) next
  m <- etas_trigger_means(fit_info$par, cfg$BETA_GR, cfg$SEM_T_TRUNC_DAYS)
  for (proc in c("Control", "Treated")) {
    diag_rows[[length(diag_rows) + 1]] <- data.frame(
      Model = fit_info$long,
      Process = proc,
      `Mean trigger time (days)` = if (is.finite(m$mean_time)) round(m$mean_time, 3) else Inf,
      `Mean trigger distance (km)` = if (is.finite(m$mean_distance)) round(m$mean_distance, 3) else Inf,
      check.names = FALSE
    )
    if (is.finite(m$mean_time_trunc) && !is.na(m$mean_time_trunc)) {
      diag_rows[[length(diag_rows) + 1]] <- data.frame(
        Model = fit_info$long,
        Process = paste0(proc, " (t<=t_trunc)"),
        `Mean trigger time (days)` = round(m$mean_time_trunc, 3),
        `Mean trigger distance (km)` = if (is.finite(m$mean_distance)) round(m$mean_distance, 3) else Inf,
        check.names = FALSE
      )
    }
  }
}
if (length(diag_rows) > 0) {
  diag_df <- do.call(rbind, diag_rows)
  kable(diag_df, align = "llcc",
        caption = paste0(
          "Model-implied ETAS triggering means across all available fits. ",
          "Rows marked '(t<=t_trunc)' report temporal means under the SEM truncation window. ",
          "Infinite values indicate heavy-tail regimes where the corresponding mean does not exist ",
          "(time: p <= 2, space: q <= 1.5)."
        )) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
}
```

# All-or-Nothing Treatment Effect

## County Partition

```{r tbl-ate-helpers}
#| label: tbl-ate-helpers
#| include: false
n_cty <- suppressWarnings(as.integer(unlist(cfg$n_counties, use.names = FALSE)))
if (length(n_cty) < 1L || !is.finite(n_cty[[1]]) || is.na(n_cty[[1]]) || n_cty[[1]] < 1L) {
  n_cty <- suppressWarnings(as.integer(unlist(results$counties$n_counties, use.names = FALSE)))
}
if (length(n_cty) < 1L || !is.finite(n_cty[[1]]) || is.na(n_cty[[1]]) || n_cty[[1]] < 1L) {
  n_cty <- 1L
}
n_cty <- as.integer(n_cty[[1]])
ate_days <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 365
obs_post_n <- NA_integer_
if (!is.null(results$pp_data) && !is.null(results$pp_data$pp_post)) {
  pp_post_df <- coerce_pp_df(results$pp_data$pp_post)
  if (nrow(pp_post_df) > 0 && "t" %in% names(pp_post_df)) {
    t_num <- suppressWarnings(as.numeric(pp_post_df$t))
    if (all(!is.na(t_num))) {
      obs_post_n <- as.integer(sum(t_num >= 0 & t_num <= ate_days))
    } else {
      obs_post_n <- as.integer(nrow(pp_post_df))
    }
  } else {
    obs_post_n <- as.integer(nrow(pp_post_df))
  }
}

get_total_effect_sim <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  if (!is.finite(n_tiles) || is.na(n_tiles) || n_tiles < 1) n_tiles <- 1
  sim_df <- ate_obj$all_nothing_sim
  if ("total_saved" %in% names(sim_df)) return(sim_df$total_saved)
  if ("total_effect" %in% names(sim_df)) return(sim_df$total_effect)
  sim_df$ATE * n_tiles
}

get_expected_totals <- function(ate_obj, n_tiles = n_cty) {
  if (is.null(ate_obj) || is.null(ate_obj$all_nothing_sim)) return(NULL)
  sim_df <- ate_obj$all_nothing_sim
  if (all(c("c_total", "t_total") %in% names(sim_df))) {
    return(list(c_total = sim_df$c_total, t_total = sim_df$t_total))
  }
  list(c_total = sim_df$c_mean * n_tiles, t_total = sim_df$t_mean * n_tiles)
}

paper_ok_counts_row <- function(method_label, ate_obj, estimand_label = NA_character_) {
  expected <- tryCatch(get_expected_totals(ate_obj, n_tiles = n_cty), error = function(e) NULL)
  if (is.null(expected)) return(NULL)
  c_mean <- mean(expected$c_total, na.rm = TRUE)
  t_mean <- mean(expected$t_total, na.rm = TRUE)
  if (!is.finite(c_mean) || !is.finite(t_mean)) return(NULL)
  data.frame(
    Method = method_label,
    Estimand = estimand_label,
    `E[N(control)]` = round(c_mean, 0),
    `E[N(right)]` = round(t_mean, 0),
    `Delta = control - right` = round(c_mean - t_mean, 0),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}
pick_ate_contrast_early <- function(ate_obj, contrast = c("all_or_nothing", "observed")) {
  contrast <- match.arg(contrast)
  if (is.null(ate_obj)) return(NULL)
  bc <- ate_obj$by_contrast
  if (is.list(bc) && !is.null(bc[[contrast]])) return(bc[[contrast]])
  if (!is.null(ate_obj$contrast) && identical(as.character(ate_obj$contrast)[1], contrast)) {
    return(ate_obj)
  }
  if (identical(contrast, "all_or_nothing") && !is.null(ate_obj$all_nothing_sim)) {
    return(ate_obj)
  }
  NULL
}
```

### Paper-style counts: univariate Naive vs SEM (G–J)

Compact analogue of manuscript Table~13 (`tab:ok_counts`) for univariate fits,
with both all-or-nothing and observed-vs-control estimands when available.

```{r tbl-ok-counts-univ}
#| label: tbl-ok-counts-univ
univ_fit_specs <- list(
  list(lab = "Naive (G, univariate+KDE)", ate = if (!is.null(fit_G_obj)) fit_G_obj$ate else NULL),
  list(lab = "SEM (H, univariate+KDE)", ate = if (!is.null(fit_H_obj)) fit_H_obj$ate else NULL),
  list(lab = "Naive (I, univariate, no KDE)", ate = if (!is.null(fit_I_obj)) fit_I_obj$ate else NULL),
  list(lab = "SEM (J, univariate, no KDE)", ate = if (!is.null(fit_J_obj)) fit_J_obj$ate else NULL)
)
univ_count_rows <- Filter(Negate(is.null), unlist(lapply(univ_fit_specs, function(sp) {
  list(
    paper_ok_counts_row(sp$lab, pick_ate_contrast_early(sp$ate, "all_or_nothing"), "All-or-nothing"),
    paper_ok_counts_row(sp$lab, pick_ate_contrast_early(sp$ate, "observed"), "Observed vs control")
  )
}), recursive = FALSE))
if (length(univ_count_rows) > 0L) {
  univ_counts <- do.call(rbind, univ_count_rows)
  kable(
    univ_counts,
    align = "llrrr",
    caption = paste0(
      "Expected post-treatment earthquake counts for univariate fits I–L under both estimands ",
      "(All-or-nothing: right = treated everywhere; Observed: right = AOI regime; ",
      ate_days, "-day window). Point estimates only; bootstrap uncertainty omitted."
    )
  ) |>
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = FALSE)
} else {
  cat("Univariate I–L ATE simulations are unavailable in this results file.\n")
}
```

### Full model comparison (all available fits)

```{r tbl-ate-biv}
#| label: tbl-ate-biv

biv_rows <- list(
  list(lab = "A: Naive Bivariate", ate = if (!is.null(fit_A_obj)) fit_A_obj$ate else NULL, louis = NULL),
  list(lab = "B: SEM Bivariate",   ate = if (!is.null(fit_B_obj)) fit_B_obj$ate else NULL, louis = if (!is.null(fit_B_obj)) fit_B_obj$louis else NULL),
  list(lab = "C: Naive Biv+KDE (all-free)",   ate = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL, louis = NULL),
  list(lab = "D: SEM Biv+KDE (all-free)",     ate = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL, louis = if (!is.null(fit_D_obj)) fit_D_obj$louis else NULL),
  list(lab = "E: Naive Biv+KDE (struct-fixed)", ate = if (!is.null(fit_E_obj)) fit_E_obj$ate else NULL, louis = NULL),
  list(lab = "F: SEM Biv+KDE (struct-fixed)",   ate = if (!is.null(fit_F_obj)) fit_F_obj$ate else NULL, louis = if (!is.null(fit_F_obj)) fit_F_obj$louis else NULL))
if (!is.null(fit_G_obj) && !is.null(fit_G_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "G: Naive Univ+KDE", ate = fit_G_obj$ate, louis = NULL)
}
if (!is.null(fit_H_obj) && !is.null(fit_H_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "H: SEM Univ+KDE", ate = fit_H_obj$ate, louis = NULL)
}
if (!is.null(fit_I_obj) && !is.null(fit_I_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "I: Naive Univ (no KDE)", ate = fit_I_obj$ate, louis = NULL)
}
if (!is.null(fit_J_obj) && !is.null(fit_J_obj$ate)) {
  biv_rows[[length(biv_rows) + 1]] <- list(lab = "J: SEM Univ (no KDE)", ate = fit_J_obj$ate, louis = NULL)
}

boot_obj <- results$bootstrap_ate
# Primary bootstrap pair is C/D (all_free). Accept legacy E/F target tags.
boot_primary_ids <- c("C", "D")
if (!is.null(boot_obj) && !is.null(boot_obj$config) && !is.null(boot_obj$config$targets)) {
  tgt <- unique(toupper(trimws(as.character(boot_obj$config$targets))))
  if (length(tgt) >= 2 && all(c("C", "D") %in% tgt)) {
    boot_primary_ids <- c("C", "D")
  } else if (length(tgt) >= 2 && all(c("E", "F") %in% tgt)) {
    boot_primary_ids <- c("C", "D")
  }
}
model_label_by_id <- function(id) {
  switch(
    toupper(id),
    "C" = "C: Naive Biv+KDE (all-free)",
    "D" = "D: SEM Biv+KDE (all-free)",
    "E" = "E: Naive Biv+KDE (struct-fixed)",
    "F" = "F: SEM Biv+KDE (struct-fixed)",
    paste0(id, ": model")
  )
}
get_boot_df <- function(id) {
  if (is.null(boot_obj)) return(NULL)
  stable_rows <- function(df) {
    df <- as.data.frame(df)
    if ("is_stable" %in% names(df)) {
      df <- df[!is.na(df$is_stable) & df$is_stable, , drop = FALSE]
    }
    df
  }
  key <- paste0("fit_", toupper(id))
  if (!is.null(boot_obj[[key]]) &&
      !is.null(boot_obj[[key]]$replicate_summary) &&
      nrow(boot_obj[[key]]$replicate_summary) > 0) {
    return(stable_rows(boot_obj[[key]]$replicate_summary))
  }
  # Cross-alias: legacy primary stored under fit_E/fit_F maps to C/D.
  alias <- switch(toupper(id), "C" = "fit_E", "D" = "fit_F", NA_character_)
  if (is.character(alias) && !is.na(alias) &&
      !is.null(boot_obj[[alias]]) &&
      !is.null(boot_obj[[alias]]$replicate_summary) &&
      nrow(boot_obj[[alias]]$replicate_summary) > 0) {
    return(stable_rows(boot_obj[[alias]]$replicate_summary))
  }
  NULL
}
boot_naive_id <- boot_primary_ids[[1]]
boot_sem_id <- boot_primary_ids[[2]]
boot_naive_df <- get_boot_df(boot_naive_id)
boot_sem_df <- get_boot_df(boot_sem_id)
boot_naive_label <- model_label_by_id(boot_naive_id)
boot_sem_label <- model_label_by_id(boot_sem_id)
boot_pair_label <- paste0(boot_naive_id, "/", boot_sem_id)

# Explosive refits are removed in the analysis pipeline before these summaries.
# Retain every finite ATE from the stable remainder; do not apply an ATE-tail
# filter, because a large but stable estimate is part of parameter uncertainty.
retain_boot_vals <- function(x) {
  x <- suppressWarnings(as.numeric(x))
  ok <- is.finite(x)
  list(
    vals = x[ok],
    n_in = length(x),
    n_out = as.integer(sum(!ok)),
    dropped = x[!ok]
  )
}

get_boot_summary <- function(model_label) {
  vals <- NULL
  if (model_label == boot_naive_label && !is.null(boot_naive_df)) vals <- boot_naive_df$ate_total_mean
  if (model_label == boot_sem_label && !is.null(boot_sem_df)) vals <- boot_sem_df$ate_total_mean
  # Compatibility fallback if table labels differ from selected pair labels.
  if (is.null(vals) && model_label == "C: Naive Biv+KDE (all-free)" && !is.null(get_boot_df("C"))) vals <- get_boot_df("C")$ate_total_mean
  if (is.null(vals) && model_label == "D: SEM Biv+KDE (all-free)" && !is.null(get_boot_df("D"))) vals <- get_boot_df("D")$ate_total_mean
  if (is.null(vals)) return(NULL)
  vals <- retain_boot_vals(vals)$vals
  if (length(vals) < 2) return(NULL)
  list(
    sd = stats::sd(vals, na.rm = TRUE),
    lo = as.numeric(stats::quantile(vals, probs = 0.025, na.rm = TRUE)),
    hi = as.numeric(stats::quantile(vals, probs = 0.975, na.rm = TRUE)),
    n = length(vals),
    vals = vals
  )
}

boot_ci_efron_bc <- function(boot_vals, theta_hat, conf = 0.95) {
  x <- suppressWarnings(as.numeric(boot_vals))
  x <- x[is.finite(x)]
  theta_hat <- suppressWarnings(as.numeric(theta_hat))
  if (length(x) < 5L || !is.finite(theta_hat) || is.na(theta_hat)) {
    return(NULL)
  }
  alpha <- (1 - conf) / 2
  q_pct <- stats::quantile(x, probs = c(alpha, 1 - alpha), na.rm = TRUE, names = FALSE)
  p_less <- mean(x < theta_hat, na.rm = TRUE)
  # Guard against p=0 or p=1 leading to infinite z0.
  eps <- 1 / (length(x) + 1)
  p_less <- min(max(p_less, eps), 1 - eps)
  z0 <- stats::qnorm(p_less)
  z_lo <- stats::qnorm(alpha)
  z_hi <- stats::qnorm(1 - alpha)
  a_lo <- stats::pnorm(2 * z0 + z_lo)
  a_hi <- stats::pnorm(2 * z0 + z_hi)
  a_lo <- min(max(a_lo, eps), 1 - eps)
  a_hi <- min(max(a_hi, eps), 1 - eps)
  if (a_lo > a_hi) {
    tmp <- a_lo; a_lo <- a_hi; a_hi <- tmp
  }
  # If BC collapses to a numerically degenerate interval, use percentile fallback.
  if (!is.finite(a_lo) || !is.finite(a_hi) || is.na(a_lo) || is.na(a_hi) ||
      (a_hi - a_lo) < (1 / length(x))) {
    return(list(
      lo = as.numeric(q_pct[[1]]),
      hi = as.numeric(q_pct[[2]]),
      z0 = z0,
      p_less = p_less,
      alpha_lo = alpha,
      alpha_hi = 1 - alpha,
      method = "percentile_fallback"
    ))
  }
  lo <- as.numeric(stats::quantile(x, probs = a_lo, na.rm = TRUE, names = FALSE))
  hi <- as.numeric(stats::quantile(x, probs = a_hi, na.rm = TRUE, names = FALSE))
  list(lo = lo, hi = hi, z0 = z0, p_less = p_less, alpha_lo = a_lo, alpha_hi = a_hi, method = "bc")
}

boot_ci_quantile_length <- function(boot_vals, theta_hat, conf = 0.95, center_stat = c("median", "mean")) {
  center_stat <- match.arg(center_stat)
  x <- suppressWarnings(as.numeric(boot_vals))
  x <- x[is.finite(x)]
  theta_hat <- suppressWarnings(as.numeric(theta_hat))
  if (length(x) < 5L || !is.finite(theta_hat) || is.na(theta_hat)) return(NULL)
  alpha <- (1 - conf) / 2
  q_lo <- as.numeric(stats::quantile(x, probs = alpha, na.rm = TRUE, names = FALSE))
  q_hi <- as.numeric(stats::quantile(x, probs = 1 - alpha, na.rm = TRUE, names = FALSE))
  center <- if (identical(center_stat, "median")) stats::median(x, na.rm = TRUE) else mean(x, na.rm = TRUE)
  if (!is.finite(q_lo) || !is.finite(q_hi) || !is.finite(center)) return(NULL)
  left_len <- center - q_lo
  right_len <- q_hi - center
  if (!is.finite(left_len) || !is.finite(right_len)) return(NULL)
  list(
    lo = theta_hat - left_len,
    hi = theta_hat + right_len,
    q_lo = q_lo,
    q_hi = q_hi,
    center = center
  )
}

format_partition_label <- function(part_id) {
  if (is.na(part_id) || is.null(part_id)) return(part_id)
  if (part_id == "county") return("County")
  if (part_id == "aoi_region") return("AOI region")
  m <- regexec("^grid_([0-9.]+)R$", part_id)
  g <- regmatches(part_id, m)[[1]]
  if (length(g) == 2) {
    mult <- sub("\\.0$", "", g[2])
    return(paste0("Grid (", mult, "x triggering range)"))
  }
  part_id
}

# Prefer contrast-specific payload from $by_contrast when present (new dual-estimand runs).
pick_ate_contrast <- function(ate_obj, contrast = c("all_or_nothing", "observed")) {
  contrast <- match.arg(contrast)
  if (is.null(ate_obj)) return(NULL)
  bc <- ate_obj$by_contrast
  if (is.list(bc) && !is.null(bc[[contrast]])) return(bc[[contrast]])
  if (!is.null(ate_obj$contrast) && identical(as.character(ate_obj$contrast)[1], contrast)) {
    return(ate_obj)
  }
  # Legacy single-contrast objects: only reuse for all_or_nothing if sims are present.
  if (identical(contrast, "all_or_nothing") && !is.null(ate_obj$all_nothing_sim)) {
    return(ate_obj)
  }
  NULL
}

estimand_specs <- list(
  list(
    id = "all_or_nothing",
    label = "All-or-nothing",
    right_lab = "E[N(treated everywhere)]"
  ),
  list(
    id = "observed",
    label = "Observed vs control",
    right_lab = "E[N(actual treatment regime)]"
  )
)

ate_row_for <- function(r, est) {
  ate_use <- pick_ate_contrast(r$ate, est$id)
  if (is.null(ate_use)) return(NULL)
  expected <- tryCatch(get_expected_totals(ate_use, n_tiles = n_cty), error = function(e) NULL)
  sim_totals <- tryCatch(get_total_effect_sim(ate_use, n_tiles = n_cty), error = function(e) NULL)
  c_mean <- if (!is.null(expected) && !is.null(expected$c_total) && length(expected$c_total) > 0) {
    mean(expected$c_total, na.rm = TRUE)
  } else { NA_real_ }
  t_mean <- if (!is.null(expected) && !is.null(expected$t_total) && length(expected$t_total) > 0) {
    mean(expected$t_total, na.rm = TRUE)
  } else { NA_real_ }
  c_sd <- if (!is.null(expected) && !is.null(expected$c_total) && length(expected$c_total) > 1L) {
    sd(expected$c_total, na.rm = TRUE)
  } else { NA_real_ }
  t_sd <- if (!is.null(expected) && !is.null(expected$t_total) && length(expected$t_total) > 1L) {
    sd(expected$t_total, na.rm = TRUE)
  } else { NA_real_ }
  saved_diff_mean <- if (is.finite(c_mean) && is.finite(t_mean)) c_mean - t_mean else NA_real_
  sim_sd_total  <- if (!is.null(sim_totals) && length(sim_totals) > 1L) sd(sim_totals, na.rm = TRUE) else NA_real_
  base_row <- data.frame(
    Model = r$lab,
    Estimand = est$label,
    `Observed N(actual regime)` = obs_post_n,
    `E[N(control everywhere)]` = round(c_mean, 0),
    `E[N(right-hand regime)]` = round(t_mean, 0),
    `SD(control everywhere)` = round(c_sd, 0),
    `SD(right-hand regime)` = round(t_sd, 0),
    `Saved = control - right` = round(saved_diff_mean, 0),
    `SD(saved)` = round(sim_sd_total, 0),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  if (using_pre_bootstrap) return(base_row)
  # Bootstrap uncertainty only for primary all-free pair (C/D) on the run's primary contrast.
  boot_models_for_table <- c("C: Naive Biv+KDE (all-free)", "D: SEM Biv+KDE (all-free)")
  primary_contrast <- tolower(as.character(cfg$ATE_CONTRAST %||% "all_or_nothing")[1])
  boot_sum <- if (r$lab %in% boot_models_for_table && identical(est$id, primary_contrast)) {
    get_boot_summary(r$lab)
  } else {
    NULL
  }
  boot_sd <- if (!is.null(boot_sum)) round(boot_sum$sd, 0) else NA
  boot_ci_q_len <- if (!is.null(boot_sum) && is.finite(saved_diff_mean)) {
    ci <- boot_ci_quantile_length(boot_sum$vals, saved_diff_mean, conf = 0.95, center_stat = "median")
    if (!is.null(ci) && is.finite(ci$lo) && is.finite(ci$hi)) {
      sprintf("[%d, %d]", round(ci$lo, 0), round(ci$hi, 0))
    } else {
      "NA"
    }
  } else {
    "NA"
  }
  base_row$`Bootstrap SD (ATE)` <- boot_sd
  base_row$`95% quantile-length CI` <- boot_ci_q_len
  base_row
}

ate_rows <- unlist(lapply(biv_rows, function(r) {
  lapply(estimand_specs, function(est) ate_row_for(r, est))
}), recursive = FALSE)
ate_rows <- Filter(Negate(is.null), ate_rows)
ate_biv <- if (length(ate_rows) > 0L) do.call(rbind, ate_rows) else NULL
if (!is.null(ate_biv) && nrow(ate_biv) > 0) {
  core_cols <- c(
    "Model",
    "Estimand",
    "Observed N(actual regime)",
    "E[N(control everywhere)]",
    "E[N(right-hand regime)]",
    "SD(control everywhere)",
    "SD(right-hand regime)",
    "Saved = control - right",
    "SD(saved)"
  )
  if (using_pre_bootstrap) {
    ate_biv <- ate_biv[, core_cols, drop = FALSE]
  } else {
    full_cols <- c(core_cols, "Bootstrap SD (ATE)", "95% quantile-length CI")
    ate_biv <- ate_biv[, full_cols, drop = FALSE]
  }
  kable(ate_biv,
        col.names = if (using_pre_bootstrap) {
          c(
            "Model", "Estimand",
            paste0("Observed N(actual regime) / ", ate_days, "d"),
            "E[N(control everywhere)]",
            "E[N(right-hand regime)]",
            "SD(control everywhere)", "SD(right-hand regime)",
            paste0("Saved = control - right / ", n_cty, " counties / ", ate_days, "d"),
            "SD(saved)"
          )
        } else {
          c(
            "Model", "Estimand",
            paste0("Observed N(actual regime) / ", ate_days, "d"),
            "E[N(control everywhere)]",
            "E[N(right-hand regime)]",
            "SD(control everywhere)", "SD(right-hand regime)",
            paste0("Saved = control - right / ", n_cty, " counties / ", ate_days, "d"),
            "SD(saved)", "Bootstrap SD (ATE)", "95% quantile-length CI"
          )
        },
        caption = paste0(
          "Expected counts over ", ate_days, " days for both estimands: ",
          "All-or-nothing (right-hand = treated everywhere) and Observed vs control ",
          "(right-hand = actual AOI regime). Saved = E[N(control everywhere)] - E[N(right-hand)]. ",
          if (using_pre_bootstrap) {
            "Point estimates only (pre-bootstrap checkpoint)."
          } else {
            paste0(
              "Bootstrap SD/CI (when present) apply to the run's primary contrast only; ",
              "quantile-length 95% intervals recenter bootstrap lengths around the point ATE."
            )
          }
        ),
        align = if (using_pre_bootstrap) "llccccccc" else "llccccccccc") |>
    kable_styling(bootstrap_options = c("striped","hover","bordered"), full_width = FALSE)
} else {
  cat(sprintf("ATE table unavailable in this checkpoint stage (%s). Run until pre-bootstrap or full stage for Table 6.1.\n",
              if (!is.null(results$checkpoint) && !is.null(results$checkpoint$stage)) results$checkpoint$stage else "unknown"))
}
```

### Combined parameter comparison (all fits A–J)

```{r tbl-params-all-fits}
#| label: tbl-params-all-fits
# Unified A–J comparison only (pair-wise tables removed).
# Bivariate fits use shared structural (c,p,D,q); univariate fits map A_k -> A_kk,
# leave cross channels blank, and keep margin-specific structural terms.
to_named_num <- function(x) {
  if (is.null(x)) return(setNames(numeric(0), character(0)))
  if (is.list(x)) x <- unlist(x, use.names = TRUE)
  v <- suppressWarnings(as.numeric(x))
  nms <- names(x)
  if (is.null(nms)) nms <- rep("", length(v))
  names(v) <- nms
  v
}
beta_gr_tbl <- suppressWarnings(as.numeric(unlist(cfg$BETA_GR, use.names = FALSE)))
beta_gr_tbl <- beta_gr_tbl[is.finite(beta_gr_tbl) & beta_gr_tbl > 0]
beta_gr_tbl <- if (length(beta_gr_tbl)) beta_gr_tbl[[1]] else 2.3
etas_branch_matrix <- function(par, beta_gr) {
  getp <- function(nm) {
    v <- suppressWarnings(as.numeric(par[[nm]]))
    if (!length(v) || !is.finite(v[[1]])) return(NA_real_)
    v[[1]]
  }
  eta_ij <- function(A, alpha) {
    if (!is.finite(A) || !is.finite(alpha) || !is.finite(beta_gr) || beta_gr <= 0) return(NA_real_)
    gap <- beta_gr - alpha
    if (!is.finite(gap) || gap <= 1e-8) return(Inf)
    A * beta_gr / gap
  }
  matrix(
    c(
      eta_ij(getp("A_00"), getp("alpha_m_00")),
      eta_ij(getp("A_01"), getp("alpha_m_01")),
      eta_ij(getp("A_10"), getp("alpha_m_10")),
      eta_ij(getp("A_11"), getp("alpha_m_11"))
    ),
    nrow = 2, byrow = TRUE,
    dimnames = list(c("from_0", "from_1"), c("to_0", "to_1"))
  )
}
spectral_radius_etas <- function(par, beta_gr) {
  M <- etas_branch_matrix(par, beta_gr)
  if (!all(is.finite(M))) return(Inf)
  max(Re(eigen(M, only.values = TRUE)$values))
}
univ_eta <- function(par, beta_gr) {
  if (is.null(par)) return(NA_real_)
  pv <- to_named_num(par)
  A <- suppressWarnings(as.numeric(pv[["A"]]))
  alpha <- suppressWarnings(as.numeric(pv[["alpha_m"]]))
  if (!length(A) || !length(alpha) || !is.finite(A[[1]]) || !is.finite(alpha[[1]]) ||
      !is.finite(beta_gr) || beta_gr <= 0) {
    return(NA_real_)
  }
  gap <- beta_gr - alpha[[1]]
  if (!is.finite(gap) || gap <= 1e-8) return(Inf)
  A[[1]] * beta_gr / gap
}
extract_univ_margin <- function(fit_obj, margin) {
  if (is.null(fit_obj) || is.null(fit_obj$params)) return(NULL)
  p <- fit_obj$params
  if (!is.null(p[[margin]])) return(p[[margin]])
  NULL
}
fmt_sigfig_all <- function(x, digits = 3L) {
  if (is.null(x) || length(x) < 1L || is.na(x)) return("—")
  if (!is.finite(x)) return("Inf")
  v <- signif(as.numeric(x), digits)
  if (!is.finite(v) || v == 0) return(format(v, scientific = FALSE, trim = TRUE))
  order <- floor(log10(abs(v)))
  dp <- max(0L, as.integer(digits) - order - 1L)
  sprintf(paste0("%.", dp, "f"), v)
}
pull_num <- function(par, nm) {
  if (is.null(par)) return(NA_real_)
  v <- to_named_num(par)
  suppressWarnings(as.numeric(v[[nm]]))
}
biv_col <- function(fit_obj) {
  p <- to_named_num(if (!is.null(fit_obj)) fit_obj$params else NULL)
  if (!length(p)) return(NULL)
  list(
    kind = "biv",
    p = p,
    rho = spectral_radius_etas(p, beta_gr_tbl)
  )
}
univ_col <- function(fit_obj) {
  ctrl <- extract_univ_margin(fit_obj, "control")
  treat <- extract_univ_margin(fit_obj, "treated")
  if (is.null(ctrl) && is.null(treat)) return(NULL)
  eta0 <- univ_eta(ctrl, beta_gr_tbl)
  eta1 <- univ_eta(treat, beta_gr_tbl)
  rho <- max(c(eta0, eta1), na.rm = TRUE)
  if (!is.finite(rho)) rho <- Inf
  list(kind = "univ", ctrl = ctrl, treat = treat, rho = rho)
}
fit_cols <- list(
  A = biv_col(fit_A_obj), B = biv_col(fit_B_obj),
  C = biv_col(fit_C_obj), D = biv_col(fit_D_obj),
  E = biv_col(fit_E_obj), F = biv_col(fit_F_obj),
  G = univ_col(fit_G_obj), H = univ_col(fit_H_obj),
  I = univ_col(fit_I_obj), J = univ_col(fit_J_obj)
)
param_rows_all <- c(
  "mu_0", "mu_1",
  "A_00", "alpha_m_00", "A_11", "alpha_m_11",
  "A_01", "alpha_m_01", "A_10", "alpha_m_10",
  "c_0", "p_0", "D_0", "q_0",
  "c_1", "p_1", "D_1", "q_1",
  "rho", "explosive?"
)
val_for <- function(col, nm) {
  if (is.null(col)) return(NA_real_)
  if (identical(nm, "rho")) return(col$rho)
  if (identical(nm, "explosive?")) {
    return(if (!is.finite(col$rho)) Inf else if (col$rho >= 1) 1 else 0)
  }
  if (identical(col$kind, "biv")) {
    base <- sub("_[01]$", "", nm)
    if (grepl("^[cpDq]_", nm) || nm %in% c("c", "p", "D", "q")) {
      # Shared structural under bivariate law.
      return(pull_num(col$p, base))
    }
    return(pull_num(col$p, nm))
  }
  # univariate
  if (grepl("_0$", nm)) {
    return(pull_num(col$ctrl, sub("_0$", "", nm)))
  }
  if (grepl("_1$", nm)) {
    return(pull_num(col$treat, sub("_1$", "", nm)))
  }
  if (grepl("_(00|11)$", nm)) {
    margin <- if (grepl("_00$", nm)) col$ctrl else col$treat
    return(pull_num(margin, sub("_[01]{2}$", "", nm)))
  }
  # Cross channels absent under independent margins.
  if (grepl("_(01|10)$", nm)) return(NA_real_)
  NA_real_
}
fmt_cell <- function(col, nm) {
  if (identical(nm, "explosive?")) {
    if (is.null(col)) return("—")
    if (!is.finite(col$rho)) return("yes (Inf)")
    if (col$rho >= 1) "yes" else "no"
  } else {
    fmt_sigfig_all(val_for(col, nm), 3L)
  }
}
present_letters <- names(fit_cols)[vapply(fit_cols, function(z) !is.null(z), logical(1))]
if (length(present_letters) > 0L) {
  all_mat <- sapply(present_letters, function(let) {
    vapply(param_rows_all, function(nm) fmt_cell(fit_cols[[let]], nm), character(1))
  }, simplify = "matrix")
  if (is.null(dim(all_mat))) {
    all_mat <- matrix(all_mat, ncol = 1, dimnames = list(param_rows_all, present_letters))
  }
  all_tbl <- data.frame(Parameter = param_rows_all, all_mat, check.names = FALSE)
  kable(
    all_tbl,
    caption = paste0(
      "Combined ETAS parameter comparison across available fits A–J ",
      "(gamma fixed at 0). Bivariate columns share structural (c,p,D,q) across margins; ",
      "univariate columns leave cross-excitation (A_01/A_10) blank and report margin-specific ",
      "structural terms. rho is the offspring spectral radius (bivariate) or max margin branching ",
      "(univariate); explosive means rho >= 1 (beta=", signif(beta_gr_tbl, 3), ")."
    ),
    align = paste0("l", paste(rep("c", length(present_letters)), collapse = ""))
  ) |>
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = FALSE) |>
    scroll_box(width = "100%", height = "640px")
} else {
  cat("Combined A–J parameter table unavailable in this results checkpoint.\n")
}
```

```{r fig-ate-biv}
#| label: fig-ate-biv
#| fig-cap: "Simulation distribution of the all-or-nothing total saved over the post-treatment horizon, facetted by model family with Naive vs SEM overlays."
#| fig-height: 4
biv_sim <- do.call(rbind, lapply(biv_rows, function(r) {
  if (is.null(r$ate)) return(NULL)
  data.frame(Model = r$lab, TotalATE = get_total_effect_sim(r$ate, n_tiles = n_cty))
}))
if (!is.null(biv_sim) && nrow(biv_sim) > 0) {
  assign_role <- function(model_label) {
    if (grepl("^A:|^C:|^E:", model_label)) return("Naive")
    if (grepl("^B:|^D:|^F:", model_label)) return("SEM")
    NA_character_
  }
  assign_pair <- function(model_label) {
    if (grepl("^A:|^B:", model_label)) return("Homogeneous (A/B)")
    if (grepl("^C:|^D:", model_label)) return("KDE all-free (C/D)")
    if (grepl("^E:|^F:", model_label)) return("KDE struct-fixed (E/F)")
    NA_character_
  }
  biv_sim$Role <- vapply(as.character(biv_sim$Model), assign_role, character(1))
  biv_sim$Pair <- vapply(as.character(biv_sim$Model), assign_pair, character(1))
  biv_sim <- biv_sim[!is.na(biv_sim$Role) & !is.na(biv_sim$Pair), , drop = FALSE]
  pair_ok <- vapply(split(biv_sim$Role, biv_sim$Pair), function(x) {
    all(c("Naive", "SEM") %in% unique(x))
  }, logical(1))
  biv_sim <- biv_sim[biv_sim$Pair %in% names(pair_ok)[pair_ok], , drop = FALSE]
  biv_sim$Role <- factor(biv_sim$Role, levels = c("Naive", "SEM"))
  biv_sim$Pair <- factor(
    biv_sim$Pair,
    levels = c("Homogeneous (A/B)", "KDE all-free (C/D)", "KDE struct-fixed (E/F)")
  )
  biv_means <- aggregate(TotalATE ~ Pair + Role, biv_sim, mean)
  ggplot(biv_sim, aes(x = TotalATE, fill = Role)) +
    geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_vline(data = biv_means, aes(xintercept = TotalATE, colour = Role),
               linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
    facet_wrap(~Pair, scales = "free") +
    scale_fill_manual(values = c(Naive = "#fdb462", SEM = "#80b1d3")) +
    scale_colour_manual(values = c(Naive = "#e6550d", SEM = "#3182bd")) +
    labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes saved /" ~ .(ate_days) ~ "days)"),
         y = "Simulation count",
         title = "All-or-Nothing Total Saved: Naive vs SEM by Model Family") +
    theme_minimal(base_size = 13) +
    theme(legend.position = "top",
          plot.title = element_text(hjust = 0.5, face = "bold"))
}
```

```{r ate-scenario-compare-note, results='asis'}
#| echo: false
if (using_pre_bootstrap || is.null(boot_obj)) {
  cat("\n### Four-scenario ATE comparison\n\n")
  cat("Bootstrap scenario panels are omitted for this pre-bootstrap / no-bootstrap render.\n\n")
} else {
  cat("\n### Four-scenario ATE comparison\n\n")
  cat("The **2×2 panel below** overlays Naive (C) and SEM (D) bias-corrected bootstrap ")
  cat("histograms for univariate vs bivariate × all-or-nothing vs observed. ")
  cat("When present, **512-rep** scenario RDS (`*_boot512`) are preferred; otherwise the ")
  cat("largest available bootstrap *n* per scenario is used. ")
  cat("Detailed companion HTML: ")
  scen_html <- file.path(dirname(result_file), "oklahoma_ate_scenarios.html")
  scen_csv <- file.path(dirname(result_file), "ate_scenarios", "scenario_summary.csv")
  if (file.exists(scen_html)) {
    cat(sprintf("[`oklahoma_ate_scenarios.html`](%s)", basename(scen_html)))
  } else {
    cat("`oklahoma_ate_scenarios.html` (render with `render_ate_scenario_compare.R`)")
  }
  cat(".\n\n")
  if (file.exists(scen_csv)) {
    scen_sum <- utils::read.csv(scen_csv, stringsAsFactors = FALSE)
    if (nrow(scen_sum) > 0) {
      knitr::kable(
        scen_sum[, intersect(c("label", "method", "point_E", "point_F",
                               "boot_bc_mean_E", "boot_bc_sd_E",
                               "boot_bc_mean_F", "boot_bc_sd_F"), names(scen_sum)),
                 drop = FALSE],
        digits = 1,
        caption = "ATE scenario summary (bias-corrected bootstrap)."
      ) |>
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
    }
  }
}
```

```{r fig-ate-scenarios-4panel}
#| label: fig-ate-scenarios-4panel
#| fig-cap: "Four-scenario bias-corrected bootstrap comparison (univariate vs bivariate × all-or-nothing vs observed). Each panel overlays Naive (C) and SEM (D), using all finite ATEs retained after failed or explosive refits are excluded. Prefers 512-attempt scenario RDS when available."
#| fig-height: 7.2
#| eval: !expr '!using_pre_bootstrap && !is.null(boot_obj)'
scen_specs <- list(
  list(id = "univ_aon", law = "Univariate", contrast = "All-or-nothing",
       files = c("for_paper_univ_aon_boot512.rds", "for_paper_univ_aon.rds",
                 file.path("ate_scenarios", "univ_aon_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "univ_aon", "for_paper.rds"))),
  list(id = "univ_obs", law = "Univariate", contrast = "Observed vs control",
       files = c("for_paper_univ_obs_boot512.rds", "for_paper_univ_obs.rds",
                 file.path("ate_scenarios", "univ_obs_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "univ_obs", "for_paper.rds"))),
  list(id = "biv_aon", law = "Bivariate", contrast = "All-or-nothing",
       files = c("for_paper_biv_aon_boot512.rds", "for_paper.rds",
                 "for_paper_biv_aon_boot256.rds", "for_paper_biv_aon.rds",
                 file.path("ate_scenarios", "biv_aon_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "biv_aon", "for_paper.rds"))),
  list(id = "biv_obs", law = "Bivariate", contrast = "Observed vs control",
       files = c("for_paper_biv_obs_boot512.rds", "for_paper_biv_obs.rds",
                 "for_paper_biv_obs_boot256.rds",
                 file.path("ate_scenarios", "biv_obs_boot512", "for_paper.rds"),
                 file.path("ate_scenarios", "biv_obs", "for_paper.rds"),
                 file.path("ate_scenarios", "biv_obs_boot256", "for_paper.rds")))
)
scen_primary_boot <- function(boot) {
  if (is.null(boot)) return(list(naive = NULL, sem = NULL))
  list(
    naive = if (!is.null(boot$fit_C)) boot$fit_C else boot$fit_E,
    sem = if (!is.null(boot$fit_D)) boot$fit_D else boot$fit_F
  )
}
scen_primary_fits <- function(fn) {
  if (is.null(fn)) return(list(naive = NULL, sem = NULL))
  list(
    naive = if (!is.null(fn$C)) fn$C else fn$E,
    sem = if (!is.null(fn$D)) fn$D else fn$F
  )
}
resolve_scen_file <- function(rel_files) {
  cands <- unique(unlist(lapply(c(OUT_DIR, candidate_out_dirs), function(d) {
    file.path(d, rel_files)
  }), use.names = FALSE))
  # Prefer the largest bootstrap n among existing candidates.
  existing <- cands[file.exists(cands)]
  if (!length(existing)) return(NA_character_)
  n_boot <- vapply(existing, function(f) {
    r <- tryCatch(normalize_oklahoma_fit_letters(readRDS(f)), error = function(e) NULL)
    blk <- scen_primary_boot(r$bootstrap_ate)$sem
    if (is.null(blk) || is.null(blk$replicate_summary)) return(0L)
    as.integer(nrow(blk$replicate_summary))
  }, integer(1))
  existing[[which.max(n_boot)]]
}
recenter_boot_scen <- function(boot_vals, target_mean) {
  bv <- suppressWarnings(as.numeric(boot_vals))
  if (!is.finite(target_mean) || is.na(target_mean)) return(bv)
  bmean <- mean(bv, na.rm = TRUE)
  if (!is.finite(bmean) || is.na(bmean)) return(bv)
  bv + (target_mean - bmean)
}
point_mean_scen <- function(fit_obj) {
  if (is.null(fit_obj) || is.null(fit_obj$ate)) return(NA_real_)
  vals <- get_total_effect_sim(fit_obj$ate, n_tiles = n_cty)
  vals <- suppressWarnings(as.numeric(vals))
  vals <- vals[is.finite(vals)]
  if (!length(vals)) return(NA_real_)
  mean(vals, na.rm = TRUE)
}
scen_boot_rows <- list()
scen_n_note <- character(0)
for (sp in scen_specs) {
  fp <- resolve_scen_file(sp$files)
  if (is.na(fp)) next
  rr <- tryCatch(normalize_oklahoma_fit_letters(readRDS(fp)), error = function(e) NULL)
  boot_pair <- scen_primary_boot(rr$bootstrap_ate)
  if (is.null(rr) || is.null(boot_pair$naive) || is.null(boot_pair$sem)) next
  be_df <- as.data.frame(boot_pair$naive$replicate_summary)
  bf_df <- as.data.frame(boot_pair$sem$replicate_summary)
  if ("is_stable" %in% names(be_df)) {
    be_df <- be_df[!is.na(be_df$is_stable) & be_df$is_stable, , drop = FALSE]
  }
  if ("is_stable" %in% names(bf_df)) {
    bf_df <- bf_df[!is.na(bf_df$is_stable) & bf_df$is_stable, , drop = FALSE]
  }
  be_raw <- be_df$ate_total_mean
  bf_raw <- bf_df$ate_total_mean
  fit_pair <- scen_primary_fits(rr$fits_named)
  pe <- point_mean_scen(fit_pair$naive)
  pf <- point_mean_scen(fit_pair$sem)
  be <- retain_boot_vals(be_raw)
  bf <- retain_boot_vals(bf_raw)
  attempted_C <- boot_pair$naive$n_attempted %||%
    rr$bootstrap_ate$config$reps %||% be$n_in
  attempted_D <- boot_pair$sem$n_attempted %||%
    rr$bootstrap_ate$config$reps %||% bf$n_in
  scen_n_note <- c(
    scen_n_note,
    sprintf(
      "%s retained C=%d/%d, D=%d/%d",
      sp$id, length(be$vals), attempted_C, length(bf$vals), attempted_D
    )
  )
  scen_boot_rows[[length(scen_boot_rows) + 1L]] <- rbind(
    data.frame(
      Law = sp$law, Contrast = sp$contrast, Model = "Naive (C)",
      TotalATE = recenter_boot_scen(be$vals, pe), stringsAsFactors = FALSE
    ),
    data.frame(
      Law = sp$law, Contrast = sp$contrast, Model = "SEM (D)",
      TotalATE = recenter_boot_scen(bf$vals, pf), stringsAsFactors = FALSE
    )
  )
}
if (length(scen_boot_rows) > 0) {
  scen_boot_df <- do.call(rbind, scen_boot_rows)
  scen_boot_df <- scen_boot_df[is.finite(scen_boot_df$TotalATE), , drop = FALSE]
  scen_boot_df$Law <- factor(scen_boot_df$Law, levels = c("Univariate", "Bivariate"))
  scen_boot_df$Contrast <- factor(scen_boot_df$Contrast, levels = c("All-or-nothing", "Observed vs control"))
  scen_boot_df$Model <- factor(scen_boot_df$Model, levels = c("Naive (C)", "SEM (D)"))
  scen_means <- aggregate(TotalATE ~ Law + Contrast + Model, scen_boot_df, mean)
  ggplot(scen_boot_df, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_vline(data = scen_means, aes(xintercept = TotalATE, colour = Model),
               linewidth = 0.9, show.legend = FALSE) +
    facet_grid(Law ~ Contrast, scales = "free_y") +
    scale_fill_manual(values = c("Naive (C)" = "#fdb462", "SEM (D)" = "#80b1d3")) +
    scale_colour_manual(values = c("Naive (C)" = "#e6550d", "SEM (D)" = "#3182bd")) +
    coord_cartesian(xlim = c(-1000, 1000)) +
    labs(
      x = bquote("Bias-corrected bootstrap " ~ hat(Delta) ~ "(total saved /" ~ .(ate_days) ~ "days)"),
      y = "Bootstrap replicate count",
      fill = "Model",
      title = "Four-scenario ATE bootstrap comparison",
      subtitle = paste(scen_n_note, collapse = " · ")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 9),
      strip.text = element_text(face = "bold")
    )
} else {
  plot.new()
  text(0.5, 0.5, "Four-scenario RDS files not found under oklahoma output.")
}
```

```{r fig-ate-bootstrap-cd}
#| label: fig-ate-bootstrap-cd
#| fig-cap: "Bias-corrected bootstrap replicate distribution of total saved for the selected KDE bootstrap pair, using all finite ATEs retained after failed or explosive refits are excluded. Shape/spread preserved, recentered to the point ATE."
#| fig-height: 4
#| eval: !expr '!using_pre_bootstrap && !is.null(boot_naive_df) && !is.null(boot_sem_df)'
# Labels / centers follow results$config$ATE_METHOD for this RDS.
if (!is.null(boot_naive_df) && !is.null(boot_sem_df) &&
    nrow(boot_naive_df) > 0 && nrow(boot_sem_df) > 0 &&
    "ate_total_mean" %in% names(boot_naive_df) && "ate_total_mean" %in% names(boot_sem_df)) {
  get_fit_obj_by_id <- function(id) {
    switch(
      toupper(id),
      "C" = fit_C_obj,
      "D" = fit_D_obj,
      "E" = fit_E_obj,
      "F" = fit_F_obj,
      NULL
    )
  }
  obs_ate_mean <- function(id) {
    fo <- get_fit_obj_by_id(id)
    if (is.null(fo) || is.null(fo$ate)) return(NA_real_)
    vals <- get_total_effect_sim(fo$ate, n_tiles = n_cty)
    vals <- suppressWarnings(as.numeric(vals))
    vals <- vals[is.finite(vals)]
    if (length(vals) < 1L) return(NA_real_)
    mean(vals, na.rm = TRUE)
  }
  recenter_boot <- function(boot_vals, target_mean) {
    bv <- suppressWarnings(as.numeric(boot_vals))
    if (!is.finite(target_mean) || is.na(target_mean)) return(bv)
    bmean <- mean(bv, na.rm = TRUE)
    if (!is.finite(bmean) || is.na(bmean)) return(bv)
    bv + (target_mean - bmean)
  }

  naive_obs_mean <- obs_ate_mean(boot_naive_id)
  sem_obs_mean <- obs_ate_mean(boot_sem_id)
  naive_filt <- retain_boot_vals(boot_naive_df$ate_total_mean)
  sem_filt <- retain_boot_vals(boot_sem_df$ate_total_mean)
  naive_bc_vals <- recenter_boot(naive_filt$vals, naive_obs_mean)
  sem_bc_vals <- recenter_boot(sem_filt$vals, sem_obs_mean)

  boot_cd <- rbind(
    data.frame(Model = boot_naive_label, TotalATE = naive_bc_vals),
    data.frame(Model = boot_sem_label, TotalATE = sem_bc_vals)
  )
  boot_cd <- boot_cd[is.finite(boot_cd$TotalATE), , drop = FALSE]
  if (nrow(boot_cd) > 0) {
    boot_means <- aggregate(TotalATE ~ Model, boot_cd, mean)
    n_drop <- naive_filt$n_out + sem_filt$n_out
    ggplot(boot_cd, aes(x = TotalATE, fill = Model)) +
      geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
      geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
      geom_vline(data = boot_means, aes(xintercept = TotalATE, colour = Model),
                 linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
      scale_fill_manual(values = setNames(c("#fdb462", "#80b1d3"), c(boot_naive_label, boot_sem_label))) +
      scale_colour_manual(values = setNames(c("#e6550d", "#3182bd"), c(boot_naive_label, boot_sem_label))) +
      coord_cartesian(xlim = c(-1000, 1000)) +
      labs(x = bquote(hat(Delta)[AoN] ~ "(bias-corrected bootstrap replicate mean total saved /" ~ .(ate_days) ~ "days)"),
           y = "Bootstrap replicate count",
           title = paste0("Bias-corrected bootstrap AoN total-saved distribution (", boot_pair_label, ")"),
           subtitle = sprintf(
             "Centers: E=%.1f, F=%.1f; retained stable finite ATEs %d/%d (E) and %d/%d (F); %d non-finite ATE(s) omitted",
             naive_obs_mean, sem_obs_mean,
             length(naive_filt$vals), naive_filt$n_in,
             length(sem_filt$vals), sem_filt$n_in,
             n_drop
           )) +
      theme_minimal(base_size = 13) +
      theme(legend.position = "top",
            plot.title = element_text(hjust = 0.5, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, size = 10))
  } else {
    plot.new()
    text(0.5, 0.5, sprintf("Bias-corrected bootstrap %s replicate means are present but non-finite.", boot_pair_label))
  }
} else {
  plot.new()
  text(0.5, 0.5, sprintf("Bias-corrected bootstrap %s results are unavailable in this results checkpoint.", boot_pair_label))
}
```

```{r fig-ate-bootstrap-biv-obs}
#| label: fig-ate-bootstrap-biv-obs
#| fig-cap: "Bias-corrected bootstrap replicate distribution for the bivariate observed-vs-control-everywhere contrast (C/D), using all finite ATEs retained after failed or explosive refits are excluded. Same display conventions as the all-or-nothing bootstrap figure."
#| fig-height: 4
#| eval: !expr '!using_pre_bootstrap && !is.null(boot_obj)'
# Companion scenario RDS written by launch_ate_scenarios_nesi.sh / bootstrap-only jobs.
biv_obs_candidates <- unique(c(
  file.path(OUT_DIR, "for_paper_biv_obs.rds"),
  file.path(OUT_DIR, "ate_scenarios", "biv_obs", "for_paper.rds"),
  unlist(lapply(candidate_out_dirs, function(d) {
    c(
      file.path(d, "for_paper_biv_obs.rds"),
      file.path(d, "ate_scenarios", "biv_obs", "for_paper.rds")
    )
  }), use.names = FALSE)
))
biv_obs_file <- biv_obs_candidates[file.exists(biv_obs_candidates)][1]
if (!using_pre_bootstrap && !is.na(biv_obs_file) && nzchar(biv_obs_file)) {
  res_obs <- normalize_oklahoma_fit_letters(readRDS(biv_obs_file))
  boot_obs <- res_obs$bootstrap_ate
  fit_C_obs <- if (!is.null(res_obs$fits_named)) res_obs$fits_named$C else NULL
  fit_D_obs <- if (!is.null(res_obs$fits_named)) res_obs$fits_named$D else NULL
  boot_C_blk <- if (!is.null(boot_obs$fit_C)) boot_obs$fit_C else boot_obs$fit_E
  boot_D_blk <- if (!is.null(boot_obs$fit_D)) boot_obs$fit_D else boot_obs$fit_F
  boot_C_obs <- if (!is.null(boot_C_blk)) boot_C_blk$replicate_summary else NULL
  boot_D_obs <- if (!is.null(boot_D_blk)) boot_D_blk$replicate_summary else NULL
  if (!is.null(boot_C_obs) && "is_stable" %in% names(boot_C_obs)) {
    boot_C_obs <- boot_C_obs[
      !is.na(boot_C_obs$is_stable) & boot_C_obs$is_stable, , drop = FALSE
    ]
  }
  if (!is.null(boot_D_obs) && "is_stable" %in% names(boot_D_obs)) {
    boot_D_obs <- boot_D_obs[
      !is.na(boot_D_obs$is_stable) & boot_D_obs$is_stable, , drop = FALSE
    ]
  }
  if (!is.null(boot_C_obs) && !is.null(boot_D_obs) &&
      nrow(boot_C_obs) > 0 && nrow(boot_D_obs) > 0 &&
      "ate_total_mean" %in% names(boot_C_obs) && "ate_total_mean" %in% names(boot_D_obs)) {
    obs_ate_mean_from_fit <- function(fit_obj) {
      if (is.null(fit_obj) || is.null(fit_obj$ate)) return(NA_real_)
      vals <- get_total_effect_sim(fit_obj$ate, n_tiles = n_cty)
      vals <- suppressWarnings(as.numeric(vals))
      vals <- vals[is.finite(vals)]
      if (length(vals) < 1L) return(NA_real_)
      mean(vals, na.rm = TRUE)
    }
    recenter_boot_obs <- function(boot_vals, target_mean) {
      bv <- suppressWarnings(as.numeric(boot_vals))
      if (!is.finite(target_mean) || is.na(target_mean)) return(bv)
      bmean <- mean(bv, na.rm = TRUE)
      if (!is.finite(bmean) || is.na(bmean)) return(bv)
      bv + (target_mean - bmean)
    }
    lab_C <- "C: Naive Biv+KDE (all-free)"
    lab_D <- "D: SEM Biv+KDE (all-free)"
    mean_C_obs <- obs_ate_mean_from_fit(fit_C_obs)
    mean_D_obs <- obs_ate_mean_from_fit(fit_D_obs)
    filt_C_obs <- retain_boot_vals(boot_C_obs$ate_total_mean)
    filt_D_obs <- retain_boot_vals(boot_D_obs$ate_total_mean)
    boot_cd_obs <- rbind(
      data.frame(Model = lab_C, TotalATE = recenter_boot_obs(filt_C_obs$vals, mean_C_obs)),
      data.frame(Model = lab_D, TotalATE = recenter_boot_obs(filt_D_obs$vals, mean_D_obs))
    )
    boot_cd_obs <- boot_cd_obs[is.finite(boot_cd_obs$TotalATE), , drop = FALSE]
    if (nrow(boot_cd_obs) > 0) {
      boot_means_obs <- aggregate(TotalATE ~ Model, boot_cd_obs, mean)
      n_drop_obs <- filt_C_obs$n_out + filt_D_obs$n_out
      ate_days_obs <- suppressWarnings(as.numeric(res_obs$config$ATE_WINDOW_DAYS))
      if (!is.finite(ate_days_obs)) ate_days_obs <- ate_days
      ggplot(boot_cd_obs, aes(x = TotalATE, fill = Model)) +
        geom_histogram(bins = 30, alpha = 0.55, position = "identity") +
        geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
        geom_vline(data = boot_means_obs, aes(xintercept = TotalATE, colour = Model),
                   linetype = "solid", linewidth = 0.9, show.legend = FALSE) +
        scale_fill_manual(values = setNames(c("#fdb462", "#80b1d3"), c(lab_C, lab_D))) +
        scale_colour_manual(values = setNames(c("#e6550d", "#3182bd"), c(lab_C, lab_D))) +
        coord_cartesian(xlim = c(-1000, 1000)) +
        labs(
          x = bquote(hat(Delta)[obs] ~ "(bias-corrected bootstrap replicate mean total saved /" ~ .(ate_days_obs) ~ "days)"),
          y = "Bootstrap replicate count",
          title = "Bias-corrected bootstrap observed-vs-control distribution (C/D, bivariate)",
          subtitle = sprintf(
            "Contrast: bivariate/observed; centers C=%.1f, D=%.1f; retained stable finite ATEs %d/%d (C) and %d/%d (D); omitted %d non-finite; source %s",
            mean_C_obs, mean_D_obs,
            length(filt_C_obs$vals), filt_C_obs$n_in,
            length(filt_D_obs$vals), filt_D_obs$n_in,
            n_drop_obs,
            basename(biv_obs_file)
          )
        ) +
        theme_minimal(base_size = 13) +
        theme(
          legend.position = "top",
          plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, size = 10)
        )
    } else {
      plot.new()
      text(0.5, 0.5, "Bivariate observed bootstrap replicate means are present but non-finite.")
    }
  } else {
    plot.new()
    text(0.5, 0.5, sprintf("Bivariate observed bootstrap C/D summaries missing in %s.", basename(biv_obs_file)))
  }
} else {
  plot.new()
  text(0.5, 0.5, "Bivariate observed scenario RDS not found (expected for_paper_biv_obs.rds).")
}
```

# Partition Sensitivity

The analysis is repeated under alternative spatial partitions to assess
sensitivity to spatial resolution:

- **Grid partitions** with cell diameters equal to 1, 2, and 5
  times the estimated triggering range
  (`r fmt_scaled_num(if (!is.null(results$kde_info)) results$kde_info$trigger_range_km else NULL)` km).
- **AOI region** partition: the AOI polygon defines one treated region,
  everything else is control (2 tiles total).

For each partition, both the naive bivariate + KDE (C-type, all-free)
and SEM bivariate + KDE (D-type, all-free) models are fitted.

```{r fig-partition-all}
#| label: fig-partition-all
#| fig-cap: "Partition maps for all partition schemes evaluated in the run."
partition_ids <- names(results$partition_results)
partition_plot_files <- character(0)
partition_plot_labels <- character(0)
for (pid in partition_ids) {
  pfile <- resolve_plot_file(paste0("partition_map_", pid, ".png"))
  if (file.exists(pfile)) {
    partition_plot_files <- c(partition_plot_files, pfile)
    partition_plot_labels <- c(partition_plot_labels, format_partition_label(pid))
  }
}
if (length(partition_plot_files) > 0) {
  knitr::kable(
    data.frame(`Panels (top to bottom)` = partition_plot_labels),
    align = "l"
  )
  knitr::include_graphics(partition_plot_files)
}
```

```{r tbl-partition-sensitivity}
#| label: tbl-partition-sensitivity
pr <- results$partition_results
ap <- results$ate_partitions

part_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (info$label == "county") {
    list(ate_E = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL,
         ate_F = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL)
  } else {
    ap[[pname]]
  }

  for (fit_type in list(
    list(lab = "C: Naive Biv+KDE (all-free)", par = info$E_params, ate = ate_info$ate_E, sem_obj = NULL),
    list(lab = "D: SEM Biv+KDE (all-free)",   par = info$F_params, ate = ate_info$ate_F, sem_obj = info$semF)
  )) {
    expected <- if (!is.null(fit_type$ate)) {
      get_expected_totals(fit_type$ate, n_tiles = info$n_tiles)
    } else { NULL }
    sim_totals <- if (!is.null(fit_type$ate)) {
      get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles)
    } else { NULL }
    c_mean <- if (!is.null(expected) && !is.null(expected$c_total)) {
      mean(expected$c_total, na.rm = TRUE)
    } else { NA_real_ }
    t_mean <- if (!is.null(expected) && !is.null(expected$t_total)) {
      mean(expected$t_total, na.rm = TRUE)
    } else { NA_real_ }
    t_all_mean <- if (!is.null(fit_type$ate$analytic) && !is.null(fit_type$ate$analytic$eta_treat)) {
      as.numeric(fit_type$ate$analytic$eta_treat) * info$n_tiles
    } else { NA_real_ }
    saved_diff_mean <- if (is.finite(c_mean) && is.finite(t_mean)) c_mean - t_mean else NA_real_
    sim_sd_total <- if (!is.null(sim_totals) && length(sim_totals) > 1L) {
      sd(sim_totals, na.rm = TRUE)
    } else { NA_real_ }
    spill_total <- if (!is.null(fit_type$ate$total_spillover)) {
      fit_type$ate$total_spillover
    } else if (!is.null(fit_type$ate$ATE_spillover)) {
      fit_type$ate$ATE_spillover * info$n_tiles
    } else { NA_real_ }
    n_relabel <- if (!is.null(fit_type$sem_obj) &&
                     !is.null(fit_type$sem_obj$adaptive$adaptive_labelling)) {
      lp <- fit_type$sem_obj$adaptive$adaptive_labelling
      lp <- lp[lp$t >= 0, , drop = FALSE]
      sum(lp$location_process != lp$inferred_process, na.rm = TRUE)
    } else if (identical(fit_type$lab, "D: SEM Biv+KDE (all-free)") &&
               !is.null(info$n_relabel) && is.finite(info$n_relabel)) {
      as.integer(info$n_relabel)
    } else { 0 }

    part_rows[[length(part_rows) + 1]] <- data.frame(
      PartitionID = info$label,
      Partition = format_partition_label(info$label),
      Model = fit_type$lab,
      Tiles = info$n_tiles,
      Treated = info$n_treated,
      `Relabelled points` = n_relabel,
      `E[N(control everywhere)]` = round(c_mean, 0),
      `E[N(actual treatment regime)]` = round(t_mean, 0),
      `E[N(treated everywhere)]` = round(t_all_mean, 0),
      `Saved = control - actual` = round(saved_diff_mean, 0),
      `SD(saved)` = round(sim_sd_total, 0),
      `Spillover` = round(spill_total, 0),
      mu_0 = round(fit_type$par[["mu_0"]], 4),
      mu_1 = round(fit_type$par[["mu_1"]], 4),
      A_0 = round(fit_type$par[["A_00"]], 4),
      A_1 = round(fit_type$par[["A_11"]], 4),
      alpha_m_0 = round(fit_type$par[["alpha_m_00"]], 4),
      alpha_m_1 = round(fit_type$par[["alpha_m_11"]], 4),
      check.names = FALSE)
  }
}

if (length(part_rows) > 0) {
  part_df <- do.call(rbind, part_rows)
  # Keep county first, then all remaining partitions alphabetically by ID.
  part_df$PartitionID <- factor(part_df$PartitionID,
                                levels = c("county",
                                           sort(setdiff(unique(as.character(part_df$PartitionID)),
                                                        "county"))))
  part_df <- part_df[order(part_df$PartitionID, part_df$Model), ]
  part_df$Partition <- factor(as.character(part_df$Partition),
                              levels = unique(as.character(part_df$Partition)))
  part_df_print <- subset(part_df, select = -PartitionID)
  kable(part_df_print, align = "llcccccccccccccccc",
        caption = paste0("Partition comparison over ", ate_days, " days using the same expected-count and ATE metrics as Section 6.1 (plus marginal ETAS parameters).")) |>
    kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE,
                  font_size = 9, latex_options = c("scale_down", "hold_position"))
}
```

```{r fig-partition-params}
#| label: fig-partition-params
#| fig-cap: "Marginal ETAS parameters (mu, A, alpha_m) across partition schemes. County is highlighted to compare against grid-based partitions."
#| fig-height: 6
if (exists("part_df") && nrow(part_df) > 0) {
  plot_df <- part_df
  plot_df$is_county <- as.character(plot_df$Partition) == "County"

  param_long <- rbind(
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "control", mu = plot_df$mu_0, A = plot_df$A_0, alpha_m = plot_df$alpha_m_0),
    data.frame(plot_df[, c("Partition", "Model", "is_county")],
               Process = "treated", mu = plot_df$mu_1, A = plot_df$A_1, alpha_m = plot_df$alpha_m_1)
  )
  param_long <- pivot_longer(param_long, cols = c(mu, A, alpha_m),
                             names_to = "Parameter", values_to = "Value")
  param_long$Process <- factor(param_long$Process, levels = c("control", "treated"))

  param_long$Parameter <- factor(
    param_long$Parameter,
    levels = c("mu", "A", "alpha_m"),
    labels = c(expression(mu), expression(A), expression(alpha[m]))
  )

  ggplot(param_long, aes(x = Partition, y = Value, colour = Model, group = Model)) +
    geom_line(alpha = 0.45) +
    geom_point(aes(size = is_county), alpha = 0.9) +
    scale_size_manual(values = c(`FALSE` = 2.1, `TRUE` = 3.6), guide = "none") +
    scale_colour_manual(values = c("C: Naive Biv+KDE (all-free)" = "#fdb462", "D: SEM Biv+KDE (all-free)" = "#80b1d3")) +
    facet_grid(Parameter ~ Process, scales = "free_y", labeller = label_parsed) +
    labs(x = "Partition scheme", y = "Estimate", colour = "Model") +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      axis.text.x = element_text(angle = 30, hjust = 1),
      strip.text = element_text(face = "bold")
    )
}
```

```{r fig-partition-ate}
#| label: fig-partition-ate
#| fig-cap: "Partition-sensitivity total-saved distributions (same units and x-axis scale as Section 6.1)."
if (exists("biv_sim") && !is.null(biv_sim) && nrow(biv_sim) > 0) {
  biv_xlim <- range(biv_sim$TotalATE, na.rm = TRUE)
} else {
  biv_xlim <- NULL
}

part_sim_rows <- list()
for (pname in names(pr)) {
  info <- pr[[pname]]
  ate_info <- if (info$label == "county") {
    list(ate_E = if (!is.null(fit_C_obj)) fit_C_obj$ate else NULL,
         ate_F = if (!is.null(fit_D_obj)) fit_D_obj$ate else NULL)
  } else {
    ap[[pname]]
  }
  for (fit_type in list(list(lab = "C: Naive Biv+KDE (all-free)", ate = ate_info$ate_E),
                        list(lab = "D: SEM Biv+KDE (all-free)",   ate = ate_info$ate_F))) {
    if (!is.null(fit_type$ate) && !is.null(fit_type$ate$all_nothing_sim)) {
      part_sim_rows[[length(part_sim_rows) + 1]] <- data.frame(
        Partition = info$label,
        Model = fit_type$lab,
        TotalATE = get_total_effect_sim(fit_type$ate, n_tiles = info$n_tiles)
      )
    }
  }
}

if (length(part_sim_rows) > 0) {
  part_sim <- do.call(rbind, part_sim_rows)
  part_sim$PartitionID <- part_sim$Partition
  part_sim$Partition <- vapply(part_sim$Partition, format_partition_label, character(1))
  part_sim$PartitionID <- factor(part_sim$PartitionID,
                                 levels = c("county",
                                            sort(setdiff(unique(as.character(part_sim$PartitionID)),
                                                         "county"))))
  part_order <- unique(as.character(part_sim[order(part_sim$PartitionID), "Partition"]))
  part_sim$Partition <- factor(part_sim$Partition, levels = part_order)
  p_part <- ggplot(part_sim, aes(x = TotalATE, fill = Model)) +
    geom_histogram(bins = 25, alpha = 0.55, position = "identity") +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
    facet_wrap(~Partition, scales = "fixed") +
    scale_fill_manual(values = c("C: Naive Biv+KDE (all-free)" = "#fdb462", "D: SEM Biv+KDE (all-free)" = "#80b1d3")) +
    labs(x = bquote(hat(Delta)[AoN] ~ "(total earthquakes saved /" ~ .(ate_days) ~ "days)"),
         y = "Simulation count") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          strip.text = element_text(face = "bold"))
  if (!is.null(biv_xlim) && all(is.finite(biv_xlim))) {
    p_part <- p_part + coord_cartesian(xlim = biv_xlim)
  }
  print(p_part)
}
```

# Interpretation

```{r summaries, include=FALSE}
n_cty_interp <- cfg$n_counties
ate_days_interp <- if (!is.null(cfg$ATE_WINDOW_DAYS)) cfg$ATE_WINDOW_DAYS else 100

safe_mu_ratio <- function(par) {
  if (is.null(par)) return(NA_real_)
  denom <- safe_bc(par, "mu")
  numer <- safe_bt(par, "mu")
  if (!is.finite(denom) || is.na(denom) || abs(denom) < .Machine$double.eps) return(NA_real_)
  numer / denom
}
mu_ratio_C <- safe_mu_ratio(C_par)
mu_ratio_D <- safe_mu_ratio(D_par)
mu_ratio_E <- safe_mu_ratio(E_par)
mu_ratio_F <- safe_mu_ratio(F_par)
mu_ratio_G <- safe_mu_ratio(G_par)
mu_ratio_H <- safe_mu_ratio(H_par)

total_C_1yr <- if (!is.null(fit_C_obj) && !is.null(fit_C_obj$ate)) round(mean(get_total_effect_sim(fit_C_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_D_1yr <- if (!is.null(fit_D_obj) && !is.null(fit_D_obj$ate)) round(mean(get_total_effect_sim(fit_D_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_E_1yr <- if (!is.null(fit_E_obj) && !is.null(fit_E_obj$ate)) round(mean(get_total_effect_sim(fit_E_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_F_1yr <- if (!is.null(fit_F_obj) && !is.null(fit_F_obj$ate)) round(mean(get_total_effect_sim(fit_F_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_G_1yr <- if (!is.null(fit_G_obj) && !is.null(fit_G_obj$ate)) round(mean(get_total_effect_sim(fit_G_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_H_1yr <- if (!is.null(fit_H_obj) && !is.null(fit_H_obj$ate)) round(mean(get_total_effect_sim(fit_H_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA
total_J_1yr <- if (!is.null(fit_J_obj) && !is.null(fit_J_obj$ate)) round(mean(get_total_effect_sim(fit_J_obj$ate, n_tiles = n_cty_interp), na.rm=TRUE), 0) else NA

effect_direction <- function(x, tol = 10) {
  if (is.na(x)) return("unavailable")
  if (x > tol) return("worse under treatment")
  if (x < -tol) return("beneficial under treatment")
  "close to no detectable difference"
}
```

**Inhomogeneous background rate focus.** Fits C and D (primary all-free)
use a spatially inhomogeneous background estimated from the first half of
control pre-treatment data. This provides an independent estimate of
the baseline seismicity pattern without using post-treatment data.

**Background contrast (inhomogeneous variants).** The treated-to-control
background ratio $\mu_1/\mu_0$ is:
`r paste(sprintf("C=%.1fx, D=%.1fx, E=%.1fx, F=%.1fx, G=%.1fx, H=%.1fx",
                 mu_ratio_C, mu_ratio_D, mu_ratio_E, mu_ratio_F, mu_ratio_G, mu_ratio_H), collapse = "")`.

**All-or-nothing total saved (`r ate_days_interp`-day horizon).**

```{r interp-effects, results='asis'}
effect_lines <- c(
  paste0("- **Fit C (naive+KDE, all-free):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_C_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit D (SEM+KDE, all-free):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_D_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit E (naive+KDE, struct-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_E_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  paste0("- **Fit F (SEM+KDE, struct-fixed):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_F_1yr, "** total earthquakes saved / ", ate_days_interp, " days"),
  {
    sem_totals <- c(total_D_1yr, total_F_1yr)
    sem_totals <- sem_totals[is.finite(sem_totals)]
    if (length(sem_totals) > 0L) {
      paste0("- **Interpretation:** Across KDE bivariate SEM fits (D/F), effects range from **",
             effect_direction(min(sem_totals)),
             "** to **",
             effect_direction(max(sem_totals)),
             "**.")
    } else {
      "- **Interpretation:** KDE SEM variant effects are unavailable in this results file."
    }
  }
)
if (!is.na(total_H_1yr)) {
  effect_lines <- c(effect_lines, paste0("- **Fit H (SEM univariate+KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_H_1yr, "** total earthquakes saved / ", ate_days_interp, " days"))
}
if (!is.na(total_J_1yr)) {
  effect_lines <- c(effect_lines, paste0("- **Fit J (SEM univariate, no KDE):** $\\hat\\Delta_{\\text{AoN}}=$ **", total_J_1yr, "** total earthquakes saved / ", ate_days_interp, " days"))
}
cat(paste(effect_lines, collapse = "\n"), "\n")
```

**Partition sensitivity.** The alternative partition schemes (grids at
different scales and the AOI-as-single-region partition) show how the
treatment effect estimate varies with spatial resolution. Coarser
partitions aggregate more spatial heterogeneity, while finer grids
may better capture localized effects but have fewer events per cell.

# Configuration

```{r tbl-config}
#| label: tbl-config
scalar_chr <- function(x, default = "N/A") {
  if (is.null(x)) return(default)
  xv <- unlist(x, use.names = FALSE)
  if (length(xv) < 1L) return(default)
  x1 <- xv[[1]]
  if (is.na(x1)) return(default)
  as.character(x1)
}
ate_win <- scalar_chr(cfg$ATE_WINDOW_DAYS)
sem_inner_props <- scalar_chr(cfg$SEM_INNER_PROPS %||% cfg$SEM_N_LABELLINGS)
grid_vals <- suppressWarnings(as.numeric(unlist(cfg$grid_diameters, use.names = FALSE)))
grid_vals <- grid_vals[is.finite(grid_vals)]
grid_str <- if (length(grid_vals) > 0L) {
  paste(round(grid_vals, 2), "km", collapse = ", ")
} else { "N/A" }
cdf <- data.frame(
  Setting = c("Magnitude threshold","GR \u03b2","Counties (total / treated)",
              "SEM outer iters","SEM inner iters","SEM inner proposals","SEM labellings",
              "SEM change factor","SEM max relabel step frac","SEM forced refit flip frac","Fitting window (days)",
              "SEM t_trunc (days)","SEM t_trunc source","SEM t_trunc rel level",
              "ATE simulation window (days)","Simulation replicates",
              "Spatial scale D (km)", "Estimated triggering range (km)", "Grid diameters",
              "Test mode"),
  Value = c(scalar_chr(cfg$ETAS_M0), scalar_chr(cfg$BETA_GR),
            paste0(scalar_chr(cfg$n_counties), " / ", scalar_chr(cfg$n_treated)),
            scalar_chr(cfg$SEM_N_ITER), scalar_chr(cfg$SEM_INNER_ITER), sem_inner_props, scalar_chr(cfg$SEM_N_LABELLINGS),
            scalar_chr(cfg$SEM_CHANGE_FACTOR),
            scalar_chr(cfg$SEM_MAX_RELABEL_STEP_FRAC),
            scalar_chr(cfg$SEM_FORCE_PARAM_UPDATE_FLIP_FRAC),
            if (!is.null(cfg$windowT_post)) paste(unlist(cfg$windowT_post, use.names = FALSE), collapse=" \u2013 ") else "N/A",
            scalar_chr(cfg$SEM_T_TRUNC_DAYS),
            scalar_chr(cfg$SEM_T_TRUNC_SOURCE),
            scalar_chr(cfg$SEM_T_TRUNC_REL),
            ate_win, scalar_chr(cfg$ATE_N_SIMS),
            fmt_scaled_num(cfg$spatial_scale_D),
            fmt_scaled_num(if (!is.null(results$kde_info)) results$kde_info$trigger_range_km else NULL),
            grid_str,
            scalar_chr(cfg$TEST_MODE)),
  stringsAsFactors = FALSE
)
kable(cdf, align="lc", caption="Analysis configuration.") |>
  kable_styling(bootstrap_options="striped", full_width=FALSE)
```

# References

- Ogata, Y. (1988). Statistical models for earthquake occurrences.
  *JASA*, 83(401), 9--27.
- Zhuang, J., Ogata, Y. & Vere-Jones, D. (2002). Stochastic
  declustering of space-time earthquake occurrences. *JASA*, 97(458),
  369--380.
- Helmstetter, A. & Sornette, D. (2003). Importance of direct and
  indirect triggered seismicity in the ETAS model of seismicity.
  *Geophysical Research Letters*, 30(11), 1576.
- Langenbruch, C. & Zoback, M. D. (2016). How will induced seismicity
  in Oklahoma respond to decreased saltwater injection rates? *Science
  Advances*, 2(11), e1601542.

---
*Report generated by PPDisentangle.*
